### 2015-11-18 acta_manuscript.tex
first commit of journal

\documentclass[preprint]{elsarticle}

\usepackage{acta_astronautica_packages}
\input{macros}

\journal{Acta Astronautica}

\begin{document}

\begin{frontmatter}

\title{Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem}

%% Group authors per affiliation:
\author{Shankar Kulumani\fnref{kulumanifootnote}}
\ead{skulumani@gwu.edu}
\address{George Washington University, 800 22nd St NW, Washington, DC 20052}

\author{Taeyoung Lee\fnref{leefootnote}}
\ead{tylee@gwu.edu}
\address{George Washington University, 800 22nd St NW, Washington, DC 20052}

\fntext[kulumanifootnote]{Doctoral Student, Mechanical and Aerospace Engineering}
\fntext[leefootnote]{Associate Professor, Mechanical and Aerospace Engineering}

\begin{abstract}
A computational approach is developed for the design of continuous low thrust transfers in the planar circular restricted three-body problem.
The transfer design method of invariant manifolds is extended with the addition of continuous low thrust propulsion.
A reachable region is generated and it is used to determine transfer opportunities, analogous to the intersection of invariant manifolds.
The reachable set is developed on a lower dimensional \Poincare section and used to design transfer trajectories. 
%Computational geometric mechanics allows for the development of variational integrators and a discrete optimal control problem.
This is solved numerically as a discrete optimal control problem using a variational integrator.
This provides for a geometrically exact and numerically efficient method for the motion in the three-body problem.
A numerical simulation is provided developing a transfer from a \( L_1 \) periodic orbit in  the Earth-Moon system to a target orbit about the Moon.
\end{abstract}

\begin{keyword}
Keywords
\end{keyword}

\end{frontmatter}

\section{Introduction}\label{sec:introduction}
% MOTIVATION FOR LOW THRUST ORBITAL TRANSFER 
Designing spacecraft trajectories is a classic and ongoing topic of research.
With the current fiscal constraints there is an increased focus on technologies with a critical impact on mass.
Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
Electric propulsion systems offer a much greater specific impulse than chemical systems and are able to operate for extended periods of time.
However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore must operate over longer durations in order to impart the desired momentum change.
Recent developments in miniature electric propulsion offer the potential for new research opportunities for small spacecraft~\cite{haque2013}.
With reduced development intervals and decreased launch costs, small satellites have gained increased attention as a cost effective means of scientific and technologic development. 
The merger of small satellites with miniature electric propulsion enables inexpensive and responsive missions requiring large changes in orbital energy or extended mission lifespan.
With the potential for more demanding missions, even greater importance is placed on the mission design to ensure that optimal trajectories satisfy mission requirements. 
In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{koon2000}.
Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional conic approach~\cite{ross2006,gomez2001}.
Optimally combining the dynamical structure of the three-body problem with low-thrust propulsion systems is vital for future mission success.

% PRIOR WORK AND CHALLENGES
There has been extensive research focused on optimal control for spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}.
Typically, the optimal control problem is solved via direct methods, which approximate the continuous problem as a parameter optimization problem.
The state and/or control trajectories are parameterized and solved in the form of a nonlinear optimization problem.
References~\citenum{mingotti2011} and~\citenum{grebow2011} use this direct approach in designing low-thrust transfers in the three-body problem.
Alternatively, indirect methods apply calculus of variations to derive the necessary conditions for optimality. 
This yields a lower dimensioned problem than the direct approach and algebraic conditions that, when satisfied, guarantee local optimality in contrast to direct methods which result in sub-optimal solutions.

The application of optimal control methods for orbital trajectory design is nontrivial.
The three-body system dynamics are nonlinear and exhibit chaotic behaviors. 
Small changes in initial conditions result in large variations of the resulting system trajectory. 
Therefore, any optimization routine is highly sensitive to the initial guess.
In addition, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge which achieve satisfactory results.
Efficient numerical implementation is dependent on correct initial conditions as well as accurate numerical integration.

Additionally, References~\citenum{mingotti2011} and~\citenum{grebow2011} implement the solutions using conventional Runge-Kutta integration techniques.
These techniques suffer from numerical instability and energy drift behaviors which make them ill-suited for long-term propagation.
These dissipative effects are even more detrimental with the addition of low-thrust propulsion to the dynamic equations of motion.
Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system.
As a result, the long term effects of low-thrust on the spacecraft trajectory are not accurately captured.

References~\citenum{koon2000} and~\citenum{ross2006} have illustrated the rich structure that exists in the three-body problem.
Within the three-body problem, a spacecraft's feasible region of motion is constrained by its energy, or Jacobi integral. 
It has been shown that there exist multi-dimensional tubes, or invariant manifolds, of constant energy trajectories that span the state space. 
Associated with periodic solutions of the three-body dynamics, these invariant manifolds allow for the spacecraft to traverse vast expanses of the state space with zero energy change. 
However, the results presented are highly case specific and difficult to generalize to arbitrary transfers.
Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system.
In addition, transfer orbits along an invariant manifold require a longer time of flight which may be undesirable for time critical missions.
The addition of low-thrust propulsion offers the potential of reduced transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.

% OBJECTIVE OF THIS WORK
In order to address these issues, the authors propose an accurate and numerically stable method for long-term optimal orbital transfers.
This approach will avoid the instability and dissipative effects of conventional integration schemes.
In addition, the effects of low-thrust propulsion will be directly included in the integration method to ensure that extended duration optimal trajectories are computed correctly.
This improved method will enable the derivation of a systematic method of generating optimal transfer orbits between arbitrary states.
Indirect optimal control, based on the calculus of variations, will be implemented in order to generate optimal orbital transfers.
This will avoid the approximation issues inherent in the previous work, which utilized direct optimal control methods.
With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low-thrust propulsion systems.

% TECHNICAL DESCRIPTION OF APPROACH
To achieve these objectives, computational geometric optimal control techniques are applied. 
The dynamics of the three-body system are derived from the discrete Lagrangian, which approximates the integral of the continuous time Lagrangian over a fixed discrete step.
Application of the discrete Euler-Lagrange equations, or the discrete Legendre transform, results in the discrete equations of motion.
This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long time periods.
A discrete optimal control problem is formulated from the discrete equations of motion.
This approach, where explicit discretization occurs prior to optimization,  is in contrast to the typical method, where the equations of motion are implicitly discretized during the optimization procedure.
Formulating the problem in this manner results in more stable and accurate optimal solutions. 
In indirect methods the optimal control problem is expressed as a two-point boundary value problem.
Optimal solutions are generally sensitive to small variations in the initial multipliers.
As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance. 
The use of geometric integrators, which do not suffer the numerical dissipation of conventional integration methods, results in a more robust and efficient solution.

A discrete optimal control problem is formulated to determine the reachability set on a \Poincare section.
Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft. 
Generation of this reachability set allows for the extension of the previous control-free missions in the three-body problem.
In addition, the generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer. 
The addition of low-thrust propulsion affords the ability to enlarge the reachable set as compared to the control-free case.
Maximization of the reachability set, on an appropriately chosen \Poincare section, allows for a greater space of potential transfer trajectories.
The use of the \Poincare section allows for design on a lower dimensional space and simplifies the design process.
Once an intersection is determined on the \Poincare section a transfer is calculated.

% SUMMARY OF CONTRIBUTIONS
In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
Previous results in the design of optimal transfers have relied on suboptimal direct optimization methods and conventional integration techniques.
This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
The use of a geometric integrators ensures numerical stability for long-duration orbit transfers.
A numerical example is presented demonstrating a transfer trajectory from \( L_1 \) to the vicinity of the Moon and compared to the control-free design methodology.
\section{System Model}\label{sec:pcrtbp}
The system model is based on the planar circular restricted three body problem~(PCRTBP).
The Earth is assumed to be the more massive primary, \( m_1 \), while the Moon is the second, smaller primary \( m_2\).
The equations of motion are developed in a rotating reference frame which allows for much greater insight into the structure of the dynamics.
The \( \hat{x} \) axis is directed along the vector from the Earth to the Moon.
The \( \hat{y} \) axis lies in the orbital plane and is orthogonal to \( \hat{x} \).
The rotating reference frame is centered at the system barycenter.
It is assumed that the \(\left( \hat{x}, \hat{y}\right)\) rotates with a constant angular velocity equal to the mean motion of the Moon.
Following convention, the system is also non-dimensionalized by the characteristic mass, length, and time~\cite{koon2000}.
As a result, the system can be characterized by a single mass parameter \( \mu \),
\begin{equation}
	\mu = \frac{m_2}{m_1+m_2} \, .
	\label{eq:mass_param}
\end{equation}
The larger primary, \(m_1\), is located at \( \left(  -\mu , 0 \right)\) and the smaller \( m_2\) is located at \( \left( 1-\mu , 0 \right)\).
In the rotating reference frame the Lagrangian is given by
\begin{equation}
	L = \frac{1}{2} \left( \left( \dot{x} -y \right)^2 + \left( \dot{y} + x \right)^2 \right) + \frac{1-\mu}{r_1} + \frac{\mu}{r_2}\, ,
	\label{eq:lagrangian}
\end{equation}
where the distance \(r_1\) and \(r_2\) of the spacecraft to each primary is defined in rotating coordinates as
\begin{align}
	r_1 &= \sqrt{\left( x + \mu\right)^2 + y^2}\, , \\
	r_2 &= \sqrt{\left( x - 1 + \mu\right)^2 + y^2}\, .
	\label{eq:distances}
\end{align}
Application of the Euler-Lagrange equations results in following equations of motion defined in the rotating reference frame
\begin{align}
	\ddot{x} - 2 \dot{y} + \frac{\partial}{\partial x} U &= u_x \, ,\nonumber \\
	\ddot{y} + 2 \dot{y} + \frac{\partial}{\partial y} U &= u_y \, ,
	\label{eq:cont_eom}
\end{align}
where the effective potential \( U\) is defined as
\begin{equation}
	U = \frac{1}{2} \left( x^2 + y^2\right) + \frac{1-\mu}{r_1} + \frac{\mu}{r_2}\, ,
	\label{eq:eff_pot}
\end{equation}
and the control inputs are defined as \( u_x\) and \(u_y\).
The state is defined as \( \bar{x} = \begin{bmatrix}\bar{r} &\bar{v} \end{bmatrix}\) with \(\bar{r} \in \R^{2\times1}\) and \(\bar{v} \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
The equations of motion may be rewritten in state space form as
\begin{equation}
	\left[\begin{array}{c} \dot{\bar{r}} \\ \dot{\bar{v}} \end{array} \right] = 
	\left[ \begin{array}{c} \bar{v} \\ A \bar{v} + \nabla U + \bar{u}(t) \end{array} \right] = f\left( t,x, u\right) \, ,
\end{equation}
where the matrix \( A \) and psuedo gravitational potential gradient \( \nabla U\) are
\begin{equation}\label{eq:A_mat}
	A = \left[ \begin{array}{ccc} 0 & 2 & 0 \\ -2 & 0 & 0 \\ 0 & 0 & 0 \end{array} \right] \, ,
\end{equation}
\begin{equation} \label{eq:grav_pot}
	\nabla U = \left[ \begin{array}{c} x - \frac{ \left(1 - \mu\right) \left(x + \mu\right)}{r_1^3} - \frac{\mu \left( x - 1 + \mu \right)}{r_2^3} \\
											y - \frac{ \left(1 - \mu\right) y}{r_1^3} - \frac{\mu y}{r_2^3} \\
											- \frac{ \left(1 - \mu\right) z}{r_1^3} - \frac{\mu z}{r_2^3}\end{array}\right]
					= \left[\begin{array}{c} U_x \\ U_y \\ U_z\end{array} \right] \, .
\end{equation}

\subsection{Jacobi Integral}
There exists a single integral, or constant of motion for the three-body problem~\cite{lanczos1970,szebehely1967}.
This energy constant is analogous to the total mechanical energy, however is a non-physical quantity arising from the problem formulation~\cite{szebehely1967}.
Also known as the Jacobi constant, it is defined as a function of the position and velocity in the rotating frame and given by
\begin{equation}
	E\left( \bar{r} , \bar{v} \right) = \frac{1}{2}\left( \dot{x}^2 + \dot{y}^2\right) - U\left(x,y \right) \, .
	\label{eq:jacobi}
\end{equation}
\Cref{eq:jacobi} divides the phase space into distinct regions based on the energy level of the spacecraft.
Fixing the Jacobi integral to a constant defines zero velocity curves, which are the locus of points where the kinetic energy, and hence velocity vanishes.
As seen in Figure~\ref{fig:energy_contour}, the phase space is divided into distinct realms based on the energy level.
In the vicinity of \( m_1\) or \(m_2\) there exists a potential well. 
As the energy level increases there are five critical points of the effective potential~\cref{eq:eff_pot} where the slope is zero.
Three collinear saddle points on the \(x\) axis and two equilateral points.
These equilibrium, or Lagrange points, are labeled \( L_i, i = 1, \hdots, 5 \) and are shown in Figure~\ref{fig:energy_contour}.
The Jacobi integral is a valuable invariant property of the three-body system that allows for greater insight into the motion of the spacecraft.
\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.5\textwidth]{energy_contours}
	\caption{Contour Plot of Jacobi Integral}
	\label{fig:energy_contour}
\end{figure} 
%\subsection{Periodic Orbits and Invariant Manifolds}
%Used to generate a Periodic orbit about a lagrange point
%Periodic orbits about the lagrange points
%\subsection{\Poincare Section}
%Intersection of states with a lower dimensional plane
%Introduce invariant manifolds

\section{Variational Integrator}\label{sec:discrete_var}
% description of variational integrator development
Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invaraint properties and symplecticity.
Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations~\cite{marsden2001}.
As a result, integrators developed in this manner have the desirable properties that they are symplectic and momentum preserving.
In addition, they exhibit improved energy behavior over long integration periods.
A short background on the variational principle for mechanical systems is presented. 
A discrete approximation of the action integral is presented and allows for construction of a variational integrator for the PCRTBP.

\subsection{Variational Principle}
Consider a continuous mechanical system described by the Lagrangian, \( L( q, \dot{q} ) \), for the generalized position, \( q\), and velocity, \( \dot{q} \).
In the standard approach of variational mechanics the action integral is formed by integrating the continuous Lagrangian along a path \( q(t) \) that the system follows from time \( t = 0 \) to \( t = T \)~\cite{greenwood1988}.
In the continuous time the action integral is defined as
\begin{align}\label{eq:action_integral}
	S = \int_{0}^T L\left( q, \dot{q}\right) \, dt \, .
\end{align}
Hamilton's principle states that the actual path followed by a holonomic system results in a stationary action integral with respect to path variations for fixed endpoints.
Taking the variation of~\cref{eq:action_integral} gives
\begin{subequations}\label{eq:var_principle}
\begin{align}
	\delta S &= \int_{0}^T \deriv{L}{q} \delta q + \deriv{L}{\dot{q}} \delta \dot{q} \, dt \\
		&= \int_{0}^T \deriv{L}{q} \delta q - \frac{d}{dt} \left( \deriv{L}{\dot{q}}\right) \delta q \, dt - \left. \left[ \deriv{L}{\dot{q}} \delta q\right] \right|_0^T \\
	&= \int_{0}^T \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}}	\right) \, dt \, ,
\end{align}
\end{subequations} 
where we have used integration by parts and the conditions \( \delta q(0) = 0 \) and \( \delta q(T) = 0\).
For Hamilton's principle to be valid for all admissible variations \( \delta q \), the integrand of~\cref{eq:var_principle} must be zero for all \( t\), giving the continuous Euler-Lagrange Equations~\cite{lanczos1970}.
\begin{align}\label{eq:euler-lagrange}
	0 = \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}} \right) \, .
\end{align}
Hamilton's equations are derivable through the use of the Legendre transformation which is a mapping \( \left( q, \dot{q},t\right) \rightarrow \left(q, p, t \right) \) where \( p_i\) is the generalized momenta.
\begin{align}\label{eq:legendre_transform}
	p_i = \deriv{L}{\dot{q}_i} \, .
\end{align}
In the continuous time case the Hamiltonian is defined as
\begin{align}\label{eq:hamiltonian}
	H &= \sum_{i = 1}^N p_i \dot{q}_i - L \left( q_i,\dot{q}_i, t \right) \, .
\end{align}
Applying~\cref{eq:legendre_transform} and taking the variation of~\cref{eq:hamiltonian} allows us to derive the equations of motion in Hamiltonian form
\begin{subequations}\label{eq:hamilton_eq}
\begin{align}
	\dot{q}_i &= \deriv{H}{p_i} \, ,\\
	\dot{p}_i &= - \deriv{H}{q_i} \, , \\
	\deriv{L}{t} &= -\deriv{H}{t} \, .
\end{align}
\end{subequations}
Both~\cref{eq:euler-lagrange,eq:hamilton_eq} result in equations of motion for the mechanical system and are equivalent via the Legendre transform.
\Cref{eq:euler-lagrange} results in \( n \) second order differential equations while~\cref{eq:hamilton_eq} results in \( 2n \) first order differential equations.
\subsection{Discrete Variational Mechanics}
A discrete analogue of Hamilton's principle and the action integral is formed.
Rather than taking a position, \( q \), and velocity, \( \dot{q} \), consider two positions \( q_0 \) and \( q_1 \) and a fixed time step \( h \in \R \).
The two positions are points on the curve \( q(t) \) such that \( q_0 \approx q(0) \) and \( q_1 \approx q(h) \).
A discrete time Lagrangian \( L_d( q_0, q_1) \) is formed which approximates the action integral between \( q_0 \) and \( q_1 \) as 
\begin{align}\label{eq:discrete_lagrangian}
	L_d\left( q_0 , q_1 \right) \approx \int_{0}^{h} L \left( q , \dot{q} \right) \, dt \, .
\end{align}
Since~\cref{eq:discrete_lagrangian} is calculated as a numerical integral, an appropriate quadrature rule is required.
There are multiple possible methods one can use to approximate the integral in~\cref{eq:discrete_lagrangian}.
An appropriate approximation rule is determined based on the ease of implementation and accuracy desired.
\begin{table}[htbp]
\caption{Selected Quadrature Rules\label{tab:quadrature}}
\begin{center}
\begin{tabular}{l|l}Rectangle & \( L_d(q_0,q_1) =L(q_0,\frac{q_1-q_0}{h}) h \)  \\ \hline
Midpoint & \( L_d(q_0,q_1) = L(\frac{q_0 + q_1}{2},\frac{q_1 - q_0}{h}) h \) \\ \hline
Trapezoidal & \( L_d(q_0, q_1) = \frac{1}{2} \left[ L(q_0, \frac{q_1 - q_0}{h} ) + L(q_1, \frac{q_1 - q_0 }{h} )\right] h \)
\end{tabular} 
\end{center}
\end{table}
\Cref{tab:quadrature} shows several possible approximation rules that are typically applied.
The rectangle rule is a first order accurate method and offers a straightforward implementation.
The midpoint and trapezoidal rules are both second order accurate methods. 
However, the midpoint rule results in an implicit form which adds further complexity to the equations of motion.
In this work, the trapezoidal approximation is applied to the PCRTBP.

Once an appropriate discrete Lagrangian is formed a discrete action sum is formed as the discrete analogue of~\cref{eq:action_integral}
\begin{align}\label{eq:action_sum}
	S_d = \sum_{k=0}^{N-1} L_d(q_k, q_{k+1}) \, .
\end{align}
Once again a discrete version of Hamilton's principle is applied to~\cref{eq:action_sum}.
Applying summation by parts yields
\begin{subequations}
\begin{align}\label{eq:dis_var_principle}
	\delta S_d &= \sum_{k=0}^{N-1} \deriv{L_d(q_k, q_{k+1})}{q_k} \delta q_k + \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \delta q_{k+1} \\
	&= \sum_{k=1}^{N-1} \bracket{ \deriv{L_d(q_k, q_{k+1})}{q_k} + \deriv{L_d(q_{k-1}, q_k)}{q_{k+1}}} \delta q_k \, .
\end{align}
\end{subequations}
For the discrete action sum to be stationary with respect to all admissible path variations, with fixed endpoints, the discrete Euler-Lagrange equations must be satisfied for \( k = 1, \cdots, N-1 \) resulting in
\begin{align}\label{eq:discrete_euler-lagrange}
	0 = \deriv{L_d(q_k, q_{k+1})}{q_k} + \deriv{L_d(q_{k-1}, q_k)}{q_{k+1}} \, .
\end{align}
A discrete version of the Legendre transformation, referred to as a discrete fiber derivative, results in the equivalent Hamiltonian form expression.
The discrete fiber derivative is given as 
\begin{subequations}\label{eq:discrete_legendre}
\begin{align}
	p_k &= \deriv{L_d(q_{k-1},q_k)}{q_{k+1}} = - \deriv{L_d(q_k, q_{k+1})}{q_k} \, ,\\
	p_{k+1} &= \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \, .
\end{align}
\end{subequations}
This yields a discrete Hamiltonian map \( (q_k, p_k) \to (q_{k+1}, p_{k+1}) \).
A more extensive development of variational integrators can be found in Reference~\citenum{marsden2001}.

\subsection{Discrete Equations of Motion}
The discrete equations of motion for the PCRTBP are derived by choosing an appropriate quadrature rule to discretize the Lagrangian in~\cref{eq:lagrangian}. 
In this work, the trapezoidal approximation is applied.
The trapezoid rule allows for an explicit second order accurate approximation.
The discrete Lagrangian is given by
\begin{align}\label{eq:discrete_lagrangian}
	L_d = &\frac{h}{2} \left( \frac{1}{2} \bracket{\left(  \frac{\xkp - \xk}{h} -\yk \right)^2 + \left( \frac{\ykp - \yk}{h} + \xk \right)^2} + \frac{1 - \mu}{r_{1_k}} + \frac{\mu}{r_{2_k}} \right. \nonumber \\ 
		& + \left. \frac{1}{2} \bracket{\left(  \frac{\xkp - \xk}{h} -\ykp \right)^2 + \left( \frac{\ykp - \yk}{h} + \xkp \right)^2} + \frac{1-\mu}{r_{1_{k+1}}} + \frac{\mu}{r_{2_{k+1}}}  \right) \, .
\end{align}
Applying a discrete version of the Lagrange-d'Alembert principle allows for inclusion of an external control force on the system~\cite{marsden2001}.
Using~\cref{eq:discrete_legendre,eq:discrete_lagrangian} and some manipulation, the equations of motion are given by
\begin{subequations}\label{eq:discrete_eoms}
\begin{align}
	\xkp &= \frac{1}{1+ h^2} \bracket{h \dot{x}_k + h^2 \dot{y}_k  + \xk \parenth{1+ \frac{3h^2}{2}} + \frac{h^3}{2} \yk - \frac{h^3}{2} U_{y_k} - \frac{h^2}{2} U_{x_k} } \label{eq:xkp} \, ,\\
	\ykp &= h \dot{y}_k + h \xk - h \xkp + \yk + \frac{h^2 \yk}{2} - \frac{h^2 }{2} U_{y_k} \label{eq:ykp} \, ,\\
	\dot{x}_{k+1} &= \dot{x}_k - 2 \yk + 2 \ykp + \frac{h}{2} \parenth{\xkp + \xk} - \frac{h}{2} U_{\xkp} - \frac{h}{2} U_{\xk} + h u_x \label{eq:xdotkp}\, ,\\
	\dot{y}_{k+1} &= \dot{y}_{k} + 2 \xk - 2 \xkp + \frac{h}{2} \parenth{\ykp + \yk} - \frac{h}{2} U_{\ykp} - \frac{h}{2} U_{\yk} + h u_y \label{eq:ydotkp} \, .
\end{align}
\end{subequations}
The discrete equations of motion are given in the Lagrangian form after applying the discrete fiber derivative from~\cref{eq:discrete_legendre} as \( p_{\xk} = \dot{x}_k - \yk \) and \( p_{\yk} = \dot{y}_k + \xk \).
The state is defined as \( \bar{ x}_k = \begin{bmatrix} \xk & \yk & \dot{x}_k & \dot{y}_k \end{bmatrix}^T\) and the control input is \( \bar{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \).
The discrete potential gradients are given by
\begin{subequations}\label{eq:discrete_potential_grad}
\begin{align}
	U_{\xk} &= \frac{\parenth{1 -\mu} \parenth{\xk + \mu}}{\distonek^3} + \frac{ \mu \parenth{\xk -1 + \mu}}{\disttwok^3} \label{eq:Uxk} \, ,\\
	U_{\yk} &= \frac{\parenth{1 -\mu} \yk}{\disttwok^3} + \frac{ \mu \yk}{\disttwok^3} \label{eq:Uyk} \, ,\\
	U_{\xkp} &= \frac{\parenth{1 -\mu} \parenth{\xkp + \mu}}{\distonekp^3} + \frac{ \mu \parenth{\xkp -1 + \mu}}{\disttwokp^3} \label{eq:Uxkp}\, ,\\
	U_{\ykp} &= \frac{\parenth{1 -\mu} \ykp}{\distonekp^3} + \frac{ \mu \ykp}{\disttwokp^3} \label{eq:Uykp} \, .
\end{align}	
\end{subequations}
The distances to each primary are defined as
\begin{subequations}\label{eq:discrete_distance}
\begin{align}
	\distonek &= \sqrt{\left( \xk + \mu\right)^2 + \yk^2} \label{eq:distonek}\, ,\\
	\disttwok &= \sqrt{\left( \xk - 1 + \mu\right)^2 + \yk^2} \label{eq:disttwok}\, ,\\
	\distonekp &= \sqrt{\left( \xkp + \mu\right)^2 + \ykp^2} \label{eq:distonekp}\, ,\\
	\disttwokp &= \sqrt{\left( \xkp - 1 + \mu\right)^2 + \ykp^2} \label{eq:disttwokp} \, .
\end{align}
\end{subequations}
Care must be taken during the implementation of~\cref{eq:discrete_eoms}.
As~\cref{eq:discrete_potential_grad,eq:discrete_distance} are defined at both step \( k \) and \( k+1 \) they must be evaluated at both time steps.
\Cref{eq:discrete_eoms} is implemented by first defining an initial state \( \bar{x}_k \) and control \( \bar{u}_k \).
The distances and gravitational potential at step \( k \) are evaluated from~\cref{eq:distonek,eq:disttwok,eq:Uxk,eq:Uyk}.
The discrete update steps in~\cref{eq:xkp,eq:ykp} are evaluated to generate \( \xkp \) and \( \ykp\).
Next, the distances and gravitational potential at step \( k+1 \) are evaluated from~\cref{eq:distonekp,eq:disttwokp,eq:Uxkp,eq:Uykp}. 
Finally, the update steps in~\cref{eq:xdotkp,eq:ydotkp} are evaluated.
This results in the complete discrete update map \( \bar{x}_k \to \bar{x}_{k+1} \) given \( \bar{u}_k \).

\begin{figure} 
	\centering 
	\begin{subfigure}[h]{0.3\textwidth} 
		\includegraphics[width=\textwidth]{trajectory} 
		\caption{Trajectory} \label{fig:compare_trajectory} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.3\textwidth} 
		\includegraphics[width=\textwidth]{components} 
		\caption{State \( \parenth{x , y, \dot{x}, \dot{y}}\)} \label{fig:compare_components} 
	\end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.3\textwidth} 
		\includegraphics[width=\textwidth]{energy} 
		\caption{Jacobi Integral} \label{fig:compare_energy} 
	\end{subfigure} 
	\caption{Integrator Comparison}
	\label{fig:integrator_compare} 
\end{figure}
A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
A particle is simulated from an initial condition of \( \bar{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
The variational integrator uses a step size of \SI{47.22}{\second} while the Runge-Kutta method uses a variable step size implemented via ODE45 in Matlab.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference frame for both integration schemes.
Both integration schemes result in trajectories that are initially nearly identical.
The discrete equations of motion are an accurate approximation for the continuous dynamics as they closely match the solution of ODE45 over the initial portion of the simulation.
However, as time progresses the trajectories begin to diverge due to the differences in system energy.
\Cref{fig:compare_energy} shows the evolution of the Jacobi integral.
The variational integrator exhibits a bounded behavior about the initial energy with a mean variation of \num{4.2522e-20}.
However, the conventional Runge-Kutta method demonstrates a clear energy drift of \num{4.2814e-8}. 
Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques.

\section{Invariant Manifolds} \label{sec:invariant_manifold}
% introduce \Poincare section idea
There exists a rich structure of tubes, or invariant manifolds, in the three-body problem~\cite{koon2000,conley1968}.
The manifold structure associated with periodic orbits about the \( L_1 \) and \( L_2 \) Lagrange points are critical to the understanding of the motion of spacecraft as well as comets/asteroids.
In addition, the stable and unstable manifolds serve as the boundaries of the phase space region that allow for the transport between realms in a single three-body system or between multiple three-body systems.
These invariant manifolds only exist as a result of the dynamic formulation of the three-body problem in a rotating reference frame. 
In this way, it is possible to design trajectories that intersect the invariant manifolds and connect widely separated regions of the state space. 
\Cref{fig:invariant_manifolds} shows some of the tubes, projected from the phase space onto the configuration space, associated with \( L_1 \) and \( L_2 \) periodic orbits in the Earth-Moon system. 
This technique has been heavily investigated in Reference~\citenum{koon2000}, applied to operational missions in Reference~\citenum{koon1999}, and applied to potential multi-moon orbiter missions in Reference~\citenum{tanaka2011}.
In addition, much recent work has focused on potential return missions to the Moon~\cite{zanzottera2012,campagnola2012,mingotti2011,ozimek2010a,mingotti2009}.
\begin{figure}
     \centering
        \begin{subfigure}[b]{0.25\textwidth}
                \includegraphics[width=\columnwidth]{U1_Manifolds}
                \caption{U1 Manifolds}
                \label{fig:u1_manifolds}
        \end{subfigure}%
        ~%add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[b]{0.25\textwidth}
                \includegraphics[width=\columnwidth]{U2_Manifolds}
                \caption{U2 Manifolds}
                \label{fig:u2_manifolds}
        \end{subfigure}~
        \begin{subfigure}[b]{0.25\textwidth}
                \includegraphics[width=\columnwidth]{U3_Manifolds}
                \caption{U3 Manifolds}
                \label{fig:u3_manifolds}
        \end{subfigure}%
        ~%add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[b]{0.25\textwidth}
                \includegraphics[width=\columnwidth]{U4_Manifolds}
                \caption{U4 Manifolds}
                \label{fig:u4_manifolds}
        \end{subfigure}
        \caption{Invariant Manifolds for Planar Earth-Moon three-body system}
	\label{fig:invariant_manifolds}
\end{figure}
\begin{figure}
     \centering
        \begin{subfigure}[b]{0.25\textwidth}
                \includegraphics[width=\columnwidth]{U1_poincare}
                \caption{U1 Section}
                \label{fig:u1_poincare}
        \end{subfigure}%
        ~%add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[b]{0.25\textwidth}
                \includegraphics[width=\columnwidth]{U2_poincare}
                \caption{U2 Section}
                \label{fig:u2_poincare}
        \end{subfigure}~
        \begin{subfigure}[b]{0.25\textwidth}
                \includegraphics[width=\columnwidth]{U3_poincare}
                \caption{U3 Section}
                \label{fig:u3_poincare}
        \end{subfigure}%
        ~%add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[b]{0.25\textwidth}
                \includegraphics[width=\columnwidth]{U4_poincare}
                \caption{U4 Section}
                \label{fig:u4_poincare}
        \end{subfigure}
        \caption{\Poincare Sections for Planar Earth-Moon three-body system}
	\label{fig:poincare_sections}
\end{figure}

Another useful technique in the analysis of the free motion of dynamical systems is that of the \Poincare section.
The \Poincare section is the intersection of a periodic orbit of a dynamical system with a lower dimensional sub-space transverse to the flow.
This allows for greater insight into the stability and dynamics of periodic solutions of dynamic system. 
Points are drawn as the periodic solution intersects the \Poincare section.
Great insight and structure into the dynamical system is typically available through use of \Poincare section.
For example, \cref{fig:poincare_sections} shows the intersections of the invariant manifolds of~\cref{fig:invariant_manifolds} with the \Poincare sections defined by the black line segments.
The \Poincare section and the Jacobi integral reduce the state space from four to two dimensions for the planar three-body problem~\cite{koon2001}.
This greatly eases the design process and allows for a simple planar visualization of the intersection of the invariant manifolds.
Intersecting states are easy to determine and allow for transfers between invariant manifolds.

The intersection of the invariant manifolds on a \Poincare section is central to determining the desired control-free trajectory.
However, the results previously developed are highly case specific and difficult to generalize to arbitrary transfers.
Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system.
In addition, transfer orbits along an invariant manifold require large time of flights which may be undesirable for time critical missions.
The addition of low-thrust propulsion offers the potential of reduced transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities. 
In this work, an optimal control problem is formulated to generate a reachable set of the spacecraft.
The reachable set is computed on an appropriate \Poincare section and is used to design an appropriate transfer trajectory.
\section{Optimal Control Formulation for Reachability Set}\label{sec:optimal_control}
% introduce method
In this section, an optimal control formulation is presented to determine and design transfers within the three-body problem.
The application of variational integrators to optimal control problems is referred to as computational geometric optimal control.
This formulation is based on the concept of the reachability set on a \Poincare section.
This method allows for one to easily determine potential transfer opportunities by finding set intersections on a lower dimensional space and greatly reduces the design process.
The addition of continuous low thrust propulsion extends the control free design process presented in Reference~\citenum{koon2000} and allows for a greater range of potential transfers with a reduced time of flight.

% discuss reachability and application to space system
Reachability theory provides a framework to evaluate control capability and safety.  
The reachable set contains all possible trajectories that are achievable over a fixed time horizon from a defined initial condition, subject to the operational constraints of the system.
Reachability theory has been applied to aerospace systems such as collision avoidance, safety planning, and performance characterization.
The theory formally supporting reachability has been extensively developed and is directly derivable from optimal control theory~\cite{varaiya2000,lygeros2002,lygeros2004}.
Computation of the reachable set for a system involves solving the Hamilton-Jacobi partial differential equation or satisfying a dynamic programming principle.
Analytical computation of reachable sets is an ongoing problem and is only possible for certain classes of systems.
Typically, numerical methods are used to generate approximations of the reachability set, but are generally limited by the dimensionality of the problem.

Reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a}.
Computation of reachable sets is critical to space situational awareness, rendezvous and proximity operations, and orbit determination operations.
Specifically maintaining accurate estimates of a spacecraft state over extended periods is not trivial.
The challenge is increased for multiple spacecraft operating in close proximity or when there are long periods of time between measurements.
Coupling the ability for continuous low-thrust propulsion between measurements increases the measurement association complexity.
Computing the reachability set given estimated states and control authorities allows one to better correlate subsequent measurements or determine sensor pointing regions in the event of a lost spacecraft.

The cost function is defined as
\begin{equation}
	J = -\frac{1}{2} \left( \bar{x}(N) - \bar{x}_{n}(N)\right)^T 
	\left[
	\begin{array}{cccc}
		1 & 0& 0& 0 \\
		 0& 0& 0& 0\\
		 0 & 0 & 1 &0\\
		 0 & 0& 0& 0
	\end{array}
	\right]
	\left( \bar{x}(N) - \bar{x}_{n}(N)\right) \, .
	\label{eq:cost}
\end{equation}
The term \( \bar{x}_n(N) \) is the final state of a control-free trajectory while the term \( \bar{x}(N) \) is the final state under the influence of the control input.
In this fashion, the aim is to maximize the distance of the final state from that of the control-free trajectory. 
A chosen \Poincare section is defined through the use of appropriate terminal constraints given by
\begin{subequations}
\begin{align}
    m_1 &= 0 = \frac{y(N) - L_{1y}}{x(N) - L_{1x}} - \tan{\alpha_d} \, , \\ 
    m_2&= 0 = \frac{\dot{x}(N) - \dot{x_n}(N) }{x(N) -x_n(N) } - \tan{\theta_d} \, , \\
	 0 &\geq\bar{u}^T \bar{u} - u_{max}^2 \, ,
\end{align}
    \label{eq:constraints}
\end{subequations}
where the angles \( \alpha_d\) and \( \theta_d\) define the \Poincare section and a specific direction upon the section, respectively. 
The goal is to determine the control input \( \bar{u}_k\) such that the cost function~\cref{eq:cost} is minimized subject to the state equations of motion~\cref{eq:discrete_eoms} and constraints~\cref{eq:constraints}.
Application of the Euler-Lagrange equations allows us to derive the necessary conditions for optimality~\cite{bryson1975}.
The discrete variational integrator in~\cref{eq:discrete_eoms} is used rather than the continuous time counterpart. 
This results in a discrete optimal control problem and the discrete necessary conditions are given as
\begin{subequations}\label{eq:necc_cond}
\begin{align}
	\lambda_{k+1}^T &= \lambda_k^T  \deriv{f_k}{\bar{x}_k}^{-1} \, , \\
	0 &=  \deriv{H_k}{\bar{u}_k} \, ,\\
	0 &= \deriv{\phi}{\bar{x}_k}^T + \deriv{m}{\bar{x}_k}^T\beta  - \lambda^T(N) \, ,  
\end{align}
\end{subequations}
where the Hamiltonian \(H\) is defined as
\begin{equation}
	H_k = \lambda_k^T f(\bar{x}_k, \bar{u}_k) \, ,
	\label{eq:hamiltonian_opt}
\end{equation}
and \(\lambda \in \R^{4 \times 1}\) is the costate and \(\beta \in \R^{2 \times 1} \) are the additional Lagrange multipliers associated with the terminal constraints in~\cref{eq:constraints}.
This indirect optimal control formulation leads to a two point boundary value problem with split boundary conditions. 
By sweeping the angle \( \theta_d \) one can approximate the reachable set on the \Poincare section subject to the bounded control input.

The costate equation of motion requires the Jacobian of~\cref{eq:discrete_eoms} and is given by
\begin{align}\label{eq:costate_eom}
	\lambda_{k+1}^T = \lambda_k^T
	\begin{bmatrix} 
		f_{1_x} & f_{1_y} & f_{1_{\dot{x}}} & f_{1_{\dot{y}}} \\
		f_{2_x} & f_{2_y} & f_{2_{\dot{x}}} & f_{2_{\dot{y}}} \\
		f_{3_x} & f_{3_y} & f_{3_{\dot{x}}} & f_{3_{\dot{y}}} \\
		f_{4_x} & f_{4_y} & f_{4_{\dot{x}}} & f_{4_{\dot{y}}}
	\end{bmatrix} ^ {-1} \, .
\end{align}
The derivation of~\cref{eq:costate_eom} is given in Appendix A.
In addition, the computation of~\cref{eq:costate_eom} requires inversion of the Jacobian matrix.
This is a computationally expensive operation that is prone to numerical error and instability.
A method is presented in Appendix B to avoid this inversion and determine an explicit update map \( \lambda_k \to \lambda_{k+1} \).

The optimal control formulation presented in this section results in a two point boundary value problem (TPBVP). 
There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{bryson1975}.
In this work, a multiple shooting method is implemented.
Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement.
In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs.
Additional interior constraints are used to ensure a continuous solution.
This has the benefit of decreasing the numerical sensitivity of the final states to changes in the initial conditions along each sub-arc.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
A numerical simulation is presented to demonstrate the transfer procedure.
The goal is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix} \) with \( \mu = 0.0125 \).
The initial state is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
   \caption{Target Orbit Region}
   \label{fig:moon_orbit}
\end{figure}
\Cref{fig:moon_orbit} shows that the desired trajectories remain in the vicinity of the Moon in the rotating reference frame. 
This type of orbit would be useful for a variety of mission scenarios.
For example, a series of communication satellites could be placed in this type of orbit. 
The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.

As a baseline, the method introduced in Reference~\citenum{koon2000} is implemented.
The method is based on the invariant manifolds associated with the periodic orbits of the three body system.
These invariant manifolds are a set of trajectories that either asymptotically arrive or depart the periodic orbit. 
Linking the invariant manifolds of several periodic orbits, or coupled three body systems, has been used to generate orbital transfers.
\Cref{fig:manifold_trajectory} shows the unstable invariant manifold associated with the \( L_1\) periodic orbit. 
The invariant manifolds are globalized using the eigenvectors of the Monodromy matrix along the periodic orbit.
The eigenvectors serve as a local approximation of the stable and unstable directions. 
Perturbations along these directions serve to approximate the invariant manifolds.

The unstable invariant manifold is numerically propagated to the same \Poincare section defined on the \( \hat{x} \) axis.
The intersection of the invariant manifold and the \Poincare section is denoted by the green markers in~\cref{fig:poincare_compare}.
Only a single branch of the invariant manifold intersects with the ascending region of the target orbit.
There are no intersections of the invariant manifold with the descending region of the target orbit.
The numerical values associated with the green points denote the required time of flight along the invariant manifold in non-dimensional units.
A transfer along the invariant manifold requires on average \( t_f \approx 3.1 \) as compared to \( t_f \approx 1.4 \) for a transfer using low thrust propulsion and the reachable set.
Finally, it should be noted that an additional maneuver would be required for a transfer using the invariant manifold.
The intersection on the \Poincare section only shows that the \( x \text{ and } \dot{x} \) components intersect.
An additional instantaneous \( \Delta V \) would be required to transfer from the invariant manifold to the target.

A periodic orbit is generated about the \( L_1 \) Lagrange point. 
A \Poincare section is chosen to allow for design on a lower dimensional subspace.
The section is defined along the \( \hat{x} \) axis, or defined with \( \alpha = \ang{0} \) and intersects both the initial and target orbits.
From the periodic orbit, a series of optimal trajectories are generated to approximate the reachable set.
The trajectories are generated from a fixed initial state of \( x_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix} \) over a fixed time span of \( t_f = 1.4 \).
By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:transfer_orbit}.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{reach_trajectory} 
		\caption{Controlled Trajectories} \label{fig:reach_trajectory} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{manifold_trajectory} 
		\caption{Invariant Manifold} \label{fig:manifold_trajectory} 
	\end{subfigure}

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare_compare} 
		\caption{Poincar\`e Section} \label{fig:poincare_compare} 
	\end{subfigure} 
	\caption{Transfer Trajectory}
	\label{fig:transfer_orbit} 
\end{figure}
The optimal trajectories, under the influence of the control input \( \bar{u} \), are shown in~\cref{fig:reach_trajectory}.
Initially, the spacecraft is assumed to lie on the periodic orbit.
As a result, the intersection of this periodic orbit with the \Poincare section are two points corresponding to the two crossing of the orbit.
The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
The reachable set is an ellipsoidal region with a major axis aligned along \( \theta \approx \ang{70} \).

The blue points in~\cref{fig:poincare_compare} are the intersections of the target Moon orbit and the \Poincare section.
The two circular regions are the ascending (right) and descending (left) intersections of the target orbit and \Poincare section.
\Cref{fig:poincare_compare} shows that the reachable set and those of the descending target region intersect.
As both regions are discretely approximated a linear interpolation is used to determine the exact intersection state on the \Poincare section.
This intersection generates a partial target state of \( x_t \text{ and } \dot{x}_t \).
Using~\cref{eq:jacobi} and the intersection state the final component \( \dot{y} \) is calculated. 
This results in a complete target state \( \bar{x}_t \) which lies on the reachable set and on the target orbit. 
A final optimal trajectory is generated such that the \( \bar{x}(N) = \bar{x}_t \).
This transfer trajectory is denoted by the green path in~\cref{fig:reach_trajectory}.

%Determine reachable set on \Poincare Section
%
%Intersection of the sets on section
%
%Minimization to a target state
%
%Show transfer to the moon
%
%Compare with the Ross work and the invariant manifold ( long time of flight)
\section{Conclusions}\label{sec:conclusion}
In this paper, an optimal transfer process which combines concepts of reachability and \Poincare section is used to generate transfer between planar periodic orbits in the three-body problem.
The \Poincare section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
This analysis has also assumed perfect attitude control and the ability to orient thrust in any direction.
The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability.

\appendix
\section*{Appendix A: Costate Equations of Motion}\label{sec:costate_appendix}
The development of the costate equations of motions begins with determining the second order partial derivatives of the gravitational potential. 
Due to the symmetry of partial derivatives only three terms are required and are given by
\begin{align}\label{eq:second_discrete_potential_grad}
	U_{x\xk} &= \parenth{1-\mu} \bracket{\frac{1}{\distonek^3} - \frac{3 \parenth{\xk +\mu}^2}{\distonek^5}} + \mu \bracket{\frac{1}{\disttwok^3} - \frac{3 \parenth{\xk -1 + \mu}^2}{\disttwok^5}} \, ,\\
	U_{y\yk} &= \parenth{1-\mu} \bracket{\frac{1}{\distonek^3} - \frac{3 \yk^2}{\distonek^5}} + \mu \bracket{\frac{1}{\disttwok^3} - \frac{3 \yk^2}{\disttwok^5}} \, ,\\
	U_{x\yk} &= U_{y\xk} =  \frac{-3 \parenth{1-\mu} \parenth{\xk +\mu} \yk}{\distonek^3} - \frac{3\mu\yk\parenth{\xk-1+\mu}}{\disttwok^5} \, .
\end{align}
The gradient of~\cref{eq:xkp} is given as
\begin{subequations}\label{eq:xkp_grad}
\begin{align}
	f_{1_x} &= \frac{1}{1+h^2} \bracket{h^2 + 1 + \frac{h^2}{2} -\frac{h^3}{2} U_{y\xk} - \frac{h^2}{2}U_{x\xk}} \, ,\\
	f_{1_y} &= \frac{1}{1+h^2} \bracket{ \frac{h^3}{2} -\frac{h^3}{2} U_{y\yk} - \frac{h^2}{2}U_{x\yk}} \, ,\\
	f_{1_{\dot{x}}} &= \frac{h}{1+h^2} \, ,\\
	f_{1_{\dot{y}}} &= \frac{h^2}{1+h^2} \, .
\end{align}
\end{subequations}
The gradient of~\cref{eq:ykp} is given as
\begin{subequations}\label{eq:ykp_grad}
\begin{align}
	f_{2_x} &= h -h f_{1_x} - \frac{h^2}{2} U_{y\xk} \, , \\
	f_{2_y} &= -h f_{1_y} + 1 + \frac{h^2}{2} - \frac{h^2}{2} U_{y\yk} \, ,\\
	f_{2_{\dot{x}}} &= -h f_{1_{\dot{x}}} \, ,\\
	f_{2_{\dot{y}}} &= h - h f_{1_{\dot{y}}} \, .
\end{align}
\end{subequations}
The gradients of~\cref{eq:distonekp,eq:disttwokp} are given as
\begin{subequations}\label{eq:distkp_grad}
\begin{align}
	\deriv{\distonekp}{\bar{x}} &= \parenth{\left( \xkp + \mu\right)^2 + \ykp^2}^{-\frac{1}{2}} \bracket{\parenth{\xkp + \mu} f_{1_{\bar{x}}} + \ykp f_{2_{\bar{x}}}} \, ,\\
	\deriv{\disttwokp}{\bar{x}} &= \parenth{\left( \xkp - 1 + \mu\right)^2 + \ykp^2}^{-\frac{1}{2}} \bracket{\parenth{\xkp -1 + \mu} f_{1_{\bar{x}}} + \ykp f_{2_{\bar{x}}}}  \, .
\end{align}
\end{subequations}
The second order partial derivatives of the gravitational potential at \( k+1\) are given as
\begin{subequations}
\begin{align}
	\deriv{U_{x\xkp}}{\bar{x}} &= \parenth{1-\mu}\bracket{\frac{1}{\distonekp^3} f_{1_{\bar{x}}} - \frac{3 \parenth{\xkp +mu}}{\distonekp^4} \deriv{\distonekp}{\bar{x}}} + \mu \bracket{\frac{1}{\disttwokp^3} f_{1_{\bar{x}}} - \frac{-3 \parenth{\xkp -1 + \mu}}{\disttwokp^4} \deriv{\disttwokp}{\bar{x}}} \, ,\\
	\deriv{U_{y\xkp}}{\bar{x}} &= \parenth{1-\mu}\bracket{\frac{1}{\distonekp^3} f_{2_{\bar{x}}} - \frac{3 \ykp}{\distonekp^4} \deriv{\distonekp}{\bar{x}}} + \mu \bracket{\frac{1}{\disttwokp^3} f_{2_{\bar{x}}} - \frac{-3 \ykp}{\disttwokp^4} \deriv{\disttwokp}{\bar{x}}} \, .
\end{align}
\end{subequations}
The gradient of~\cref{eq:xdotkp,eq:ydotkp} are given as
\begin{subequations}\label{eq:xdotkp_grad}
\begin{align}
	f_{3_x} &= 2 f_{2_x} + \frac{h}{2} \parenth{f_{1_x} + 1} - \frac{h}{2} U_{x\xkp} - \frac{h}{2} U_{x\xk} \, ,\\
	f_{3_y} &= -2 + 2 f_{2_y} + \frac{h}{2} f_{1_y} - \frac{h}{2} U_{x\ykp} -\frac{h}{2} U_{x\yk} \, ,\\
	f_{3_{\dot{x}}} &= 1 + 2 f_{2_{\dot{x}}} + \frac{h}{2} f_{1_{\dot{x}}} - \frac{h}{2} U_{x\xdotkp}\, ,\\
	f_{3_{\dot{y}}} &= 2 f_{2_{\dot{y}}}\, ,
\end{align}
\end{subequations}
\begin{subequations}\label{eq:ydotkp_grad}
\begin{align}
	f_{4_x} &= 2 - 2 f_{1_x} + \frac{h}{2} f_{2_x}  - \frac{h}{2} U_{y\xkp} - \frac{h}{2} U_{y\xk} \, ,\\
	f_{4_y} &= -2f_{1_y}  - \frac{h}{2} \parenth{f_{2_y} + 1} - \frac{h}{2} U_{y\ykp} -\frac{h}{2} U_{y\yk} \, ,\\
	f_{4_{\dot{x}}} &= - 2 f_{1_{\dot{x}}} +  \frac{h}{2} f_{2_{\dot{x}}} - \frac{h}{2} U_{y\xdotkp}\, ,\\
	f_{4_{\dot{y}}} &= 1 - 2 f_{1_{\dot{y}}} + \frac{h}{2} f_{2_{\dot{y}}} - \frac{h}{2} U_{y\ydotkp}\, .
\end{align}
\end{subequations}
These gradient equations are in a cascade type structure.
\Cref{eq:ydotkp_grad,eq:xdotkp_grad} are functions of~\cref{eq:xdotkp,eq:ydotkp}.
As a result, the accuracy of the Jacobian will tend to decrease as the first order approximation errors accumulate.

\section*{Appendix B: Gauss Jordan Elimination}\label{sec:costate_gauss_jordan}
The costate equations of motion are given by~\cref{eq:costate_eom} and repeated here as
\begin{align}\label{eq:costate_eom_transpose}
	\begin{bmatrix} 
		\fonex & \ftwox & \fthreex & \ffourx \\
		\foney & \ftwoy & \fthreey & \ffoury \\
		\fonexd & \ftwoxd & \fthreexd & \ffourxd \\
		\foneyd & \ftwoyd & \fthreeyd & \ffouryd
	\end{bmatrix}
	\begin{bmatrix} \lambda_{\xkp} \\ \lambda_{\ykp} \\ \lambda_{\xdotkp} \\ \lambda_{\ydotkp} \end{bmatrix}
	=
	\begin{bmatrix} \lambda_{\xk} \\ \lambda_{\yk} \\ \lambda_{\xdotk} \\ \lambda_{\ydotk} \end{bmatrix} \, .
\end{align}
To determine the discrete update map \( \lambda_k \to \lambda_{k+1}\) the inverse of the Jacobian matrix is required.
In order to avoid the need of an explicit inversion a Gauss Jordan method is implemented.
To begin, several terms are defined which are required to carry out the row operations and are defined as
\begin{subequations}\label{eq:ref_scale}
\begin{align}
	a &= -\frac{\foney}{\fonex} \, ,\\
	b &= -\frac{\fonexd}{\fonex} \, , \\
	c &= -\frac{\foneyd}{\fonex} \, ,\\
	e &= -\frac{\ftwoxd + b \ftwox}{\ftwoy + a \ftwox} \, ,\\
	f &= -\frac{\ftwoyd + c \ftwox}{\ftwoy + a \ftwox} \, ,\\
	g &= -\frac{\fthreeyd + c \fthreex + f \parenth{\fthreey + a \fthreex}}{\fthreexd + b \fthreex + e \parenth{\fthreey + a \fthreex}}\, .
\end{align}
\end{subequations}
\Cref{eq:costate_eom_transpose} is transformed to row echelon form using elementary row operations and is defined as
\begin{align}\label{eq:costate_ref}
	\begin{bmatrix} 
		\alpha_{11} & \alpha_{12} & \alpha_{13} & \alpha_{14} \\
		0 & \alpha_{22} & \alpha_{23} & \alpha_{24} \\
		0 & 0 & \alpha_{33} & \alpha_{34} \\
		0 & 0 & 0 & \alpha_{44}
	\end{bmatrix}
	\begin{bmatrix} \lambda_{\xkp} \\ \lambda_{\ykp} \\ \lambda_{\xdotkp} \\ \lambda_{\ydotkp} \end{bmatrix}
	=
	\begin{bmatrix} \beta_1 \\ \beta_2 \\ \beta_3 \\ \beta_4 \end{bmatrix} \, ,
\end{align}
where the terms \( \alpha_{ij} \) and \( \beta_{i} \) are defined as follows
\begin{subequations}
\begin{align}
	\alpha_{11} &= \fonex \, ,\\
	\alpha_{12} &= \ftwox \, ,\\
	\alpha_{13} &= \fthreex \, ,\\
	\alpha_{14} &= \ffourx \, ,\\
	\alpha_{22} &= \ftwoy + a \ftwox \, ,\\
	\alpha_{23} &= \fthreey + a\fthreex \, ,\\
	\alpha_{24} &= \ffoury + a \ffourx \, ,\\
	\alpha_{33} &= \fthreexd + b \fthreex + e \parenth{\fthreey + a \fthreex}\, ,\\
	\alpha_{34} &= \ffourxd + b \ffourx + e \parenth{\ffoury + a \ffourx} \, ,\\
	\alpha_{44} &= \ffouryd + c \ffourx + f \parenth{\ffoury + a \ffourx} + g \parenth{\ffourxd + b \ffourx + e \parenth{\ffoury + a \ffourx}} \, ,\\
	\beta_1 &= \lambda_{\xk} \, ,\\
	\beta_2 &= \lambda_{\yk} + a \lambda_{\xk} \, ,\\
	\beta_3 &= \lambda_{\xdotk} + b \lambda_{\xk} + e \parenth{\lambda_{\yk} + a \lambda_{\xk}} \, ,\\
	\beta_4 &= \lambda_{\ydotk} + c \lambda_{\xk} + f \parenth{\lambda_{\yk}+a \lambda_{\xk}} + g \parenth{\lambda_{\xdotk} + b \lambda_{\xk} + e \parenth{\lambda_{\yk} + a \lambda_{\xk}}} \, .
\end{align}
\end{subequations}
Finally, backsubstituion is used to determine explicit equations for the discrete update map \( \lambda_k \to \lambda_{k+1} \) which is defined as
\begin{subequations}\label{eq:costate_update}
\begin{align}
	\lambda_{\ydotkp} &= \frac{\beta_4}{\alpha_{44}} \, ,\\
	\lambda_{\xdotkp} &= \frac{\beta_3}{\alpha_{33}} - \frac{\alpha_{34}}{\alpha_{33}} \lambda_{\ydotkp} \, ,\\
	\lambda_{\ykp} &= \frac{\beta_2}{\alpha_22} - \frac{\alpha_{33}}{\alpha_{22}}\lambda_{\xdotkp} - \frac{\alpha_{24}}{\alpha_{22}} \lambda_{\ydotkp} \, ,\\
	\lambda_{\xkp} &= \frac{\beta_1}{\alpha_{11}} - \frac{\alpha_{12}}{\alpha_{11}} \lambda_{\ykp} - \frac{\alpha_{13}}{\alpha_{11}} \lambda_{\xdotkp} - \frac{\alpha_{14}}{\alpha_{11}} \lambda_{\ydotkp} \, .
\end{align}
\end{subequations}
%Once again, these equations exhibit a cascade type relationship. 
%As as result any errors will compound and tend to accumulate most prevalently in \( \lambda_{\xkp} \).
\section*{References}

\bibliography{library}
\bibliographystyle{elsarticle-num}
\end{document}

### 2015-11-22 acta_manuscript.tex
added geo transfer results

@@
 
-\usepackage{acta_astronautica_packages}
-\input{macros}
+\usepackage{acta_astronautica_packages} % packages and referencing customization
+\input{macros} % defines custom macros
 
@@
 \begin{keyword}
-Keywords
+Low-thrust, Reachability, Optimal Control, Variational Integrator
 \end{keyword}
@@
 \section{Numerical Example}\label{sec:simulation}
-A numerical simulation is presented to demonstrate the transfer procedure.
-The goal is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
+\subsection{Periodic Orbit transfer}
+We present two numerical simulations to demonstrate the transfer procedure.
+The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
 The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix} \) with \( \mu = 0.0125 \).
@@
 
-%Determine reachable set on \Poincare Section
-%
-%Intersection of the sets on section
-%
-%Minimization to a target state
-%
-%Show transfer to the moon
-%
-%Compare with the Ross work and the invariant manifold ( long time of flight)
+\subsection{Geostationary Transfer}
+This example demonstrates the feasibility of iteratively computing the reachable set. 
+In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
+Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
+It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
+The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.
+\begin{figure}[htbp]
+   \centering
+   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
+   \caption{Transfer target}
+   \label{fig:geo_transfer_target}
+\end{figure}
+In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
+This type of low energy transfer is best suited for autonomous spacecraft missions as the long time of flight is undesirable for manned spacecraft. 
+Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
+Low energy transfers from the Earth to the Lagrange points are necessary for future missions.
+
+We generate the stable invariant manifold associated with the periodic orbit.
+The stable manifold is propagated to the \Poincare section directed along the positive \( \hat{x} \) axis, or \( \alpha =\ang{0} \).
+The reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section.
+A final state is chosen from the reachable set to initialize subsequent iterations. 
+The state which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, is used as the initial condition for the next reachable set computation.
+With each reachable set we move the final trajectory closer to the stable manifold.
+\Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
+Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
+\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory found, with the final trajectory shown which ensures the intersection with the stable invariant manifold.
+Once at the stable manifold, no further control input is required and the vehicle will coast towards the target orbit.
+\begin{figure} 
+	\centering 
+	\begin{subfigure}[htbp]{0.5\textwidth} 
+		\includegraphics[width=\textwidth]{geo_transfer_full} 
+		\caption{Transfer} \label{fig:geo_transfer_full} 
+	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+	\begin{subfigure}[htbp]{0.5\textwidth} 
+		\includegraphics[width=\textwidth]{geo_transfer_zoom} 
+		\caption{State Trajectory} \label{fig:geo_transfer_zoom} 
+	\end{subfigure}
+
+	\begin{subfigure}[htbp]{0.5\textwidth} 
+		\includegraphics[width=\textwidth]{poincare} 
+		\caption{\Poincare section} \label{fig:geo_transfer_poincare} 
+	\end{subfigure} 
+	\caption{Geostationary Transfer}
+	\label{fig:geo_transfer} 
+\end{figure}
+
 \section{Conclusions}\label{sec:conclusion}

### 2015-11-22 acta_manuscript.tex
modified referencing format

@@
 A numerical example is presented demonstrating a transfer trajectory from \( L_1 \) to the vicinity of the Moon and compared to the control-free design methodology.
+
 \section{System Model}\label{sec:pcrtbp}
@@
 The reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section.
-A final state is chosen from the reachable set to initialize subsequent iterations. 
-The state which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, is used as the initial condition for the next reachable set computation.
+The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) defined on the \Poincare section.
+This allows us to approximate the set of states that are achievable in the \( \hat{x} \text{ vs. } \hat{\dot{x}} \) space.
+A final state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
 With each reachable set we move the final trajectory closer to the stable manifold.
@@
 Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
-\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory found, with the final trajectory shown which ensures the intersection with the stable invariant manifold.
-Once at the stable manifold, no further control input is required and the vehicle will coast towards the target orbit.
+\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory, with the final trajectory shown in red which ensures the intersection with the stable invariant manifold.
+Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
 \begin{figure} 

### 2015-11-23 acta_manuscript.tex
added more detail on geo transfer

@@
 \author{Shankar Kulumani\fnref{kulumanifootnote}}
-\ead{skulumani@gwu.edu}
 \address{George Washington University, 800 22nd St NW, Washington, DC 20052}
@@
 \author{Taeyoung Lee\fnref{leefootnote}}
-\ead{tylee@gwu.edu}
 \address{George Washington University, 800 22nd St NW, Washington, DC 20052}
@@
 This provides for a geometrically exact and numerically efficient method for the motion in the three-body problem.
-A numerical simulation is provided developing a transfer from a \( L_1 \) periodic orbit in  the Earth-Moon system to a target orbit about the Moon.
+Two numerical simulations are provided developing a transfer from a \( L_1 \) periodic orbit in  the Earth-Moon system to a target orbit about the Moon as well as a transfer from geostationary orbit to a \( L_1 \) periodic orbit.
 \end{abstract}
@@
 In addition, the generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer. 
+Rather than relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on a \Poincare section.
+This simple methodology allows for extended transfer trajectories which iteratively approach a target region.
 The addition of low-thrust propulsion affords the ability to enlarge the reachable set as compared to the control-free case.
@@
 
-The unstable invariant manifold is numerically propagated to the same \Poincare section defined on the \( \hat{x} \) axis.
-The intersection of the invariant manifold and the \Poincare section is denoted by the green markers in~\cref{fig:poincare_compare}.
-Only a single branch of the invariant manifold intersects with the ascending region of the target orbit.
-There are no intersections of the invariant manifold with the descending region of the target orbit.
-The numerical values associated with the green points denote the required time of flight along the invariant manifold in non-dimensional units.
-A transfer along the invariant manifold requires on average \( t_f \approx 3.1 \) as compared to \( t_f \approx 1.4 \) for a transfer using low thrust propulsion and the reachable set.
-Finally, it should be noted that an additional maneuver would be required for a transfer using the invariant manifold.
-The intersection on the \Poincare section only shows that the \( x \text{ and } \dot{x} \) components intersect.
-An additional instantaneous \( \Delta V \) would be required to transfer from the invariant manifold to the target.
-
-A periodic orbit is generated about the \( L_1 \) Lagrange point. 
-A \Poincare section is chosen to allow for design on a lower dimensional subspace.
-The section is defined along the \( \hat{x} \) axis, or defined with \( \alpha = \ang{0} \) and intersects both the initial and target orbits.
-From the periodic orbit, a series of optimal trajectories are generated to approximate the reachable set.
-The trajectories are generated from a fixed initial state of \( x_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix} \) over a fixed time span of \( t_f = 1.4 \).
-By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
-The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:transfer_orbit}.
 \begin{figure} 
@@
 \end{figure}
+
+The unstable invariant manifold is numerically propagated to the same \Poincare section defined on the \( \hat{x} \) axis.
+The intersection of the invariant manifold and the \Poincare section is denoted by the green markers in~\cref{fig:poincare_compare}.
+Only a single branch of the invariant manifold intersects with the ascending region of the target orbit.
+There are no intersections of the invariant manifold with the descending region of the target orbit.
+The numerical values associated with the green points denote the required time of flight along the invariant manifold in non-dimensional units.
+A transfer along the invariant manifold requires on average \( t_f \approx 3.1 \) as compared to \( t_f \approx 1.4 \) for a transfer using low thrust propulsion and the reachable set.
+Finally, it should be noted that an additional maneuver would be required for a transfer using the invariant manifold.
+The intersection on the \Poincare section only shows that the \( x \text{ and } \dot{x} \) components intersect.
+An additional instantaneous \( \Delta V \) would be required to transfer from the invariant manifold to the target.
+
+A periodic orbit is generated about the \( L_1 \) Lagrange point. 
+A \Poincare section is chosen to allow for design on a lower dimensional subspace.
+The section is defined along the \( \hat{x} \) axis, or defined with \( \alpha = \ang{0} \) and intersects both the initial and target orbits.
+From the periodic orbit, a series of optimal trajectories are generated to approximate the reachable set.
+The trajectories are generated from a fixed initial state of \( x_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix} \) over a fixed time span of \( t_f = 1.4 \).
+By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
+The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:transfer_orbit}.
+
 The optimal trajectories, under the influence of the control input \( \bar{u} \), are shown in~\cref{fig:reach_trajectory}.
@@
 \subsection{Geostationary Transfer}
-This example demonstrates the feasibility of iteratively computing the reachable set. 
-In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
-Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
-It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
-The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.
 \begin{figure}[htbp]
@@
 \end{figure}
+
+This example demonstrates the feasibility of iteratively computing the reachable set. 
+In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
+Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
+It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
+The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.
+
 In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
@@
 We generate the stable invariant manifold associated with the periodic orbit.
-The stable manifold is propagated to the \Poincare section directed along the positive \( \hat{x} \) axis, or \( \alpha =\ang{0} \).
-The reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section.
-The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) defined on the \Poincare section.
+The stable manifold is propagated to the \Poincare section directed along the positive \( \hat{x} \) axis, or \( \alpha =\ang{0} \) denoted by the horizontal black line.
+The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section.
+The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eqn:constraints} defined on the \Poincare section.
 This allows us to approximate the set of states that are achievable in the \( \hat{x} \text{ vs. } \hat{\dot{x}} \) space.

### 2015-12-01 acta_manuscript.tex
working on updating the introduction with comments from dr. lee

@@
 \begin{abstract}
-A computational approach is developed for the design of continuous low thrust transfers in the planar circular restricted three-body problem.
-The transfer design method of invariant manifolds is extended with the addition of continuous low thrust propulsion.
-A reachable region is generated and it is used to determine transfer opportunities, analogous to the intersection of invariant manifolds.
+We develop a computational approach for the design of continuous low thrust transfers in the planar circular restricted three-body problem.
+The use of low thrust propulsion allows the spacecraft to depart from the natural dynamics and enables a wider range of transfers.
+We generate the reachable set of the spacecraft and use this to determine transfer opportunities, analogous to the intersection of control-free invariant manifolds.
 The reachable set is developed on a lower dimensional \Poincare section and used to design transfer trajectories. 
 %Computational geometric mechanics allows for the development of variational integrators and a discrete optimal control problem.
-This is solved numerically as a discrete optimal control problem using a variational integrator.
-This provides for a geometrically exact and numerically efficient method for the motion in the three-body problem.
-Two numerical simulations are provided developing a transfer from a \( L_1 \) periodic orbit in  the Earth-Moon system to a target orbit about the Moon as well as a transfer from geostationary orbit to a \( L_1 \) periodic orbit.
+This is solved numerically as a discrete optimal control problem using a variational integrator, which preserve the geometric structure of the motion in the three-body problem.
+We demonstrate our approach with two numerical simulations of transfers in the Earth-Moon three-body system.
 \end{abstract}
@@
 Designing spacecraft trajectories is a classic and ongoing topic of research.
-With the current fiscal constraints there is an increased focus on technologies with a critical impact on mass.
+%With the current fiscal constraints there is an increased focus on technologies with a critical impact on mass.
 Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
@@
 In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{koon2000}.
-Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional conic approach~\cite{ross2006,gomez2001}.
+Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional patched-conic approach~\cite{ross2006,gomez2001}.
 Optimally combining the dynamical structure of the three-body problem with low-thrust propulsion systems is vital for future mission success.
 
-% PRIOR WORK AND CHALLENGES
+%------------------ PRIOR WORK AND CHALLENGES -----------------------------------%
+% DIRECT OPTIMAL CONTROL
 There has been extensive research focused on optimal control for spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}.
@@
 
+% CHAOTIC SYSTEM AND INITIAL GUESSES FOR OPTIMIZATION
 The application of optimal control methods for orbital trajectory design is nontrivial.
-The three-body system dynamics are nonlinear and exhibit chaotic behaviors. 
-Small changes in initial conditions result in large variations of the resulting system trajectory. 
-Therefore, any optimization routine is highly sensitive to the initial guess.
-In addition, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge which achieve satisfactory results.
+In order to implement any optimization method a sufficiently accurate ``initial guess'' is required.
+Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the optimal solution.
+However, the three-body system dynamics are nonlinear and exhibit chaotic behaviors. 
+Popularly refered to as the ``butterfly effect'' small changes in initial conditions result in large variations of the resulting system trajectory. 
+This makes solving the optimization problem highly dependent on the initial condition.
+In addition, there is no closed form analytical solution for the three-body problem~\cite{szebehely1967}.
+Since there is an insufficient number of analytical constants, or integral of motion, numerical methods must be used to investigate solutions to the three-body problem.
+As a result, accurate numerical methods are required to determine optimal solutions.
+
+%For example, analytical design is possible for complicated trajectories using a Keplerian system model via the patched-conic approach.
+
+% NUMERICAL INTEGRATION
 Efficient numerical implementation is dependent on correct initial conditions as well as accurate numerical integration.
-
-Additionally, References~\citenum{mingotti2011} and~\citenum{grebow2011} implement the solutions using conventional Runge-Kutta integration techniques.
+References~\citenum{mingotti2011} and~\citenum{grebow2011} implement the solutions using conventional Runge-Kutta integration techniques.
 These techniques suffer from numerical instability and energy drift behaviors which make them ill-suited for long-term propagation.
 These dissipative effects are even more detrimental with the addition of low-thrust propulsion to the dynamic equations of motion.
-Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system.
+Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system~\cite{hairer2006}.
 As a result, the long term effects of low-thrust on the spacecraft trajectory are not accurately captured.
 
+% DYNAMICAL STRUCTURE
 References~\citenum{koon2000} and~\citenum{ross2006} have illustrated the rich structure that exists in the three-body problem.
@@
 The addition of low-thrust propulsion offers the potential of reduced transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities. 
+%------------------ PRIOR WORK AND CHALLENGES -----------------------------------%
 
-% OBJECTIVE OF THIS WORK
-In order to address these issues, the authors propose an accurate and numerically stable method for long-term optimal orbital transfers.
-This approach will avoid the instability and dissipative effects of conventional integration schemes.
-In addition, the effects of low-thrust propulsion will be directly included in the integration method to ensure that extended duration optimal trajectories are computed correctly.
-This improved method will enable the derivation of a systematic method of generating optimal transfer orbits between arbitrary states.
-Indirect optimal control, based on the calculus of variations, will be implemented in order to generate optimal orbital transfers.
-This will avoid the approximation issues inherent in the previous work, which utilized direct optimal control methods.
+%-------------------- OBJECTIVES OF THIS WORK-----------------------------%
+% SYSTEMATIC DESIGN TO AVOID DIFFICULTIES IN CHOOSING A GOOD INITIAL GUESS
+% CAPTURE LONG TERM EFFECTS OF LOW THRUST ACCURATELY IN NUMERICAL SIMULATION
+In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem.
+Our approach avoids the difficulties in selecting an appropriate initial guess for optimization.
+We utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
+In addition, we accurately capture the effects of low thrust on the system dynamics in the numerical simulation. 
+This approach avoids the instability and dissipative effects of conventional integration schemes.
 With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low-thrust propulsion systems.
@@
 
-% SUMMARY OF CONTRIBUTIONS
+%---------------------------- EXPLICIT SUMMARY OF CONTRIBUTIONS--------------------------%
 In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.

### 2015-12-02 acta_manuscript.tex
working on intro

@@
 Our approach avoids the difficulties in selecting an appropriate initial guess for optimization.
-We utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
+Rather than We utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
 In addition, we accurately capture the effects of low thrust on the system dynamics in the numerical simulation. 
@@
 
-% TECHNICAL DESCRIPTION OF APPROACH
+% HIGH LEVEL TECHNICAL DESCRIPTION OF APPROACH
+% Application of low thrust to invariant manifold transfers. 
+% Use reachability set to determine transfer opportunities.
 To achieve these objectives, computational geometric optimal control techniques are applied. 

### 2015-12-03 acta_manuscript.tex
updating detail on invariant manifolds

--- 2015-12-01 acta_manuscript.tex
@@
 Our approach avoids the difficulties in selecting an appropriate initial guess for optimization.
-We utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
-In addition, we accurately capture the effects of low thrust on the system dynamics in the numerical simulation. 
+Instead, we utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
+In addition, through the use geometric integrators we accurately capture the effects of low thrust on the system dynamics in the numerical simulation. 
 This approach avoids the instability and dissipative effects of conventional integration schemes.
@@
 
-% TECHNICAL DESCRIPTION OF APPROACH
+% HIGH LEVEL TECHNICAL DESCRIPTION OF APPROACH
+% Generalization of invariant manifold transfers  
+% Use reachability set to determine transfer opportunities with low thrust control input
+
 To achieve these objectives, computational geometric optimal control techniques are applied. 
+We generalize the previous work of determining orbital transfers using invariant manifolds~\cite{koon2000}.
+We incorporate the effects of continuous low thrust propulsion to enable departure from the system dynamics and enable increased transfer opportunities.
 The dynamics of the three-body system are derived from the discrete Lagrangian, which approximates the integral of the continuous time Lagrangian over a fixed discrete step.
 Application of the discrete Euler-Lagrange equations, or the discrete Legendre transform, results in the discrete equations of motion.
-This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long time periods.
+This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs.
 A discrete optimal control problem is formulated from the discrete equations of motion.
@@
 Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft. 
-Generation of this reachability set allows for the extension of the previous control-free missions in the three-body problem.
+Generation of this reachability set allows for the extension of the previous methods based on invariant manifolds in the three-body problem.
 In addition, the generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer. 
-Rather than relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on a \Poincare section.
+The \Poincare section reduces the dimensionality of the system dyanmics to the study of a related discrete update map.  
+This allows for the study of the complex dynamics of the three-body problem to a lower dimensional space.
+Rather than relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on an appropriate \Poincare section.
 This simple methodology allows for extended transfer trajectories which iteratively approach a target region.
@@
 The use of the \Poincare section allows for design on a lower dimensional space and simplifies the design process.
-Once an intersection is determined on the \Poincare section a transfer is calculated.
+%Once an intersection is determined on the \Poincare section a transfer is calculated.
 
@@
 A numerical example is presented demonstrating a transfer trajectory from \( L_1 \) to the vicinity of the Moon and compared to the control-free design methodology.
+\begin{itemize}
+	\item generalize invariant manifold transfers with continuous control input
+	\item Low thrust propulsion rather than impulsive manuevers
+	\item Combination of reachable set and \Poincare section
+	\item Geometric integrators for accurate numerical simulation
+\end{itemize}
 
-\section{System Model}\label{sec:pcrtbp}
+\section{Planar Circular Restricted Three Body Problem}\label{sec:pcrtbp}
 The system model is based on the planar circular restricted three body problem~(PCRTBP).
@@
 The equations of motion are developed in a rotating reference frame which allows for much greater insight into the structure of the dynamics.
-The \( \hat{x} \) axis is directed along the vector from the Earth to the Moon.
-The \( \hat{y} \) axis lies in the orbital plane and is orthogonal to \( \hat{x} \).
+The \( \hat{e}_1 \) axis is directed along the vector from the Earth to the Moon.
+The \( \hat{e}_2 \) axis lies in the orbital plane and is orthogonal to \( \hat{e}_1 \).
 The rotating reference frame is centered at the system barycenter.
-It is assumed that the \(\left( \hat{x}, \hat{y}\right)\) rotates with a constant angular velocity equal to the mean motion of the Moon.
+It is assumed that the \(\left( \hat{e}_1, \hat{e}_2\right)\) frame rotates with a constant angular velocity equal to the mean motion of the Moon.
 Following convention, the system is also non-dimensionalized by the characteristic mass, length, and time~\cite{koon2000}.
@@
 \end{equation}
-The larger primary, \(m_1\), is located at \( \left(  -\mu , 0 \right)\) and the smaller \( m_2\) is located at \( \left( 1-\mu , 0 \right)\).
+The larger primary, \(m_1\), is located at \( \left(  -\mu , 0 \right)\) and the smaller \( m_2\) is located at \( \left( 1-\mu , 0 \right)\), while the spacecraft is located at \( x \hat{e}_1 + y \hat{e}_2 \).
 In the rotating reference frame the Lagrangian is given by
@@
 \end{equation}
-where the distance \(r_1\) and \(r_2\) of the spacecraft to each primary is defined in rotating coordinates as
+where the distance \( r_1, r_2 \in \R \) of the spacecraft to each primary is defined in rotating coordinates as
 \begin{align}
@@
 \begin{align}
-	\ddot{x} - 2 \dot{y} + \frac{\partial}{\partial x} U &= u_x \, ,\nonumber \\
-	\ddot{y} + 2 \dot{y} + \frac{\partial}{\partial y} U &= u_y \, ,
+	\ddot{x} - 2 \dot{y} + \deriv{U}{x} &= u_x \, ,\nonumber \\
+	\ddot{y} + 2 \dot{y} + \deriv{U}{y} &= u_y \, ,
 	\label{eq:cont_eom}
@@
 \end{equation}
-and the control inputs are defined as \( u_x\) and \(u_y\).
-The state is defined as \( \bar{x} = \begin{bmatrix}\bar{r} &\bar{v} \end{bmatrix}\) with \(\bar{r} \in \R^{2\times1}\) and \(\bar{v} \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
+and the control inputs are defined as \( u_x \text{ and } u_y \in \R \).
+The state is defined as \( \bar{x} = \begin{bmatrix}\bar{r} &\bar{v} \end{bmatrix}^T\) with \(\bar{r} \in \R^{2\times1}\) and \(\bar{v} \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
 The equations of motion may be rewritten in state space form as
-\begin{equation}
+\begin{equation}\label{eqn:cont_dyn}
 	\left[\begin{array}{c} \dot{\bar{r}} \\ \dot{\bar{v}} \end{array} \right] = 
@@
 
-\subsection{Jacobi Integral}
+\subsection{Jacobi Integral}\label{sec:jacobi}
 There exists a single integral, or constant of motion for the three-body problem~\cite{lanczos1970,szebehely1967}.
@@
 	\includegraphics[width=0.5\textwidth]{energy_contours}
-	\caption{Contour Plot of Jacobi Integral}
+	\caption{Contour plot of Jacobi integral}
 	\label{fig:energy_contour}
@@
 Taking the variation of~\cref{eq:action_integral} gives
-\begin{subequations}\label{eq:var_principle}
-\begin{align}
-	\delta S &= \int_{0}^T \deriv{L}{q} \delta q + \deriv{L}{\dot{q}} \delta \dot{q} \, dt \\
-		&= \int_{0}^T \deriv{L}{q} \delta q - \frac{d}{dt} \left( \deriv{L}{\dot{q}}\right) \delta q \, dt - \left. \left[ \deriv{L}{\dot{q}} \delta q\right] \right|_0^T \\
+\begin{align}\label{eq:var_principle}
+	\delta S %&= \int_{0}^T \deriv{L}{q} \delta q + \deriv{L}{\dot{q}} \delta \dot{q} \, dt \\
+		%&= \int_{0}^T \deriv{L}{q} \delta q - \frac{d}{dt} \left( \deriv{L}{\dot{q}}\right) \delta q \, dt - \left. \left[ \deriv{L}{\dot{q}} \delta q\right] \right|_0^T \\
 	&= \int_{0}^T \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}}	\right) \, dt \, ,
 \end{align}
-\end{subequations} 
+
 where we have used integration by parts and the conditions \( \delta q(0) = 0 \) and \( \delta q(T) = 0\).
@@
 However, the midpoint rule results in an implicit form which adds further complexity to the equations of motion.
-In this work, the trapezoidal approximation is applied to the PCRTBP.
+%In this work, the trapezoidal approximation is applied to the PCRTBP.
 
@@
 Once again a discrete version of Hamilton's principle is applied to~\cref{eq:action_sum}.
-Applying summation by parts yields
-\begin{subequations}
+This yields a discrete time counterpart to~\cref{eq:var_principle} as
 \begin{align}\label{eq:dis_var_principle}
-	\delta S_d &= \sum_{k=0}^{N-1} \deriv{L_d(q_k, q_{k+1})}{q_k} \delta q_k + \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \delta q_{k+1} \\
+	\delta S_d %&= \sum_{k=0}^{N-1} \deriv{L_d(q_k, q_{k+1})}{q_k} \delta q_k + \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \delta q_{k+1} \\
 	&= \sum_{k=1}^{N-1} \bracket{ \deriv{L_d(q_k, q_{k+1})}{q_k} + \deriv{L_d(q_{k-1}, q_k)}{q_{k+1}}} \delta q_k \, .
 \end{align}
-\end{subequations}
+
 For the discrete action sum to be stationary with respect to all admissible path variations, with fixed endpoints, the discrete Euler-Lagrange equations must be satisfied for \( k = 1, \cdots, N-1 \) resulting in
@@
 The discrete equations of motion are given in the Lagrangian form after applying the discrete fiber derivative from~\cref{eq:discrete_legendre} as \( p_{\xk} = \dot{x}_k - \yk \) and \( p_{\yk} = \dot{y}_k + \xk \).
-The state is defined as \( \bar{ x}_k = \begin{bmatrix} \xk & \yk & \dot{x}_k & \dot{y}_k \end{bmatrix}^T\) and the control input is \( \bar{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \).
+The state is defined as \( \vec{x}_k = \begin{bmatrix} \xk & \yk & \dot{x}_k & \dot{y}_k \end{bmatrix}^T\) and the control input is \( \vec{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \).
+This results in a discrete update map \( f_k: \vec{x}_k \to \vec{x}_{k+1} \) which preserves the same properties of the continuous-time dynamics in~\cref{eqn:cont_dyn} such as invariants, symplecticity, and the configuration manifold.
 The discrete potential gradients are given by
@@
 Finally, the update steps in~\cref{eq:xdotkp,eq:ydotkp} are evaluated.
-This results in the complete discrete update map \( \bar{x}_k \to \bar{x}_{k+1} \) given \( \bar{u}_k \).
+This results in the complete discrete update map \( \vec{x}_k \to \vec{x}_{k+1} \) given \( \vec{u}_k \).
 
@@
 	\centering 
-	\begin{subfigure}[h]{0.3\textwidth} 
+	\begin{subfigure}[h]{0.5\textwidth} 
 		\includegraphics[width=\textwidth]{trajectory} 
@@
 	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-	\begin{subfigure}[htbp]{0.3\textwidth} 
-		\includegraphics[width=\textwidth]{components} 
-		\caption{State \( \parenth{x , y, \dot{x}, \dot{y}}\)} \label{fig:compare_components} 
-	\end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-	\begin{subfigure}[htbp]{0.3\textwidth} 
+%	\begin{subfigure}[htbp]{0.3\textwidth} 
+%		\includegraphics[width=\textwidth]{components} 
+%		\caption{State \( \parenth{x , y, \dot{x}, \dot{y}}\)} \label{fig:compare_components} 
+%	\end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+	\begin{subfigure}[htbp]{0.5\textwidth} 
 		\includegraphics[width=\textwidth]{energy} 
@@
 A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
-A particle is simulated from an initial condition of \( \bar{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
+A particle is simulated from an initial condition of \( \vec{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
 The variational integrator uses a step size of \SI{47.22}{\second} while the Runge-Kutta method uses a variable step size implemented via ODE45 in Matlab.
@@
 
-\section{Invariant Manifolds} \label{sec:invariant_manifold}
-% introduce \Poincare section idea
-There exists a rich structure of tubes, or invariant manifolds, in the three-body problem~\cite{koon2000,conley1968}.
+\section{Invariant Manifolds and \Poincare Map} \label{sec:invariant_manifold}
+% discuss usefulness and basic generation of invariant manifolds
+
+%Planar problem has four dimensional phase space
+%associated with periodic solutions about lagrange equilibrium points
+%tubes, invariant manifolds are four dimenionsal surface which serve to separate classes of trajectories
+%Localized linear analysis, Monodromy matrix, is used to study and compute invariant manifold
+
+% intro to invariant manifolds
+There exist five equilibrium points in the equations of motion for the three-body problem as introduced in~\cref{sec:jacobi}.
+It has been shown that the local orbit structure near the Lagrange points give rise to families of periodic orbits as well as the stable and unstable manifolds of these periodic orbits.
+This rich structure is globally connected and gives rise to a dynamical chain which allows trajectories to pass through the phase space~\cite{koon2000,conley1968}.
 The manifold structure associated with periodic orbits about the \( L_1 \) and \( L_2 \) Lagrange points are critical to the understanding of the motion of spacecraft as well as comets/asteroids.
@@
 These invariant manifolds only exist as a result of the dynamic formulation of the three-body problem in a rotating reference frame.
-In this way, it is possible to design trajectories that intersect the invariant manifolds and connect widely separated regions of the state space. 
-\Cref{fig:invariant_manifolds} shows some of the tubes, projected from the phase space onto the configuration space, associated with \( L_1 \) and \( L_2 \) periodic orbits in the Earth-Moon system. 
-This technique has been heavily investigated in Reference~\citenum{koon2000}, applied to operational missions in Reference~\citenum{koon1999}, and applied to potential multi-moon orbiter missions in Reference~\citenum{tanaka2011}.
-In addition, much recent work has focused on potential return missions to the Moon~\cite{zanzottera2012,campagnola2012,mingotti2011,ozimek2010a,mingotti2009}.
+
+% technical definition of manifold and some equations
+Invariant manifolds serve as a higher dimensional generalization of the concept of seperatrices from linear systems to the case of nonlinear systems. 
+To compute these manifolds first one must determine an appropriate periodic orbit. 
+The process of differential correction and numerical continuation are used to determine a periodic orbit with a desired energy level. 
+Let the trajectories of the equations of motion defined in~\cref{eqn:cont_dyn} be defined by the flow map \( \phi(t;\vec{x}_0) \) originating from the initial condition \( \vec{x}(t_0) = \vec{x}_0 \) and and a periodic orbit denoted by \( \gamma\).
+The invariant manifolds: 
+\begin{align*}
+	W^s ( \vec{x}_0 ) &= \braces{ \vec{x} : \phi(t;\vec{x}_0) \to \gamma \text{ as } t \to \infty} \\
+	W^u ( \vec{x}_0 ) &= \braces{ \vec{x} : \phi(t;\vec{x}_0) \to \gamma \text{ as } t \to -\infty} 
+\end{align*}
+are the set of trajectories that asymptotically depart or approach a periodic orbit.
+For the nonlinear system~\cref{eqn:cont_dyn} or~\cref{eq:discrete_eoms}, we can determine the state transition matrix \( \Phi \) along the periodic orbit, from \( 0 \to T\), by numerical integration of the variational equations.
+In the neighborhood of the periodic orbit \( \gamma \), the tangent spaces to the stable and unstable manifolds are provided by the generalized eigenspaces \( E^s \) and \( E^u \) of the linearization \( A = \D f(\vec{x}_0) \) where \( \vec{x}_0 \) is on the periodic orbit.
+The eigenvectors of the monodromy matrix, \( M = \Phi(T) \), serve to locally approximate the directions of the stable and unstable manifolds. 
+A perturbation in the direction of the unstable eigenvector, outside the unit circle, is used to generate an initial condition on the unstable manifold. 
+Similarly, a perturbation in the direction of the stable eigenvector, inside the unit circle, is used to generate an initial condition on the stable manifold. 
+Each periodic orbit has two stable and unstable manifold sets associated with both a postive or negative perturbation along the respective eigenvector.
+This idea is illustrated by~\cref{fig:invariant_manifold}.
 \begin{figure}
 	\centering
-        \begin{subfigure}[b]{0.25\textwidth}
-                \includegraphics[width=\columnwidth]{U1_Manifolds}
-                \caption{U1 Manifolds}
-                \label{fig:u1_manifolds}
-        \end{subfigure}%
-        ~%add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
-          %(or a blank line to force the subfigure onto a new line)
-        \begin{subfigure}[b]{0.25\textwidth}
-                \includegraphics[width=\columnwidth]{U2_Manifolds}
-                \caption{U2 Manifolds}
-                \label{fig:u2_manifolds}
-        \end{subfigure}~
-        \begin{subfigure}[b]{0.25\textwidth}
-                \includegraphics[width=\columnwidth]{U3_Manifolds}
-                \caption{U3 Manifolds}
-                \label{fig:u3_manifolds}
-        \end{subfigure}%
-        ~%add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
-          %(or a blank line to force the subfigure onto a new line)
-        \begin{subfigure}[b]{0.25\textwidth}
-                \includegraphics[width=\columnwidth]{U4_Manifolds}
-                \caption{U4 Manifolds}
-                \label{fig:u4_manifolds}
-        \end{subfigure}
-        \caption{Invariant Manifolds for Planar Earth-Moon three-body system}
-	\label{fig:invariant_manifolds}
-\end{figure}
-\begin{figure}
-     \centering
-        \begin{subfigure}[b]{0.25\textwidth}
-                \includegraphics[width=\columnwidth]{U1_poincare}
-                \caption{U1 Section}
-                \label{fig:u1_poincare}
-        \end{subfigure}%
-        ~%add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
-          %(or a blank line to force the subfigure onto a new line)
-        \begin{subfigure}[b]{0.25\textwidth}
-                \includegraphics[width=\columnwidth]{U2_poincare}
-                \caption{U2 Section}
-                \label{fig:u2_poincare}
-        \end{subfigure}~
-        \begin{subfigure}[b]{0.25\textwidth}
-                \includegraphics[width=\columnwidth]{U3_poincare}
-                \caption{U3 Section}
-                \label{fig:u3_poincare}
-        \end{subfigure}%
-        ~%add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
-          %(or a blank line to force the subfigure onto a new line)
-        \begin{subfigure}[b]{0.25\textwidth}
-                \includegraphics[width=\columnwidth]{U4_poincare}
-                \caption{U4 Section}
-                \label{fig:u4_poincare}
-        \end{subfigure}
-        \caption{\Poincare Sections for Planar Earth-Moon three-body system}
-	\label{fig:poincare_sections}
+	\includegraphics[width=0.5\textwidth]{invariant_manifold}
+	\caption{Stable and unstable manifold of a periodic orbit\label{fig:invariant_manifold}}
 \end{figure}
 
+% introduce \Poincare section theory and math
 Another useful technique in the analysis of the free motion of dynamical systems is that of the \Poincare section.
@@
 Intersecting states are easy to determine and allow for transfers between invariant manifolds.
-
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth]{poincare_map}
+	\caption{\Poincare map\label{fig:poincare_map}}
+\end{figure}
+% intro to how you can determine intersection on a poincare section for transfers
 The intersection of the invariant manifolds on a \Poincare section is central to determining the desired control-free trajectory.
@@
 The reachable set is computed on an appropriate \Poincare section and is used to design an appropriate transfer trajectory.
+
+% detail on how invariant manifolds can be used for transfers (example of koon work)
+In this way, it is possible to design trajectories that intersect the invariant manifolds and connect widely separated regions of the state space. 
+\Cref{fig:invariant_manifolds} shows some of the tubes, projected from the phase space onto the configuration space, associated with \( L_1 \) and \( L_2 \) periodic orbits in the Earth-Moon system. 
+This technique has been heavily investigated in Reference~\citenum{koon2000}, applied to operational missions in Reference~\citenum{koon1999}, and applied to potential multi-moon orbiter missions in Reference~\citenum{tanaka2011}.
+In addition, much recent work has focused on potential return missions to the Moon~\cite{zanzottera2012,campagnola2012,mingotti2011,ozimek2010a,mingotti2009}.
+\begin{figure}
+     \centering
+        \begin{subfigure}[b]{0.5\textwidth}
+                \includegraphics[width=\columnwidth]{U1_Manifolds}
+                \caption{U1 Manifolds}
+                \label{fig:u1_manifolds}
+        \end{subfigure}%
+        ~
+        \begin{subfigure}[b]{0.5\textwidth}
+                \includegraphics[width=\columnwidth]{U1_poincare}
+                \caption{U1 Section}
+                \label{fig:u1_poincare}
+        \end{subfigure}
+        \caption{Invariant manifold}
+	\label{fig:u1}
+\end{figure}
+
 \section{Optimal Control Formulation for Reachability Set}\label{sec:optimal_control}
@@
 Computing the reachability set given estimated states and control authorities allows one to better correlate subsequent measurements or determine sensor pointing regions in the event of a lost spacecraft.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth]{reachability_set}
+	\caption{Reachability set\label{fig:reachability_set}}
+\end{figure}
 

### 2015-12-04 acta_manuscript.tex
working on poincare section detail

@@
 With the potential for more demanding missions, even greater importance is placed on the mission design to ensure that optimal trajectories satisfy mission requirements. 
-In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{koon2000}.
+In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{koon2011}.
 Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional patched-conic approach~\cite{ross2006,gomez2001}.
@@
 To achieve these objectives, computational geometric optimal control techniques are applied. 
-We generalize the previous work of determining orbital transfers using invariant manifolds~\cite{koon2000}.
+We generalize the previous work of determining orbital transfers using invariant manifolds~\cite{koon2011}.
 We incorporate the effects of continuous low thrust propulsion to enable departure from the system dynamics and enable increased transfer opportunities.
@@
 It is assumed that the \(\left( \hat{e}_1, \hat{e}_2\right)\) frame rotates with a constant angular velocity equal to the mean motion of the Moon.
-Following convention, the system is also non-dimensionalized by the characteristic mass, length, and time~\cite{koon2000}.
+Following convention, the system is also non-dimensionalized by the characteristic mass, length, and time~\cite{koon2011}.
 As a result, the system can be characterized by a single mass parameter \( \mu \),
@@
 The process of differential correction and numerical continuation are used to determine a periodic orbit with a desired energy level. 
-Let the trajectories of the equations of motion defined in~\cref{eqn:cont_dyn} be defined by the flow map \( \phi(t;\vec{x}_0) \) originating from the initial condition \( \vec{x}(t_0) = \vec{x}_0 \) and and a periodic orbit denoted by \( \gamma\).
+Let the trajectories of the equations of motion defined in~\cref{eqn:cont_dyn} be defined by the flow map \( \phi(t;\vec{x}_0) \) originating from \( \vec{x}(t_0) = \vec{x}_0 \) and we define a periodic orbit as \( \gamma\).
 The invariant manifolds: 
@@
 For the nonlinear system~\cref{eqn:cont_dyn} or~\cref{eq:discrete_eoms}, we can determine the state transition matrix \( \Phi \) along the periodic orbit, from \( 0 \to T\), by numerical integration of the variational equations.
-In the neighborhood of the periodic orbit \( \gamma \), the tangent spaces to the stable and unstable manifolds are provided by the generalized eigenspaces \( E^s \) and \( E^u \) of the linearization \( A = \D f(\vec{x}_0) \) where \( \vec{x}_0 \) is on the periodic orbit.
+In the neighborhood of the periodic orbit \( \gamma \), the tangent spaces to the stable and unstable manifolds are provided by the generalized eigenspaces \( E^s \) and \( E^u \) of the linearization \( A = \D f(\vec{x}_0) \), where \( \vec{x}_0 \) is on the periodic orbit.
 The eigenvectors of the monodromy matrix, \( M = \Phi(T) \), serve to locally approximate the directions of the stable and unstable manifolds. 
-A perturbation in the direction of the unstable eigenvector, outside the unit circle, is used to generate an initial condition on the unstable manifold. 
-Similarly, a perturbation in the direction of the stable eigenvector, inside the unit circle, is used to generate an initial condition on the stable manifold. 
-Each periodic orbit has two stable and unstable manifold sets associated with both a postive or negative perturbation along the respective eigenvector.
-This idea is illustrated by~\cref{fig:invariant_manifold}.
+A perturbation in the direction of the unstable eigenvector, with \( \norm{\lambda} > 1 \), is used to generate an initial condition on the unstable manifold. 
+Similarly, a perturbation in the direction of the stable eigenvector, with \( \norm{\lambda} < 1 \), is used to generate an initial condition on the stable manifold. 
+Each periodic orbit has two stable and unstable manifold sets associated with both a positive or negative perturbation along the respective eigenvector.
+This is technique of determining the invariant manifolds is graphically presented in~\cref{fig:invariant_manifold}.
 \begin{figure}
@@
 % introduce \Poincare section theory and math
-Another useful technique in the analysis of the free motion of dynamical systems is that of the \Poincare section.
-The \Poincare section is the intersection of a periodic orbit of a dynamical system with a lower dimensional sub-space transverse to the flow.
-This allows for greater insight into the stability and dynamics of periodic solutions of dynamic system. 
-Points are drawn as the periodic solution intersects the \Poincare section.
-Great insight and structure into the dynamical system is typically available through use of \Poincare section.
-For example, \cref{fig:poincare_sections} shows the intersections of the invariant manifolds of~\cref{fig:invariant_manifolds} with the \Poincare sections defined by the black line segments.
-The \Poincare section and the Jacobi integral reduce the state space from four to two dimensions for the planar three-body problem~\cite{koon2001}.
-This greatly eases the design process and allows for a simple planar visualization of the intersection of the invariant manifolds.
-Intersecting states are easy to determine and allow for transfers between invariant manifolds.
+\Poincare maps are a useful tool in the analysis of the flow near periodic orbits in the three-body problem.
+We let \( \Sigma \) be a hypersurface of section chosen such that all trajectories in the vicinity of a state \( q \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
+A \Poincare map, \( P(q) = \phi(T;q) \), maps the state of a trajectory from one intersection to the next.
+Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
+This allows for greater insight into the stability and dynamics of periodic solutions of dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories.
+For example, \Poincare maps have been used to prove the existence of homoclinic orbits, orbits which are both forward and backward asymptotic to a single unstable periodic orbit, and heteroclinic orbits, which join different periodic orbits~\cite{conley1968,koon2000b}.
+These dynamic features have been shown to play a vital role in the movement of natural bodies as well as critical for spacecraft missions~\cite{lo1997,gomez2001}.
 \begin{figure}
@@
 \end{figure}
-% intro to how you can determine intersection on a poincare section for transfers
+
+% intersect invariant manifolds on poincare section
+% find intersection between stable and unstable manifolds
+% define the sections used in Koon2011 (placed strategically near heteroclinic points) - rotation of manifolds
+Combining invariant manifolds and an appropriate \Poincare section provides a simple manner to determine trajectories which cross wide regions of the phase space.
 The intersection of the invariant manifolds on a \Poincare section is central to determining the desired control-free trajectory.
@@
 
-% detail on how invariant manifolds can be used for transfers (example of koon work)
 In this way, it is possible to design trajectories that intersect the invariant manifolds and connect widely separated regions of the state space. 

### 2015-12-04 acta_manuscript.tex
starting on revising the results section

@@
 % define the sections used in Koon2011 (placed strategically near heteroclinic points) - rotation of manifolds
-Combining invariant manifolds and an appropriate \Poincare section provides a simple manner to determine trajectories which cross wide regions of the phase space.
-The intersection of the invariant manifolds on a \Poincare section is central to determining the desired control-free trajectory.
-However, the results previously developed are highly case specific and difficult to generalize to arbitrary transfers.
-Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system.
-In addition, transfer orbits along an invariant manifold require large time of flights which may be undesirable for time critical missions.
-The addition of low-thrust propulsion offers the potential of reduced transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities. 
-In this work, an optimal control problem is formulated to generate a reachable set of the spacecraft.
-The reachable set is computed on an appropriate \Poincare section and is used to design an appropriate transfer trajectory.
+Combining invariant manifolds and an appropriate \Poincare section provides a conceptually simple manner to determine trajectories which connect wide regions of the phase space.
+\Cref{fig:manifold_transfer_example} demonstrates how a \Poincare map may be used to determine a connection between periodic orbits.
+The initial orbit is a planar periodic orbit about \( L_2 \) in the Earth-Moon system while the final orbit is a periodic orbit about \( L_1 \).
+Both periodic orbits have the Jacobi energy constant and as a result it is possible to transfer between the orbits without any energy change.
+\Cref{fig:manifolds} shows the unstable manifold of the initial orbit and the stable manifold of the desired orbit. 
+Both invariant manifolds are globalized and propogated to the first intersection with the \Poincare surface of section located at the position of the Moon, \( 1 - \mu \).
+The \Poincare section in~\cref{fig:manifold_transfer_example} is defined as
+\begin{align*}
+	\Sigma = \braces{ ( y, \dot{y}) \, | \, x = 1- \mu, y > 0} , 
+\end{align*}
+and located to allow an intersection between both invariant manifolds.
+The configuration space of the invariant manifold is \( \vec{x} \in \R^4 \).
+The surface of section, and the resulting intersection of the four-dimensional space with the hyperplane \( \Sigma \), reduces the resulting intersection to a surface in \( \R^3\).
+Since the two orbits share the same Jacobi constant, the intersection phase space is further reduced to \( \R^2\).
+The intersection of the invariant manifolds on the \Poincare section is central to determining transfer connections between the initial and final orbits.
 
+% describe this specific poincare section and why it's placed and the axes
+The position, \( y \), and velocity, \( \dot{y} \), at the \Poincare section are plotted in~\cref{fig:poincare}.
+The surface of section is placed at a strategic location in order to allow for a cross section of the flow within the three-dimensional energy surface. 
+In a frame aligned with the rotating frame, \( ( \hat{e}_1, \hat{e}_2 ) \), but centered on the equilibrium point \( L_1 \) or \( L_2\), the unstable manifolds are locally moving in forward time in the second or forth quadrant.
+Similarly, the stable manifold branches are locally moving in backward time in the first and third quadrants. 
+As a result, the location of the section in~\cref{fig:poincare} is chosen to intersect both the stable manifolds of \( L_1 \) periodic orbits as well as unstable manifolds of \( L_2 \) periodic orbits.
+Additional \Poincare sections are located at other heteroclinic points in the phase space to enable design throughout the phase space~\cite{koon2011}.
 In this way, it is possible to design trajectories that intersect the invariant manifolds and connect widely separated regions of the state space. 
-\Cref{fig:invariant_manifolds} shows some of the tubes, projected from the phase space onto the configuration space, associated with \( L_1 \) and \( L_2 \) periodic orbits in the Earth-Moon system. 
 This technique has been heavily investigated in Reference~\citenum{koon2000}, applied to operational missions in Reference~\citenum{koon1999}, and applied to potential multi-moon orbiter missions in Reference~\citenum{tanaka2011}.
@@
         \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\columnwidth]{U1_Manifolds}
-                \caption{U1 Manifolds}
-                \label{fig:u1_manifolds}
+                \includegraphics[width=\columnwidth]{U2_Manifolds}
+                \caption{Invariant manifolds}
+                \label{fig:manifolds}
         \end{subfigure}%
@@
         \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\columnwidth]{U1_poincare}
+                \includegraphics[width=\columnwidth]{U2_poincare}
                 \caption{U1 Section}
-                \label{fig:u1_poincare}
+                \label{fig:poincare}
         \end{subfigure}
-        \caption{Invariant manifold}
-	\label{fig:u1}
+        \caption{Illustration of process to identify transfer trajectories}
+	\label{fig:manifold_transfer_example}
 \end{figure}
+
+% drawbacks of this approach on how we seek to improve upon it
+However, the results previously developed are highly case specific and difficult to generalize to arbitrary transfers.
+Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system.
+In addition, transfer orbits along an invariant manifold require large time of flights which may be undesirable for time critical missions.
+The addition of low-thrust propulsion offers the potential of reduced transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities. 
+In this paper, an optimal control problem is formulated to generate a reachable set of the spacecraft.
+The reachable set is computed on an appropriate \Poincare section and is used to design an appropriate transfer trajectory.
 
@@
 This method allows for one to easily determine potential transfer opportunities by finding set intersections on a lower dimensional space and greatly reduces the design process.
-The addition of continuous low thrust propulsion extends the control free design process presented in Reference~\citenum{koon2000} and allows for a greater range of potential transfers with a reduced time of flight.
+The addition of continuous low thrust propulsion extends the control free design process developed previously and allows for a greater range of potential transfers with a reduced time of flight.
 
@@
 Reachability theory has been applied to aerospace systems such as collision avoidance, safety planning, and performance characterization.
-The theory formally supporting reachability has been extensively developed and is directly derivable from optimal control theory~\cite{varaiya2000,lygeros2002,lygeros2004}.
+The theory formally supporting reachability has been extensively developed and is directly derivable from optimal control theory~\cite{varaiya2000,lygeros2002,lygeros2004,holzinger2009,komendera2012a}.
 Computation of the reachable set for a system involves solving the Hamilton-Jacobi partial differential equation or satisfying a dynamic programming principle.
@@
 
-Reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a}.
+Reachability theory has recently been applied to space systems~\cite{}.
 Computation of reachable sets is critical to space situational awareness, rendezvous and proximity operations, and orbit determination operations.

### 2015-12-05 acta_manuscript.tex
working on updating the results section

@@
 The application of variational integrators to optimal control problems is referred to as computational geometric optimal control.
-This formulation is based on the concept of the reachability set on a \Poincare section.
+Our formulation is based on the concept of the reachability set on a \Poincare section.
 This method allows for one to easily determine potential transfer opportunities by finding set intersections on a lower dimensional space and greatly reduces the design process.
@@
 Reachability theory has been applied to aerospace systems such as collision avoidance, safety planning, and performance characterization.
-The theory formally supporting reachability has been extensively developed and is directly derivable from optimal control theory~\cite{varaiya2000,lygeros2002,lygeros2004,holzinger2009,komendera2012a}.
+The theory formally supporting reachability has been extensively developed and is directly derivable from optimal control theory~\cite{varaiya2000,lygeros2002,lygeros2004}.
+Reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a}.
 Computation of the reachable set for a system involves solving the Hamilton-Jacobi partial differential equation or satisfying a dynamic programming principle.
@@
 
-Reachability theory has recently been applied to space systems~\cite{}.
 Computation of reachable sets is critical to space situational awareness, rendezvous and proximity operations, and orbit determination operations.
-Specifically maintaining accurate estimates of a spacecraft state over extended periods is not trivial.
+Specifically, maintaining accurate estimates of a spacecraft state over extended periods is not trivial.
 The challenge is increased for multiple spacecraft operating in close proximity or when there are long periods of time between measurements.
@@
 Computing the reachability set given estimated states and control authorities allows one to better correlate subsequent measurements or determine sensor pointing regions in the event of a lost spacecraft.
+
+In this paper, we seek to approximate the reachability set on a \Poincare set by solving a related optimal control problem. 
+We choose our \Poincare section in a similar manner to those used previously for the design of transfers via invariant manifolds.
+The \Poincare section is chosen to intersect transversally with trajectories emanating from the initial orbit. 
+In the case of a periodic orbit the trajectories will cross the \Poincare section at two distinct fixed points every half period.
+The addition of low thrust propulsion allows us to enlarge the set of trajectories achievable. 
+\Cref{fig:reachability_set} illustrates how, without any control input, trajectories will intersect with the \Poincare section at \( \vec{x}_n \). 
+However, the addition of low thrust propulsion allows the spacecraft to depart from the natural dynamics and intersect the \Poincare section at a different location.
+We use a cost function to define a distance metric on the \Poincare section from the control-free intersection to an intersection under the influence of the control input.
+Maximization of this distance enables us to generate the largest reachability set under the bounded control input.
 \begin{figure}
@@
 
-The cost function is defined as
+We define a  cost function as
 \begin{equation}
-	J = -\frac{1}{2} \left( \bar{x}(N) - \bar{x}_{n}(N)\right)^T 
+	J = -\frac{1}{2} \left( \vec{x}(N) - \vec{x}_{n}(N)\right)^T 
 	\left[
@@
 	\right]
-	\left( \bar{x}(N) - \bar{x}_{n}(N)\right) \, .
+	\left( \vec{x}(N) - \vec{x}_{n}(N)\right) \, .
 	\label{eq:cost}
 \end{equation}
-The term \( \bar{x}_n(N) \) is the final state of a control-free trajectory while the term \( \bar{x}(N) \) is the final state under the influence of the control input.
-In this fashion, the aim is to maximize the distance of the final state from that of the control-free trajectory. 
-A chosen \Poincare section is defined through the use of appropriate terminal constraints given by
+The term \( \vec{x}_n(N) \) is the final state of a control-free trajectory while the term \( \vec{x}(N)\) is the final state under the influence of the control input.
+Maximization of the distance between \( \vec{x}_n \) and \(\vec{x} \) is equivalent to the minimization of \( J \).
+The \Poincare section is defined through the use of appropriate terminal constraints given by
 \begin{subequations}
 \begin{align}
-    m_1 &= 0 = \frac{y(N) - L_{1y}}{x(N) - L_{1x}} - \tan{\alpha_d} \, , \\ 
-    m_2&= 0 = \frac{\dot{x}(N) - \dot{x_n}(N) }{x(N) -x_n(N) } - \tan{\theta_d} \, , \\
-	 0 &\geq\bar{u}^T \bar{u} - u_{max}^2 \, ,
+    m_1( \vec{x}(N) ) &= 0 = y(N) \, , \\ 
+    m_2 ( \vec{x}(N) )&= 0 = \frac{\dot{x}(N) - \dot{x}_n(N) }{x(N) -x_n(N) } - \tan{\theta_d} \, , \\
+	 0 &\geq\vec{u}^T \vec{u} - u_{max}^2 \, ,
 \end{align}
@@
 \end{subequations}
-where the angles \( \alpha_d\) and \( \theta_d\) define the \Poincare section and a specific direction upon the section, respectively. 
-The goal is to determine the control input \( \bar{u}_k\) such that the cost function~\cref{eq:cost} is minimized subject to the state equations of motion~\cref{eq:discrete_eoms} and constraints~\cref{eq:constraints}.
+where the angle \( \theta_d\) a direction to maximize the reachability set on the \Poincare section.
+We define the \Poincare section along the horizontal axis and equivalent to the  surface \( y(N) = 0 \).
+This is similar to the previous work in determining homoclinic orbits in the three-body problem~\cite{llibre1985,koon2011}.
+Previous analytical results have shown that homoclinic orbits intersect transversally in the \( (x, \dot{x} ) \) space on the plane \( y = 0 \).
+We seek to compliment these results with the addition of low thrust propulsion.
+The maximum control thrust magnitude is defined by \( u_max \) and is non-dimensionalized by the characteristic length,mass, and time.
+The goal is to determine the control input \( \vec{u}_k\) such that the cost function~\cref{eq:cost} is minimized subject to the state equations of motion~\cref{eq:discrete_eoms} and constraints~\cref{eq:constraints}.
+
 Application of the Euler-Lagrange equations allows us to derive the necessary conditions for optimality~\cite{bryson1975}.
@@
 \begin{align}
-	\lambda_{k+1}^T &= \lambda_k^T  \deriv{f_k}{\bar{x}_k}^{-1} \, , \\
-	0 &=  \deriv{H_k}{\bar{u}_k} \, ,\\
-	0 &= \deriv{\phi}{\bar{x}_k}^T + \deriv{m}{\bar{x}_k}^T\beta  - \lambda^T(N) \, ,  
+	\vec{\lambda}_{k+1}^T &= \vec{\lambda}_k^T  \deriv{\vec{f}_k}{\vec{x}_k}^{-1} \, , \\
+	0 &=  \deriv{H_k}{\vec{u}_k} \, , \label{eqn:control_necc_cond}\\
+	0 &= \deriv{\phi}{\vec{x}_k}^T + \deriv{\vec{m}}{\vec{x}_k}^T \vec{\beta}  - \vec{\lambda}^T(N) \, ,  
 \end{align}
@@
 \begin{equation}
-	H_k = \lambda_k^T f(\bar{x}_k, \bar{u}_k) \, ,
+	H_k = \vec{\lambda}_k^T \vec{f}(\vec{x}_k, \vec{u}_k) \, ,
 	\label{eq:hamiltonian_opt}
 \end{equation}
-and \(\lambda \in \R^{4 \times 1}\) is the costate and \(\beta \in \R^{2 \times 1} \) are the additional Lagrange multipliers associated with the terminal constraints in~\cref{eq:constraints}.
+and \( \vec{\lambda} \in \R^{4 \times 1} \) is the costate and \(\vec{\beta} \in \R^{2 \times 1} \) are the additional Lagrange multipliers associated with the terminal constraints in~\cref{eq:constraints}.
+The state dynamics are represented by \( \vec{f}(\vec{x}_k, \vec{\lambda}_k ) \) after substituting~\cref{eqn:control_necc_cond} into~\cref{eq:distrete_eoms}.
 This indirect optimal control formulation leads to a two point boundary value problem with split boundary conditions. 
@@
 \begin{align}\label{eq:costate_eom}
-	\lambda_{k+1}^T = \lambda_k^T
+	\vec{\lambda}_{k+1}^T = \vec{\lambda}_k^T
 	\begin{bmatrix} 
@@
 This is a computationally expensive operation that is prone to numerical error and instability.
-A method is presented in Appendix B to avoid this inversion and determine an explicit update map \( \lambda_k \to \lambda_{k+1} \).
+A method is presented in Appendix B to avoid this inversion and determine an explicit update map \( \vec{\lambda}_k \to \vec{\lambda}_{k+1} \).
 
@@
 There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{bryson1975}.
-In this work, a multiple shooting method is implemented.
-Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement.
+Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement~\cite{bryson1975}.
 In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
 Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs.
+The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches.
 Additional interior constraints are used to ensure a continuous solution.

### 2015-12-06 acta_manuscript.tex
working on periodic orbit transfer

@@
 \section{Numerical Example}\label{sec:simulation}
-\subsection{Periodic Orbit transfer}
-We present two numerical simulations to demonstrate the transfer procedure.
+\subsection{Periodic Orbit transfer}\label{sec:periodic_orbit_transfer}
+% define problem setup (initial and target orbits)
+We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
 The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
 The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix} \) with \( \mu = 0.0125 \).
-The initial state is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
+The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
 \begin{figure}[htbp]
@@
 \end{figure}
-\Cref{fig:moon_orbit} shows that the desired trajectories remain in the vicinity of the Moon in the rotating reference frame. 
+\Cref{fig:moon_orbit} shows that the target set remain in the vicinity of the Moon in the rotating reference frame. 
 This type of orbit would be useful for a variety of mission scenarios.
@@
 The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
+The initial set is a planar periodic orbit about \( L_1\).
+We generate the periodic orbit by using a linear approximation to first determine periodic solutions.
+This approximate periodic solution from the linear system is then implemented in the nonlinear dynamics through the method of differential correction, or single shooting~\cite{koon2011}.
+The initial \( \dot{y} \) is varied until the the periodic orbit makes a perpendicular crossing, \( \dot{x} = 0 \), after one half period.
+Due to the symmetry of the three-body problem, we are ensured that the trajectory will return to the initial state after a single period.
 
-As a baseline, the method introduced in Reference~\citenum{koon2000} is implemented.
-The method is based on the invariant manifolds associated with the periodic orbits of the three body system.
-These invariant manifolds are a set of trajectories that either asymptotically arrive or depart the periodic orbit. 
-Linking the invariant manifolds of several periodic orbits, or coupled three body systems, has been used to generate orbital transfers.
-\Cref{fig:manifold_trajectory} shows the unstable invariant manifold associated with the \( L_1\) periodic orbit. 
-The invariant manifolds are globalized using the eigenvectors of the Monodromy matrix along the periodic orbit.
-The eigenvectors serve as a local approximation of the stable and unstable directions. 
-Perturbations along these directions serve to approximate the invariant manifolds.
-
+% invariant manifold transfer
+As a source for comparison, the method of using invariant manifolds, introduced in Reference~\citenum{koon2000}, is implemented.
+As described in~\cref{sec:invariant_manifolds}, these invariant manifolds are the set of trajectories that either asymptotically arrive or depart the periodic orbit. 
+We generate the unstable manifold associated with the initial planar periodic orbit.
+We numerical propagate the unstable manifold forward in time until the trajectories intersect the \Poincare section \( y = 0 \).
+\Cref{fig:manifold_trajectory} shows the unstable invariant manifold generated from the initial \( L_1\) periodic orbit. 
+The blue points in~\cref{fig:manifold_poincare} are the intersections of the target Moon orbit and the \Poincare section.
+The two circular regions are the ascending (right) and descending (left) intersections of the target orbit and \Poincare section.
+The green points in~\cref{fig:manifold_poincare} are intersections of the unstable manifold from~\cref{fig:manifold_trajectory} with the \Poincare section \( y = 0 \).
+Only a single branch of the invariant manifold intersects with the ascending region of the target orbit.
+There are no intersections of the invariant manifold with the descending region of the target orbit.
+The numerical values associated with the green points denote the required time of flight along the invariant manifold in non-dimensional units.
 \begin{figure} 
@@
 	\begin{subfigure}[htbp]{0.5\textwidth} 
-		\includegraphics[width=\textwidth]{reach_trajectory} 
-		\caption{Controlled Trajectories} \label{fig:reach_trajectory} 
-	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-	\begin{subfigure}[htbp]{0.5\textwidth} 
 		\includegraphics[width=\textwidth]{manifold_trajectory} 
 		\caption{Invariant Manifold} \label{fig:manifold_trajectory} 
+	\end{subfigure}~ %
+	\begin{subfigure}[htbp]{0.5\textwidth} 
+		\includegraphics[width=\textwidth]{manifold_poincare} 
+		\caption{Poincar\`e Section} \label{fig:manifold_poincare} 
 	\end{subfigure} 
-
-	\begin{subfigure}[htbp]{0.5\textwidth} 
-		\includegraphics[width=\textwidth]{poincare_compare} 
-		\caption{Poincar\`e Section} \label{fig:poincare_compare} 
-	\end{subfigure} 
-	\caption{Transfer Trajectory}
-	\label{fig:transfer_orbit} 
+	\caption{Invariant manifold transfer}
+	\label{fig:invariant_manifold} 
 \end{figure}
 
-The unstable invariant manifold is numerically propagated to the same \Poincare section defined on the \( \hat{x} \) axis.
-The intersection of the invariant manifold and the \Poincare section is denoted by the green markers in~\cref{fig:poincare_compare}.
-Only a single branch of the invariant manifold intersects with the ascending region of the target orbit.
-There are no intersections of the invariant manifold with the descending region of the target orbit.
-The numerical values associated with the green points denote the required time of flight along the invariant manifold in non-dimensional units.
-A transfer along the invariant manifold requires on average \( t_f \approx 3.1 \) as compared to \( t_f \approx 1.4 \) for a transfer using low thrust propulsion and the reachable set.
-Finally, it should be noted that an additional maneuver would be required for a transfer using the invariant manifold.
-The intersection on the \Poincare section only shows that the \( x \text{ and } \dot{x} \) components intersect.
-An additional instantaneous \( \Delta V \) would be required to transfer from the invariant manifold to the target.
-
+% reachable set transfer
 A periodic orbit is generated about the \( L_1 \) Lagrange point. 
@@
 
-The blue points in~\cref{fig:poincare_compare} are the intersections of the target Moon orbit and the \Poincare section.
-The two circular regions are the ascending (right) and descending (left) intersections of the target orbit and \Poincare section.
+\begin{figure} 
+	\centering 
+	\begin{subfigure}[htbp]{0.5\textwidth} 
+		\includegraphics[width=\textwidth]{reach_trajectory} 
+		\caption{Reachability trajectories} \label{fig:reach_trajectory} 
+	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+	\begin{subfigure}[htbp]{0.5\textwidth} 
+		\includegraphics[width=\textwidth]{poincare_compare} 
+		\caption{Poincar\`e Section} \label{fig:poincare_compare} 
+	\end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+	\caption{Reachability set transfer}
+	\label{fig:reachability_set_transfer} 
+\end{figure}
+
 \Cref{fig:poincare_compare} shows that the reachable set and those of the descending target region intersect.
@@
 
-\subsection{Geostationary Transfer}
+% compare reachable set transfer with the invariant manifold transfer
+A transfer along the invariant manifold requires on average \( t_f \approx 3.1 \) as compared to \( t_f \approx 1.4 \) for a transfer using low thrust propulsion and the reachable set.
+Finally, it should be noted that an additional maneuver would be required for a transfer using the invariant manifold.
+The intersection on the \Poincare section only shows that the \( x \text{ and } \dot{x} \) components intersect.
+An additional instantaneous \( \Delta V \) would be required to transfer from the invariant manifold to the target.
+
+\subsection{Geostationary Transfer}\label{sec:geo_transfer}
 \begin{figure}[htbp]
@@
 \end{subequations}
-Finally, backsubstituion is used to determine explicit equations for the discrete update map \( \lambda_k \to \lambda_{k+1} \) which is defined as
+Finally, backsubstituion is used to determine explicit equations for the discrete update map \( \vec{\lambda}_k \to \vec{\lambda}_{k+1} \) which is defined as
 \begin{subequations}\label{eq:costate_update}

### 2015-12-06 acta_manuscript.tex
updated periodic orbit transfer

@@
 % reachable set transfer
-A periodic orbit is generated about the \( L_1 \) Lagrange point. 
-A \Poincare section is chosen to allow for design on a lower dimensional subspace.
-The section is defined along the \( \hat{x} \) axis, or defined with \( \alpha = \ang{0} \) and intersects both the initial and target orbits.
-From the periodic orbit, a series of optimal trajectories are generated to approximate the reachable set.
-The trajectories are generated from a fixed initial state of \( x_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix} \) over a fixed time span of \( t_f = 1.4 \).
+Next, we determine the reachability set with the addition of a low thrust control input.
+We define a maximum magnitude of the thrust as \( u_max = 0.5 \) and assume we can point the thrust in any direction on the plane. 
+This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
+We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
+The trajectories are generated from a fixed initial state of \( \vec{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix} \) over a fixed time span of \( t_f = 1.4 \).
+This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit.
+
+From the initial state on the periodic orbit, a series of optimal trajectories are generated to determine the reachable set.
 By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
+We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
 The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:transfer_orbit}.
 
-The optimal trajectories, under the influence of the control input \( \bar{u} \), are shown in~\cref{fig:reach_trajectory}.
+% discuss reachability set computation
+The optimal trajectories, under the influence of the control input \( \vec{u} \), are plotted in red in~\cref{fig:reach_trajectory}.
 Initially, the spacecraft is assumed to lie on the periodic orbit.
@@
 The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
-The reachable set is an ellipsoidal region with a major axis aligned along \( \theta \approx \ang{70} \).
-
+The reachable set is an ellipsoidal region with a major axis aligned along \( \theta \approx \ang{70} \) as compared to a fixed point without any control input.
 \begin{figure} 
@@
 
+% generate the optimal transfer
 \Cref{fig:poincare_compare} shows that the reachable set and those of the descending target region intersect.
@@
 Using~\cref{eq:jacobi} and the intersection state the final component \( \dot{y} \) is calculated. 
-This results in a complete target state \( \bar{x}_t \) which lies on the reachable set and on the target orbit. 
-A final optimal trajectory is generated such that the \( \bar{x}(N) = \bar{x}_t \).
+This results in a complete target state \( \vec{x}_t \) which lies on the reachable set and on the target orbit. 
+A final optimal trajectory is generated such that the \( \vec{x}(N) = \vec{x}_t \).
 This transfer trajectory is denoted by the green path in~\cref{fig:reach_trajectory}.
 
-% compare reachable set transfer with the invariant manifold transfer
+% compare the invariant manifold method and reachability set
 A transfer along the invariant manifold requires on average \( t_f \approx 3.1 \) as compared to \( t_f \approx 1.4 \) for a transfer using low thrust propulsion and the reachable set.
-Finally, it should be noted that an additional maneuver would be required for a transfer using the invariant manifold.
-The intersection on the \Poincare section only shows that the \( x \text{ and } \dot{x} \) components intersect.
-An additional instantaneous \( \Delta V \) would be required to transfer from the invariant manifold to the target.
+This long time of flight is typical of transfers using invariant manifolds.
+The unstable invariant manifold traverses a large region of the phase space and is dependent on the system dynamics. 
+In addition, the invariant manifolds asymptotically arrive and depart from the periodic orbit. 
+As a result, it may take an arbitrarily long period of time to depart from the vicinity of the periodic orbit.
+In addition, only a small portion of the invariant manifold intersects with the target Moon orbit.
+In contrast, with the low thrust control input we are able to enlarge the reachability set from a single point to a larger ellipsoidal region in~\cref{fig:poincare_compare}.
+This achieves an intersection with the target orbit with a much lower time of flight as compared to the invariant manifold method.
+In addition, by generating the reachability set we are able to compute the required control input to exactly intersect the target orbit.
+This avoids having to compute and accomplish a secondary impulsive manuever to transition from the invariant manifold to the target orbit.
+%Since The intersection on the \Poincare section only shows that the \( x \text{ and } \dot{x} \) components intersect.
+%An additional instantaneous \( \Delta V \) would be required to transfer from the invariant manifold to the target.
 
 \subsection{Geostationary Transfer}\label{sec:geo_transfer}
+
+% motivation - may need to compute several reachability sets
+
+% discuss difficulties in applying direct optimization or the invariant manifold method
+
+% describe initial and target states for this example
+This example demonstrates the feasibility of iteratively computing the reachable set. 
+In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
+Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
+It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
+The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.
 \begin{figure}[htbp]
@@
 \end{figure}
-
-This example demonstrates the feasibility of iteratively computing the reachable set. 
-In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
-Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
-It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
-The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.
 

### 2015-12-07 acta_manuscript.tex
finished updating the text. need to generate new pretty figures

@@
 This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
-The use of a geometric integrators ensures numerical stability for long-duration orbit transfers.
-A numerical example is presented demonstrating a transfer trajectory from \( L_1 \) to the vicinity of the Moon and compared to the control-free design methodology.
-\begin{itemize}
-	\item generalize invariant manifold transfers with continuous control input
-	\item Low thrust propulsion rather than impulsive manuevers
-	\item Combination of reachable set and \Poincare section
-	\item Geometric integrators for accurate numerical simulation
-\end{itemize}
+We derive and solve the Euler-Lagrange equations to generate the necessary conditions for an optimal solution.
+This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods.
+In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
+Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
+We avoid the issues inherent in selecting a valid initial condition for optimization.
+Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
+We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.
 
@@
 % motivation - may need to compute several reachability sets
+There are many situations where a more complicated and extensive orbital transfer is desired. 
+For example, we present a simulation of transferring from a geostationary orbit to the a periodic orbit about the Moon.
+This type of low energy transfer would be most suitable for unmanned spacecraft transitioning from the Earth to the Moon.
+The long time of flight would make such a transfer unsuitable for manned missions due to life support constraints.
+%This type of low energy transfer is best suited for autonomous spacecraft missions as the long time of flight is undesirable for manned spacecraft. 
+Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
+Low energy transfers from the Earth to the Lagrange points are necessary for future missions.
+
+The previous example demonstrated the capability of determining an orbital transfer after determining the intersection of the reachability set and a target orbit.
+However, it may not be possible to achieve an intersection on the \Poincare section after a single iteration. 
+Since the spacecraft has an upper bounded control input and time of flight the reachability set is finite in size. 
+As a result, we present a method of performing several iterations of the reachability set computation. 
+A straightforward method is presented which allows for a series of reachability sets to be computed which progressively move the trajectory towards the target.
+In this manner, it is possible to determine more complicated transfers by a simple selection of states from the reachability set.
 
 % discuss difficulties in applying direct optimization or the invariant manifold method
+It is possible to design arbitrary transfers using either a direct optimization or invariant manifold based approach.
+The direct optimization method transforms the optimal control problem into a nonlinear programming problem.
+Instead of solving the Euler-Lagrange equations the state and control histories are parameterized and solved through any number of mathematical programming methods.
+However, due to this parameterization only an approximate solution, which approximates the true optimal solution in the limit, is feasible. 
+On the other hand, our method applies an indirect optimization method.
+The necessary conditions for optimality are computed and directly solved in generating the reachability set. 
+The use of the reachability set also avoids the issues of selecting a valid initial condition.
+We select a state on the maximum reachability set which minimizes the distance toward the target. 
+This straightforward approach achieves an optimal trajectory and is used to generate general transfers.
+
+The invariant manifold method is difficult to apply to general orbital transfers. 
+The manifolds are associated with periodic orbits in the three-body system. 
+In this case, an appropriate periodic orbit must first be determined prior to generating the invariant manifold.
+Furthermore, there is not guarantee that the invariant manifold will pass through a desired region of space. 
+For example, in the Earth-Moon system the unstable manifolds of periodic orbits about \( L_1 \) do not pass close to the Earth, but rather are beyond the geostationary orbit altitude. 
+In addition, determining the intersections between various invariant manifolds is not trivial. 
+It requires an appropriate \Poincare section and the generation of several invariant manifolds.
+There is no clear method of selecting the periodic orbits required based on a the type of transfer desired. 
+Determining an intersection between these invariant manifolds generally requires extended flight times, involving several orbits of the primaries, before an appropriate intersection is found.
+As a result, it is difficult to generalize this method to arbitrary transfers in the three-body problem.
 
 % describe initial and target states for this example
-This example demonstrates the feasibility of iteratively computing the reachable set. 
-In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
-Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
+This numerical simulation demonstrates the ability of using multiple iterations of the reachability set to achieve a more complicated transfer.
+In this manner, it is possible to design arbitrary transfers which are not possible using a single reachability computation.
+Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional Earth-Moon three-body system. 
 It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
-The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.
+The initial and target orbits are illustrated in~\cref{fig:geo_transfer_target}.
 \begin{figure}[htbp]
@@
 \end{figure}
+In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
+We seek to generate a transfer from the geostationary orbit onto the stable invariant manfiold.
+Once on the manifold, the spacecraft will coast in an uncontrolled fashion and asymptotically arrive at the desired periodic orbit.
 
-In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
-This type of low energy transfer is best suited for autonomous spacecraft missions as the long time of flight is undesirable for manned spacecraft. 
-Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
-Low energy transfers from the Earth to the Lagrange points are necessary for future missions.
+\begin{figure}[htbp]
+   \centering
+   \includegraphics[width=0.5\textwidth]{third_reachability_set} % requires the graphicx package
+   \caption{Third reachability set}
+   \label{fig:third_reachability_set}
+\end{figure}
+We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
+The stable manifold is propagated to the \Poincare section at \( y = 0 \) denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
+The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_max = 0.75 \).
+The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eq:constraints} defined on the \Poincare section.
+This allows us to approximate the set of states that are achievable in the \( x \text{ vs. } \dot{x} \) space.
+After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds.
+%A state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
+For example,~\cref{fig:third_reachability_set} shows the reachable set in red and the minimum distance to the stable manifold in green.
 
-We generate the stable invariant manifold associated with the periodic orbit.
-The stable manifold is propagated to the \Poincare section directed along the positive \( \hat{x} \) axis, or \( \alpha =\ang{0} \) denoted by the horizontal black line.
-The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section.
-The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eqn:constraints} defined on the \Poincare section.
-This allows us to approximate the set of states that are achievable in the \( \hat{x} \text{ vs. } \hat{\dot{x}} \) space.
-A final state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
-With each reachable set we move the final trajectory closer to the stable manifold.
+With each reachable set we move the controlled trajectory closer to the stable manifold.
 \Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
@@
 		\caption{\Poincare section} \label{fig:geo_transfer_poincare} 
+	\end{subfigure}~%
+	\begin{subfigure}[htbp]{0.5\textwidth} 
+		\includegraphics[width=\textwidth]{control_input_geo} 
+		\caption{Control input} \label{fig:control_input_geo} 
 	\end{subfigure} 
@@
 \end{figure}
+
+This approach demonstrates the ability to link several computations of the reachability set to enable a more general transfer.
+We use eight iterations in order to transfer from the geostationary orbit to the stable manifold.
+This approach allows for a larger class of potential transfers which leverage the capabilities of low-thrust propulsion systems.
+We are able to achieve a transfer that is not possible via the standard invariant manifold approach. 
+In addition, this example illustrates a straightforward method to depart from the natural dynamics and transfer to a large region of the phase space that is not accessible via invariant manifolds alone.
 

### 2015-12-07 acta_manuscript.tex
added tikz figures

@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{poincare_map}
-	\caption{\Poincare map\label{fig:poincare_map}}
+	\begin{scaletikzpicturetowidth}{0.5\textwidth}
+	\begin{tikzpicture}[scale=\tikzscale]
+		\coordinate [label=left:\textcolor{black}{\LARGE \(\vec{q}_0\)}] (x0) at (0,0);
+		\coordinate [label=right:\textcolor{black}{\LARGE  \(\vec{q}_k\)}] (xn) at (-1,-1);
+		\coordinate [label=left:\textcolor{black}{\LARGE  \(\Sigma\)}] (sigma) at (-4,3);
+		\coordinate [label=below:\textcolor{black}{\LARGE  \(P(\vec{q})\)}] (P) at (0,-3.5);
+		% define the path of the flow with coordinates
+		\coordinate [label=right:\textcolor{blue}{}] (f1) at (5,-2);
+		\coordinate [label=below:\textcolor{black}{\LARGE  \(\phi(t,\vec{q})\)}] (f2) at (2,-5);
+		\coordinate [label=right:\textcolor{blue}{}] (f3) at (-4,-4);
+		\coordinate [label=right:\textcolor{blue}{}] (f4) at (-4,-1);
+
+	%	\draw[help lines] (-10,-10) grid (10,10); %grid
+		\filldraw [black] (x0) circle [radius=3pt];
+		\filldraw [black] (xn) circle [radius=3pt];
+
+		\draw [ultra thick,black,->-](x0) to[out=20,in=90,distance=2cm] (f1) to[out=-90,in=0,distance=2cm] (f2) to[out=180,in=-45,distance=2cm] (f3) to[out=135,in=-135,distance=2cm] (f4) ;
+		\draw [ultra thick, black,dashed] (f4) to[out=45,in=120,distance=1cm] (xn);
+
+		\draw [ultra thick] plot [smooth cycle, tension=0.1, rotate=5] coordinates { (-4,-3) (4,-3) (4,3) (-4,3) };
+	\end{tikzpicture}
+	\end{scaletikzpicturetowidth}
+	\caption{Diagram of the \Poincare map\label{fig:poincare_map}}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{reachability_set}
-	\caption{Reachability set\label{fig:reachability_set}}
+%	\includegraphics[width=0.5\textwidth]{reachability_set_hand}
+	\begin{scaletikzpicturetowidth}{0.5\textwidth}
+	\begin{tikzpicture}[scale=\tikzscale]
+		\coordinate [label=left:\textcolor{black}{\LARGE \(\vec{x}_0\)}] (x0) at (-1,-2);
+		\coordinate [label=below:\textcolor{black}{\LARGE  \(\vec{x}_n\)}] (xn) at (1,1);
+		\coordinate [label=left:\textcolor{black}{\LARGE  \(\Sigma\)}] (sigma) at (-4,3);
+		\coordinate [label=below:\textcolor{black}{\LARGE  \(P(\vec{x})\)}] (P) at (0,-3.5);
+		% define the path of the flow with coordinates
+		\coordinate [label=right:\textcolor{black}{}] (f1) at (5,-2);
+		\coordinate [label=below:\textcolor{black}{\LARGE  \(\phi(t,\vec{x}_0)\)}] (f2) at (2,-5);
+		\coordinate [label=right:\textcolor{black}{}] (f3) at (-4,-4);
+		\coordinate [label=right:\textcolor{black}{}] (f4) at (-4,-1);
+
+	%	\draw[help lines] (-10,-10) grid (10,10); %grid
+		\filldraw [black] (x0) circle [radius=3pt];
+		\filldraw [black] (xn) circle [radius=3pt];
+
+		\draw [ultra thick,black,->-](x0) to[out=20,in=90,distance=2cm] (f1) to[out=-90,in=0,distance=2cm] (f2) to[out=180,in=-45,distance=2cm] (f3) to[out=135,in=-135,distance=2cm] (f4) ;
+		\draw [ultra thick, black,dashed,->] (f4) to[out=45,in=180,distance=1cm] ($(xn)-(2,0)$);
+
+		\draw [ultra thick] plot [smooth cycle, tension=0.1, rotate=5] coordinates { (-4,-3) (4,-3) (4,3) (-4,3) };
+
+		\draw [thick,dashed] (xn) circle [radius=2cm]; % reachability set
+
+		\draw [thick,->] (xn) -- ($(xn) + (2.5,0)$);
+		\draw [thick,rotate=45,->] (xn) -- ($(xn) + (2.5,0)$);
+		\draw ($(xn) + (1,0)$) arc [start angle=0,end angle=45, radius=1];
+		\node [draw=none] at (2.2,1.5) {\Large \(\theta\)};
+		\draw [decorate,decoration={brace,amplitude=5pt},rotate=45] (xn) -- ($(xn) + (2,0)$);
+		\node [draw=none] at ($ (xn) + (0,1) $) {\Large \( J \)};
+	\end{tikzpicture}
+	\end{scaletikzpicturetowidth}
+	\caption{Reachability set on a \Poincare section\label{fig:reachability_set}}
 \end{figure}

### 2015-12-08 acta_manuscript.tex
corrected some typos and errors

@@
 Designing spacecraft trajectories is a classic and ongoing topic of research.
+There has been significant research into the design of orbital transfers for space vehicles.
 %With the current fiscal constraints there is an increased focus on technologies with a critical impact on mass.
@@
 As a result, accurate numerical methods are required to determine optimal solutions.
-
+These methods are critically dependent on accurate initial guess in order to allow for convergence to an optimal solution.
 %For example, analytical design is possible for complicated trajectories using a Keplerian system model via the patched-conic approach.
@@
 % DYNAMICAL STRUCTURE
-References~\citenum{koon2000} and~\citenum{ross2006} have illustrated the rich structure that exists in the three-body problem.
+References~\citenum{koon2011} and~\citenum{ross2006} have illustrated the rich structure that exists in the three-body problem.
 Within the three-body problem, a spacecraft's feasible region of motion is constrained by its energy, or Jacobi integral. 
@@
 Instead, we utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
-In addition, through the use geometric integrators we accurately capture the effects of low thrust on the system dynamics in the numerical simulation. 
+In addition, through the use of geometric integrators we accurately capture the effects of low thrust on the system dynamics in the numerical simulation. 
 This approach avoids the instability and dissipative effects of conventional integration schemes.
-With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low-thrust propulsion systems.
+With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low thrust propulsion systems.
 
@@
 
-To achieve these objectives, computational geometric optimal control techniques are applied. 
+To achieve these objectives, we apply computational geometric optimal control techniques. 
 We generalize the previous work of determining orbital transfers using invariant manifolds~\cite{koon2011}.
@@
 The dynamics of the three-body system are derived from the discrete Lagrangian, which approximates the integral of the continuous time Lagrangian over a fixed discrete step.
-Application of the discrete Euler-Lagrange equations, or the discrete Legendre transform, results in the discrete equations of motion.
+Application of the discrete Euler-Lagrange equations, or equivalently the discrete Legendre transform, results in the discrete equations of motion.
 This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs.
@@
 The \Poincare section reduces the dimensionality of the system dyanmics to the study of a related discrete update map.  
-This allows for the study of the complex dynamics of the three-body problem to a lower dimensional space.
+This allows for the study of the complex dynamics of the three-body problem on a lower dimensional space.
 Rather than relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on an appropriate \Poincare section.
 This simple methodology allows for extended transfer trajectories which iteratively approach a target region.
-The addition of low-thrust propulsion affords the ability to enlarge the reachable set as compared to the control-free case.
+We utilize the low thrust control input to enlarge the reachability set.
 Maximization of the reachability set, on an appropriately chosen \Poincare section, allows for a greater space of potential transfer trajectories.
@@
 It is assumed that the \(\left( \hat{e}_1, \hat{e}_2\right)\) frame rotates with a constant angular velocity equal to the mean motion of the Moon.
-Following convention, the system is also non-dimensionalized by the characteristic mass, length, and time~\cite{koon2011}.
-As a result, the system can be characterized by a single mass parameter \( \mu \),
+Following convention, the system is also non-dimensionalized by the characteristic units of length, mass, and time~\cite{koon2011}.
+As a result, the system can be characterized by a single mass ratio parameter \( \mu \),
 \begin{equation}
@@
 \end{equation}
-and the control inputs are defined as \( u_x \text{ and } u_y \in \R \).
-The state is defined as \( \bar{x} = \begin{bmatrix}\bar{r} &\bar{v} \end{bmatrix}^T\) with \(\bar{r} \in \R^{2\times1}\) and \(\bar{v} \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
+and the control inputs are defined as \( \vec{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \).
+The state is defined as \( \vec{x} = \begin{bmatrix}\vec{r} &\vec{v} \end{bmatrix}^T\) with \(\vec{r} = \begin{bmatrix} x & y \end{bmatrix}^T \in \R^{2\times1}\) and \(\vec{v}= \begin{bmatrix} \dot{x} & \dot{y} \end{bmatrix}^T \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
 The equations of motion may be rewritten in state space form as
 \begin{equation}\label{eqn:cont_dyn}
-	\left[\begin{array}{c} \dot{\bar{r}} \\ \dot{\bar{v}} \end{array} \right] = 
-	\left[ \begin{array}{c} \bar{v} \\ A \bar{v} + \nabla U + \bar{u}(t) \end{array} \right] = f\left( t,x, u\right) \, ,
+	\left[\begin{array}{c} \dot{\vec{r}} \\ \dot{\vec{v}} \end{array} \right] = 
+	\left[ \begin{array}{c} \vec{v} \\ A \vec{v} + \nabla U + \vec{u} \end{array} \right] = f\left( t,\vec{x}, \vec{u}\right) \, ,
 \end{equation}
@@
 There exists a single integral, or constant of motion for the three-body problem~\cite{lanczos1970,szebehely1967}.
-This energy constant is analogous to the total mechanical energy, however is a non-physical quantity arising from the problem formulation~\cite{szebehely1967}.
+This energy constant is analogous to the total mechanical energy, however it is a non-physical quantity arising from the problem formulation~\cite{szebehely1967}.
 Also known as the Jacobi constant, it is defined as a function of the position and velocity in the rotating frame and given by
 \begin{equation}
-	E\left( \bar{r} , \bar{v} \right) = \frac{1}{2}\left( \dot{x}^2 + \dot{y}^2\right) - U\left(x,y \right) \, .
+	E\left( \vec{r} , \vec{v} \right) = \frac{1}{2}\left( \dot{x}^2 + \dot{y}^2\right) - U\left(x,y \right) \, .
 	\label{eq:jacobi}
 \end{equation}
-\Cref{eq:jacobi} divides the phase space into distinct regions based on the energy level of the spacecraft.
+\Cref{eq:jacobi} divides the phase space into distinct regions of allowable motion based on the energy level of the spacecraft.
 Fixing the Jacobi integral to a constant defines zero velocity curves, which are the locus of points where the kinetic energy, and hence velocity vanishes.
@@
 As the energy level increases there are five critical points of the effective potential~\cref{eq:eff_pot} where the slope is zero.
-Three collinear saddle points on the \(x\) axis and two equilateral points.
+Three collinear saddle points on the \(\hat{e}_1\) axis and two equilateral points.
 These equilibrium, or Lagrange points, are labeled \( L_i, i = 1, \hdots, 5 \) and are shown in Figure~\ref{fig:energy_contour}.
@@
 In addition, they exhibit improved energy behavior over long integration periods.
-A short background on the variational principle for mechanical systems is presented. 
-A discrete approximation of the action integral is presented and allows for construction of a variational integrator for the PCRTBP.
+Next, we present a short background on the variational principle for mechanical systems. 
+We then develop a discrete approximation of the action integral and construct a variational integrator for the PCRTBP.
 
@@
 \begin{align}\label{eq:hamiltonian}
-	H &= \sum_{i = 1}^N p_i \dot{q}_i - L \left( q_i,\dot{q}_i, t \right) \, .
+	H &= \sum_{i = 1}^n p_i \dot{q}_i - L \left( q_i,\dot{q}_i, t \right) \, .
 \end{align}
@@
 Care must be taken during the implementation of~\cref{eq:discrete_eoms}.
-As~\cref{eq:discrete_potential_grad,eq:discrete_distance} are defined at both step \( k \) and \( k+1 \) they must be evaluated at both time steps.
-\Cref{eq:discrete_eoms} is implemented by first defining an initial state \( \bar{x}_k \) and control \( \bar{u}_k \).
+As~\cref{eq:discrete_potential_grad,eq:discrete_distance} are defined at both step \( k \) and \( k+1 \) they must be evaluated at both time instances.
+\Cref{eq:discrete_eoms} is implemented by first defining an initial state \( \vec{x}_k \) and control \( \vec{u}_k \).
 The distances and gravitational potential at step \( k \) are evaluated from~\cref{eq:distonek,eq:disttwok,eq:Uxk,eq:Uyk}.
@@
 There exist five equilibrium points in the equations of motion for the three-body problem as introduced in~\cref{sec:jacobi}.
-It has been shown that the local orbit structure near the Lagrange points give rise to families of periodic orbits as well as the stable and unstable manifolds of these periodic orbits.
-This rich structure is globally connected and gives rise to a dynamical chain which allows trajectories to pass through the phase space~\cite{koon2000,conley1968}.
+It has been shown that the local orbit structure near the Lagrange points gives rise to families of periodic orbits as well as the stable and unstable manifolds of these periodic orbits.
+This rich structure is globally connected and gives rise to a dynamical chain which allows trajectories to pass through the phase space~\cite{koon2011,conley1968}.
 The manifold structure associated with periodic orbits about the \( L_1 \) and \( L_2 \) Lagrange points are critical to the understanding of the motion of spacecraft as well as comets/asteroids.
@@
 % technical definition of manifold and some equations
-Invariant manifolds serve as a higher dimensional generalization of the concept of seperatrices from linear systems to the case of nonlinear systems. 
+Invariant manifolds serve as a higher dimensional generalization of the concept of seperatrices from linear systems as applied to the case of nonlinear systems. 
 To compute these manifolds first one must determine an appropriate periodic orbit. 
@@
 are the set of trajectories that asymptotically depart or approach a periodic orbit.
-For the nonlinear system~\cref{eqn:cont_dyn} or~\cref{eq:discrete_eoms}, we can determine the state transition matrix \( \Phi \) along the periodic orbit, from \( 0 \to T\), by numerical integration of the variational equations.
+For the nonlinear system, defined by~\cref{eqn:cont_dyn} or~\cref{eq:discrete_eoms}, we can determine the state transition matrix \( \Phi \) along the periodic orbit, from \( 0 \to T\), by numerical integration of the variational equations.
 In the neighborhood of the periodic orbit \( \gamma \), the tangent spaces to the stable and unstable manifolds are provided by the generalized eigenspaces \( E^s \) and \( E^u \) of the linearization \( A = \D f(\vec{x}_0) \), where \( \vec{x}_0 \) is on the periodic orbit.
@@
 \Poincare maps are a useful tool in the analysis of the flow near periodic orbits in the three-body problem.
-We let \( \Sigma \) be a hypersurface of section chosen such that all trajectories in the vicinity of a state \( q \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
-A \Poincare map, \( P(q) = \phi(T;q) \), maps the state of a trajectory from one intersection to the next.
+We let \( \Sigma \) define a hypersurface of section chosen such that all trajectories in the vicinity of a state \( \vec{q} \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
+A \Poincare map, \( P(\vec{q}) = \phi(T;\vec{q}) \), maps the state of a trajectory from one intersection to the next.
 Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
 This allows for greater insight into the stability and dynamics of periodic solutions of dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories.
-For example, \Poincare maps have been used to prove the existence of homoclinic orbits, orbits which are both forward and backward asymptotic to a single unstable periodic orbit, and heteroclinic orbits, which join different periodic orbits~\cite{conley1968,koon2000b}.
+For example, \Poincare maps have been used to prove the existence of homoclinic orbits, which are orbits both forward and backward asymptotic to a single unstable periodic orbit, and heteroclinic orbits, which join different periodic orbits~\cite{conley1968,koon2000b}.
 These dynamic features have been shown to play a vital role in the movement of natural bodies as well as critical for spacecraft missions~\cite{lo1997,gomez2001}.
@@
 	\begin{tikzpicture}[scale=\tikzscale]
-		\coordinate [label=left:\textcolor{black}{\LARGE \(\vec{q}_0\)}] (x0) at (0,0);
-		\coordinate [label=right:\textcolor{black}{\LARGE  \(\vec{q}_k\)}] (xn) at (-1,-1);
-		\coordinate [label=left:\textcolor{black}{\LARGE  \(\Sigma\)}] (sigma) at (-4,3);
-		\coordinate [label=below:\textcolor{black}{\LARGE  \(P(\vec{q})\)}] (P) at (0,-3.5);
+		\coordinate [label=above:\textcolor{black}{\Large \(\vec{q}_0\)}] (x0) at (0,0);
+		\coordinate [label=below:\textcolor{black}{\Large  \(\vec{q}_k\)}] (xn) at (-1,-1);
+		\coordinate [label=left:\textcolor{black}{\Large  \(\Sigma\)}] (sigma) at (-4,3);
+		\coordinate [label=below:\textcolor{black}{\Large  \(P(\vec{q})\)}] (P) at (0,-3.5);
 		% define the path of the flow with coordinates
 		\coordinate [label=right:\textcolor{blue}{}] (f1) at (5,-2);
-		\coordinate [label=below:\textcolor{black}{\LARGE  \(\phi(t,\vec{q})\)}] (f2) at (2,-5);
+		\coordinate [label=below:\textcolor{black}{\Large  \(\phi(t,\vec{q})\)}] (f2) at (2,-5);
 		\coordinate [label=right:\textcolor{blue}{}] (f3) at (-4,-4);
@@
 The initial orbit is a planar periodic orbit about \( L_2 \) in the Earth-Moon system while the final orbit is a periodic orbit about \( L_1 \).
-Both periodic orbits have the Jacobi energy constant and as a result it is possible to transfer between the orbits without any energy change.
+Both periodic orbits have the same Jacobi energy constant and as a result it is possible to transfer between the orbits without any energy change.
 \Cref{fig:manifolds} shows the unstable manifold of the initial orbit and the stable manifold of the desired orbit. 
-Both invariant manifolds are globalized and propogated to the first intersection with the \Poincare surface of section located at the position of the Moon, \( 1 - \mu \).
+Both invariant manifolds are globalized and propogated to the first intersection with the \Poincare surface of section located at the position of the Moon, \( x = 1 - \mu \).
 The \Poincare section in~\cref{fig:manifold_transfer_example} is defined as
@@
 In this way, it is possible to design trajectories that intersect the invariant manifolds and connect widely separated regions of the state space. 
-This technique has been heavily investigated in Reference~\citenum{koon2000}, applied to operational missions in Reference~\citenum{koon1999}, and applied to potential multi-moon orbiter missions in Reference~\citenum{tanaka2011}.
+This technique has been heavily investigated in Reference~\citenum{koon2011}, applied to operational missions in Reference~\citenum{koon1999}, and applied to potential multi-moon orbiter missions in Reference~\citenum{tanaka2011}.
 In addition, much recent work has focused on potential return missions to the Moon~\cite{zanzottera2012,campagnola2012,mingotti2011,ozimek2010a,mingotti2009}.
@@
                 \includegraphics[width=\columnwidth]{U2_poincare}
-                \caption{U1 Section}
+                \caption{\(U_2\) Section}
                 \label{fig:poincare}
@@
 The addition of low-thrust propulsion offers the potential of reduced transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities. 
-In this paper, an optimal control problem is formulated to generate a reachable set of the spacecraft.
-The reachable set is computed on an appropriate \Poincare section and is used to design an appropriate transfer trajectory.
+In this paper, we formulate an optimal control problem to generate the reachable set of the spacecraft.
+We compute the reachable set on an appropriate \Poincare section and use this to design a transfer trajectory.
 
@@
 Our formulation is based on the concept of the reachability set on a \Poincare section.
-This method allows for one to easily determine potential transfer opportunities by finding set intersections on a lower dimensional space and greatly reduces the design process.
+This method allows one to easily determine potential transfer opportunities by finding set intersections on a lower dimensional space and greatly reduces the design process.
 The addition of continuous low thrust propulsion extends the control free design process developed previously and allows for a greater range of potential transfers with a reduced time of flight.
@@
 The theory formally supporting reachability has been extensively developed and is directly derivable from optimal control theory~\cite{varaiya2000,lygeros2002,lygeros2004}.
-Reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a}.
+More recently, reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a}.
 Computation of the reachable set for a system involves solving the Hamilton-Jacobi partial differential equation or satisfying a dynamic programming principle.
@@
 	\begin{tikzpicture}[scale=\tikzscale]
-		\coordinate [label=left:\textcolor{black}{\LARGE \(\vec{x}_0\)}] (x0) at (-1,-2);
-		\coordinate [label=below:\textcolor{black}{\LARGE  \(\vec{x}_n\)}] (xn) at (1,1);
-		\coordinate [label=left:\textcolor{black}{\LARGE  \(\Sigma\)}] (sigma) at (-4,3);
-		\coordinate [label=below:\textcolor{black}{\LARGE  \(P(\vec{x})\)}] (P) at (0,-3.5);
+		\coordinate [label=left:\textcolor{black}{\Large \(\vec{x}_0\)}] (x0) at (-1,-2);
+		\coordinate [label=below:\textcolor{black}{\Large  \(\vec{x}_n\)}] (xn) at (1,1);
+		\coordinate [label=left:\textcolor{black}{\Large  \(\Sigma\)}] (sigma) at (-4,3);
+		\coordinate [label=below:\textcolor{black}{\Large  \(P(\vec{x})\)}] (P) at (0,-3.5);
 		% define the path of the flow with coordinates
 		\coordinate [label=right:\textcolor{black}{}] (f1) at (5,-2);
-		\coordinate [label=below:\textcolor{black}{\LARGE  \(\phi(t,\vec{x}_0)\)}] (f2) at (2,-5);
+		\coordinate [label=below:\textcolor{black}{\Large  \(\phi(t,\vec{x}_0)\)}] (f2) at (2,-5);
 		\coordinate [label=right:\textcolor{black}{}] (f3) at (-4,-4);
@@
 \end{subequations}
-where the angle \( \theta_d\) a direction to maximize the reachability set on the \Poincare section.
-We define the \Poincare section along the horizontal axis and equivalent to the  surface \( y(N) = 0 \).
+where the angle \( \theta_d\) defines a direction in which we wish to maximize the reachability set on the \Poincare section.
+We define the \Poincare section along the horizontal axis and which is equivalent to the  surface \( y(N) = 0 \).
 This is similar to the previous work in determining homoclinic orbits in the three-body problem~\cite{llibre1985,koon2011}.
@@
 We seek to compliment these results with the addition of low thrust propulsion.
-The maximum control thrust magnitude is defined by \( u_max \) and is non-dimensionalized by the characteristic length,mass, and time.
+The maximum control thrust magnitude is defined by \( u_{max} \) and is non-dimensionalized by the characteristic units of length, mass, and time.
 The goal is to determine the control input \( \vec{u}_k\) such that the cost function~\cref{eq:cost} is minimized subject to the state equations of motion~\cref{eq:discrete_eoms} and constraints~\cref{eq:constraints}.
@@
 \begin{align}
-	\vec{\lambda}_{k+1}^T &= \vec{\lambda}_k^T  \deriv{\vec{f}_k}{\vec{x}_k}^{-1} \, , \\
+	\vec{\lambda}_{k+1}^T &= \vec{\lambda}_k^T  \parenth{\deriv{\vec{f}_k}{\vec{x}_k}}^{-1} \, , \\
 	0 &=  \deriv{H_k}{\vec{u}_k} \, , \label{eqn:control_necc_cond}\\
@@
 and \( \vec{\lambda} \in \R^{4 \times 1} \) is the costate and \(\vec{\beta} \in \R^{2 \times 1} \) are the additional Lagrange multipliers associated with the terminal constraints in~\cref{eq:constraints}.
-The state dynamics are represented by \( \vec{f}(\vec{x}_k, \vec{\lambda}_k ) \) after substituting~\cref{eqn:control_necc_cond} into~\cref{eq:distrete_eoms}.
+The state dynamics are represented by \( \vec{f}(\vec{x}_k, \vec{\lambda}_k ) \) after substituting~\cref{eqn:control_necc_cond} into~\cref{eq:discrete_eoms}.
 This indirect optimal control formulation leads to a two point boundary value problem with split boundary conditions. 
@@
 The optimal control formulation presented in this section results in a two point boundary value problem (TPBVP). 
-There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{bryson1975}.
-Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement~\cite{bryson1975}.
+There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{kirk2012,bryson1975}.
+Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement.
 In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
@@
 The initial \( \dot{y} \) is varied until the the periodic orbit makes a perpendicular crossing, \( \dot{x} = 0 \), after one half period.
-Due to the symmetry of the three-body problem, we are ensured that the trajectory will return to the initial state after a single period.
+Due to the symmetry of the three-body problem, this ensures that the trajectory will return to the initial state after a single period.
 
 % invariant manifold transfer
-As a source for comparison, the method of using invariant manifolds, introduced in Reference~\citenum{koon2000}, is implemented.
-As described in~\cref{sec:invariant_manifolds}, these invariant manifolds are the set of trajectories that either asymptotically arrive or depart the periodic orbit. 
+As a source for comparison, the method of using invariant manifolds, introduced in Reference~\citenum{koon2011}, is implemented.
+As described in~\cref{sec:invariant_manifold}, these invariant manifolds are the set of trajectories that either asymptotically arrive or depart the periodic orbit. 
 We generate the unstable manifold associated with the initial planar periodic orbit.
-We numerical propagate the unstable manifold forward in time until the trajectories intersect the \Poincare section \( y = 0 \).
+We numerically propagate the unstable manifold forward in time until the trajectories intersect the \Poincare section \( y = 0 \).
 \Cref{fig:manifold_trajectory} shows the unstable invariant manifold generated from the initial \( L_1\) periodic orbit. 
@@
 	\caption{Invariant manifold transfer}
-	\label{fig:invariant_manifold} 
+	\label{fig:invariant_manifold_transfer} 
 \end{figure}
@@
 Next, we determine the reachability set with the addition of a low thrust control input.
-We define a maximum magnitude of the thrust as \( u_max = 0.5 \) and assume we can point the thrust in any direction on the plane. 
+We define a maximum magnitude of the thrust as \( u_{max} = 0.5 \) and assume we can point the thrust in any direction within the plane. 
 This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
@@
 We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
-The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:transfer_orbit}.
+The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:invariant_manifold_transfer}.
 
@@
 In addition, only a small portion of the invariant manifold intersects with the target Moon orbit.
-In contrast, with the low thrust control input we are able to enlarge the reachability set from a single point to a larger ellipsoidal region in~\cref{fig:poincare_compare}.
+In contrast, the low thrust control input we are able to enlarge the reachability set from a single point to a larger ellipsoidal region in~\cref{fig:poincare_compare}.
 This achieves an intersection with the target orbit with a much lower time of flight as compared to the invariant manifold method.
@@
 In this case, an appropriate periodic orbit must first be determined prior to generating the invariant manifold.
-Furthermore, there is not guarantee that the invariant manifold will pass through a desired region of space. 
+Furthermore, there is no guarantee that the invariant manifold will pass through a desired region of space. 
 For example, in the Earth-Moon system the unstable manifolds of periodic orbits about \( L_1 \) do not pass close to the Earth, but rather are beyond the geostationary orbit altitude. 
@@
 The stable manifold is propagated to the \Poincare section at \( y = 0 \) denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
-The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_max = 0.75 \).
+The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_{max} = 0.75 \).
 The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eq:constraints} defined on the \Poincare section.
-This allows us to approximate the set of states that are achievable in the \( x \text{ vs. } \dot{x} \) space.
+This allows us to approximate the set of states that are achievable in the \( \parenth{x ,\dot{x}} \) space.
 After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds.
 %A state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
-For example,~\cref{fig:third_reachability_set} shows the reachable set in red and the minimum distance to the stable manifold in green.
+For example,~\cref{fig:third_reachability_set} shows the reachable set after three iterations in red and the minimum distance to the stable manifold in green.
 

### 2015-12-09 acta_manuscript.tex
updating with corrections from DR. Lee

@@
 % DIRECT OPTIMAL CONTROL
-There has been extensive research focused on optimal control for spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}.
+There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}.
 Typically, the optimal control problem is solved via direct methods, which approximate the continuous problem as a parameter optimization problem.
@@
 
-To achieve these objectives, we apply computational geometric optimal control techniques. 
+To achieve these objectives, an optimal control problem is formulated to determine the reachability set on a \Poincare section.
+Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft. 
+Generation of this reachability set allows for the extension of the previous methods based on invariant manifolds in the three-body problem.
+In addition, the generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer. 
+The \Poincare section reduces the dimensionality of the system dyanmics to the study of a related discrete update map.  
+This allows for the study of the complex dynamics of the three-body problem on a lower dimensional space.
+Rather than relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on an appropriate \Poincare section.
+This simple methodology allows for extended transfer trajectories which iteratively approach a target region.
+We utilize the low thrust control input to enlarge the reachability set.
+Maximization of the reachability set, on an appropriately chosen \Poincare section, allows for a greater space of potential transfer trajectories.
+The use of the \Poincare section allows for design on a lower dimensional space and simplifies the design process.
+%Once an intersection is determined on the \Poincare section a transfer is calculated.
+
+To address the issues associated with conventional numerical integration, we apply computational geometric optimal control techniques. 
 We generalize the previous work of determining orbital transfers using invariant manifolds~\cite{koon2011}.
@@
 The use of geometric integrators, which do not suffer the numerical dissipation of conventional integration methods, results in a more robust and efficient solution.
-
-A discrete optimal control problem is formulated to determine the reachability set on a \Poincare section.
-Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft. 
-Generation of this reachability set allows for the extension of the previous methods based on invariant manifolds in the three-body problem.
-In addition, the generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer. 
-The \Poincare section reduces the dimensionality of the system dyanmics to the study of a related discrete update map.  
-This allows for the study of the complex dynamics of the three-body problem on a lower dimensional space.
-Rather than relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on an appropriate \Poincare section.
-This simple methodology allows for extended transfer trajectories which iteratively approach a target region.
-We utilize the low thrust control input to enlarge the reachability set.
-Maximization of the reachability set, on an appropriately chosen \Poincare section, allows for a greater space of potential transfer trajectories.
-The use of the \Poincare section allows for design on a lower dimensional space and simplifies the design process.
-%Once an intersection is determined on the \Poincare section a transfer is calculated.
 
@@
 \end{equation}
-where the distance \( r_1, r_2 \in \R \) of the spacecraft to each primary is defined in rotating coordinates as
+where the distance \( r_1, r_2 \in \R \) of the spacecraft to each primary is defined in rotating reference frame as
 \begin{align}
@@
 In addition, they exhibit improved energy behavior over long integration periods.
-Next, we present a short background on the variational principle for mechanical systems. 
+We present a short background on the variational principle for mechanical systems. 
 We then develop a discrete approximation of the action integral and construct a variational integrator for the PCRTBP.
@@
 \begin{align}\label{eq:hamiltonian}
-	H &= \sum_{i = 1}^n p_i \dot{q}_i - L \left( q_i,\dot{q}_i, t \right) \, .
+	H &= \sum_{i = 1}^n p_i \dot{q}_i - L \left( q,\dot{q}, t \right) \, .
 \end{align}
@@
 This results in the complete discrete update map \( \vec{x}_k \to \vec{x}_{k+1} \) given \( \vec{u}_k \).
-
+\subsection{Numerical Example}\label{sec:variational_example}
 \begin{figure} 
@@
 
-\section{Invariant Manifolds and \Poincare Map} \label{sec:invariant_manifold}
+\section{Invariant Manifolds and \Poincare Map} 
 % discuss usefulness and basic generation of invariant manifolds
@@
 % intro to invariant manifolds
-There exist five equilibrium points in the equations of motion for the three-body problem as introduced in~\cref{sec:jacobi}.
+Dynamics systems theory has been applied to design control-free maneuvers in the restricted three-body problem~\cite{koon2011}.
+As previously introduced in~\cref{sec:jacobi}, there exist five equilibrium points in the equations of motion for the three-body problem.
 It has been shown that the local orbit structure near the Lagrange points gives rise to families of periodic orbits as well as the stable and unstable manifolds of these periodic orbits.
@@
 
+\subsection{Invariant Manifolds}\label{sec:invariant_manifold}
 % technical definition of manifold and some equations
@@
 Let the trajectories of the equations of motion defined in~\cref{eqn:cont_dyn} be defined by the flow map \( \phi(t;\vec{x}_0) \) originating from \( \vec{x}(t_0) = \vec{x}_0 \) and we define a periodic orbit as \( \gamma\).
-The invariant manifolds: 
+The invariant manifolds are defined as : 
 \begin{align*}
-	W^s ( \vec{x}_0 ) &= \braces{ \vec{x} : \phi(t;\vec{x}_0) \to \gamma \text{ as } t \to \infty} \\
-	W^u ( \vec{x}_0 ) &= \braces{ \vec{x} : \phi(t;\vec{x}_0) \to \gamma \text{ as } t \to -\infty} 
+	W^s ( \vec{x}_0 ) &= \braces{ \vec{x} : \phi(t;\vec{x}_0) \to \gamma \text{ as } t \to \infty} , \\
+	W^u ( \vec{x}_0 ) &= \braces{ \vec{x} : \phi(t;\vec{x}_0) \to \gamma \text{ as } t \to -\infty} .
 \end{align*}
-are the set of trajectories that asymptotically depart or approach a periodic orbit.
+The stable manifold, \( W^s(\vec{x}_0) \), is the set of trajectories which asymptotically approach the periodic orbit.
+Similarly, the unstable manifold, \( W^u(\vec{x}_0) \), is the set of trajectories which asymptotically depart the periodic orbit.
 For the nonlinear system, defined by~\cref{eqn:cont_dyn} or~\cref{eq:discrete_eoms}, we can determine the state transition matrix \( \Phi \) along the periodic orbit, from \( 0 \to T\), by numerical integration of the variational equations.
@@
 Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
-This allows for greater insight into the stability and dynamics of periodic solutions of dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories.
+This allows for greater insight into the stability and dynamics of periodic solutions of a dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories.
 For example, \Poincare maps have been used to prove the existence of homoclinic orbits, which are orbits both forward and backward asymptotic to a single unstable periodic orbit, and heteroclinic orbits, which join different periodic orbits~\cite{conley1968,koon2000b}.
@@
 	\begin{tikzpicture}[scale=\tikzscale]
-		\coordinate [label=above:\textcolor{black}{\Large \(\vec{q}_0\)}] (x0) at (0,0);
-		\coordinate [label=below:\textcolor{black}{\Large  \(\vec{q}_k\)}] (xn) at (-1,-1);
-		\coordinate [label=left:\textcolor{black}{\Large  \(\Sigma\)}] (sigma) at (-4,3);
-		\coordinate [label=below:\textcolor{black}{\Large  \(P(\vec{q})\)}] (P) at (0,-3.5);
+		\coordinate [label=above:\textcolor{black}{\large \(\vec{q}_0\)}] (x0) at (0,0);
+		\coordinate [label=below:\textcolor{black}{\large   \(P(\vec{q}_0)\)}] (xn) at (-1,-1);
+		\coordinate [label=left:\textcolor{black}{\large  \(\Sigma\)}] (sigma) at (-4,3);
+		\coordinate [label=below:\textcolor{black}{}] (P) at (0,-3.5);
 		% define the path of the flow with coordinates
 		\coordinate [label=right:\textcolor{blue}{}] (f1) at (5,-2);
-		\coordinate [label=below:\textcolor{black}{\Large  \(\phi(t,\vec{q})\)}] (f2) at (2,-5);
+		\coordinate [label=below:\textcolor{black}{}] (f2) at (2,-5);
 		\coordinate [label=right:\textcolor{blue}{}] (f3) at (-4,-4);
@@
 
+\subsection{Control-free Orbital Transfers}\label{sec:control-free_transfers}
 % intersect invariant manifolds on poincare section
@@
 In a frame aligned with the rotating frame, \( ( \hat{e}_1, \hat{e}_2 ) \), but centered on the equilibrium point \( L_1 \) or \( L_2\), the unstable manifolds are locally moving in forward time in the second or forth quadrant.
-Similarly, the stable manifold branches are locally moving in backward time in the first and third quadrants. 
+Similarly, the stable manifold branches are locally moving in backward time in the first and third quadrants as illustrated in~\cref{fig:invariant_manifold}. 
 As a result, the location of the section in~\cref{fig:poincare} is chosen to intersect both the stable manifolds of \( L_1 \) periodic orbits as well as unstable manifolds of \( L_2 \) periodic orbits.
@@
                 \includegraphics[width=\columnwidth]{U2_Manifolds}
-                \caption{Invariant manifolds}
+                \caption{Invariant manifolds }
                 \label{fig:manifolds}
@@
                 \includegraphics[width=\columnwidth]{U2_poincare}
-                \caption{\(U_2\) Section}
+                \caption{\(U_2\) \Poincare section}
                 \label{fig:poincare}
@@
 
-\section{Optimal Control Formulation for Reachability Set}\label{sec:optimal_control}
+\section{Optimal Control Formulation for Reachability Set}
 % introduce method
@@
 
+\subsection{Reachability Set}\label{sec:reachability_set}
 % discuss reachability and application to space system
@@
 
-In this paper, we seek to approximate the reachability set on a \Poincare set by solving a related optimal control problem. 
+\subsection{Optimal Control Formulation}\label{sec:optimal_control}
+In this paper, we seek to approximate the reachability set on a \Poincare section by solving a related optimal control problem. 
 We choose our \Poincare section in a similar manner to those used previously for the design of transfers via invariant manifolds.
@@
 In the case of a periodic orbit the trajectories will cross the \Poincare section at two distinct fixed points every half period.
-The addition of low thrust propulsion allows us to enlarge the set of trajectories achievable. 
+The main idea is that the addition of low thrust propulsion allows us to enlarge the set of trajectories achievable in the \Poincare section. 
 \Cref{fig:reachability_set} illustrates how, without any control input, trajectories will intersect with the \Poincare section at \( \vec{x}_n \). 
@@
 We use a cost function to define a distance metric on the \Poincare section from the control-free intersection to an intersection under the influence of the control input.
-Maximization of this distance enables us to generate the largest reachability set under the bounded control input.
+Maximization of this distance along varying directions enables us to generate the largest reachability set under the bounded control input.
 \begin{figure}
@@
 	\begin{tikzpicture}[scale=\tikzscale]
-		\coordinate [label=left:\textcolor{black}{\Large \(\vec{x}_0\)}] (x0) at (-1,-2);
-		\coordinate [label=below:\textcolor{black}{\Large  \(\vec{x}_n\)}] (xn) at (1,1);
-		\coordinate [label=left:\textcolor{black}{\Large  \(\Sigma\)}] (sigma) at (-4,3);
-		\coordinate [label=below:\textcolor{black}{\Large  \(P(\vec{x})\)}] (P) at (0,-3.5);
+		\coordinate [label=left:\textcolor{black}{\large \(\vec{x}_0\)}] (x0) at (-1,-2);
+		\coordinate [label=below:\textcolor{black}{\large  \(\vec{x}_n\)}] (xn) at (1,1);
+		\coordinate [label=left:\textcolor{black}{\large  \(\Sigma\)}] (sigma) at (-4,3);
+		%\coordinate [label=below:\textcolor{black}{\large  \(P(\vec{x})\)}] (P) at (0,-3.5);
 		% define the path of the flow with coordinates
 		\coordinate [label=right:\textcolor{black}{}] (f1) at (5,-2);
-		\coordinate [label=below:\textcolor{black}{\Large  \(\phi(t,\vec{x}_0)\)}] (f2) at (2,-5);
+		\coordinate [label=below:\textcolor{black}{\large  \(\phi(t,\vec{x}_0)\)}] (f2) at (2,-5);
 		\coordinate [label=right:\textcolor{black}{}] (f3) at (-4,-4);
@@
 		\draw ($(xn) + (1,0)$) arc [start angle=0,end angle=45, radius=1];
-		\node [draw=none] at (2.2,1.5) {\Large \(\theta\)};
+		\node [draw=none] at (2.4,1.5) {\Large \(\theta_d\)};
 		\draw [decorate,decoration={brace,amplitude=5pt},rotate=45] (xn) -- ($(xn) + (2,0)$);
@@
 
-We define a  cost function as
+We define the \Poincare section along the horizontal axis, which is equivalent to the surface \( y(N) = 0 \), and given by
+\begin{align}
+	\Sigma = \braces{ ( x, \dot{x}) \, | \, y(N) = 0} . 
+	\label{eqn:poincare_section}
+\end{align}
+This is similar to the previous work in determining homoclinic orbits in the three-body problem~\cite{llibre1985,koon2011}.
+Previous analytical results have shown that homoclinic orbits intersect transversally in the \( (x, \dot{x} ) \) space on the plane \( y = 0 \).
+We seek to compliment these results with the addition of low thrust propulsion to maximize the reachable set on the \Poincare section.
+Placing our section at \( y = 0\) ensures that all trajectories will intersect our section.
+An optimal control problem is defined by a  cost function
 \begin{equation}
@@
 The term \( \vec{x}_n(N) \) is the final state of a control-free trajectory while the term \( \vec{x}(N)\) is the final state under the influence of the control input.
-Maximization of the distance between \( \vec{x}_n \) and \(\vec{x} \) is equivalent to the minimization of \( J \).
+Maximization of the distance between \( \vec{x}_n \) and \(\vec{x} \), in the \Poincare section defined in~\cref{eqn:poincare_section}, is equivalent to the minimization of \( J \).
 The \Poincare section is defined through the use of appropriate terminal constraints given by
@@
 where the angle \( \theta_d\) defines a direction in which we wish to maximize the reachability set on the \Poincare section.
-We define the \Poincare section along the horizontal axis and which is equivalent to the  surface \( y(N) = 0 \).
-This is similar to the previous work in determining homoclinic orbits in the three-body problem~\cite{llibre1985,koon2011}.
-Previous analytical results have shown that homoclinic orbits intersect transversally in the \( (x, \dot{x} ) \) space on the plane \( y = 0 \).
-We seek to compliment these results with the addition of low thrust propulsion.
 The maximum control thrust magnitude is defined by \( u_{max} \) and is non-dimensionalized by the characteristic units of length, mass, and time.
@@
 The optimal control formulation presented in this section results in a two point boundary value problem (TPBVP). 
-There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{kirk2012,bryson1975}.
+There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{bryson1975,kirk2012}.
 Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement.
@@
 \section{Numerical Example}\label{sec:simulation}
+We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
+The simulations demonstrate the ability of using the reachable set to generate orbital transfer trajectories.
 \subsection{Periodic Orbit transfer}\label{sec:periodic_orbit_transfer}
 % define problem setup (initial and target orbits)
-We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
 The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
-The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix} \) with \( \mu = 0.0125 \).
+The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
 The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
@@
 \end{figure}
-\Cref{fig:moon_orbit} shows that the target set remain in the vicinity of the Moon in the rotating reference frame. 
+\Cref{fig:moon_orbit} shows that the target set remain in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
 This type of orbit would be useful for a variety of mission scenarios.
@@
 The numerical values associated with the green points denote the required time of flight along the invariant manifold in non-dimensional units.
+The required travel time for a transfer using the unstable invariant manifold is \( t_f \approx 3.1\).
 \begin{figure} 
@@
 		\includegraphics[width=\textwidth]{reach_trajectory} 
-		\caption{Reachability trajectories} \label{fig:reach_trajectory} 
+		\caption{Reachability set trajectories} \label{fig:reach_trajectory} 
 	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
@@
 		\caption{Poincar\`e Section} \label{fig:poincare_compare} 
-	\end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+	\end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
+
+	\begin{subfigure}[htbp]{0.5\textwidth} 
+		\includegraphics[width=\textwidth]{poincare_compare} 
+		\caption{Transfer trajectory} \label{fig:reach_transfer} 
+	\end{subfigure}~ 
+	\begin{subfigure}[htbp]{0.5\textwidth} 
+		\includegraphics[width=\textwidth]{control_input_l1} 
+		\caption{Control input} \label{fig:control_l1} 
+	\end{subfigure}~
 	\caption{Reachability set transfer}
@@
 We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
-The stable manifold is propagated to the \Poincare section at \( y = 0 \) denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
+The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
 The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_{max} = 0.75 \).

### 2015-12-10 acta_manuscript.tex
made second set of corrections from dr. lee. regenerated some of the plots

@@
 
-To address the issues associated with conventional numerical integration, we apply computational geometric optimal control techniques. 
+In order to address the issues associated with conventional numerical integration, we apply computational geometric optimal control techniques. 
 We generalize the previous work of determining orbital transfers using invariant manifolds~\cite{koon2011}.
@@
 \subsection{Jacobi Integral}\label{sec:jacobi}
-There exists a single integral, or constant of motion for the three-body problem~\cite{lanczos1970,szebehely1967}.
+There exists a single integral, or constant of motion for the three-body problem~\cite{szebehely1967,lanczos1970}.
 This energy constant is analogous to the total mechanical energy, however it is a non-physical quantity arising from the problem formulation~\cite{szebehely1967}.
@@
 The term \( \vec{x}_n(N) \) is the final state of a control-free trajectory while the term \( \vec{x}(N)\) is the final state under the influence of the control input.
-Maximization of the distance between \( \vec{x}_n \) and \(\vec{x} \), in the \Poincare section defined in~\cref{eqn:poincare_section}, is equivalent to the minimization of \( J \).
+Maximization of the distance between \( \vec{x}_n \) and \(\vec{x} \), on the \Poincare section defined in~\cref{eqn:poincare_section}, is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
 The \Poincare section is defined through the use of appropriate terminal constraints given by
@@
 \begin{align}
-    m_1( \vec{x}(N) ) &= 0 = y(N) \, , \\ 
-    m_2 ( \vec{x}(N) )&= 0 = \frac{\dot{x}(N) - \dot{x}_n(N) }{x(N) -x_n(N) } - \tan{\theta_d} \, , \\
+    m_1( \vec{x}(N) ) &= y(N) = 0 \, , \\ 
+    m_2 ( \vec{x}(N) )&=  \frac{\dot{x}(N) - \dot{x}_n(N) }{x(N) -x_n(N) } - \tan{\theta_d} = 0\, , \\
 	 0 &\geq\vec{u}^T \vec{u} - u_{max}^2 \, ,
@@
 In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
-Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs.
+Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs~\cite{stoer2013}.
 The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches.
@@
 We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
-The simulations demonstrate the ability of using the reachable set to generate orbital transfer trajectories.
+These simulations enable the spacecraft to depart from the natural dynamics through the use of low-thrust propulsion.
+The reachability set on the \Poincare section allows for a straightforward method of determining transfer opportunities.
+
 \subsection{Periodic Orbit transfer}\label{sec:periodic_orbit_transfer}
@@
 		\includegraphics[width=\textwidth]{poincare_compare} 
-		\caption{Poincar\`e Section} \label{fig:poincare_compare} 
+		\caption{\Poincare section} \label{fig:poincare_compare} 
 	\end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
@@
 	\begin{subfigure}[htbp]{0.5\textwidth} 
-		\includegraphics[width=\textwidth]{poincare_compare} 
-		\caption{Transfer trajectory} \label{fig:reach_transfer} 
+		\includegraphics[width=\textwidth]{reach_transfer} 
+		\caption{Final transfer trajectory} \label{fig:reach_transfer} 
 	\end{subfigure}~ 
@@
 % describe initial and target states for this example
-This numerical simulation demonstrates the ability of using multiple iterations of the reachability set to achieve a more complicated transfer.
+This numerical simulation demonstrates the ability of the proposed approach. 
+We use multiple iterations of the reachability set to achieve a more complicated transfer.
 In this manner, it is possible to design arbitrary transfers which are not possible using a single reachability computation.
@@
    \includegraphics[width=0.5\textwidth]{third_reachability_set} % requires the graphicx package
-   \caption{Third reachability set}
+   \caption{Reachable set after three iterations}
    \label{fig:third_reachability_set}
@@
 		\includegraphics[width=\textwidth]{geo_transfer_full} 
-		\caption{Transfer} \label{fig:geo_transfer_full} 
+		\caption{Completed transfer trajectory} \label{fig:geo_transfer_full} 
 	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
@@
 		\includegraphics[width=\textwidth]{geo_transfer_zoom} 
-		\caption{State Trajectory} \label{fig:geo_transfer_zoom} 
+		\caption{Detailed view of transfer} \label{fig:geo_transfer_zoom} 
 	\end{subfigure}
@@
 	\end{subfigure} 
-	\caption{Geostationary Transfer}
+	\caption{Transfer from geostationary orbit to the \( L_1 \) stable manifold using multiple iterations of the reachability set}
 	\label{fig:geo_transfer} 

### 2015-12-10 acta_manuscript.tex
made some updates and corrections to paper

@@
 \end{equation}
-The larger primary, \(m_1\), is located at \( \left(  -\mu , 0 \right)\) and the smaller \( m_2\) is located at \( \left( 1-\mu , 0 \right)\), while the spacecraft is located at \( x \hat{e}_1 + y \hat{e}_2 \).
+The larger primary, \(m_1\), is located at \( \left(  -\mu , 0 \right)\) and the smaller \( m_2\) is located at \( \left( 1-\mu , 0 \right)\), while the spacecraft is located at \( (x ,y) \).
 In the rotating reference frame the Lagrangian is given by
@@
 \end{equation}
-where the distance \( r_1, r_2 \in \R \) of the spacecraft to each primary is defined in rotating reference frame as
+where the distance \( r_1, r_2 \in \R \) of the spacecraft to each primary is as
 \begin{align}
@@
 \end{equation}
-and the control inputs are defined as \( \vec{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \).
+and the control inputs is defined as \( \vec{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \).
 The state is defined as \( \vec{x} = \begin{bmatrix}\vec{r} &\vec{v} \end{bmatrix}^T\) with \(\vec{r} = \begin{bmatrix} x & y \end{bmatrix}^T \in \R^{2\times1}\) and \(\vec{v}= \begin{bmatrix} \dot{x} & \dot{y} \end{bmatrix}^T \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
@@
 \end{align}
-
 where we have used integration by parts and the conditions \( \delta q(0) = 0 \) and \( \delta q(T) = 0\).
@@
 		\includegraphics[width=\textwidth]{trajectory} 
-		\caption{Trajectory} \label{fig:compare_trajectory} 
+		\caption{Trajectory comparison} \label{fig:compare_trajectory} 
 	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
@@
 		\includegraphics[width=\textwidth]{energy} 
-		\caption{Jacobi Integral} \label{fig:compare_energy} 
+		\caption{Jacobi Integral comparison} \label{fig:compare_energy} 
 	\end{subfigure} 
@@
 Each periodic orbit has two stable and unstable manifold sets associated with both a positive or negative perturbation along the respective eigenvector.
-This is technique of determining the invariant manifolds is graphically presented in~\cref{fig:invariant_manifold}.
+This technique of determining the invariant manifolds is graphically presented in~\cref{fig:invariant_manifold}.
 \begin{figure}
@@
 For example, \Poincare maps have been used to prove the existence of homoclinic orbits, which are orbits both forward and backward asymptotic to a single unstable periodic orbit, and heteroclinic orbits, which join different periodic orbits~\cite{conley1968,koon2000b}.
-These dynamic features have been shown to play a vital role in the movement of natural bodies as well as critical for spacecraft missions~\cite{lo1997,gomez2001}.
+These dynamic features have been shown to play a vital role in the movement of natural bodies as well as critical for spacecraft missions~\cite{gomez2001,lo1997}.
 \begin{figure}
@@
 The surface of section is placed at a strategic location in order to allow for a cross section of the flow within the three-dimensional energy surface. 
-In a frame aligned with the rotating frame, \( ( \hat{e}_1, \hat{e}_2 ) \), but centered on the equilibrium point \( L_1 \) or \( L_2\), the unstable manifolds are locally moving in forward time in the second or forth quadrant.
+In a frame aligned with the rotating frame, \( ( \hat{e}_1, \hat{e}_2 ) \), but centered on the equilibrium point \( L_1 \) or \( L_2\), the unstable manifolds are locally moving in forward time in the second and forth quadrant.
 Similarly, the stable manifold branches are locally moving in backward time in the first and third quadrants as illustrated in~\cref{fig:invariant_manifold}. 
@@
 Next, we determine the reachability set with the addition of a low thrust control input.
-We define a maximum magnitude of the thrust as \( u_{max} = 0.5 \) and assume we can point the thrust in any direction within the plane. 
+We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
 This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
 We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
-The trajectories are generated from a fixed initial state of \( \vec{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix} \) over a fixed time span of \( t_f = 1.4 \).
+The trajectories are generated from a fixed initial state of \( \vec{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
 This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit.
@@
 As a result, the intersection of this periodic orbit with the \Poincare section are two points corresponding to the two crossing of the orbit.
+We show the control-free intersection, \( \vec{x}_n \), of the periodic orbit on the \Poincare section in~\cref{fig:manifold_poincare,fig:poincare_compare}
 The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
@@
 A final optimal trajectory is generated such that the \( \vec{x}(N) = \vec{x}_t \).
-This transfer trajectory is denoted by the green path in~\cref{fig:reach_trajectory}.
+This transfer trajectory is denoted by the green path in~\cref{fig:reach_transfer}.
+The optimal control input is shown in~\cref{fig:control_l1}. 
+The spacecraft achieves the desired target while satisfying the bounded control input.
 
@@
 Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
+\Cref{fig:control_input_geo} shows the control input during the powered portion of the transfer. 
+The spacecraft maintains a bounded control magnitude during the transfer to the stable manifold.
 \begin{figure} 

### 2017-01-11 setup_repo.sh
Add setup repo script for remotes

@@ -1,0 +1,20 @@
+#!/bin/bash
+
+# add the correct remote repositories
+git_repo="git@github.com:skulumani/2015_Acta_manuscript.git"
+bb_repo="git@bitbucket.org:shankarkulumani/2015_acta_manscript.git"
+
+
+printf "Setting the origin remote to point to Github and Bitbucket:\n\n"
+printf "Github: $git_repo\n"
+printf "Bitbucket: $bb_repo\n\n"
+
+
+git remote add origin $git_repo
+git remote set-url origin --push --add $bb_repo
+git remote set-url origin --push --add $git_repo
+
+printf "Finished! The new remotes are listed.\n"
+printf "\n"
+
+git remote -v

### 2017-01-11 setup_repo.sh
Update setup script with links to repos

@@ -1,9 +1,8 @@
 #!/bin/bash
 
 # add the correct remote repositories
-git_repo="git@github.com:skulumani/2015_Acta_manuscript.git"
-bb_repo="git@bitbucket.org:shankarkulumani/2015_acta_manscript.git"
-
+git_repo="git@github.com:skulumani/2017_JAS.git"
+bb_repo="git@bitbucket.org:shankarkulumani/2017_jas.git"
 
 printf "Setting the origin remote to point to Github and Bitbucket:\n\n"
 printf "Github: $git_repo\n"

### 2017-01-11 manuscript.tex
Use JAS style
Issue #1

--- 2015-12-10 acta_manuscript.tex
@@
-\documentclass[preprint]{elsarticle}
-
-\usepackage{acta_astronautica_packages} % packages and referencing customization
+\documentclass[letterpaper, submit]{AAS}
+
+\usepackage{jas_packages} % packages and referencing customization
 \input{macros} % defines custom macros
 
-\journal{Acta Astronautica}
-
 \begin{document}
 
-\begin{frontmatter}
-
 \title{Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem}
 
-%% Group authors per affiliation:
-\author{Shankar Kulumani\fnref{kulumanifootnote}}
-\address{George Washington University, 800 22nd St NW, Washington, DC 20052}
-
-\author{Taeyoung Lee\fnref{leefootnote}}
-\address{George Washington University, 800 22nd St NW, Washington, DC 20052}
-
-\fntext[kulumanifootnote]{Doctoral Student, Mechanical and Aerospace Engineering}
-\fntext[leefootnote]{Associate Professor, Mechanical and Aerospace Engineering}
+\author{Shankar Kulumani\thanks{Doctoral Student, Mechanical \& Aerospace Engineering, George Washington University, 800 22nd St NW, Washington, DC 20052.},  
+Taeyoung Lee\thanks{Associate Professor, Mechanical \& Aerospace Engineering, George Washington University, 800 22nd St NW, Washington, DC 20052.}
+}
+
+\maketitle{}
 
@@
 \end{abstract}
-
-\begin{keyword}
-Low-thrust, Reachability, Optimal Control, Variational Integrator
-\end{keyword}
-
-\end{frontmatter}
 
@@
 \bibliography{library}
-\bibliographystyle{elsarticle-num}
+\bibliographystyle{AAS_publication}
 \end{document}

### 2017-01-12 manuscript.tex
Add updated and much better tikz plots
Closes #2

@@
 \usepackage{jas_packages} % packages and referencing customization
+\usepackage{jas_tikz_packages}
+\tdplotsetmaincoords{60}{125} % view angle in spherical coordinates
 \input{macros} % defines custom macros
@@
 	\begin{scaletikzpicturetowidth}{0.5\textwidth}
-	\begin{tikzpicture}[scale=\tikzscale]
-		\coordinate [label=above:\textcolor{black}{\large \(\vec{q}_0\)}] (x0) at (0,0);
-		\coordinate [label=below:\textcolor{black}{\large   \(P(\vec{q}_0)\)}] (xn) at (-1,-1);
-		\coordinate [label=left:\textcolor{black}{\large  \(\Sigma\)}] (sigma) at (-4,3);
-		\coordinate [label=below:\textcolor{black}{}] (P) at (0,-3.5);
-		% define the path of the flow with coordinates
-		\coordinate [label=right:\textcolor{blue}{}] (f1) at (5,-2);
-		\coordinate [label=below:\textcolor{black}{}] (f2) at (2,-5);
-		\coordinate [label=right:\textcolor{blue}{}] (f3) at (-4,-4);
-		\coordinate [label=right:\textcolor{blue}{}] (f4) at (-4,-1);
-
-	%	\draw[help lines] (-10,-10) grid (10,10); %grid
-		\filldraw [black] (x0) circle [radius=3pt];
-		\filldraw [black] (xn) circle [radius=3pt];
-
-		\draw [ultra thick,black,->-](x0) to[out=20,in=90,distance=2cm] (f1) to[out=-90,in=0,distance=2cm] (f2) to[out=180,in=-45,distance=2cm] (f3) to[out=135,in=-135,distance=2cm] (f4) ;
-		\draw [ultra thick, black,dashed] (f4) to[out=45,in=120,distance=1cm] (xn);
-
-		\draw [ultra thick] plot [smooth cycle, tension=0.1, rotate=5] coordinates { (-4,-3) (4,-3) (4,3) (-4,3) };
+    \begin{tikzpicture}[tdplot_main_coords,
+      poincare/.style={opacity=.2,very thick,fill=blue},
+      orbit/.style={very thick,black},
+      orbit hidden/.style={very thick,dashed},
+      grid/.style={very thin,gray!50},
+      axis/.style={->,blue,thick}, scale=\tikzscale]
+
+    % nodes for the poincare section
+    \node[label=above:\(\Sigma\)] (upper_right) at (0,5,5) {};
+    \node[] (upper_left) at (0,1,5) {};
+    \node[] (lower_left) at (0,1,0) {};
+    \node[] (lower_right) at (0,5,0) {};
+
+    % draw poincare section
+    \draw[poincare] (upper_right.center) -- (upper_left.center) -- (lower_left.center) -- (lower_right.center) -- (upper_right.center);
+
+    % draw a periodic orbit
+    \coordinate (center) at (0,0,2);
+    \node[below right] (x0) at (0,2,2) {\(\vecbf{x}_0\)};
+    \filldraw (0,2,2) circle (3pt);
+
+    \node[below right] (x1) at (0,3,2) {\(\vecbf{x}_1\)};
+    \filldraw (0,3,2) circle (3pt);
+
+    \tdplotdrawarc[orbit hidden]{(center)}{2}{90}{190}{}{};
+    \tdplotdrawarc[orbit,<-]{(center)}{2}{-170}{90}{}{};
+
+    \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{199}{}{};
+    \tdplotdrawarc[orbit,<-]{(center)}{3}{-161}{90}{}{};
+
 	\end{tikzpicture}
@@
 	\begin{scaletikzpicturetowidth}{0.5\textwidth}
-	\begin{tikzpicture}[scale=\tikzscale]
-		\coordinate [label=left:\textcolor{black}{\large \(\vec{x}_0\)}] (x0) at (-1,-2);
-		\coordinate [label=below:\textcolor{black}{\large  \(\vec{x}_n\)}] (xn) at (1,1);
-		\coordinate [label=left:\textcolor{black}{\large  \(\Sigma\)}] (sigma) at (-4,3);
-		%\coordinate [label=below:\textcolor{black}{\large  \(P(\vec{x})\)}] (P) at (0,-3.5);
-		% define the path of the flow with coordinates
-		\coordinate [label=right:\textcolor{black}{}] (f1) at (5,-2);
-		\coordinate [label=below:\textcolor{black}{\large  \(\phi(t,\vec{x}_0)\)}] (f2) at (2,-5);
-		\coordinate [label=right:\textcolor{black}{}] (f3) at (-4,-4);
-		\coordinate [label=right:\textcolor{black}{}] (f4) at (-4,-1);
-
-	%	\draw[help lines] (-10,-10) grid (10,10); %grid
-		\filldraw [black] (x0) circle [radius=3pt];
-		\filldraw [black] (xn) circle [radius=3pt];
-
-		\draw [ultra thick,black,->-](x0) to[out=20,in=90,distance=2cm] (f1) to[out=-90,in=0,distance=2cm] (f2) to[out=180,in=-45,distance=2cm] (f3) to[out=135,in=-135,distance=2cm] (f4) ;
-		\draw [ultra thick, black,dashed,->] (f4) to[out=45,in=180,distance=1cm] ($(xn)-(2,0)$);
-
-		\draw [ultra thick] plot [smooth cycle, tension=0.1, rotate=5] coordinates { (-4,-3) (4,-3) (4,3) (-4,3) };
-
-		\draw [thick,dashed] (xn) circle [radius=2cm]; % reachability set
-
-		\draw [thick,->] (xn) -- ($(xn) + (2.5,0)$);
-		\draw [thick,rotate=45,->] (xn) -- ($(xn) + (2.5,0)$);
-		\draw ($(xn) + (1,0)$) arc [start angle=0,end angle=45, radius=1];
-		\node [draw=none] at (2.4,1.5) {\Large \(\theta_d\)};
-		\draw [decorate,decoration={brace,amplitude=5pt},rotate=45] (xn) -- ($(xn) + (2,0)$);
-		\node [draw=none] at ($ (xn) + (0,1) $) {\Large \( J \)};
+	\begin{tikzpicture}[tdplot_main_coords,
+          poincare/.style={opacity=.2,very thick,fill=blue},
+          orbit/.style={very thick,black},
+          orbit hidden/.style={very thick,dashed},
+          grid/.style={very thin,black},
+          axis/.style={->,blue,thick},
+          reachability/.style={thick,blue},scale=\tikzscale]
+
+        % nodes for the poincare section
+        \node[label=above:\(\Sigma\)] (upper_right) at (0,5,5) {};
+        \node[] (upper_left) at (0,1,5) {};
+        \node[] (lower_left) at (0,1,0) {};
+        \node[] (lower_right) at (0,5,0) {};
+
+        % draw poincare section
+        \draw[poincare] (upper_right.center) -- (upper_left.center) -- (lower_left.center) -- (lower_right.center) -- (upper_right.center);
+
+        % draw a periodic orbit
+        \coordinate (center) at (0,0,2);
+        \node[label=below:\(\vecbf{x}_0\)] (x0) at (0,3,2) {};
+        % \node[label=below:\(\vecbf{x}_n\)] at (x0) {};
+        \filldraw (x0) circle (3pt);
+
+        % \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{200}{}{};
+        \tdplotdrawarc[orbit,<-]{(center)}{3}{-160}{90}{}{};
+
+        % draw reachability set on the poincare section
+        \coordinate (reach) at (0,4.5,2);
+        \tdplotsetthetaplanecoords{90}
+
+        \draw[tdplot_rotated_coords,grid] (x0) -- (reach);
+        \draw[tdplot_rotated_coords,grid] (x0) -- ++(-45:1.5);
+
+        \tdplotdrawarc[tdplot_rotated_coords,grid]{(x0)}{0.5}{-45}{90}{above}{\(\phi\)};
+
+        % draw terminal state on reachability set
+        \node[tdplot_rotated_coords,label=above:\(\vecbf{x}_f\)] (xf) at ($ (x0)+(-45:1.5) $) {};
+        \filldraw (xf) circle (3pt);
+
+        \node[tdplot_rotated_coords,label=below:\(J\)] at (xf) {};
+
+        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
+        % place
     \end{tikzpicture}

### 2017-01-13 manuscript.tex
Cref to a section isn't working
Use the actual JAS template
Issue #1
Add formatting specific to springer journal

@@
-\documentclass[letterpaper, submit]{AAS}
+\makeatletter
+\def\cl@chapter{}
+\makeatother
+
+\RequirePackage{fix-cm}
+
+\documentclass[smallcondensed]{svjour3}
 
 \usepackage{jas_packages} % packages and referencing customization
+\usepackage{mathptmx}
 \usepackage{jas_tikz_packages}
@@
 
-\title{Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem}
-
-\author{Shankar Kulumani\thanks{Doctoral Student, Mechanical \& Aerospace Engineering, George Washington University, 800 22nd St NW, Washington, DC 20052.},  
-Taeyoung Lee\thanks{Associate Professor, Mechanical \& Aerospace Engineering, George Washington University, 800 22nd St NW, Washington, DC 20052.}
+\title{Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem
+\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
 }
 
+\author{Shankar Kulumani \and Taeyoung Lee}
+
+\institute{S. Kulumani \and T. Lee \at
+		Department of Mechanical \& Aerospace Engineering,\\
+		George Washington University,\\
+		800 22nd St NW, Washington, DC 20052.\\
+		\email{skulumani@gwu.edu}\\
+		\email{tylee@gwu.edu}
+		}
+
+\date{Received: date / Accepted: date}
+
 \maketitle{}        
-
 \begin{abstract}
@@
 We demonstrate our approach with two numerical simulations of transfers in the Earth-Moon three-body system.
+\keywords{First keyword \and Second keyword \and More}
+% \PACS{PACS code1 \and PACS code2 \and more}
+% \subclass{MSC code1 \and MSC code2 \and more}
 \end{abstract}
@@
 Dynamics systems theory has been applied to design control-free maneuvers in the restricted three-body problem~\cite{koon2011}.
-As previously introduced in~\cref{sec:jacobi}, there exist five equilibrium points in the equations of motion for the three-body problem.
+As previously introduced in~\Cref{sec:jacobi}, there exist five equilibrium points in the equations of motion for the three-body problem.
 It has been shown that the local orbit structure near the Lagrange points gives rise to families of periodic orbits as well as the stable and unstable manifolds of these periodic orbits.
@@
 %As as result any errors will compound and tend to accumulate most prevalently in \( \lambda_{\xkp} \).
-\section*{References}
+
+%\begin{acknowledgements}
+%If you'd like to thank anyone, place your comments here
+%and remove the percent signs.
+%\end{acknowledgements}
 
 \bibliography{library}
-\bibliographystyle{AAS_publication}
+\bibliographystyle{spmpsci}
 \end{document}

### 2017-01-13 manuscript.tex
Use cite instead of citenum

@@
 The state and/or control trajectories are parameterized and solved in the form of a nonlinear optimization problem.
-References~\citenum{mingotti2011} and~\citenum{grebow2011} use this direct approach in designing low-thrust transfers in the three-body problem.
+The work in~\cite{mingotti2011,grebow2011} use this direct approach in designing low-thrust transfers in the three-body problem.
 Alternatively, indirect methods apply calculus of variations to derive the necessary conditions for optimality. 
@@
 Efficient numerical implementation is dependent on correct initial conditions as well as accurate numerical integration.
-References~\citenum{mingotti2011} and~\citenum{grebow2011} implement the solutions using conventional Runge-Kutta integration techniques.
+\cite{mingotti2011,grebow2011} implement the solutions using conventional Runge-Kutta integration techniques.
 These techniques suffer from numerical instability and energy drift behaviors which make them ill-suited for long-term propagation.
@@
 % DYNAMICAL STRUCTURE
-References~\citenum{koon2011} and~\citenum{ross2006} have illustrated the rich structure that exists in the three-body problem.
+\cite{koon2011,ross2006} have illustrated the rich structure that exists in the three-body problem.
 Within the three-body problem, a spacecraft's feasible region of motion is constrained by its energy, or Jacobi integral. 
@@
 This yields a discrete Hamiltonian map \( (q_k, p_k) \to (q_{k+1}, p_{k+1}) \).
-A more extensive development of variational integrators can be found in Reference~\citenum{marsden2001}.
+A more extensive development of variational integrators can be found in~\cite{marsden2001}.
 
@@
 In this way, it is possible to design trajectories that intersect the invariant manifolds and connect widely separated regions of the state space. 
-This technique has been heavily investigated in Reference~\citenum{koon2011}, applied to operational missions in Reference~\citenum{koon1999}, and applied to potential multi-moon orbiter missions in Reference~\citenum{tanaka2011}.
+This technique has been heavily investigated in~\cite{koon2011}, applied to operational missions in~\cite{koon1999}, and applied to potential multi-moon orbiter missions in~\cite{tanaka2011}.
 In addition, much recent work has focused on potential return missions to the Moon~\cite{zanzottera2012,campagnola2012,mingotti2011,ozimek2010a,mingotti2009}.
@@
 % invariant manifold transfer
-As a source for comparison, the method of using invariant manifolds, introduced in Reference~\citenum{koon2011}, is implemented.
+As a source for comparison, the method of using invariant manifolds, introduced in~\cite{koon2011}, is implemented.
 As described in~\cref{sec:invariant_manifold}, these invariant manifolds are the set of trajectories that either asymptotically arrive or depart the periodic orbit. 

### 2017-01-17 setup_repo.sh
Add repo remote setup script

@@ -3,15 +3,16 @@
 # add the correct remote repositories
 git_repo="git@github.com:skulumani/2017_JAS.git"
 bb_repo="git@bitbucket.org:shankarkulumani/2017_jas.git"
+fdcl_repo="git@github.com:fdcl-gwu/2017_JAS.git"
 
 printf "Setting the origin remote to point to Github and Bitbucket:\n\n"
 printf "Github: $git_repo\n"
 printf "Bitbucket: $bb_repo\n\n"
 
 
-git remote add origin $git_repo
 git remote set-url origin --push --add $bb_repo
 git remote set-url origin --push --add $git_repo
+git remote add fdcl $fdcl_repo
 
 printf "Finished! The new remotes are listed.\n"
 printf "\n"

### 2017-01-17 manuscript.tex
Use \vecbf instead of \vec which is already defined by amsmath

@@
 \end{equation}
-and the control inputs is defined as \( \vec{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \).
-The state is defined as \( \vec{x} = \begin{bmatrix}\vec{r} &\vec{v} \end{bmatrix}^T\) with \(\vec{r} = \begin{bmatrix} x & y \end{bmatrix}^T \in \R^{2\times1}\) and \(\vec{v}= \begin{bmatrix} \dot{x} & \dot{y} \end{bmatrix}^T \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
+and the control inputs is defined as \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \).
+The state is defined as \( \vecbf{x} = \begin{bmatrix}\vecbf{r} &\vecbf{v} \end{bmatrix}^T\) with \(\vecbf{r} = \begin{bmatrix} x & y \end{bmatrix}^T \in \R^{2\times1}\) and \(\vecbf{v}= \begin{bmatrix} \dot{x} & \dot{y} \end{bmatrix}^T \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
 The equations of motion may be rewritten in state space form as
 \begin{equation}\label{eqn:cont_dyn}
-	\left[\begin{array}{c} \dot{\vec{r}} \\ \dot{\vec{v}} \end{array} \right] = 
-	\left[ \begin{array}{c} \vec{v} \\ A \vec{v} + \nabla U + \vec{u} \end{array} \right] = f\left( t,\vec{x}, \vec{u}\right) \, ,
+	\left[\begin{array}{c} \dot{\vecbf{r}} \\ \dot{\vecbf{v}} \end{array} \right] = 
+	\left[ \begin{array}{c} \vecbf{v} \\ A \vecbf{v} + \nabla U + \vecbf{u} \end{array} \right] = f\left( t,\vecbf{x}, \vecbf{u}\right) \, ,
 \end{equation}
@@
 \begin{equation}
-	E\left( \vec{r} , \vec{v} \right) = \frac{1}{2}\left( \dot{x}^2 + \dot{y}^2\right) - U\left(x,y \right) \, .
+	E\left( \vecbf{r} , \vecbf{v} \right) = \frac{1}{2}\left( \dot{x}^2 + \dot{y}^2\right) - U\left(x,y \right) \, .
 	\label{eq:jacobi}
@@
 A discrete time Lagrangian \( L_d( q_0, q_1) \) is formed which approximates the action integral between \( q_0 \) and \( q_1 \) as 
-\begin{align}\label{eq:discrete_lagrangian}
+\begin{align}\label{eq:discrete_lagrangian_gen}
 	L_d\left( q_0 , q_1 \right) \approx \int_{0}^{h} L \left( q , \dot{q} \right) \, dt \, .
 \end{align}
-Since~\cref{eq:discrete_lagrangian} is calculated as a numerical integral, an appropriate quadrature rule is required.
-There are multiple possible methods one can use to approximate the integral in~\cref{eq:discrete_lagrangian}.
+Since~\cref{eq:discrete_lagrangian_gen} is calculated as a numerical integral, an appropriate quadrature rule is required.
+There are multiple possible methods one can use to approximate the integral in~\cref{eq:discrete_lagrangian_gen}.
 An appropriate approximation rule is determined based on the ease of implementation and accuracy desired.
@@
 The discrete equations of motion are given in the Lagrangian form after applying the discrete fiber derivative from~\cref{eq:discrete_legendre} as \( p_{\xk} = \dot{x}_k - \yk \) and \( p_{\yk} = \dot{y}_k + \xk \).
-The state is defined as \( \vec{x}_k = \begin{bmatrix} \xk & \yk & \dot{x}_k & \dot{y}_k \end{bmatrix}^T\) and the control input is \( \vec{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \).
-This results in a discrete update map \( f_k: \vec{x}_k \to \vec{x}_{k+1} \) which preserves the same properties of the continuous-time dynamics in~\cref{eqn:cont_dyn} such as invariants, symplecticity, and the configuration manifold.
+The state is defined as \( \vecbf{x}_k = \begin{bmatrix} \xk & \yk & \dot{x}_k & \dot{y}_k \end{bmatrix}^T\) and the control input is \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \).
+This results in a discrete update map \( f_k: \vecbf{x}_k \to \vecbf{x}_{k+1} \) which preserves the same properties of the continuous-time dynamics in~\cref{eqn:cont_dyn} such as invariants, symplecticity, and the configuration manifold.
 The discrete potential gradients are given by
@@
 As~\cref{eq:discrete_potential_grad,eq:discrete_distance} are defined at both step \( k \) and \( k+1 \) they must be evaluated at both time instances.
-\Cref{eq:discrete_eoms} is implemented by first defining an initial state \( \vec{x}_k \) and control \( \vec{u}_k \).
+\Cref{eq:discrete_eoms} is implemented by first defining an initial state \( \vecbf{x}_k \) and control \( \vecbf{u}_k \).
 The distances and gravitational potential at step \( k \) are evaluated from~\cref{eq:distonek,eq:disttwok,eq:Uxk,eq:Uyk}.
@@
 Finally, the update steps in~\cref{eq:xdotkp,eq:ydotkp} are evaluated.
-This results in the complete discrete update map \( \vec{x}_k \to \vec{x}_{k+1} \) given \( \vec{u}_k \).
+This results in the complete discrete update map \( \vecbf{x}_k \to \vecbf{x}_{k+1} \) given \( \vecbf{u}_k \).
 \subsection{Numerical Example}\label{sec:variational_example}
@@
 A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
-A particle is simulated from an initial condition of \( \vec{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
+A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
 The variational integrator uses a step size of \SI{47.22}{\second} while the Runge-Kutta method uses a variable step size implemented via ODE45 in Matlab.
@@
 The process of differential correction and numerical continuation are used to determine a periodic orbit with a desired energy level. 
-Let the trajectories of the equations of motion defined in~\cref{eqn:cont_dyn} be defined by the flow map \( \phi(t;\vec{x}_0) \) originating from \( \vec{x}(t_0) = \vec{x}_0 \) and we define a periodic orbit as \( \gamma\).
+Let the trajectories of the equations of motion defined in~\cref{eqn:cont_dyn} be defined by the flow map \( \phi(t;\vecbf{x}_0) \) originating from \( \vecbf{x}(t_0) = \vecbf{x}_0 \) and we define a periodic orbit as \( \gamma\).
 The invariant manifolds are defined as : 
 \begin{align*}
-	W^s ( \vec{x}_0 ) &= \braces{ \vec{x} : \phi(t;\vec{x}_0) \to \gamma \text{ as } t \to \infty} , \\
-	W^u ( \vec{x}_0 ) &= \braces{ \vec{x} : \phi(t;\vec{x}_0) \to \gamma \text{ as } t \to -\infty} .
+	W^s ( \vecbf{x}_0 ) &= \braces{ \vecbf{x} : \phi(t;\vecbf{x}_0) \to \gamma \text{ as } t \to \infty} , \\
+	W^u ( \vecbf{x}_0 ) &= \braces{ \vecbf{x} : \phi(t;\vecbf{x}_0) \to \gamma \text{ as } t \to -\infty} .
 \end{align*}
-The stable manifold, \( W^s(\vec{x}_0) \), is the set of trajectories which asymptotically approach the periodic orbit.
-Similarly, the unstable manifold, \( W^u(\vec{x}_0) \), is the set of trajectories which asymptotically depart the periodic orbit.
+The stable manifold, \( W^s(\vecbf{x}_0) \), is the set of trajectories which asymptotically approach the periodic orbit.
+Similarly, the unstable manifold, \( W^u(\vecbf{x}_0) \), is the set of trajectories which asymptotically depart the periodic orbit.
 For the nonlinear system, defined by~\cref{eqn:cont_dyn} or~\cref{eq:discrete_eoms}, we can determine the state transition matrix \( \Phi \) along the periodic orbit, from \( 0 \to T\), by numerical integration of the variational equations.
-In the neighborhood of the periodic orbit \( \gamma \), the tangent spaces to the stable and unstable manifolds are provided by the generalized eigenspaces \( E^s \) and \( E^u \) of the linearization \( A = \D f(\vec{x}_0) \), where \( \vec{x}_0 \) is on the periodic orbit.
+In the neighborhood of the periodic orbit \( \gamma \), the tangent spaces to the stable and unstable manifolds are provided by the generalized eigenspaces \( E^s \) and \( E^u \) of the linearization \( A = \D f(\vecbf{x}_0) \), where \( \vecbf{x}_0 \) is on the periodic orbit.
 The eigenvectors of the monodromy matrix, \( M = \Phi(T) \), serve to locally approximate the directions of the stable and unstable manifolds. 
@@
 \Poincare maps are a useful tool in the analysis of the flow near periodic orbits in the three-body problem.
-We let \( \Sigma \) define a hypersurface of section chosen such that all trajectories in the vicinity of a state \( \vec{q} \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
-A \Poincare map, \( P(\vec{q}) = \phi(T;\vec{q}) \), maps the state of a trajectory from one intersection to the next.
+We let \( \Sigma \) define a hypersurface of section chosen such that all trajectories in the vicinity of a state \( \vecbf{q} \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
+A \Poincare map, \( P(\vecbf{q}) = \phi(T;\vecbf{q}) \), maps the state of a trajectory from one intersection to the next.
 Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
@@
 and located to allow an intersection between both invariant manifolds.
-The configuration space of the invariant manifold is \( \vec{x} \in \R^4 \).
+The configuration space of the invariant manifold is \( \vecbf{x} \in \R^4 \).
 The surface of section, and the resulting intersection of the four-dimensional space with the hyperplane \( \Sigma \), reduces the resulting intersection to a surface in \( \R^3\).
@@
 The main idea is that the addition of low thrust propulsion allows us to enlarge the set of trajectories achievable in the \Poincare section. 
-\Cref{fig:reachability_set} illustrates how, without any control input, trajectories will intersect with the \Poincare section at \( \vec{x}_n \). 
+\Cref{fig:reachability_set} illustrates how, without any control input, trajectories will intersect with the \Poincare section at \( \vecbf{x}_n \). 
 However, the addition of low thrust propulsion allows the spacecraft to depart from the natural dynamics and intersect the \Poincare section at a different location.
@@
 \begin{equation}
-	J = -\frac{1}{2} \left( \vec{x}(N) - \vec{x}_{n}(N)\right)^T 
+	J = -\frac{1}{2} \left( \vecbf{x}(N) - \vecbf{x}_{n}(N)\right)^T 
 	\left[
@@
 	\right]
-	\left( \vec{x}(N) - \vec{x}_{n}(N)\right) \, .
+	\left( \vecbf{x}(N) - \vecbf{x}_{n}(N)\right) \, .
 	\label{eq:cost}
 \end{equation}
-The term \( \vec{x}_n(N) \) is the final state of a control-free trajectory while the term \( \vec{x}(N)\) is the final state under the influence of the control input.
-Maximization of the distance between \( \vec{x}_n \) and \(\vec{x} \), on the \Poincare section defined in~\cref{eqn:poincare_section}, is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
+The term \( \vecbf{x}_n(N) \) is the final state of a control-free trajectory while the term \( \vecbf{x}(N)\) is the final state under the influence of the control input.
+Maximization of the distance between \( \vecbf{x}_n \) and \(\vecbf{x} \), on the \Poincare section defined in~\cref{eqn:poincare_section}, is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
 The \Poincare section is defined through the use of appropriate terminal constraints given by
@@
 \begin{align}
-    m_1( \vec{x}(N) ) &= y(N) = 0 \, , \\ 
-    m_2 ( \vec{x}(N) )&=  \frac{\dot{x}(N) - \dot{x}_n(N) }{x(N) -x_n(N) } - \tan{\theta_d} = 0\, , \\
-	 0 &\geq\vec{u}^T \vec{u} - u_{max}^2 \, ,
+    m_1( \vecbf{x}(N) ) &= y(N) = 0 \, , \\ 
+    m_2 ( \vecbf{x}(N) )&=  \frac{\dot{x}(N) - \dot{x}_n(N) }{x(N) -x_n(N) } - \tan{\theta_d} = 0\, , \\
+	 0 &\geq\vecbf{u}^T \vecbf{u} - u_{max}^2 \, ,
 \end{align}
@@
 The maximum control thrust magnitude is defined by \( u_{max} \) and is non-dimensionalized by the characteristic units of length, mass, and time.
-The goal is to determine the control input \( \vec{u}_k\) such that the cost function~\cref{eq:cost} is minimized subject to the state equations of motion~\cref{eq:discrete_eoms} and constraints~\cref{eq:constraints}.
+The goal is to determine the control input \( \vecbf{u}_k\) such that the cost function~\cref{eq:cost} is minimized subject to the state equations of motion~\cref{eq:discrete_eoms} and constraints~\cref{eq:constraints}.
 
@@
 \begin{align}
-	\vec{\lambda}_{k+1}^T &= \vec{\lambda}_k^T  \parenth{\deriv{\vec{f}_k}{\vec{x}_k}}^{-1} \, , \\
-	0 &=  \deriv{H_k}{\vec{u}_k} \, , \label{eqn:control_necc_cond}\\
-	0 &= \deriv{\phi}{\vec{x}_k}^T + \deriv{\vec{m}}{\vec{x}_k}^T \vec{\beta}  - \vec{\lambda}^T(N) \, ,  
+	\vecbf{\lambda}_{k+1}^T &= \vecbf{\lambda}_k^T  \parenth{\deriv{\vecbf{f}_k}{\vecbf{x}_k}}^{-1} \, , \\
+	0 &=  \deriv{H_k}{\vecbf{u}_k} \, , \label{eqn:control_necc_cond}\\
+	0 &= \deriv{\phi}{\vecbf{x}_k}^T + \deriv{\vecbf{m}}{\vecbf{x}_k}^T \vecbf{\beta}  - \vecbf{\lambda}^T(N) \, ,  
 \end{align}
@@
 \begin{equation}
-	H_k = \vec{\lambda}_k^T \vec{f}(\vec{x}_k, \vec{u}_k) \, ,
+	H_k = \vecbf{\lambda}_k^T \vecbf{f}(\vecbf{x}_k, \vecbf{u}_k) \, ,
 	\label{eq:hamiltonian_opt}
 \end{equation}
-and \( \vec{\lambda} \in \R^{4 \times 1} \) is the costate and \(\vec{\beta} \in \R^{2 \times 1} \) are the additional Lagrange multipliers associated with the terminal constraints in~\cref{eq:constraints}.
-The state dynamics are represented by \( \vec{f}(\vec{x}_k, \vec{\lambda}_k ) \) after substituting~\cref{eqn:control_necc_cond} into~\cref{eq:discrete_eoms}.
+and \( \vecbf{\lambda} \in \R^{4 \times 1} \) is the costate and \(\vecbf{\beta} \in \R^{2 \times 1} \) are the additional Lagrange multipliers associated with the terminal constraints in~\cref{eq:constraints}.
+The state dynamics are represented by \( \vecbf{f}(\vecbf{x}_k, \vecbf{\lambda}_k ) \) after substituting~\cref{eqn:control_necc_cond} into~\cref{eq:discrete_eoms}.
 This indirect optimal control formulation leads to a two point boundary value problem with split boundary conditions. 
@@
 \begin{align}\label{eq:costate_eom}
-	\vec{\lambda}_{k+1}^T = \vec{\lambda}_k^T
+	\vecbf{\lambda}_{k+1}^T = \vecbf{\lambda}_k^T
 	\begin{bmatrix} 
@@
 This is a computationally expensive operation that is prone to numerical error and instability.
-A method is presented in Appendix B to avoid this inversion and determine an explicit update map \( \vec{\lambda}_k \to \vec{\lambda}_{k+1} \).
+A method is presented in Appendix B to avoid this inversion and determine an explicit update map \( \vecbf{\lambda}_k \to \vecbf{\lambda}_{k+1} \).
 
@@
 We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
-The trajectories are generated from a fixed initial state of \( \vec{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
+The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
 This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit.
@@
 % discuss reachability set computation
-The optimal trajectories, under the influence of the control input \( \vec{u} \), are plotted in red in~\cref{fig:reach_trajectory}.
+The optimal trajectories, under the influence of the control input \( \vecbf{u} \), are plotted in red in~\cref{fig:reach_trajectory}.
 Initially, the spacecraft is assumed to lie on the periodic orbit.
 As a result, the intersection of this periodic orbit with the \Poincare section are two points corresponding to the two crossing of the orbit.
-We show the control-free intersection, \( \vec{x}_n \), of the periodic orbit on the \Poincare section in~\cref{fig:manifold_poincare,fig:poincare_compare}
+We show the control-free intersection, \( \vecbf{x}_n \), of the periodic orbit on the \Poincare section in~\cref{fig:manifold_poincare,fig:poincare_compare}
 The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
@@
 Using~\cref{eq:jacobi} and the intersection state the final component \( \dot{y} \) is calculated. 
-This results in a complete target state \( \vec{x}_t \) which lies on the reachable set and on the target orbit. 
-A final optimal trajectory is generated such that the \( \vec{x}(N) = \vec{x}_t \).
+This results in a complete target state \( \vecbf{x}_t \) which lies on the reachable set and on the target orbit. 
+A final optimal trajectory is generated such that the \( \vecbf{x}(N) = \vecbf{x}_t \).
 This transfer trajectory is denoted by the green path in~\cref{fig:reach_transfer}.
@@
 \end{subequations}
-Finally, backsubstituion is used to determine explicit equations for the discrete update map \( \vec{\lambda}_k \to \vec{\lambda}_{k+1} \) which is defined as
+Finally, backsubstituion is used to determine explicit equations for the discrete update map \( \vecbf{\lambda}_k \to \vecbf{\lambda}_{k+1} \) which is defined as
 \begin{subequations}\label{eq:costate_update}

### 2017-01-17 manuscript.tex
Remove mathptmx package because of errors
Address some of the reviewer comments

@@
 
-\RequirePackage{fix-cm}
-
 \documentclass[smallcondensed]{svjour3}
@@
 \usepackage{jas_packages} % packages and referencing customization
-\usepackage{mathptmx}
+\usepackage{fix-cm}
 \usepackage{jas_tikz_packages}
@@
 
-We define the \Poincare section along the horizontal axis, which is equivalent to the surface \( y(N) = 0 \), and given by
-\begin{align}
-	\Sigma = \braces{ ( x, \dot{x}) \, | \, y(N) = 0} . 
+We define the \Poincare section along the horizontal axis, which is equivalent to the surface \( y = 0 \), and given by
+\begin{align}
+	\Sigma = \braces{ ( x, \dot{x}) \, | \, y = 0} . 
 	\label{eqn:poincare_section}
@@
 	\right]
-	\left( \vecbf{x}(N) - \vecbf{x}_{n}(N)\right) \, .
+	\left( \vecbf{x}(N) - \vecbf{x}_{n}(N)\right) = \phi(\vecbf{x}(N),\vecbf{x}_n(N)) \, .
 	\label{eq:cost}
@@
 The term \( \vecbf{x}_n(N) \) is the final state of a control-free trajectory while the term \( \vecbf{x}(N)\) is the final state under the influence of the control input.
-Maximization of the distance between \( \vecbf{x}_n \) and \(\vecbf{x} \), on the \Poincare section defined in~\cref{eqn:poincare_section}, is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
+Maximization of the distance between \( \vecbf{x}_n \) and \(\vecbf{x} \), on the \Poincare section defined in~\cref{eqn:poincare_section} at the terminal time \( t_f = N \), is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
 The \Poincare section is defined through the use of appropriate terminal constraints given by
@@
 \begin{align}
-    m_1( \vecbf{x}(N) ) &= y(N) = 0 \, , \\ 
-    m_2 ( \vecbf{x}(N) )&=  \frac{\dot{x}(N) - \dot{x}_n(N) }{x(N) -x_n(N) } - \tan{\theta_d} = 0\, , \\
+    v_1( \vecbf{x}(N) ) &= y(N) = 0 \, , \\ 
+    v_2 ( \vecbf{x}(N) )&=  \frac{\dot{x}(N) - \dot{x}_n(N) }{x(N) -x_n(N) } - \tan{\theta_d} = 0\, , \\
 	 0 &\geq\vecbf{u}^T \vecbf{u} - u_{max}^2 \, ,
@@
 	0 &=  \deriv{H_k}{\vecbf{u}_k} \, , \label{eqn:control_necc_cond}\\
-	0 &= \deriv{\phi}{\vecbf{x}_k}^T + \deriv{\vecbf{m}}{\vecbf{x}_k}^T \vecbf{\beta}  - \vecbf{\lambda}^T(N) \, ,  
+	0 &= \deriv{\phi}{\vecbf{x}_k}^T + \deriv{\vecbf{v}}{\vecbf{x}_k}^T \vecbf{\beta}  - \vecbf{\lambda}^T(N) \, ,  
 \end{align}
@@
 This intersection generates a partial target state of \( x_t \text{ and } \dot{x}_t \).
-Using~\cref{eq:jacobi} and the intersection state the final component \( \dot{y} \) is calculated. 
+Using the energy level of the target region, defined by~\cref{eq:jacobi}, and the intersection state we can calculate the final component \( \dot{y} \) is calculated. 
 This results in a complete target state \( \vecbf{x}_t \) which lies on the reachable set and on the target orbit. 

### 2017-01-17 manuscript.tex
Add multiple shooting tikz plot
Issue #4

@@
 The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches.
+\Cref{fig:multiple shooting} shows a schematic representation of the multiple shooting procedure.
+
+\begin{figure}
+	\centering
+	\begin{scaletikzpicturetowidth}{1\textwidth}
+		\begin{tikzpicture}[scale=\tikzscale]
+		    % \draw[help lines] (0,0) grid (20,5) ; %grid
+
+		    \node[label=west:\(\vecbf{x}\)] (x0m) at (0,5) {};
+		    \node[] (x1m) at (5,5) {};
+		    \node[] (x2m) at (10,5) {};
+		    \node[] (x3m) at (15,5) {};
+		    \node[] (xnm) at (20,5) {};
+
+		    \node[label=below:\(t_0\)] (t0) at (0,0) {};
+		    \node[label=below:\(t_1\)] (t1) at (5,0) {};
+		    \node[label=below:\(t_2\)] (t2) at (10,0) {};
+		    \node[label=below:\(t_{n-1}\)] (t3) at (15,0) {};
+		    \node[label=below:\(t_n\)] (tn) at (20,0) {};
+
+		    % nodes for each subsegment
+		    \node[label=left:\(\vecbf{x}_0\)] (x0) at (0,1) {\pgfuseplotmark{*}};
+
+		    \node[label=right:\(\vecbf{x}_1^{-}\)] (x1minus) at (5,2) {\pgfuseplotmark{*}};
+		    \node[label=left:\(\vecbf{x}_1^{+}\)] (x1plus) at (5,3) {\pgfuseplotmark{o}};
+
+		    \node[label=right:\(\vecbf{x}_2^{-}\)] (x2minus) at (10,2) {\pgfuseplotmark{*}};
+		    \node[label=left:\(\vecbf{x}_2^{+}\)] (x2plus) at (10,3) {\pgfuseplotmark{o}};
+
+		    \node[label=right:\(\vecbf{x}_{n-1}^{-}\)] (x3minus) at (15,2) {\pgfuseplotmark{*}};
+		    \node[label=left:\(\vecbf{x}_{n-1}^{+}\)] (x3plus) at (15,3) {\pgfuseplotmark{o}};
+
+		    \node[label=right:\(\vecbf{x}_n\)] (xnminus) at (20,3) {\pgfuseplotmark{*}};
+		    % draw axes
+		    \draw [<->,thick] (tn.center) -- (t0.center) -- (x0m.center);
+
+		    % draw segement dividers
+		    \draw [dashed,thick] (t1.center) -- (x1m.center);
+		    \draw [dashed,thick] (t2.center) -- (x2m.center);
+		    \draw [dashed,thick] (t3.center) -- (x3m.center);
+		    \draw [dashed,thick] (tn.center) -- (xnm.center);
+
+		    % draw markers at each subsegment
+
+		    % draw the subsegments
+		    \draw [->] (x0.center) to [bend left=30] (x1minus.center);
+		    \draw [->] (x1plus.center) to [bend left=30] (x2minus.center);
+		    % dotted lines here
+		    \draw[decorate sep={2mm}{10mm},fill] (x2plus.east) to [bend left=30] (x3minus.west);
+
+		    \draw [->] (x3plus.center) to [bend left=30] (xnminus.center);
+		\end{tikzpicture}
+	\end{scaletikzpicturetowidth}
+	\caption{Multiple Shooting method\label{fig:multiple shooting}}
+\end{figure}
+
+More detail on how the multiple shooting works
+Explain the image
 Additional interior constraints are used to ensure a continuous solution.

### 2017-01-18 manuscript.tex
Add some more discussion about multiple shooting
Issue #4

@@
 The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches.
-\Cref{fig:multiple shooting} shows a schematic representation of the multiple shooting procedure.
 
@@
 
-		    \node[label=below:\(t_0\)] (t0) at (0,0) {};
-		    \node[label=below:\(t_1\)] (t1) at (5,0) {};
-		    \node[label=below:\(t_2\)] (t2) at (10,0) {};
-		    \node[label=below:\(t_{n-1}\)] (t3) at (15,0) {};
-		    \node[label=below:\(t_n\)] (tn) at (20,0) {};
+		    \node[label=below:\(k_0\)] (t0) at (0,0) {};
+		    \node[label=below:\(k_1\)] (t1) at (5,0) {};
+		    \node[label=below:\(k_2\)] (t2) at (10,0) {};
+		    \node[label=below:\(k_{n-1}\)] (t3) at (15,0) {};
+		    \node[label=below:\(k_n\)] (tn) at (20,0) {};
 
@@
 		    \node[label=left:\(\vecbf{x}_0\)] (x0) at (0,1) {\pgfuseplotmark{*}};
+		    \node[label=below left:\(\vecbf{\lambda}_0\)] at (x0) {};
 
 		    \node[label=right:\(\vecbf{x}_1^{-}\)] (x1minus) at (5,2) {\pgfuseplotmark{*}};
-		    \node[label=left:\(\vecbf{x}_1^{+}\)] (x1plus) at (5,3) {\pgfuseplotmark{o}};
+		    \node[label=below right:\(\vecbf{\lambda}_1^{-}\)] at (x1minus) {};
+
+		    \node[label=above left:\(\vecbf{x}_1^{+}\)] (x1plus) at (5,3) {\pgfuseplotmark{o}};
+		    \node[label=left:\(\vecbf{\lambda}_1^{+}\)] at (x1plus) {};
 
 		    \node[label=right:\(\vecbf{x}_2^{-}\)] (x2minus) at (10,2) {\pgfuseplotmark{*}};
-		    \node[label=left:\(\vecbf{x}_2^{+}\)] (x2plus) at (10,3) {\pgfuseplotmark{o}};
+		    \node[label=below right:\(\vecbf{\lambda}_2^{-}\)] at (x2minus) {};
+
+		    \node[label=above left:\(\vecbf{x}_2^{+}\)] (x2plus) at (10,3) {\pgfuseplotmark{o}};
+		    \node[label=left:\(\vecbf{\lambda}_2^{+}\)] at (x2plus) {};
 
 		    \node[label=right:\(\vecbf{x}_{n-1}^{-}\)] (x3minus) at (15,2) {\pgfuseplotmark{*}};
-		    \node[label=left:\(\vecbf{x}_{n-1}^{+}\)] (x3plus) at (15,3) {\pgfuseplotmark{o}};
+		    \node[label=below right:\(\vecbf{\lambda}_{n-1}^{-}\)] at (x3minus) {};
+
+		    \node[label=above left:\(\vecbf{x}_{n-1}^{+}\)] (x3plus) at (15,3) {\pgfuseplotmark{o}};
+		    \node[label=left:\(\vecbf{\lambda}_{n-1}^{+}\)] at (x3plus) {};
 
 		    \node[label=right:\(\vecbf{x}_n\)] (xnminus) at (20,3) {\pgfuseplotmark{*}};
+		    \node[label=below right:\(\vecbf{\lambda}_n\)] at (xnminus) {};
 		    % draw axes
@@
 		    % dotted lines here
-		    \draw[decorate sep={2mm}{10mm},fill] (x2plus.east) to [bend left=30] (x3minus.west);
+		    \draw[decorate sep={1mm}{5mm},fill] (x2plus.east) to [bend left=10] (x3minus.west);
 
@@
 
-More detail on how the multiple shooting works
-Explain the image
-Additional interior constraints are used to ensure a continuous solution.
-This has the benefit of decreasing the numerical sensitivity of the final states to changes in the initial conditions along each sub-arc.
+In~\cref{fig:multiple shooting} we show a schematic representation of the multiple shooting procedure.
+We split the optimal control horizon into equal length subsegments such that the length of each segment is \( \frac{k}{n} \), where \( k, n \) are the total number of steps and number of stages, respectively.
+Similarly, we divide the state and costate trajectories into \( n \) equal segments. 
+To ensure continuity, additional interior constraints are incorporated as
+\begin{subequations}
+\begin{align}\label{eq:interior_constraints}
+	\vecbf{x}_1^{-} - \vecbf{x}_1^{+} &= 0 , \\ 
+	\vecbf{\lambda}_1^{-} - \vecbf{\lambda}_1^{+} &= 0 , \\
+	\vecbf{x}_2^{-} - \vecbf{x}_2^{+} &= 0 , \\ 
+	\vecbf{\lambda}_2^{-} - \vecbf{\lambda}_2^{+} &= 0 , \\
+	&\vdots \\
+	\vecbf{x}_{n-1}^{-} - \vecbf{x}_{n-1}^{+} &= 0 , \\ 
+	\vecbf{\lambda}_{n-1}^{-} - \vecbf{\lambda}_{n-1}^{+} &= 0.
+\end{align}
+\end{subequations}
+Using the multiple shooting method reduces the sensitivity of the terminal states, \( \vecbf{x}_i^{+}, \vecbf{\lambda}_i^{+}\), to variations of the initial states, \(\vecbf{x}_{i-1}^{-}, \vecbf{\lambda}_{i-1}^{-}\).
+The constraints to be satisfied are defined by~\cref{eq:interior_constraints,eq:necc_cond} through selection of \( \vecbf{\lambda}_0, \vecbf{\beta}, \vecbf{x}_i^{+}, \text{ and } \vecbf{\lambda}_i^{+} \).
+
 % tpbvp is solved using a shooting method (fsolve in matlab)

### 2017-01-18 manuscript.tex
Add more detail about multiple shooting
Issue #4

@@
 In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
+Consider the vector of initial conditions, \( \vecbf{\chi} = \braces{\vecbf{x}_0, \vecbf{\lambda}_0}\), which is varied to satisfy some terminal constraints of the form \( \vecbf{G}(\vecbf{\chi}) = \braces{\vecbf{x}_t - \vecbf{x}_n} = 0 \).
+The free variables at the terminal time are computed by propogation of \( \vecbf{\chi} \) over the selected time horizon. 
+At the terminal time, the constraint vector is calculated and if not satisfied \( \vecbf{\chi}\) is varied.
 Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs~\cite{stoer2013}.
@@
 
-In~\cref{fig:multiple shooting} we show a schematic representation of the multiple shooting procedure.
+In~\cref{fig:multiple shooting}, we show a schematic representation of the multiple shooting procedure.
 We split the optimal control horizon into equal length subsegments such that the length of each segment is \( \frac{k}{n} \), where \( k, n \) are the total number of steps and number of stages, respectively.
@@
 Using the multiple shooting method reduces the sensitivity of the terminal states, \( \vecbf{x}_i^{+}, \vecbf{\lambda}_i^{+}\), to variations of the initial states, \(\vecbf{x}_{i-1}^{-}, \vecbf{\lambda}_{i-1}^{-}\).
-The constraints to be satisfied are defined by~\cref{eq:interior_constraints,eq:necc_cond} through selection of \( \vecbf{\lambda}_0, \vecbf{\beta}, \vecbf{x}_i^{+}, \text{ and } \vecbf{\lambda}_i^{+} \).
+As a result, the design vector \( \vecbf{\chi} \) is augmented with the additional interior initial conditions, \( \vecbf{x}_i^{-}, \vecbf{\lambda}^{-}\).
+Similarly, the constraint vector is augmented with the additional interior constraints defined in~\cref{eq:interior_constraints}.
+The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
+In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
 

### 2017-01-18 manuscript.tex
Add more detail to the solution of the first example transfer
Issue #4

@@
 From the initial state on the periodic orbit, a series of optimal trajectories are generated to determine the reachable set.
+The multiple shooting approach is implemented to sovle the optimal control problem. 
+We divide the time horizon into two equal length segments. 
+The state trajectory is intialized using the free trajectory of the periodic orbit. 
+Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propogated using the discrete equations of motion in~\cref{eq:costate_update}.
+This results in the initial guess of the design vector \( \vecbf{\chi} = \begin{bmatrix} \vecbf{\lambda}_0 & \vecbf{x}_1^{-} & \vecbf{\lambda}_1^{-} & \vecbf{\beta} \end{bmatrix}^T\).
+This design vector is then varied to ensure that the necessary conditions of optimality and the interior point constraints are satisfied.
+
 By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
 We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
+Choosing a new angle \( \theta_d \) corresponds to a different direction as well as a new optimal control problem which is again solved using the multiple shooting approach laid out previously.
 The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:invariant_manifold_transfer}.

### 2017-01-18 manuscript.tex

@@
 The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
+Next, we compute the reachable sets originating on the geo-stationary orbit and subject to the maximum control constraint. 
+The multiple shooting approach is used to solve the optimal control problem with two segments.
+The initial geo-stationary orbit is used to initialize the interior point \( \vecbf{x}_i^{-} \).
+We use the same initial guess of \( \vecbf{\lambda}_0\) as in~\cref{sec:periodic_orbit_transfer} and propogate over the fixed time horizon of the period of the geostationary orbit to initialize \( \vecbf{\lambda}_i^{-}\).
+
 The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_{max} = 0.75 \).
@@
 This allows us to approximate the set of states that are achievable in the \( \parenth{x ,\dot{x}} \) space.
-After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds.
+After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds and is given as
+\begin{align*}
+	d(\vecbf{x}(N)) = \sqrt{\parenth{x - x_t}^2 + \parenth{\dot{x} - \dot{x}_t}^2}.
+\end{align*}
+From the reachable set, the trajectory which minimizes \( d \) is used to intialize the next reachability computation, or it will intersect the stable manifold region on the \Poincare section.
 %A state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
 For example,~\cref{fig:third_reachability_set} shows the reachable set after three iterations in red and the minimum distance to the stable manifold in green.
-
-With each reachable set we move the controlled trajectory closer to the stable manifold.
+The state and costate trajectories associated with this point on the rechable set is used to initialize the following iteration. 
+In this manner, the multiple shooting algorithm solves the optimal control problem and generates reachable sets which move away from the intial orbit and towards the desired target.
+
+With each reachable set we move the controlled trajectory closer to the target stable manifold.
 \Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 

### 2017-01-19 manuscript.tex

@@
 Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional Earth-Moon three-body system. 
+The geostationary orbit about the Earth is transformed into the rotating reference frame of the Earth-Moon three body problem.
+In addition, we nondimensionalize the initial state to find \( \vecbf{x}_0 = \begin{bmatrix} 0.0972 & 0 & 0 & 3.0010\end{bmatrix}^T\) as the initial condition of the spacecraft on the geostationary orbit.
 It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
+The periodic orbit is defined by the initial condition \(\begin{bmatrix} 0.8057 & 0 & 0 & 0.2982 \end{bmatrix} \).
 The initial and target orbits are illustrated in~\cref{fig:geo_transfer_target}.
@@
 \end{figure}
-We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
+Next, we generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
 The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
@@
 The multiple shooting approach is used to solve the optimal control problem with two segments.
-The initial geo-stationary orbit is used to initialize the interior point \( \vecbf{x}_i^{-} \).
+For the first reachability set computation, the initial geostationary orbit is used to initialize the multiple shooting algorithm.
 We use the same initial guess of \( \vecbf{\lambda}_0\) as in~\cref{sec:periodic_orbit_transfer} and propogate over the fixed time horizon of the period of the geostationary orbit to initialize \( \vecbf{\lambda}_i^{-}\).
@@
 This allows us to approximate the set of states that are achievable in the \( \parenth{x ,\dot{x}} \) space.
-After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds and is given as
+After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds as the function \( d \)
 \begin{align*}
@@
 \end{align*}
-From the reachable set, the trajectory which minimizes \( d \) is used to intialize the next reachability computation, or it will intersect the stable manifold region on the \Poincare section.
+From the reachable set, the trajectory which minimizes \( d \) is used to intialize the next iteration.
+Once the reachability set intersects the stable manifold, we can achieve a complete transfer from the initial geostationary orbit to the stable manifold of \( L_1 \).
 %A state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.

### 2017-01-24 manuscript.tex
Fix typos in tikz plots

--- 2017-01-18 manuscript.tex
@@
     \coordinate (center) at (0,0,2);
-    \node[below right] (x0) at (0,2,2) {\(\vecbf{x}_0\)};
+    \node[below right] (x0) at (0,2,2) {\(\vecbf{q}_0\)};
     \filldraw (0,2,2) circle (3pt);
 
-    \node[below right] (x1) at (0,3,2) {\(\vecbf{x}_1\)};
+    \node[below right] (x1) at (0,3,2) {\(\vecbf{q}_1\)};
     \filldraw (0,3,2) circle (3pt);
@@
         \coordinate (center) at (0,0,2);
-        \node[label=below:\(\vecbf{x}_0\)] (x0) at (0,3,2) {};
+        \node[label=below:\(\vecbf{x}_n\)] (x0) at (0,3,2) {};
         % \node[label=below:\(\vecbf{x}_n\)] at (x0) {};
@@
 In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
-Consider the vector of initial conditions, \( \vecbf{\chi} = \braces{\vecbf{x}_0, \vecbf{\lambda}_0}\), which is varied to satisfy some terminal constraints of the form \( \vecbf{G}(\vecbf{\chi}) = \braces{\vecbf{x}_t - \vecbf{x}_n} = 0 \).
-The free variables at the terminal time are computed by propogation of \( \vecbf{\chi} \) over the selected time horizon. 
-At the terminal time, the constraint vector is calculated and if not satisfied \( \vecbf{\chi}\) is varied.
 Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs~\cite{stoer2013}.
@@
 
-In~\cref{fig:multiple shooting}, we show a schematic representation of the multiple shooting procedure.
+In~\cref{fig:multiple shooting} we show a schematic representation of the multiple shooting procedure.
 We split the optimal control horizon into equal length subsegments such that the length of each segment is \( \frac{k}{n} \), where \( k, n \) are the total number of steps and number of stages, respectively.
@@
 Using the multiple shooting method reduces the sensitivity of the terminal states, \( \vecbf{x}_i^{+}, \vecbf{\lambda}_i^{+}\), to variations of the initial states, \(\vecbf{x}_{i-1}^{-}, \vecbf{\lambda}_{i-1}^{-}\).
-As a result, the design vector \( \vecbf{\chi} \) is augmented with the additional interior initial conditions, \( \vecbf{x}_i^{-}, \vecbf{\lambda}^{-}\).
-Similarly, the constraint vector is augmented with the additional interior constraints defined in~\cref{eq:interior_constraints}.
-The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
-In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
+The constraints to be satisfied are defined by~\cref{eq:interior_constraints,eq:necc_cond} through selection of \( \vecbf{\lambda}_0, \vecbf{\beta}, \vecbf{x}_i^{+}, \text{ and } \vecbf{\lambda}_i^{+} \).
 
@@
 From the initial state on the periodic orbit, a series of optimal trajectories are generated to determine the reachable set.
-The multiple shooting approach is implemented to sovle the optimal control problem. 
-We divide the time horizon into two equal length segments. 
-The state trajectory is intialized using the free trajectory of the periodic orbit. 
-Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propogated using the discrete equations of motion in~\cref{eq:costate_update}.
-This results in the initial guess of the design vector \( \vecbf{\chi} = \begin{bmatrix} \vecbf{\lambda}_0 & \vecbf{x}_1^{-} & \vecbf{\lambda}_1^{-} & \vecbf{\beta} \end{bmatrix}^T\).
-This design vector is then varied to ensure that the necessary conditions of optimality and the interior point constraints are satisfied.
-
 By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
 We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
-Choosing a new angle \( \theta_d \) corresponds to a different direction as well as a new optimal control problem which is again solved using the multiple shooting approach laid out previously.
 The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:invariant_manifold_transfer}.

### 2017-06-26 Response.tex
Start work on responding to comments

\documentclass[11pt]{article}
\usepackage{jas_packages}
\usepackage{fix-cm}
\usepackage{jas_tikz_packages}
\input{macros}
\usepackage[letterpaper,margin=1in,centering]{geometry}
\usepackage{graphicx}
\usepackage{amssymb}  
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{siunitx}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage{color}
\usepackage{bm}
\usepackage{xr-hyper}
% \usepackage{nameref,zref-xr}
\usepackage{hyperref}
\usepackage[]{cleveref} % get fancy referencing

% \zexternaldocument{manuscript}
\externaldocument[]{manuscript}[manuscript.pdf]

% edit cref citations for IEEE format
\crefformat{equation}{(#2#1#3)} % no abbreviation for equation numbers
\Crefformat{equation}{Equation~(#2#1#3)} % no abbreviation for equation numbers
\crefrangeformat{equation}{(#3#1#4--#5#2#6)}
\Crefrangeformat{equation}{(#3#1#4--#5#2#6)}
\crefmultiformat{equation}{(#2#1#3)}{ and~(#2#1#3)}{, (#2#1#3)}{ and~(#2#1#3)}
% edit cref citations for IEEE format
\crefformat{figure}{Fig.~#2#1#3} % no abbreviation for equation numbers
\Crefformat{figure}{Fig.~#2#1#3} % no abbreviation for equation numbers
\crefrangeformat{figure}{Figs.~#3#1#4--#5#2#6}
\Crefrangeformat{figure}{Figs.~#3#1#4--#5#2#6}
\crefmultiformat{figure}{Figs.~#2#1#3}{ and~#2#1#3}{, #2#1#3}{ and~#2#1#3}
% for items in a list
\crefformat{enumi}{(#2#1#3)}
\Crefformat{enumi}{Item~(#2#1#3)} 
\crefrangeformat{enumi}{(#3#1#4--#5#2#6)}
\Crefrangeformat{enumi}{Items~(#3#1#4--#5#2#6)}
\crefmultiformat{enumi}{(#2#1#3)}{ and~(#2#1#3)}{, (#2#1#3)}{ and~(#2#1#3)}
% Proposition format
\crefformat{prop}{Proposition~#2#1#3} % no abbreviation for equation numbers
\Crefformat{prop}{Proposition~#2#1#3} % no abbreviation for equation numbers
\crefrangeformat{prop}{Proposition~#3#1#4--#5#2#6}
\Crefrangeformat{prop}{Propositions~#3#1#4--#5#2#6}
\crefmultiformat{prop}{Propositions~#2#1#3}{ and~#2#1#3}{, #2#1#3}{ and~#2#1#3}
% Appendix
\crefformat{appendix}{Appendix~#2#1#3} % no abbreviation for equation numbers
\Crefformat{appendix}{Appendix~#2#1#3} % no abbreviation for equation numbers
\crefrangeformat{appendix}{Appendices~#3#1#4--#5#2#6}
\Crefrangeformat{appendix}{Appendices~#3#1#4--#5#2#6}
\crefmultiformat{appendix}{Appendices~#2#1#3}{ and~#2#1#3}{, #2#1#3}{ and~#2#1#3}

\newcommand{\RNum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}
\newcommand{\RI}{\text{\RNum{1}}}
\newcommand{\RII}{\text{\RNum{2}}}
\newcommand{\RIII}{\text{\RNum{3}}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{cor}{Corollary}

\newenvironment{correction}{\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\rightmargin}{1cm}}\vspace{\parsep}\item[]``}{''\end{list}}

\begin{document}

%\pagestyle{empty}

\section*{Response to the Reviewers' Comments for JASS-D-17-00005}

I would like to thank the reviewers for their thoughtful comments, which are aimed towards improving the quality of the manuscript and the clarity of the results. 
In accordance with the comments and suggestions, the manuscript has been revised, and the answers to all comments are addressed as follows.

(In the revised manuscript, the citation numbers for equations, assumptions, propositions, and references are changed. This answer is written according to the new item numbers.)

\subsection*{Reviewer 1} 
\begin{enumerate}
    \item \textit{Text of comment}

        \begin{correction}

            My correction

        \end{correction}
\end{enumerate}
Reviewer \#1: I have some issues with this paper that, in my opinion necessitate a major revision.

These issues are:

-It extensively re-covers issues that have been covered in many papers, making it difficult to disseminate what the original contributions are.  For example, I don't think the addition of the thrust term into the dynamics necessitates re-doing every piece of analysis that has been done with various resources in the literature about variational integrators (including comparisons with RK methods).  Things that can be found in other research should be cited, and summary comments can be used to point the readers to fill in this knowledge.  The content of this paper easily be reduced considerably to maintain focus on the original contributions.

-It does not give a complete picture of the pros and cons of the proposed methodology.  This method is full of both advantages and disadvantages.  I don't think the authors need to be concerned with "selling" the work to the community, but rather showing the full and honest picture given the current state of the research.  In the conclusions, the potential cons to this research can be mitigated by setting a path forward for future research to further improve the methodology.

These pros/cons include:

-The formulation of the problem with acceleration magnitude is a major technical challenge for mission design practitioners.  The state of the art to design trajectories that depend on low-thrust engine hardware is either:
-To produce constant thrust magnitude or no thrust magnitude (typical of nuclear electric propulsion, or very simple electric propulsion systems.)
-To produce variable thrust magnitude associated with various throttle points at different power accessibility levels (solar electric propulsion).
That is not to say that useful solutions cannot be extracted from the problem, as the authors have shown.  However, a great many solutions in this approach will be unusable due to the fact that acceleration and thrust magnitude are different, i.e. a=T/m
Thus - a MAJOR focus of future work should be the introduction of a seventh state (mass).  This gives rise to a more challenging optimal control problem requiring the switching function, but it is the form that is generally needed in computing optimal open-loop trajectories that can actually be flown.

The pros and cons of the PCRTBP
-The PCRTB gives rise to exploiting Poincare sections in a systematic manner.  2D solutions might be close enough for qualitative understanding of the solution space for many problems (as long as large plane changes aren't needed).
-The PCRTB is limiting for many pragmatic departure orbits (e.g. LEO and GTO, the most common departure orbits for low-thrust transfers).

The pros and cons of variational integrators should really be acknowledged rather than completely dismissing the RK methods which are still the current state of the art in mission design (using higher-orders and suitable tolerances)
-Pros = boundedness in the energy behavior.  Very long term integration will not deviate from the known dynamical structure.
-Cons = you don't know where you are on the energy surface, just that you are on it somewhere.  This is still a problem in actual trajectory design.  Knowing the error is on an energy surface does not mean that you don't have error.

From here on out comments specific to various sections of the paper follow:

In the statement "In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential
missions at a reduced energy cost [15]" - you may want to reference, for example, the Lunar Ice Cube mission paper which absolutely requires low-thrust propulsion and non-keplerian dynamics to satisfy its mission objectives.

Folta, D., Dichmann, D., Clark, P., Haapala, A., and Howell, K., "LunarCube Transfer Trajectory Options," AAS/AIAA Space Flight Mechanics Meeting, Williamsburg, Virginia, January 11-15, 2015.

"Since there is an insufficient number of analytical constants, or integral of motion, numerical methods must be used to investigate solutions to the three-body problem."  This isn't completely true as there are also approximations to the (uncontrolled) three-body motion, such such as the Richardson-Cary approximation.  Saying numerical methods are required may be a bit strong.  May want to say "often required".

"These techniques suffer from numerical instability and energy drift behaviors which make them ill-suited for long-term propagation. These dissipative effects are even more detrimental with the addition of low-thrust propulsion to the dynamic equations of motion."  It really depends on what you mean by "long term".  If transfers are on the order of hundreds of days, then high precision runge-kutte integrator, for example a Runge-Kutte Verner 8(9), can maintain suitable accuracy such that the energy drift in the Hamiltonian is very small, as can be seen in some of the references that the authors have cited.  If, on the other hand you mean 10-100's of years or more, i.e., when you create your Poincare maps, then this drift is more apparent.  I think that this statement needs to be quantified better because it gives a dishonest portrayal of high-precision Runge-Kutte integrators without numbers attached to it.

"Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system [8]. As a result, the long term effects of low-thrust on the spacecraft trajectory are not accurately captured."  Again, I think situationally I would disagree.  Transfers on the order of 100's of days can maintain accuracies of 1e-10 (nondimensional units) in the Hamiltonian, which is generally considered a "small" amount of energy drift.

- "[15,25] have illustrated the rich structure that exists in the three-body problem." 
- "[23,6] implement the solutions using conventional Runge-Kutta integration techniques."
I don't think you can grammatically begin a sentence with reference numbers.  Please fix all instances of this.

- "It has been shown that there exist multi-dimensional tubes, or invariant manifolds, of constant energy trajectories that span the state space."  The "tube" statement is really only true for some libration point orbits at certain Jacobi Constants.  For some orbits these structures are more chaotic and don't resemble tubes.  May want to soften this language or change "tubes" to "structures".

-"In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem." Why planar?  For simplicity?  Almost all parking orbits departing the Earth are likely going to have some inclination with respect to the rotating reference plane, and will likely require some three-dimensional thrusting to target the manifold.  You may want to indicate that this is a first step in the research to exploit the Poincare sections more easily, which is why I'm assuming it was done.  It doesn't hurt to be straightforward about this, as it is a first-step towards characterizing the design space.

-"In addition, through the use of geometric integrators we accurately capture the effects of low thrust on the system dynamics in the numerical simulation."  A variational integrator will accurately capture the qualitative effects of the transfer (e.g. propellant use), however it will not necessarily accurately capture accuracy in the final state vector - it is guaranteed to be on an energy surface, but it is not guaranteed to be at exactly the right location on the energy surface.  I would soften this claim or reword.

-"The Poincare section reduces the dimensionality of the system dyanmics to the study of a related discrete update map." It does, but at the expense if a large up-front computational run that must be run well in advance to capture the problem characteristics.  Also, it is more challenging to execute in 3D, which should be noted.

-"This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs."  Yes, but the order of the variational integrator still matters.  The drift in the state vector accuracy is still important to a mission designer, even if it is guaranteed to lie on an energy surface.  I think you should indicate that this is more for preliminary analysis to characterize the transfers, when this drift in state vector accuracy is tolerable.  I.E., it's useful for the introductory phases of mission design when trying to globally characterize the design space.  Characterizing the scope of the analysis in this way would be helpful to the audience.

"Optimal solutions are generally sensitive to small variations in the initial multipliers. As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance."  I don't agree that this implies that variational integrators are better for gradient-based methods.  The primary issue is that gradient-based methods that are solving optimal control problems are using first-order approximations to iterate on the problem.  The first-order approximation is what is poor, not necessarily the gradients.  Even if these gradients are of exact accuracy (and I agree that they should always be as accurate as possible), the first-order methods in the chaotic problem are what give the difficulties.  One could use a variational or standard integrator in a first-order gradient based method and nothing will change the fact that the method will be highly sensitive, save for partitioning the sensitivities as you did with your multiple shooting approach.  At best, you can change the statement to indicate that the accuracy in the gradients may improve, which is always helpful to gradient-based methods, but it's no guarantee that it eliminates major problem sensitivity.

Authors may also want to point out the loss of accuracy in variable step integrators that can also degrade first-order gradient based methods.  I didn't see this anywhere and it's something that should be acknowledged.  Basically, variable step integrators will introduce artificial "noise" into the gradients:

Pellegrini, E.P., Russell, R.P., On the Accuracy of Trajectory State Transition Matrices, Paper AAS-15-785, AAS/AIAA Astrodynamics Specialist Conference, Vail, CO, Aug 2015

"This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods." I disagree that suboptimality is typical of direct methods.  A direct method will typically solve the KKT conditions, which can be directly transcribed into the costates in the optimal control problem.  A direct method fully satisfying the KKT conditions can map to the indirect problem. If you want to strengthen this argument, then you should argue that it applies to direct methods which rely on coarse or suboptimal representations of the control.  For example, a hybrid direct method could use continuous optimal control theory control law, but directly minimize the cost function instead of solving the full TPBVP.

"Application of the Euler-Lagrange equations results in following equations of motion defined in the rotating reference frame"  Add a reference after this statement for those who want to see the derivation.  One of the given references probably does already does this, so just pick one

It appears from equation 5 and your control input definition that you control is unbounded?  Later on in reading the paper I found that you do add bounds on the control.  You may want to indicate earlier on in the paper (closer to eq. 5) that bounds will be added to the control.

Figure 1 should probably have "axis equal" applied to it.  For that matter, please check all trajectory plots (those which plot the position state), and ensure that all plots are using "axis equal".  If not, please regenerate those plots.

Section 3.1 has been discussed in many references, the reader may want to direct them to this material.
(Choose one of Marsden or West's papers for example.)

The second-order trapezoidal rule will give significantly poorer state vector accuracy with respect to a high-precision RK integrator, even if the state is guaranteed to lie on an energy surface.  Why didn't the authors consider going to a higher-order variational integrator, or even a Taylor Integrator?  It is very likely that the integration accuracy will introduce some errors in the accuracy of the partial derivatives and in the states.

Section 3.2 - please cite a reference as to where the selected quadrature rules presented in Table 1 have been derived.

Figure 2 - x-axis has no label but I think it's nondimensional time, please indicate.  Label all axes even if they are nondimensional units.

I recommend removing appendix B.  Giving the equations and stating that the Matrix inversion is from Gauss-Jordan elimination is probably sufficient, unless there is some compelling algebraic manipulation that is very unique to the problem.

In line 14 (page 12), you define the state transition matrix over one orbital period, and then on line 18 you start referring to it as the monodromy matrix.  I understand what you're doing, but you may want to make the definition more explicit.  Some of this section can be substantially reduced and other source material could be referenced anyway.

Comparison like that of Figure 3.4 are very common in the well-known literature concerning variational integrators.  Moreover, the proper way to present such a comparison is with different delta-T's in the integration step, showing when the Runge-Kutta methods actually break down in terms of maintaining an accurate portrayal of the Energy/Hamiltonian conservation.

For example -

Figure 2.4
Lew, A., Marsden, J., Ortiz, M., West, M., "An Overview of Variational Integrators," FINITE ELEMENT METHODS: 1970's AND BEYOND L.P. Franca (Ed.) c CIMNE, Barcelona, Spain 2003.

Online copy:
http://lagrange.mechse.illinois.edu/pubs/LeMaOrWe2004/LeMaOrWe2004.pdf

It may also be generally worth acknowledging the benefits of Taylor's method within the context of variational integrators:
Schmitt, J., Shingel, T., Leok, M., "LAGRANGIAN AND HAMILTONIAN TAYLOR VARIATIONAL INTEGRATORS," eprint arXiv:1703.06599, 2017

Online copy: https://arxiv.org/pdf/1703.06599.pdf

"Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques".  I think you need to add "in which energy conservation is mandatory for characterizing the solution space."

Section 5 - This method, while very useful in its given context is highly dependent on the use of the PCRTB in the framework of the current analysis.  This needs to somehow be conveyed.  3D transfers are significantly more challenging when it comes to exploiting Poincare sections.  It can be done, but more subplots, dimensions, and data storing are necessary to view the subspace - save to say, it is a more challenging endeavor that must be acknowledged for general purpose low-thrust mission design.

Figures 10D and 13D - the authors need to present these plots in dimensional coordinates so that they may be reconciled with what existing technology can produce.  I recommend Newtons or milliNewtons.  For example, the NEXT (NASA evolutionary xenon thruster) can produce 0.25 Newtons of thrust at maximum power (~7 kW).  The XR5 Hall thruster can produce comparable thrust at 4.5 kW of power.  These are currently the highest-power single engines other than the ARM Hall thruster which is not flight ready yet.  For these examples to be reconciled against the capability of existing technology, we need to see the required forces that are being demanded from the optimal control problem.  It may be that the results are beyond the scope of current technology - that's fine - just say that they are examples and that it's ongoing work to formulate example problems that capture existing technology better.

"The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches."  Did the authors attempt single shooting for their example problems before coming to this conclusion?  With proper scaling, and problem formulation, there are examples in the literature of single shooting yielding robust results.

Example: 
Ranieri, C. and Ocampo,C..  "Indirect Optimization of Spiral Trajectories", Journal of Guidance, Control, and Dynamics, Vol. 29, No. 6 (2006), pp. 1360-1366.

So, justification for its selection may be warranted.  You could always find a reference in which this trade space was really explored thoroughly to show why you made the choice.

"In this work, we use the Matlab nonlinear solver fsolve to solve the system of nonlinear equations defined by the multiple shooting algorithm."  Can authors give some additional details?  Which fsolve algorithm? trust region dogleg,  trust region, or Levenberg-Marquardt?  Which additional options?  E.g. various tolerances.  For users of this method it's also useful to see the iteration history (more comments on this later).

Did the authors experiment with the number of multiple shooting stages in their example problems?  There must be some sort of "sweet spot" in number of stages for the example problems attempted.  If I missed it, I apologize, but I didn't see it.

Section 6.1 - I don't think you need to explain how you created the three-body orbits.  This can be found in many references - You can just cite a reference and give the orbit specifications.  (Starting with line 49, page 18, up to line 21 page 19.)

"We define a maximum magnitude of the thrust as umax = 0.75 and assume we can point the thrust in any direction within the plane." And  "we again assume a upper bound on the thrust magnitude of umax = 0.75."  What was the reason behind this? Isn't this a pretty large value of dimensional acceleration for your problem of interest?  Even if it's chosen for demonstrative purposes, there should be some sort of explanation as to why.  In the context of what electric propulsion systems can deliver, it could be arguable that this value isn't really "low-thrust" but "finite-thrust".

Is this method exceptionally difficult to adapt to create a numerical example from LEO or GTO?  These would be the most pragmatically interesting departure orbits to the mission design community.  Section 6.2 presents an example from GEO, but it is unlikely that GEO represents a "typical" parking orbit, since there is no reason to insert a spacecraft into GEO before it transfers to the destination orbit.  This should be called out in the conclusions or future work.

The authors should discuss some statistics of the convergence of their multiple shooting TPBVP - number of iterations, compute time, etc.

Can any intuition be gained from reporting the optimal converged cost function values in section 6.1 and 6.2?  One could at least compare them to compare the reachability in the two different examples.

"Low energy transfers from the Earth to the Lagrange points are necessary for future missions."  One must be careful not to confuse these with traditional "low-energy transfers" to the moon, which use Sun third-body perturbations to transfer into lunar orbit.  These are what many mission designers define as "low-energy" lunar transfers.  This is actually a four-body problem (Sun, Earth, moon, spacecraft), and represents additional delta-V savings (as was done with NASA's GRAIL mission.)

Example:

Parker, J. and Anderson, R., "Low-Energy Lunar Trajectory Design," DEEP SPACE COMMUNICATIONS AND NAVIGATION SERIES, Jet Propulsion Laboratory Pasadena, California, July 2013.

Online Copy: https://descanso.jpl.nasa.gov/monograph/series12/LunarTraj--Overall.pdf

For what it's worth, exploiting sun perturbations with low-thrust in the Earth-moon RTBP would be highly beneficial and represent a worthy avenue of future research (more on this below.)

The conclusions of the paper should acknowledge other potentially limiting factors of the presented analysis
-Omits advantage of exploiting fourth body perturbation (Sun in Earth/moon RTBP)
-Transfers are 2D only, method is more challenging to adapt to 3D
-Presented control magnitudes are dimensionally large, and more work is needed to determine feasibility of approach for lower accelerations
-Thrust and mass are completely decoupled from the analysis, which is limiting in terms of finding suitable hardware that can deliver the controls needed, as currently presented.

\section*{Reviewer 2}

Reviewer \#2: PAPER SUMMARY:
The contribution of the authors is the synthesis of reachability theory with traditional low-energy mission design approaches to ease the study of the low-thrust mission design space in the CR3BP.  A crucial point is that reachability analysis in non-integrable systems is typically limited by the dimensionality of the control parameterization, which can be extremely large for continuous-thrust problems, depending on the time discretization used.  The authors numerically probe the boundaries of the reachable space by optimizing a distance metric on the Poincare section, avoiding the extreme cost of unmotivated exploration.  Numerical experiments show how the method could be applied in two CR3BP orbit transfer scenarios.  For a multi-rev solution, the method is applied once per rev and the reachable end state with minimal distance to the target manifold is used as the basis for the next rev.

ASSESSMENT:
The central premise of the paper - using optimization of Poincare distance as a guiding motivation to make low-thrust reachability analysis more tractable - is interesting, useful, and to my knowledge reasonably novel.  Furthermore, the quality of the writing is good (at local scales) and there is an appropriate degree of literature review.  However, there are two major problems with this paper.

First, in terms of composition, large portions of the paper are spent on lengthy development of foundational work that is well-known within the astrodynamics community and available in textbooks. Secondly, the contributions and findings of the authors feel highly preliminary and do not seem to evaluate their method at significant depth.

While there are the beginnings of a good article within this manuscript, it is not sufficiently close to publishable form to be accepted.  I recommend that the paper be rejected and that the authors push their investigation a bit further.  The eventual publication should focus less on background and should provide more insight into the fundamental properties of their method, using more rigorously constructed and analyzed examples.

I can sympathize that there was quite a long road to get to the point of understanding this topic and generating results in it. I hope that the authors will not find this verdict too discouraging, as I believe that stronger results are well within their reach.

SPECIFIC FEEDBACK:

*** Bloat
In the introduction, the exposition of the problem could be considerably condensed, given the venue of publication. Additionally, the somewhat sprawling discussion jumps back and forth, and many key points seem to be repeated more than necessary.  It needs to be organized into titled subsections with specific scope.

The development of the CR3BP, invariant manifolds and Poincare maps, all go on a bit too long. These should be refreshers to the audience that expose the specific equations and concepts that you'll be leveraging, not full-on introductions to well-established areas.  Spending several pages on variational principles is way too much.

Section 4 ends on page 14 with essentially a thesis statement ("In this paper, we...").  Strive to push the boundary between background and new developments to a much earlier point in the paper; it shouldn't happen so late that you need to remind the audience what the paper is for.

A missing relevant reference: "On target for Venus — set oriented computation of energy efficient low thrust trajectories" (Dellnitz, Junge, Post, and Thiere, 2006)

*** Variational Integrators
Many of the specific claims regarding the use of a variational integrator are unsubstantiated and, I believe, incorrect. Why is energy drift "even more detrimental" with regards to "the long term effects of low-thrust"?  I would need to be convinced that a non-geometric integrator's energy drift relative to a constant energy baseline is significantly different from its drift relative to a gradually shifting baseline corresponding to low thrust. The specific claims about the numerical stability of derivatives also don't seem related to energy preservation.

Apart from these claims, I question the need for a variational integrator in the first place. The example in section 3.4 is not convincing because there is no apples-to-apples basis for the comparison; a variable-step integrator with unspecified tolerance is compared to a fixed-step integrator with an arbitrary step size.  Integrator comparisons should consider error in proportion to compute cost, which can be tuned by step size or tolerance.  Also, fixed step seems inappropriate for a path through diverse dynamic regimes, and it's not clear that the problem scope calls specifically for energy preservation (especially since the idea is to use low thrust to mitigate some time cost of exploiting the problem structure).

*** Figures
Altogether, the figures need more descriptive captions.

Figure 4 does not show what I think you intend it to show; it looks like two periodic circular orbits.

Figure 6 is also very unclear.  Is phi supposed to be \texttt{theta\_d}?  Is the bounding curve specifically circular, or should it be a more general shape?

Consider plotting the intersections of manifolds with the Poincare section as continuous curves.

Also consider zooming/scaling as necessary to show the details you discuss.

*** Method/Examples
I urge a revisiting of the claims with regards to optimality.  The optimization method is being initialized with *something*, and stationary points are *local* extrema.  More precise claims and assurances about solution quality (and therefore the validity of the computed reachable set boundary) are needed.

How damaging is the assumption of a fixed travel time?  And how about the specification of a specific departure point on the initial orbit?

In Fig 10b, it is hard to see, but appears that the low-thrust-reachable set is actually *centered* on part of the target set.  Is this related to the existence of the uncontrolled transfer opportunity?  There must be some more clear way to show how the low thrust set expands upon the zero thrust set.  Perhaps beginning with the time horizon of the zero-thrust opportunity, and successively lowering it, would help.

Perhaps there is also a better way to show what the control is doing, e.g., in some alternative parametrization?  This could help provide assurances that the solution is doing something meaningful and is not just stuck in a weird local minimum.

In the second example, what do you lose by chaining together multiple reachability-based steps?  Aren't you imposing that the trajectory routinely returns to the energy level of the Poincare section? How feasible would it be to conduct a single reachability analysis covering multiple revolutions? And can you use knowledge about directionality of movement on the Poincare section to limit the scope of your reachability search?

In general, please devote more depth to probing the strengths, weaknesses, and fundamental properties of the method.  Considering extreme cases can help with this.

\bibliography{library}
\bibliographystyle{spmpsci}
\end{document}

### 2017-06-26 manuscript.tex
Working on refactoring the introduction
More clear/less verbose
Fix some comments from reviewers
Explicitly state our thesis up front

--- 2017-01-19 manuscript.tex
@@
 \section{Introduction}\label{sec:introduction}
-% MOTIVATION FOR LOW THRUST ORBITAL TRANSFER 
-Designing spacecraft trajectories is a classic and ongoing topic of research.
-There has been significant research into the design of orbital transfers for space vehicles.
-%With the current fiscal constraints there is an increased focus on technologies with a critical impact on mass.
-Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
-Electric propulsion systems offer a much greater specific impulse than chemical systems and are able to operate for extended periods of time.
-However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore must operate over longer durations in order to impart the desired momentum change.
-Recent developments in miniature electric propulsion offer the potential for new research opportunities for small spacecraft~\cite{haque2013}.
-With reduced development intervals and decreased launch costs, small satellites have gained increased attention as a cost effective means of scientific and technologic development. 
-The merger of small satellites with miniature electric propulsion enables inexpensive and responsive missions requiring large changes in orbital energy or extended mission lifespan.
-With the potential for more demanding missions, even greater importance is placed on the mission design to ensure that optimal trajectories satisfy mission requirements. 
-In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{koon2011}.
-Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional patched-conic approach~\cite{ross2006,gomez2001}.
-Optimally combining the dynamical structure of the three-body problem with low-thrust propulsion systems is vital for future mission success.
-
-%------------------ PRIOR WORK AND CHALLENGES -----------------------------------%
-% DIRECT OPTIMAL CONTROL
-There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}.
-Typically, the optimal control problem is solved via direct methods, which approximate the continuous problem as a parameter optimization problem.
-The state and/or control trajectories are parameterized and solved in the form of a nonlinear optimization problem.
-The work in~\cite{mingotti2011,grebow2011} use this direct approach in designing low-thrust transfers in the three-body problem.
-Alternatively, indirect methods apply calculus of variations to derive the necessary conditions for optimality. 
-This yields a lower dimensioned problem than the direct approach and algebraic conditions that, when satisfied, guarantee local optimality in contrast to direct methods which result in sub-optimal solutions.
-
-% CHAOTIC SYSTEM AND INITIAL GUESSES FOR OPTIMIZATION
-The application of optimal control methods for orbital trajectory design is nontrivial.
-In order to implement any optimization method a sufficiently accurate ``initial guess'' is required.
-Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the optimal solution.
-However, the three-body system dynamics are nonlinear and exhibit chaotic behaviors. 
-Popularly refered to as the ``butterfly effect'' small changes in initial conditions result in large variations of the resulting system trajectory. 
-This makes solving the optimization problem highly dependent on the initial condition.
-In addition, there is no closed form analytical solution for the three-body problem~\cite{szebehely1967}.
-Since there is an insufficient number of analytical constants, or integral of motion, numerical methods must be used to investigate solutions to the three-body problem.
-As a result, accurate numerical methods are required to determine optimal solutions.
-These methods are critically dependent on accurate initial guess in order to allow for convergence to an optimal solution.
-%For example, analytical design is possible for complicated trajectories using a Keplerian system model via the patched-conic approach.
-
-% NUMERICAL INTEGRATION
-Efficient numerical implementation is dependent on correct initial conditions as well as accurate numerical integration.
-\cite{mingotti2011,grebow2011} implement the solutions using conventional Runge-Kutta integration techniques.
-These techniques suffer from numerical instability and energy drift behaviors which make them ill-suited for long-term propagation.
-These dissipative effects are even more detrimental with the addition of low-thrust propulsion to the dynamic equations of motion.
-Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system~\cite{hairer2006}.
-As a result, the long term effects of low-thrust on the spacecraft trajectory are not accurately captured.
-
-% DYNAMICAL STRUCTURE
-\cite{koon2011,ross2006} have illustrated the rich structure that exists in the three-body problem.
-Within the three-body problem, a spacecraft's feasible region of motion is constrained by its energy, or Jacobi integral. 
-It has been shown that there exist multi-dimensional tubes, or invariant manifolds, of constant energy trajectories that span the state space. 
-Associated with periodic solutions of the three-body dynamics, these invariant manifolds allow for the spacecraft to traverse vast expanses of the state space with zero energy change. 
-However, the results presented are highly case specific and difficult to generalize to arbitrary transfers.
-Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system.
-In addition, transfer orbits along an invariant manifold require a longer time of flight which may be undesirable for time critical missions.
-The addition of low-thrust propulsion offers the potential of reduced transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities. 
-%------------------ PRIOR WORK AND CHALLENGES -----------------------------------%
-
-%-------------------- OBJECTIVES OF THIS WORK-----------------------------%
-% SYSTEMATIC DESIGN TO AVOID DIFFICULTIES IN CHOOSING A GOOD INITIAL GUESS
-% CAPTURE LONG TERM EFFECTS OF LOW THRUST ACCURATELY IN NUMERICAL SIMULATION
-In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem.
-Our approach avoids the difficulties in selecting an appropriate initial guess for optimization.
-Instead, we utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
-In addition, through the use of geometric integrators we accurately capture the effects of low thrust on the system dynamics in the numerical simulation. 
-This approach avoids the instability and dissipative effects of conventional integration schemes.
-With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low thrust propulsion systems.
-
-% HIGH LEVEL TECHNICAL DESCRIPTION OF APPROACH
-% Generalization of invariant manifold transfers  
-% Use reachability set to determine transfer opportunities with low thrust control input
-
-To achieve these objectives, an optimal control problem is formulated to determine the reachability set on a \Poincare section.
-Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft. 
-Generation of this reachability set allows for the extension of the previous methods based on invariant manifolds in the three-body problem.
-In addition, the generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer. 
-The \Poincare section reduces the dimensionality of the system dyanmics to the study of a related discrete update map.  
-This allows for the study of the complex dynamics of the three-body problem on a lower dimensional space.
-Rather than relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on an appropriate \Poincare section.
-This simple methodology allows for extended transfer trajectories which iteratively approach a target region.
-We utilize the low thrust control input to enlarge the reachability set.
-Maximization of the reachability set, on an appropriately chosen \Poincare section, allows for a greater space of potential transfer trajectories.
-The use of the \Poincare section allows for design on a lower dimensional space and simplifies the design process.
-%Once an intersection is determined on the \Poincare section a transfer is calculated.
-
-In order to address the issues associated with conventional numerical integration, we apply computational geometric optimal control techniques. 
-We generalize the previous work of determining orbital transfers using invariant manifolds~\cite{koon2011}.
-We incorporate the effects of continuous low thrust propulsion to enable departure from the system dynamics and enable increased transfer opportunities.
-The dynamics of the three-body system are derived from the discrete Lagrangian, which approximates the integral of the continuous time Lagrangian over a fixed discrete step.
-Application of the discrete Euler-Lagrange equations, or equivalently the discrete Legendre transform, results in the discrete equations of motion.
-This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs.
-A discrete optimal control problem is formulated from the discrete equations of motion.
-This approach, where explicit discretization occurs prior to optimization,  is in contrast to the typical method, where the equations of motion are implicitly discretized during the optimization procedure.
-Formulating the problem in this manner results in more stable and accurate optimal solutions. 
-In indirect methods the optimal control problem is expressed as a two-point boundary value problem.
-Optimal solutions are generally sensitive to small variations in the initial multipliers.
-As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance. 
-The use of geometric integrators, which do not suffer the numerical dissipation of conventional integration methods, results in a more robust and efficient solution.
-
-%---------------------------- EXPLICIT SUMMARY OF CONTRIBUTIONS--------------------------%
-In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
-Previous results in the design of optimal transfers have relied on suboptimal direct optimization methods and conventional integration techniques.
-This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
-We derive and solve the Euler-Lagrange equations to generate the necessary conditions for an optimal solution.
-This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods.
-In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
-Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
-We avoid the issues inherent in selecting a valid initial condition for optimization.
-Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
-We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.
-
+
+% Motivation for low thrust orbital transfers
+
+% --------------Issues/Challenges of previous work
+
+% use of direct optimal control
+
+% chaotic systems and initial guesses for optimization
+
+% numerical integration accuracy
+
+% preserving dynamical structure
+
+% objectives of this work
+
+% description of our work - our approach
+
+% concise summary of contributions
 \section{Planar Circular Restricted Three Body Problem}\label{sec:pcrtbp}

### 2017-06-26 Response.tex
Add additional reference to cubesats which use non-keplerian orbits
Reviewer 1 Specific comment 1

@@
 
+\begin{enumerate}
+    \item
+        \begin{itshape}
             In the statement "In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential
@@
             Folta, D., Dichmann, D., Clark, P., Haapala, A., and Howell, K., "LunarCube Transfer Trajectory Options," AAS/AIAA Space Flight Mechanics Meeting, Williamsburg, Virginia, January 11-15, 2015.
+        \end{itshape}
 
+        The introductory paragraphs have been reworded and reorganized for clarity. 
+        Specifically, the first paragraph has been divided, which motivates the use of electric propulsion and the advantages of small satellites which use electric propulsion.
+        In addition, the suggested reference has been included in the paper.
+        The modified paragraphs are duplicated below.
+
+        \begin{correction}
+            Designing spacecraft trajectories is a classic and ongoing topic of research.
+            There has been significant research into the design of orbital transfers for space vehicles.
+            Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
+            Electric propulsion systems offer a much greater specific impulse than chemical systems.
+            As a result, the greatly increased effeciency allows for greater payload mass or extended duration missions.
+            However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital manuevers have a much longer time of flight.
+            In spite of this drawback, a wide variety of missions, such as communication and deep space probes, have utilized the unique benefits of low thrust electric propulsion to great effect~\cite{choueiri2009}.
+
+            With reduced development intervals and decreased launch costs, small satellites have gained increased attention as a cost effective means of scientific and technologic development. 
+            Furthermore, these cubesats are easily launched as secondary payloads rather than requiring a dedicated launch.
+            The merger of small satellites with miniature electric propulsion enables inexpensive and responsive missions requiring large changes in orbital energy or extended mission lifespan.
+            Recent developments in miniature electric propulsion offer the potential for new research opportunities for small spacecraft~\cite{haque2013}.
+            The advancements in minature electric propulsion allows for greater flexibility in the deployment of cubesats as secondary payloads.
+            Without the explicit requirement of a dedicated launch trajectory, cubesats are able to exploit a much wider range of possible mission designs~\cite{folta2015}.
+            With the potential for more demanding missions, even greater importance is placed on the mission design to ensure that optimal trajectories satisfy mission requirements. 
+            In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{folta2015,koon2011}.
+            Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional patched-conic approach~\cite{ross2006,gomez2001}.
+            Optimally combining the structure of the dynamic environment with low-thrust propulsion systems is vital for future mission success.
+        \end{correction}
+\end{enumerate}
 "Since there is an insufficient number of analytical constants, or integral of motion, numerical methods must be used to investigate solutions to the three-body problem."  This isn't completely true as there are also approximations to the (uncontrolled) three-body motion, such such as the Richardson-Cary approximation.

### 2017-06-26 manuscript.tex
Add additional reference to cubesats which use non-keplerian orbits
Reviewer 1 Specific comment 1

@@
 
-% Motivation for low thrust orbital transfers
+%%-------------- Motivation for low thrust orbital transfers
+
+% enables long duration missions with minimized propellant usage since Isp is so high
+
+% Enables different mission scenarios which are not possible through the use of conventional chemical propulsion (high thrust) systems
+Designing spacecraft trajectories is a classic and ongoing topic of research.
+There has been significant research into the design of orbital transfers for space vehicles.
+Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
+Electric propulsion systems offer a much greater specific impulse than chemical systems.
+As a result, the greatly increased effeciency allows for greater payload mass or extended duration missions.
+However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital manuevers have a much longer time of flight.
+In spite of this drawback, a wide variety of missions, such as communication and deep space probes, have utilized the unique benefits of low thrust electric propulsion to great effect~\cite{choueiri2009}.
+
+% utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
+With reduced development intervals and decreased launch costs, small satellites have gained increased attention as a cost effective means of scientific and technologic development. 
+Furthermore, these cubesats are easily launched as secondary payloads rather than requiring a dedicated launch.
+The merger of small satellites with miniature electric propulsion enables inexpensive and responsive missions requiring large changes in orbital energy or extended mission lifespan.
+Recent developments in miniature electric propulsion offer the potential for new research opportunities for small spacecraft~\cite{haque2013}.
+The advancements in minature electric propulsion allows for greater flexibility in the deployment of cubesats as secondary payloads.
+Without the explicit requirement of a dedicated launch trajectory, cubesats are able to exploit a much wider range of possible mission designs~\cite{folta2015}.
+With the potential for more demanding missions, even greater importance is placed on the mission design to ensure that optimal trajectories satisfy mission requirements. 
+In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{folta2015,koon2011}.
+Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional patched-conic approach~\cite{ross2006,gomez2001}.
+Optimally combining the structure of the dynamic environment with low-thrust propulsion systems is vital for future mission success.
 

### 2017-06-26 manuscript.tex
Fix indentation to use spaces instead of tabs

--- 2017-01-24 manuscript.tex
@@
 \nonumber \\
-		& + \left. \frac{1}{2} \bracket{\left(  \frac{\xkp - \xk}{h} -\ykp \right)^2 + \left( \frac{\ykp - \yk}{h} + \xkp \right)^2} + \frac{1-\mu}{r_{1_{k+1}}} + \frac{\mu}{r_{2_{k+1}}}  \right) \, .
+                & + \left.
+\frac{1}{2} \bracket{\left(  \frac{\xkp - \xk}{h} -\ykp \right)^2 + \left( \frac{\ykp - \yk}{h} + \xkp \right)^2} + \frac{1-\mu}{r_{1_{k+1}}} + \frac{\mu}{r_{2_{k+1}}}  \right) \, .
 \end{align}
@@
 \end{figure}
-We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
+Next, we generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
 The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
+Next, we compute the reachable sets originating on the geo-stationary orbit and subject to the maximum control constraint. 
+The multiple shooting approach is used to solve the optimal control problem with two segments.
+For the first reachability set computation, the initial geostationary orbit is used to initialize the multiple shooting algorithm.
+We use the same initial guess of \( \vecbf{\lambda}_0\) as in~\cref{sec:periodic_orbit_transfer} and propogate over the fixed time horizon of the period of the geostationary orbit to initialize \( \vecbf{\lambda}_i^{-}\).
+
 The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_{max} = 0.75 \).
@@
 This allows us to approximate the set of states that are achievable in the \( \parenth{x ,\dot{x}} \) space.
-After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds.
+After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds as the function \( d \)
+\begin{align*}
+        d(\vecbf{x}(N)) = \sqrt{\parenth{x - x_t}^2 + \parenth{\dot{x} - \dot{x}_t}^2}.
+\end{align*}
+From the reachable set, the trajectory which minimizes \( d \) is used to intialize the next iteration.
+Once the reachability set intersects the stable manifold, we can achieve a complete transfer from the initial geostationary orbit to the stable manifold of \( L_1 \).
 %A state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
 For example,~\cref{fig:third_reachability_set} shows the reachable set after three iterations in red and the minimum distance to the stable manifold in green.
-
-With each reachable set we move the controlled trajectory closer to the stable manifold.
+The state and costate trajectories associated with this point on the rechable set is used to initialize the following iteration. 
+In this manner, the multiple shooting algorithm solves the optimal control problem and generates reachable sets which move away from the intial orbit and towards the desired target.
+
+With each reachable set we move the controlled trajectory closer to the target stable manifold.
 \Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 

### 2017-06-27 Response.tex
Refactor paragraphs into one. Also address Reviewer 1 comments
Reviewer 1 Comment 2 Comment 3

@@
         \end{correction}
-\end{enumerate}
+
+\item 
+    \begin{itshape}
 "Since there is an insufficient number of analytical constants, or integral of motion, numerical methods must be used to investigate solutions to the three-body problem."  This isn't completely true as there are also approximations to the (uncontrolled) three-body motion, such such as the Richardson-Cary approximation.
@@
 May want to say "often required".
+\end{itshape}
 
+The sentence has been modified as follows.
+
+\begin{correction}
+
+Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
+As a result numerical methods are often required to investigate solutions to the three body problem.
+\end{correction}
+
+\item 
+    \begin{itshape}
 "These techniques suffer from numerical instability and energy drift behaviors which make them ill-suited for long-term propagation.
@@
 Transfers on the order of 100's of days can maintain accuracies of 1e-10 (nondimensional units) in the Hamiltonian, which is generally considered a "small" amount of energy drift.
+\end{itshape}
 
+These sentences have been modified and combined to enhance clarity and reduce some of the unecessary bloat. 
+In addition, some additional detail has been added to explicitly define the notion of ``long-term'' in the context of numerical integration.
+The modifications are copied below.
+\begin{correction}
+    Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
+    As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g.
+on the order of decades to centuries, which is typical of the three-body problem.
+\end{correction}
 - "[15,25] have illustrated the rich structure that exists in the three-body problem." 
@@
 
+\end{enumerate}
 \section*{Reviewer 2}

### 2017-06-27 manuscript.tex
Refactor paragraphs into one. Also address Reviewer 1 comments
Reviewer 1 Comment 2 Comment 3

@@
 % use of direct optimal control
-
 % chaotic systems and initial guesses for optimization
-
 % numerical integration accuracy
-
 % preserving dynamical structure
+
+There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}. 
+Inspite of the long history in astrodynamics, a number of difficulties arise when implenting optimal control techniques.
+First, the optimal control problem is typically solved using direct methods, which approximate the continuous problem as a parameter optimization problem.
+Alternatively, indirect methods apply the calculus of variations to derive the necessary conditions for optimality.
+The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which result in sub-optimal solutions.
+Second, in order to implement any optimization method a sufficiently accurate ``initial guess'' is required.
+Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system.
+Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
+As a result numerical methods are often required to investigate solutions to the three body problem.
+Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
+As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g.
+on the order of decades to centuries, which is typical of the three-body problem.
 

### 2017-06-27 Response.tex
Remove references from beginning of sentence  and don't say tubes
Reviewer 1 Comment 5 Comment 6

@@
 \end{correction}
+
+\item 
+\begin{itshape}
 - "[15,25] have illustrated the rich structure that exists in the three-body problem." 
@@
 I don't think you can grammatically begin a sentence with reference numbers.  Please fix all instances of this.
+\end{itshape}
 
-- "It has been shown that there exist multi-dimensional tubes, or invariant manifolds, of constant energy trajectories that span the state space."  The "tube" statement is really only true for some libration point orbits at certain Jacobi Constants.
+The sentences have been reworded to avoid starting with a reference number. 
+The corrections are shown below.
+
+\begin{correction}
+    \begin{itemize}
+        \item Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
+        \item There exists a rich structure in the three-body problem~\cite{koon2011,ross2006}.
+    \end{itemize}
+\end{correction}
+
+\item 
+    \begin{itshape}
+"It has been shown that there exist multi-dimensional tubes, or invariant manifolds, of constant energy trajectories that span the state space."  The "tube" statement is really only true for some libration point orbits at certain Jacobi Constants.
 For some orbits these structures are more chaotic and don't resemble tubes.
 May want to soften this language or change "tubes" to "structures".
+\end{itshape}
+
+The sentence has been modified as follows.
+
+\begin{correction}
+It has been shown that there exist multi-dimensional structures, termed invariant manifolds, of constant energy trajectories that span the state space. 
+\end{correction}
 

### 2017-06-27 manuscript.tex
Remove references from beginning of sentence  and don't say tubes
Reviewer 1 Comment 5 Comment 6

@@
 on the order of decades to centuries, which is typical of the three-body problem.
+
+% control-free trajectories in teh three-body problem 
+There exists a rich structure in the three-body problem~\cite{koon2011,ross2006}.
+Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
+It has been shown that there exist multi-dimensional structures, termed invariant manifolds, of constant energy trajectories that span the state space. 
+These invariant manifolds, which are associated with the periodic solutions of the three-body dynamics, allow for the spacecraft to traverse vast expanses of the state space with zero energy change. 
+However, the results presented are highly case specific and difficult to generalize to arbitrary transfers.
+Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system. 
+The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from teh free motion trajectory to allow for increased transfer opportunities.
 

### 2017-06-27 Response.tex
Reword another paragraph summarizing key idea
Reviewer 1 Comment 7 and Comment 8

@@
 
+\item 
+    \begin{itshape}
 -"In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem." Why planar?
@@
 It doesn't hurt to be straightforward about this, as it is a first-step towards characterizing the design space.
+\end{itshape}
 
+The introduction has been modified with an additional sentence indicating our use of the planar three body problem
+
+\begin{correction}
+Utilizing the planar assumption greatly simplifies the problem and allows for the use of established methods in determining periodic orbits and \Poincare sections.
+\end{correction}
+
+\item 
+    \begin{itshape}
 -"In addition, through the use of geometric integrators we accurately capture the effects of low thrust on the system dynamics in the numerical simulation."  A variational integrator will accurately capture the qualitative effects of the transfer (e.g.
@@
 I would soften this claim or reword.
+\end{itshape}
 
+While the geometric integrator will not exactly characterize the terminal energy or state of a dynamic system, it will better preserve the total energy variation as compared to convential integration schemes.
+In addition, the total energy deviation will remain bounded and less than than of convential schemes which experience unbounded energy deviation. 
+As a result, the terminal state/energy will be more accurately characterized using a variational integrator as compared to conventional schemes. 
+Also this difference becomes more evident over increasing time horizons.
+Some additional detail has been added stating that the increased accuracy is most evident over long time periods.
+
+\begin{correction}
+In addition, through the use of geometric integrators we more accurately capture the effects of low thrust on the system dynamics in the numerical simulation over extended periods which are typical of electric propulsion systems.
+\end{correction}
 -"The Poincare section reduces the dimensionality of the system dyanmics to the study of a related discrete update map." It does, but at the expense if a large up-front computational run that must be run well in advance to capture the problem characteristics.

### 2017-06-27 manuscript.tex
Reword another paragraph summarizing key idea
Reviewer 1 Comment 7 and Comment 8

@@
 % concise summary of contributions
+
+%-------------------- OBJECTIVES OF THIS WORK-----------------------------%
+% SYSTEMATIC DESIGN TO AVOID DIFFICULTIES IN CHOOSING A GOOD INITIAL GUESS
+% CAPTURE LONG TERM EFFECTS OF LOW THRUST ACCURATELY IN NUMERICAL SIMULATION
+In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem.
+Utilizing the planar assumption greatly simplifies the problem and allows for the use of established methods in determining periodic orbits and \Poincare~sections.
+We utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
+Through the use of a simple distance metric on a \Poincare~section greatly reduces the dimensionality of the reachability analysis and provides an approximation of the reachability set.
+In addition, through the use of geometric integrators we more accurately capture the effects of low thrust on the system dynamics in the numerical simulation over extended periods which are typical of electric propulsion systems.
+With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low thrust propulsion systems.
+
 \section{Planar Circular Restricted Three Body Problem}\label{sec:pcrtbp}

### 2017-06-27 Response.tex
Add a general description of the approach
Reviewer 1 Comment 9

@@
 \end{correction}
+
+\item 
+    \begin{itshape}
 -"The Poincare section reduces the dimensionality of the system dyanmics to the study of a related discrete update map." It does, but at the expense if a large up-front computational run that must be run well in advance to capture the problem characteristics.
 Also, it is more challenging to execute in 3D, which should be noted.
+\end{itshape}
+
+Some additional detail has been included mentioning the difficulties in applying the \Poincare~section approach to the 3D case.
+
+\begin{correction}
+While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
+\end{correction}
 

### 2017-06-27 manuscript.tex
Add a general description of the approach
Reviewer 1 Comment 9

@@
 With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low thrust propulsion systems.
+
+We defne an optimal control problem to determine the reachability set on a \Poincare~section.
+The reachable set, defined as the set of all attainable states subect to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
+Defining the reachable set on a \Poincare~section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
+While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
+Instead of relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on the \Poincare~section.
+Through the use of low-thrust control input, the reachable set on the \Poincare~section is enlarged and enables a larger space of potential transfers.
+By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare~section allows for the creation of general transfer trajectories.
 

### 2017-06-27 Response.tex
Introduction simplified and reduced
Reviewer 2 Comment 1

@@
 SPECIFIC FEEDBACK:
-
+\begin{itemize}
+    \begin{itshape}
 *** Bloat
@@
 It needs to be organized into titled subsections with specific scope.
+\end{itshape}
+
+The introduction is reduced and many of the paragraphs have been reworded and combined. 
+In addition, the extended discussion of the benefits of variational integrators is removeda.
+In the interest of space/clarity the modified introduction is not copied here.
 
@@
 In general, please devote more depth to probing the strengths, weaknesses, and fundamental properties of the method.  Considering extreme cases can help with this.
-
+\end{itemize}
 \bibliography{library}

### 2017-06-27 manuscript.tex
Introduction simplified and reduced
Reviewer 2 Comment 1

@@
 By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare~section allows for the creation of general transfer trajectories.
+
+In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
+Previous results in the design of optimal transfers have relied on suboptimal direct optimization methods and conventional integration techniques.
+This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
+We derive and solve the Euler-Lagrange equations to generate the necessary conditions for an optimal solution.
+This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods.
+In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
+Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
+We avoid the issues inherent in selecting a valid initial condition for optimization.
+Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
+We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.
 

### 2017-06-29 Response.tex
Reorganize/reduce background section
Rviewer 1 and 2 PCRTBP and Bloat comments - partial

Working to reduce the discussion on background materials.
Everythign will be under a single major section

@@
 \subsection*{Reviewer 1} 
-\begin{enumerate}
-    \item \textit{Text of comment}
+\begin{itshape}
+Reviewer \#1: I have some issues with this paper that, in my opinion necessitate a major revision.
+These issues are:
+\end{itshape}
 
-        \begin{correction}
-
-            My correction
-
-        \end{correction}
-\end{enumerate}
-Reviewer \#1: I have some issues with this paper that, in my opinion necessitate a major revision.
-
-These issues are:
-
+\begin{itemize}
+    \item 
+        \begin{itshape}
 -It extensively re-covers issues that have been covered in many papers, making it difficult to disseminate what the original contributions are.
@@
 The content of this paper easily be reduced considerably to maintain focus on the original contributions.
+    \end{itshape}
+
+    The paper is being reduced. 
+    The introduction has been reduced to avoid excessive repetition. 
+    In addition, much of the background material has been removed
 
@@
 
+\item 
+    \begin{itshape}
 The pros and cons of the PCRTBP
@@
 -The PCRTB is limiting for many pragmatic departure orbits (e.g. LEO and GTO, the most common departure orbits for low-thrust transfers).
+\end{itshape}
+
+Additional detail has been included in the problem formulation section.
+This includes some discussion on the choice of the PCRTBP and the relative strengths/weaknesses of this approach.
+The additional paragraph has been copied below.
+
+\begin{correction}
+We utilize the planar circular restricted three body problem~(PCRTBP) as the basis of our system definition. 
+It is a popular model in the preliminary analysis of multibody spacecraft trajectories. 
+In the context of Earth based mission, the PCRTBP affords a relatively simple dynamic model while still capturing the major third body perturbation of the Moon.
+In addition, this model allows for a systematic process to define and exploit \Poincare sections.
+Furthermore, the 2D solutions afforded by the PCRTBP are frequently use to gain a qualitative understanding of the trade space of transfers in the Earth-Moon system.
+The PCRTBP approach offers insight into the fundamental dynamical structure while capturing the major dynamic properties of the planar motion.
+As a result, this approach is best suited for trajectories which do not require large plane changes.
+\end{correction}
 
@@
 
+\end{itemize}
 From here on out comments specific to various sections of the paper follow:
@@
 \begin{itemize}
+    \item
     \begin{itshape}

### 2017-06-29 manuscript.tex
Reorganize/reduce background section
Rviewer 1 and 2 PCRTBP and Bloat comments - partial

Working to reduce the discussion on background materials.
Everythign will be under a single major section

@@
 
-\section{Planar Circular Restricted Three Body Problem}\label{sec:pcrtbp}
-The system model is based on the planar circular restricted three body problem~(PCRTBP).
+\section{Problem Formulation and Mathematical Background}\label{sec:background}
+
+% discuss some of the pros/cons of using PCRTBP model
+We utilize the planar circular restricted three body problem~(PCRTBP) as the basis of our system definition. 
+It is a popular model in the preliminary analysis of multibody spacecraft trajectories. 
+In the context of Earth based mission, the PCRTBP affords a relatively simple dynamic model while still capturing the major third body perturbation of the Moon.
+In addition, this model allows for a systematic process to define and exploit \Poincare sections.
+Furthermore, the 2D solutions afforded by the PCRTBP are frequently use to gain a qualitative understanding of the trade space of transfers in the Earth-Moon system.
+The PCRTBP approach offers insight into the fundamental dynamical structure while capturing the major dynamic properties of the planar motion.
+As a result, this approach is best suited for trajectories which do not require large plane changes.
+
+The research in this work utilizes a variety of disparate concepts in dynamics, control theory, and astrodynamics. 
+As a result, we briefly present some key concepts in the system model as well as established tools in the three body problem, such as invariant manifolds and \Poincare sections.
+In addition, we summarize the derivation of a variational integrator for the PCRTBP which is used in the subsequent geometric optimal control formulation.
+\subsection{Planar Circular Restricted Three Body Problem}\label{sec:pcrtbp}
 The Earth is assumed to be the more massive primary, \( m_1 \), while the Moon is the second, smaller primary \( m_2\).
 The equations of motion are developed in a rotating reference frame which allows for much greater insight into the structure of the dynamics.
-The \( \hat{e}_1 \) axis is directed along the vector from the Earth to the Moon.
-The \( \hat{e}_2 \) axis lies in the orbital plane and is orthogonal to \( \hat{e}_1 \).
-The rotating reference frame is centered at the system barycenter.
-It is assumed that the \(\left( \hat{e}_1, \hat{e}_2\right)\) frame rotates with a constant angular velocity equal to the mean motion of the Moon.
 Following convention, the system is also non-dimensionalized by the characteristic units of length, mass, and time~\cite{koon2011}.
@@
 \end{equation}
-The larger primary, \(m_1\), is located at \( \left(  -\mu , 0 \right)\) and the smaller \( m_2\) is located at \( \left( 1-\mu , 0 \right)\), while the spacecraft is located at \( (x ,y) \).
-In the rotating reference frame the Lagrangian is given by
-\begin{equation}
-        L = \frac{1}{2} \left( \left( \dot{x} -y \right)^2 + \left( \dot{y} + x \right)^2 \right) + \frac{1-\mu}{r_1} + \frac{\mu}{r_2}\, ,
-        \label{eq:lagrangian}
-\end{equation}
-where the distance \( r_1, r_2 \in \R \) of the spacecraft to each primary is as
-\begin{align}
-        r_1 &= \sqrt{\left( x + \mu\right)^2 + y^2}\, , \\
-        r_2 &= \sqrt{\left( x - 1 + \mu\right)^2 + y^2}\, .
-        \label{eq:distances}
-\end{align}
-Application of the Euler-Lagrange equations results in following equations of motion defined in the rotating reference frame
-\begin{align}
-        \ddot{x} - 2 \dot{y} + \deriv{U}{x} &= u_x \, ,\nonumber \\
-        \ddot{y} + 2 \dot{y} + \deriv{U}{y} &= u_y \, ,
-        \label{eq:cont_eom}
-\end{align}
-where the effective potential \( U\) is defined as
-\begin{equation}
-        U = \frac{1}{2} \left( x^2 + y^2\right) + \frac{1-\mu}{r_1} + \frac{\mu}{r_2}\, ,
-        \label{eq:eff_pot}
-\end{equation}
-and the control inputs is defined as \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \).
-The state is defined as \( \vecbf{x} = \begin{bmatrix}\vecbf{r} &\vecbf{v} \end{bmatrix}^T\) with \(\vecbf{r} = \begin{bmatrix} x & y \end{bmatrix}^T \in \R^{2\times1}\) and \(\vecbf{v}= \begin{bmatrix} \dot{x} & \dot{y} \end{bmatrix}^T \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
-The equations of motion may be rewritten in state space form as
+Following a straightforward application of the Euler-Lagrange equations results in the following equations of motion defined in the rotating reference frame
 \begin{equation}\label{eqn:cont_dyn}
@@
 where the matrix \( A \) and psuedo gravitational potential gradient \( \nabla U\) are
-\begin{equation}\label{eq:A_mat}
-        A = \left[ \begin{array}{ccc} 0 & 2 & 0 \\ -2 & 0 & 0 \\ 0 & 0 & 0 \end{array} \right] \, ,
-\end{equation}
-\begin{equation} \label{eq:grav_pot}
-        \nabla U = \left[ \begin{array}{c} x - \frac{ \left(1 - \mu\right) \left(x + \mu\right)}{r_1^3} - \frac{\mu \left( x - 1 + \mu \right)}{r_2^3} \\
+\begin{align}
+    A &= \left[ \begin{array}{ccc} 0 & 2 & 0 \\ -2 & 0 & 0 \\ 0 & 0 & 0 \end{array} \right], \label{eq:A_mat} \\
+    \nabla U &= \left[ \begin{array}{c} x - \frac{ \left(1 - \mu\right) \left(x + \mu\right)}{r_1^3} - \frac{\mu \left( x - 1 + \mu \right)}{r_2^3} \\
                                                                                         y - \frac{ \left(1 - \mu\right) y}{r_1^3} - \frac{\mu y}{r_2^3} \\
                                                                                         - \frac{ \left(1 - \mu\right) z}{r_1^3} - \frac{\mu z}{r_2^3}\end{array}\right]
-                                        = \left[\begin{array}{c} U_x \\ U_y \\ U_z\end{array} \right] \, .
-\end{equation}
-
-\subsection{Jacobi Integral}\label{sec:jacobi}
+                                        = \left[\begin{array}{c} U_x \\ U_y \\ U_z\end{array} \right] , \label{eq:grav_pot}
+\end{align}
+and the control inputs is defined as \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \) and assumed to be continuously variable but bounded in magnitude, i.e.
+\( \vecbf{u}^T \vecbf{u} \leq u_{max}^2 \).
+The state is defined as \( \vecbf{x} = \begin{bmatrix}\vecbf{r} &\vecbf{v} \end{bmatrix}^T\) with \(\vecbf{r} = \begin{bmatrix} x & y \end{bmatrix}^T \in \R^{2\times1}\) and \(\vecbf{v}= \begin{bmatrix} \dot{x} & \dot{y} \end{bmatrix}^T \in \R^{2\times1}\) representing the position and velocity with respect to the system barycenter, respectively.
+
+\subsubsection{Jacobi Integral}\label{sec:jacobi}
 There exists a single integral, or constant of motion for the three-body problem~\cite{szebehely1967,lanczos1970}.
@@
 %Introduce invariant manifolds
-
-\section{Variational Integrator}\label{sec:discrete_var}
-% description of variational integrator development
-Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invaraint properties and symplecticity.
-Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations~\cite{marsden2001}.
-As a result, integrators developed in this manner have the desirable properties that they are symplectic and momentum preserving.
-In addition, they exhibit improved energy behavior over long integration periods.
-We present a short background on the variational principle for mechanical systems. 
-We then develop a discrete approximation of the action integral and construct a variational integrator for the PCRTBP.
-
-\subsection{Variational Principle}
-Consider a continuous mechanical system described by the Lagrangian, \( L( q, \dot{q} ) \), for the generalized position, \( q\), and velocity, \( \dot{q} \).
-In the standard approach of variational mechanics the action integral is formed by integrating the continuous Lagrangian along a path \( q(t) \) that the system follows from time \( t = 0 \) to \( t = T \)~\cite{greenwood1988}.
-In the continuous time the action integral is defined as
-\begin{align}\label{eq:action_integral}
-        S = \int_{0}^T L\left( q, \dot{q}\right) \, dt \, .
-\end{align}
-Hamilton's principle states that the actual path followed by a holonomic system results in a stationary action integral with respect to path variations for fixed endpoints.
-Taking the variation of~\cref{eq:action_integral} gives
-\begin{align}\label{eq:var_principle}
-        \delta S %&= \int_{0}^T \deriv{L}{q} \delta q + \deriv{L}{\dot{q}} \delta \dot{q} \, dt \\
-                %&= \int_{0}^T \deriv{L}{q} \delta q - \frac{d}{dt} \left( \deriv{L}{\dot{q}}\right) \delta q \, dt - \left. \left[ \deriv{L}{\dot{q}} \delta q\right] \right|_0^T \\
-        &= \int_{0}^T \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}}     \right) \, dt \, ,
-\end{align}
-where we have used integration by parts and the conditions \( \delta q(0) = 0 \) and \( \delta q(T) = 0\).
-For Hamilton's principle to be valid for all admissible variations \( \delta q \), the integrand of~\cref{eq:var_principle} must be zero for all \( t\), giving the continuous Euler-Lagrange Equations~\cite{lanczos1970}.
-\begin{align}\label{eq:euler-lagrange}
-        0 = \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}} \right) \, .
-\end{align}
-Hamilton's equations are derivable through the use of the Legendre transformation which is a mapping \( \left( q, \dot{q},t\right) \rightarrow \left(q, p, t \right) \) where \( p_i\) is the generalized momenta.
-\begin{align}\label{eq:legendre_transform}
-        p_i = \deriv{L}{\dot{q}_i} \, .
-\end{align}
-In the continuous time case the Hamiltonian is defined as
-\begin{align}\label{eq:hamiltonian}
-        H &= \sum_{i = 1}^n p_i \dot{q}_i - L \left( q,\dot{q}, t \right) \, .
-\end{align}
-Applying~\cref{eq:legendre_transform} and taking the variation of~\cref{eq:hamiltonian} allows us to derive the equations of motion in Hamiltonian form
-\begin{subequations}\label{eq:hamilton_eq}
-\begin{align}
-        \dot{q}_i &= \deriv{H}{p_i} \, ,\\
-        \dot{p}_i &= - \deriv{H}{q_i} \, , \\
-        \deriv{L}{t} &= -\deriv{H}{t} \, .
-\end{align}
-\end{subequations}
-Both~\cref{eq:euler-lagrange,eq:hamilton_eq} result in equations of motion for the mechanical system and are equivalent via the Legendre transform.
-\Cref{eq:euler-lagrange} results in \( n \) second order differential equations while~\cref{eq:hamilton_eq} results in \( 2n \) first order differential equations.
-\subsection{Discrete Variational Mechanics}
-A discrete analogue of Hamilton's principle and the action integral is formed.
-Rather than taking a position, \( q \), and velocity, \( \dot{q} \), consider two positions \( q_0 \) and \( q_1 \) and a fixed time step \( h \in \R \).
-The two positions are points on the curve \( q(t) \) such that \( q_0 \approx q(0) \) and \( q_1 \approx q(h) \).
-A discrete time Lagrangian \( L_d( q_0, q_1) \) is formed which approximates the action integral between \( q_0 \) and \( q_1 \) as 
-\begin{align}\label{eq:discrete_lagrangian_gen}
-        L_d\left( q_0 , q_1 \right) \approx \int_{0}^{h} L \left( q , \dot{q} \right) \, dt \, .
-\end{align}
-Since~\cref{eq:discrete_lagrangian_gen} is calculated as a numerical integral, an appropriate quadrature rule is required.
-There are multiple possible methods one can use to approximate the integral in~\cref{eq:discrete_lagrangian_gen}.
-An appropriate approximation rule is determined based on the ease of implementation and accuracy desired.
-\begin{table}[htbp]
-\caption{Selected Quadrature Rules\label{tab:quadrature}}
-\begin{center}
-\begin{tabular}{l|l}Rectangle & \( L_d(q_0,q_1) =L(q_0,\frac{q_1-q_0}{h}) h \)  \\ \hline
-Midpoint & \( L_d(q_0,q_1) = L(\frac{q_0 + q_1}{2},\frac{q_1 - q_0}{h}) h \) \\ \hline
-Trapezoidal & \( L_d(q_0, q_1) = \frac{1}{2} \left[ L(q_0, \frac{q_1 - q_0}{h} ) + L(q_1, \frac{q_1 - q_0 }{h} )\right] h \)
-\end{tabular} 
-\end{center}
-\end{table}
-\Cref{tab:quadrature} shows several possible approximation rules that are typically applied.
-The rectangle rule is a first order accurate method and offers a straightforward implementation.
-The midpoint and trapezoidal rules are both second order accurate methods. 
-However, the midpoint rule results in an implicit form which adds further complexity to the equations of motion.
-%In this work, the trapezoidal approximation is applied to the PCRTBP.
-
-Once an appropriate discrete Lagrangian is formed a discrete action sum is formed as the discrete analogue of~\cref{eq:action_integral}
-\begin{align}\label{eq:action_sum}
-        S_d = \sum_{k=0}^{N-1} L_d(q_k, q_{k+1}) \, .
-\end{align}
-Once again a discrete version of Hamilton's principle is applied to~\cref{eq:action_sum}.
-This yields a discrete time counterpart to~\cref{eq:var_principle} as
-\begin{align}\label{eq:dis_var_principle}
-        \delta S_d %&= \sum_{k=0}^{N-1} \deriv{L_d(q_k, q_{k+1})}{q_k} \delta q_k + \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \delta q_{k+1} \\
-        &= \sum_{k=1}^{N-1} \bracket{ \deriv{L_d(q_k, q_{k+1})}{q_k} + \deriv{L_d(q_{k-1}, q_k)}{q_{k+1}}} \delta q_k \, .
-\end{align}
-
-For the discrete action sum to be stationary with respect to all admissible path variations, with fixed endpoints, the discrete Euler-Lagrange equations must be satisfied for \( k = 1, \cdots, N-1 \) resulting in
-\begin{align}\label{eq:discrete_euler-lagrange}
-        0 = \deriv{L_d(q_k, q_{k+1})}{q_k} + \deriv{L_d(q_{k-1}, q_k)}{q_{k+1}} \, .
-\end{align}
-A discrete version of the Legendre transformation, referred to as a discrete fiber derivative, results in the equivalent Hamiltonian form expression.
-The discrete fiber derivative is given as 
-\begin{subequations}\label{eq:discrete_legendre}
-\begin{align}
-        p_k &= \deriv{L_d(q_{k-1},q_k)}{q_{k+1}} = - \deriv{L_d(q_k, q_{k+1})}{q_k} \, ,\\
-        p_{k+1} &= \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \, .
-\end{align}
-\end{subequations}
-This yields a discrete Hamiltonian map \( (q_k, p_k) \to (q_{k+1}, p_{k+1}) \).
-A more extensive development of variational integrators can be found in~\cite{marsden2001}.
-
-\subsection{Discrete Equations of Motion}
-The discrete equations of motion for the PCRTBP are derived by choosing an appropriate quadrature rule to discretize the Lagrangian in~\cref{eq:lagrangian}. 
-In this work, the trapezoidal approximation is applied.
-The trapezoid rule allows for an explicit second order accurate approximation.
-The discrete Lagrangian is given by
-\begin{align}\label{eq:discrete_lagrangian}
-        L_d = &\frac{h}{2} \left( \frac{1}{2} \bracket{\left(  \frac{\xkp - \xk}{h} -\yk \right)^2 + \left( \frac{\ykp - \yk}{h} + \xk \right)^2} + \frac{1 - \mu}{r_{1_k}} + \frac{\mu}{r_{2_k}} \right.
-\nonumber \\
-                & + \left.
-\frac{1}{2} \bracket{\left(  \frac{\xkp - \xk}{h} -\ykp \right)^2 + \left( \frac{\ykp - \yk}{h} + \xkp \right)^2} + \frac{1-\mu}{r_{1_{k+1}}} + \frac{\mu}{r_{2_{k+1}}}  \right) \, .
-\end{align}
-Applying a discrete version of the Lagrange-d'Alembert principle allows for inclusion of an external control force on the system~\cite{marsden2001}.
-Using~\cref{eq:discrete_legendre,eq:discrete_lagrangian} and some manipulation, the equations of motion are given by
-\begin{subequations}\label{eq:discrete_eoms}
-\begin{align}
-        \xkp &= \frac{1}{1+ h^2} \bracket{h \dot{x}_k + h^2 \dot{y}_k  + \xk \parenth{1+ \frac{3h^2}{2}} + \frac{h^3}{2} \yk - \frac{h^3}{2} U_{y_k} - \frac{h^2}{2} U_{x_k} } \label{eq:xkp} \, ,\\
-        \ykp &= h \dot{y}_k + h \xk - h \xkp + \yk + \frac{h^2 \yk}{2} - \frac{h^2 }{2} U_{y_k} \label{eq:ykp} \, ,\\
-        \dot{x}_{k+1} &= \dot{x}_k - 2 \yk + 2 \ykp + \frac{h}{2} \parenth{\xkp + \xk} - \frac{h}{2} U_{\xkp} - \frac{h}{2} U_{\xk} + h u_x \label{eq:xdotkp}\, ,\\
-        \dot{y}_{k+1} &= \dot{y}_{k} + 2 \xk - 2 \xkp + \frac{h}{2} \parenth{\ykp + \yk} - \frac{h}{2} U_{\ykp} - \frac{h}{2} U_{\yk} + h u_y \label{eq:ydotkp} \, .
-\end{align}
-\end{subequations}
-The discrete equations of motion are given in the Lagrangian form after applying the discrete fiber derivative from~\cref{eq:discrete_legendre} as \( p_{\xk} = \dot{x}_k - \yk \) and \( p_{\yk} = \dot{y}_k + \xk \).
-The state is defined as \( \vecbf{x}_k = \begin{bmatrix} \xk & \yk & \dot{x}_k & \dot{y}_k \end{bmatrix}^T\) and the control input is \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \).
-This results in a discrete update map \( f_k: \vecbf{x}_k \to \vecbf{x}_{k+1} \) which preserves the same properties of the continuous-time dynamics in~\cref{eqn:cont_dyn} such as invariants, symplecticity, and the configuration manifold.
-The discrete potential gradients are given by
-\begin{subequations}\label{eq:discrete_potential_grad}
-\begin{align}
-        U_{\xk} &= \frac{\parenth{1 -\mu} \parenth{\xk + \mu}}{\distonek^3} + \frac{ \mu \parenth{\xk -1 + \mu}}{\disttwok^3} \label{eq:Uxk} \, ,\\
-        U_{\yk} &= \frac{\parenth{1 -\mu} \yk}{\disttwok^3} + \frac{ \mu \yk}{\disttwok^3} \label{eq:Uyk} \, ,\\
-        U_{\xkp} &= \frac{\parenth{1 -\mu} \parenth{\xkp + \mu}}{\distonekp^3} + \frac{ \mu \parenth{\xkp -1 + \mu}}{\disttwokp^3} \label{eq:Uxkp}\, ,\\
-        U_{\ykp} &= \frac{\parenth{1 -\mu} \ykp}{\distonekp^3} + \frac{ \mu \ykp}{\disttwokp^3} \label{eq:Uykp} \, .
-\end{align}     
-\end{subequations}
-The distances to each primary are defined as
-\begin{subequations}\label{eq:discrete_distance}
-\begin{align}
-        \distonek &= \sqrt{\left( \xk + \mu\right)^2 + \yk^2} \label{eq:distonek}\, ,\\
-        \disttwok &= \sqrt{\left( \xk - 1 + \mu\right)^2 + \yk^2} \label{eq:disttwok}\, ,\\
-        \distonekp &= \sqrt{\left( \xkp + \mu\right)^2 + \ykp^2} \label{eq:distonekp}\, ,\\
-        \disttwokp &= \sqrt{\left( \xkp - 1 + \mu\right)^2 + \ykp^2} \label{eq:disttwokp} \, .
-\end{align}
-\end{subequations}
-Care must be taken during the implementation of~\cref{eq:discrete_eoms}.
-As~\cref{eq:discrete_potential_grad,eq:discrete_distance} are defined at both step \( k \) and \( k+1 \) they must be evaluated at both time instances.
-\Cref{eq:discrete_eoms} is implemented by first defining an initial state \( \vecbf{x}_k \) and control \( \vecbf{u}_k \).
-The distances and gravitational potential at step \( k \) are evaluated from~\cref{eq:distonek,eq:disttwok,eq:Uxk,eq:Uyk}.
-The discrete update steps in~\cref{eq:xkp,eq:ykp} are evaluated to generate \( \xkp \) and \( \ykp\).
-Next, the distances and gravitational potential at step \( k+1 \) are evaluated from~\cref{eq:distonekp,eq:disttwokp,eq:Uxkp,eq:Uykp}. 
-Finally, the update steps in~\cref{eq:xdotkp,eq:ydotkp} are evaluated.
-This results in the complete discrete update map \( \vecbf{x}_k \to \vecbf{x}_{k+1} \) given \( \vecbf{u}_k \).
-\subsection{Numerical Example}\label{sec:variational_example}
-\begin{figure} 
-        \centering 
-        \begin{subfigure}[h]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{trajectory} 
-                \caption{Trajectory comparison} \label{fig:compare_trajectory} 
-        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-%       \begin{subfigure}[htbp]{0.3\textwidth} 
-%               \includegraphics[width=\textwidth]{components} 
-%               \caption{State \( \parenth{x , y, \dot{x}, \dot{y}}\)} \label{fig:compare_components} 
-%       \end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{energy} 
-                \caption{Jacobi Integral comparison} \label{fig:compare_energy} 
-        \end{subfigure} 
-        \caption{Integrator Comparison}
-        \label{fig:integrator_compare} 
-\end{figure}
-A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
-A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
-The variational integrator uses a step size of \SI{47.22}{\second} while the Runge-Kutta method uses a variable step size implemented via ODE45 in Matlab.
-\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference frame for both integration schemes.
-Both integration schemes result in trajectories that are initially nearly identical.
-The discrete equations of motion are an accurate approximation for the continuous dynamics as they closely match the solution of ODE45 over the initial portion of the simulation.
-However, as time progresses the trajectories begin to diverge due to the differences in system energy.
-\Cref{fig:compare_energy} shows the evolution of the Jacobi integral.
-The variational integrator exhibits a bounded behavior about the initial energy with a mean variation of \num{4.2522e-20}.
-However, the conventional Runge-Kutta method demonstrates a clear energy drift of \num{4.2814e-8}. 
-Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques.
-
-\section{Invariant Manifolds and \Poincare Map} 
+\subsubsection{Invariant Manifolds and \Poincare Map} 
 % discuss usefulness and basic generation of invariant manifolds
@@
 We compute the reachable set on an appropriate \Poincare section and use this to design a transfer trajectory.
+
+\section{Variational Integrator}\label{sec:discrete_var}
+% description of variational integrator development
+Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invaraint properties and symplecticity.
+Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations~\cite{marsden2001}.
+As a result, integrators developed in this manner have the desirable properties that they are symplectic and momentum preserving.
+In addition, they exhibit improved energy behavior over long integration periods.
+We present a short background on the variational principle for mechanical systems. 
+We then develop a discrete approximation of the action integral and construct a variational integrator for the PCRTBP.
+
+\subsection{Variational Principle}
+Consider a continuous mechanical system described by the Lagrangian, \( L( q, \dot{q} ) \), for the generalized position, \( q\), and velocity, \( \dot{q} \).
+In the standard approach of variational mechanics the action integral is formed by integrating the continuous Lagrangian along a path \( q(t) \) that the system follows from time \( t = 0 \) to \( t = T \)~\cite{greenwood1988}.
+In the continuous time the action integral is defined as
+\begin{align}\label{eq:action_integral}
+        S = \int_{0}^T L\left( q, \dot{q}\right) \, dt \, .
+\end{align}
+Hamilton's principle states that the actual path followed by a holonomic system results in a stationary action integral with respect to path variations for fixed endpoints.
+Taking the variation of~\cref{eq:action_integral} gives
+\begin{align}\label{eq:var_principle}
+        \delta S %&= \int_{0}^T \deriv{L}{q} \delta q + \deriv{L}{\dot{q}} \delta \dot{q} \, dt \\
+                %&= \int_{0}^T \deriv{L}{q} \delta q - \frac{d}{dt} \left( \deriv{L}{\dot{q}}\right) \delta q \, dt - \left. \left[ \deriv{L}{\dot{q}} \delta q\right] \right|_0^T \\
+        &= \int_{0}^T \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}}     \right) \, dt \, ,
+\end{align}
+where we have used integration by parts and the conditions \( \delta q(0) = 0 \) and \( \delta q(T) = 0\).
+For Hamilton's principle to be valid for all admissible variations \( \delta q \), the integrand of~\cref{eq:var_principle} must be zero for all \( t\), giving the continuous Euler-Lagrange Equations~\cite{lanczos1970}.
+\begin{align}\label{eq:euler-lagrange}
+        0 = \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}} \right) \, .
+\end{align}
+Hamilton's equations are derivable through the use of the Legendre transformation which is a mapping \( \left( q, \dot{q},t\right) \rightarrow \left(q, p, t \right) \) where \( p_i\) is the generalized momenta.
+\begin{align}\label{eq:legendre_transform}
+        p_i = \deriv{L}{\dot{q}_i} \, .
+\end{align}
+In the continuous time case the Hamiltonian is defined as
+\begin{align}\label{eq:hamiltonian}
+        H &= \sum_{i = 1}^n p_i \dot{q}_i - L \left( q,\dot{q}, t \right) \, .
+\end{align}
+Applying~\cref{eq:legendre_transform} and taking the variation of~\cref{eq:hamiltonian} allows us to derive the equations of motion in Hamiltonian form
+\begin{subequations}\label{eq:hamilton_eq}
+\begin{align}
+        \dot{q}_i &= \deriv{H}{p_i} \, ,\\
+        \dot{p}_i &= - \deriv{H}{q_i} \, , \\
+        \deriv{L}{t} &= -\deriv{H}{t} \, .
+\end{align}
+\end{subequations}
+Both~\cref{eq:euler-lagrange,eq:hamilton_eq} result in equations of motion for the mechanical system and are equivalent via the Legendre transform.
+\Cref{eq:euler-lagrange} results in \( n \) second order differential equations while~\cref{eq:hamilton_eq} results in \( 2n \) first order differential equations.
+\subsection{Discrete Variational Mechanics}
+A discrete analogue of Hamilton's principle and the action integral is formed.
+Rather than taking a position, \( q \), and velocity, \( \dot{q} \), consider two positions \( q_0 \) and \( q_1 \) and a fixed time step \( h \in \R \).
+The two positions are points on the curve \( q(t) \) such that \( q_0 \approx q(0) \) and \( q_1 \approx q(h) \).
+A discrete time Lagrangian \( L_d( q_0, q_1) \) is formed which approximates the action integral between \( q_0 \) and \( q_1 \) as 
+\begin{align}\label{eq:discrete_lagrangian_gen}
+        L_d\left( q_0 , q_1 \right) \approx \int_{0}^{h} L \left( q , \dot{q} \right) \, dt \, .
+\end{align}
+Since~\cref{eq:discrete_lagrangian_gen} is calculated as a numerical integral, an appropriate quadrature rule is required.
+There are multiple possible methods one can use to approximate the integral in~\cref{eq:discrete_lagrangian_gen}.
+An appropriate approximation rule is determined based on the ease of implementation and accuracy desired.
+\begin{table}[htbp]
+\caption{Selected Quadrature Rules\label{tab:quadrature}}
+\begin{center}
+\begin{tabular}{l|l}Rectangle & \( L_d(q_0,q_1) =L(q_0,\frac{q_1-q_0}{h}) h \)  \\ \hline
+Midpoint & \( L_d(q_0,q_1) = L(\frac{q_0 + q_1}{2},\frac{q_1 - q_0}{h}) h \) \\ \hline
+Trapezoidal & \( L_d(q_0, q_1) = \frac{1}{2} \left[ L(q_0, \frac{q_1 - q_0}{h} ) + L(q_1, \frac{q_1 - q_0 }{h} )\right] h \)
+\end{tabular} 
+\end{center}
+\end{table}
+\Cref{tab:quadrature} shows several possible approximation rules that are typically applied.
+The rectangle rule is a first order accurate method and offers a straightforward implementation.
+The midpoint and trapezoidal rules are both second order accurate methods. 
+However, the midpoint rule results in an implicit form which adds further complexity to the equations of motion.
+%In this work, the trapezoidal approximation is applied to the PCRTBP.
+
+Once an appropriate discrete Lagrangian is formed a discrete action sum is formed as the discrete analogue of~\cref{eq:action_integral}
+\begin{align}\label{eq:action_sum}
+        S_d = \sum_{k=0}^{N-1} L_d(q_k, q_{k+1}) \, .
+\end{align}
+Once again a discrete version of Hamilton's principle is applied to~\cref{eq:action_sum}.
+This yields a discrete time counterpart to~\cref{eq:var_principle} as
+\begin{align}\label{eq:dis_var_principle}
+        \delta S_d %&= \sum_{k=0}^{N-1} \deriv{L_d(q_k, q_{k+1})}{q_k} \delta q_k + \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \delta q_{k+1} \\
+        &= \sum_{k=1}^{N-1} \bracket{ \deriv{L_d(q_k, q_{k+1})}{q_k} + \deriv{L_d(q_{k-1}, q_k)}{q_{k+1}}} \delta q_k \, .
+\end{align}
+
+For the discrete action sum to be stationary with respect to all admissible path variations, with fixed endpoints, the discrete Euler-Lagrange equations must be satisfied for \( k = 1, \cdots, N-1 \) resulting in
+\begin{align}\label{eq:discrete_euler-lagrange}
+        0 = \deriv{L_d(q_k, q_{k+1})}{q_k} + \deriv{L_d(q_{k-1}, q_k)}{q_{k+1}} \, .
+\end{align}
+A discrete version of the Legendre transformation, referred to as a discrete fiber derivative, results in the equivalent Hamiltonian form expression.
+The discrete fiber derivative is given as 
+\begin{subequations}\label{eq:discrete_legendre}
+\begin{align}
+        p_k &= \deriv{L_d(q_{k-1},q_k)}{q_{k+1}} = - \deriv{L_d(q_k, q_{k+1})}{q_k} \, ,\\
+        p_{k+1} &= \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \, .
+\end{align}
+\end{subequations}
+This yields a discrete Hamiltonian map \( (q_k, p_k) \to (q_{k+1}, p_{k+1}) \).
+A more extensive development of variational integrators can be found in~\cite{marsden2001}.
+
+\subsection{Discrete Equations of Motion}
+The discrete equations of motion for the PCRTBP are derived by choosing an appropriate quadrature rule to discretize the Lagrangian in~\cref{eq:lagrangian}. 
+In this work, the trapezoidal approximation is applied.
+The trapezoid rule allows for an explicit second order accurate approximation.
+The discrete Lagrangian is given by
+\begin{align}\label{eq:discrete_lagrangian}
+        L_d = &\frac{h}{2} \left( \frac{1}{2} \bracket{\left(  \frac{\xkp - \xk}{h} -\yk \right)^2 + \left( \frac{\ykp - \yk}{h} + \xk \right)^2} + \frac{1 - \mu}{r_{1_k}} + \frac{\mu}{r_{2_k}} \right.
+\nonumber \\
+                & + \left.
+\frac{1}{2} \bracket{\left(  \frac{\xkp - \xk}{h} -\ykp \right)^2 + \left( \frac{\ykp - \yk}{h} + \xkp \right)^2} + \frac{1-\mu}{r_{1_{k+1}}} + \frac{\mu}{r_{2_{k+1}}}  \right) \, .
+\end{align}
+Applying a discrete version of the Lagrange-d'Alembert principle allows for inclusion of an external control force on the system~\cite{marsden2001}.
+Using~\cref{eq:discrete_legendre,eq:discrete_lagrangian} and some manipulation, the equations of motion are given by
+\begin{subequations}\label{eq:discrete_eoms}
+\begin{align}
+        \xkp &= \frac{1}{1+ h^2} \bracket{h \dot{x}_k + h^2 \dot{y}_k  + \xk \parenth{1+ \frac{3h^2}{2}} + \frac{h^3}{2} \yk - \frac{h^3}{2} U_{y_k} - \frac{h^2}{2} U_{x_k} } \label{eq:xkp} \, ,\\
+        \ykp &= h \dot{y}_k + h \xk - h \xkp + \yk + \frac{h^2 \yk}{2} - \frac{h^2 }{2} U_{y_k} \label{eq:ykp} \, ,\\
+        \dot{x}_{k+1} &= \dot{x}_k - 2 \yk + 2 \ykp + \frac{h}{2} \parenth{\xkp + \xk} - \frac{h}{2} U_{\xkp} - \frac{h}{2} U_{\xk} + h u_x \label{eq:xdotkp}\, ,\\
+        \dot{y}_{k+1} &= \dot{y}_{k} + 2 \xk - 2 \xkp + \frac{h}{2} \parenth{\ykp + \yk} - \frac{h}{2} U_{\ykp} - \frac{h}{2} U_{\yk} + h u_y \label{eq:ydotkp} \, .
+\end{align}
+\end{subequations}
+The discrete equations of motion are given in the Lagrangian form after applying the discrete fiber derivative from~\cref{eq:discrete_legendre} as \( p_{\xk} = \dot{x}_k - \yk \) and \( p_{\yk} = \dot{y}_k + \xk \).
+The state is defined as \( \vecbf{x}_k = \begin{bmatrix} \xk & \yk & \dot{x}_k & \dot{y}_k \end{bmatrix}^T\) and the control input is \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \).
+This results in a discrete update map \( f_k: \vecbf{x}_k \to \vecbf{x}_{k+1} \) which preserves the same properties of the continuous-time dynamics in~\cref{eqn:cont_dyn} such as invariants, symplecticity, and the configuration manifold.
+The discrete potential gradients are given by
+\begin{subequations}\label{eq:discrete_potential_grad}
+\begin{align}
+        U_{\xk} &= \frac{\parenth{1 -\mu} \parenth{\xk + \mu}}{\distonek^3} + \frac{ \mu \parenth{\xk -1 + \mu}}{\disttwok^3} \label{eq:Uxk} \, ,\\
+        U_{\yk} &= \frac{\parenth{1 -\mu} \yk}{\disttwok^3} + \frac{ \mu \yk}{\disttwok^3} \label{eq:Uyk} \, ,\\
+        U_{\xkp} &= \frac{\parenth{1 -\mu} \parenth{\xkp + \mu}}{\distonekp^3} + \frac{ \mu \parenth{\xkp -1 + \mu}}{\disttwokp^3} \label{eq:Uxkp}\, ,\\
+        U_{\ykp} &= \frac{\parenth{1 -\mu} \ykp}{\distonekp^3} + \frac{ \mu \ykp}{\disttwokp^3} \label{eq:Uykp} \, .
+\end{align}     
+\end{subequations}
+The distances to each primary are defined as
+\begin{subequations}\label{eq:discrete_distance}
+\begin{align}
+        \distonek &= \sqrt{\left( \xk + \mu\right)^2 + \yk^2} \label{eq:distonek}\, ,\\
+        \disttwok &= \sqrt{\left( \xk - 1 + \mu\right)^2 + \yk^2} \label{eq:disttwok}\, ,\\
+        \distonekp &= \sqrt{\left( \xkp + \mu\right)^2 + \ykp^2} \label{eq:distonekp}\, ,\\
+        \disttwokp &= \sqrt{\left( \xkp - 1 + \mu\right)^2 + \ykp^2} \label{eq:disttwokp} \, .
+\end{align}
+\end{subequations}
+Care must be taken during the implementation of~\cref{eq:discrete_eoms}.
+As~\cref{eq:discrete_potential_grad,eq:discrete_distance} are defined at both step \( k \) and \( k+1 \) they must be evaluated at both time instances.
+\Cref{eq:discrete_eoms} is implemented by first defining an initial state \( \vecbf{x}_k \) and control \( \vecbf{u}_k \).
+The distances and gravitational potential at step \( k \) are evaluated from~\cref{eq:distonek,eq:disttwok,eq:Uxk,eq:Uyk}.
+The discrete update steps in~\cref{eq:xkp,eq:ykp} are evaluated to generate \( \xkp \) and \( \ykp\).
+Next, the distances and gravitational potential at step \( k+1 \) are evaluated from~\cref{eq:distonekp,eq:disttwokp,eq:Uxkp,eq:Uykp}. 
+Finally, the update steps in~\cref{eq:xdotkp,eq:ydotkp} are evaluated.
+This results in the complete discrete update map \( \vecbf{x}_k \to \vecbf{x}_{k+1} \) given \( \vecbf{u}_k \).
+\subsection{Numerical Example}\label{sec:variational_example}
+\begin{figure} 
+        \centering 
+        \begin{subfigure}[h]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{trajectory} 
+                \caption{Trajectory comparison} \label{fig:compare_trajectory} 
+        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+%       \begin{subfigure}[htbp]{0.3\textwidth} 
+%               \includegraphics[width=\textwidth]{components} 
+%               \caption{State \( \parenth{x , y, \dot{x}, \dot{y}}\)} \label{fig:compare_components} 
+%       \end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+        \begin{subfigure}[htbp]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{energy} 
+                \caption{Jacobi Integral comparison} \label{fig:compare_energy} 
+        \end{subfigure} 
+        \caption{Integrator Comparison}
+        \label{fig:integrator_compare} 
+\end{figure}
+A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
+A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
+The variational integrator uses a step size of \SI{47.22}{\second} while the Runge-Kutta method uses a variable step size implemented via ODE45 in Matlab.
+\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference frame for both integration schemes.
+Both integration schemes result in trajectories that are initially nearly identical.
+The discrete equations of motion are an accurate approximation for the continuous dynamics as they closely match the solution of ODE45 over the initial portion of the simulation.
+However, as time progresses the trajectories begin to diverge due to the differences in system energy.
+\Cref{fig:compare_energy} shows the evolution of the Jacobi integral.
+The variational integrator exhibits a bounded behavior about the initial energy with a mean variation of \num{4.2522e-20}.
+However, the conventional Runge-Kutta method demonstrates a clear energy drift of \num{4.2814e-8}. 
+Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques.
 

### 2017-06-30 Response.tex
Add reference to derivation of three body problem
Reviewer 1 comment 10

@@
 
+\item 
+    \begin{itshape}
 "Application of the Euler-Lagrange equations results in following equations of motion defined in the rotating reference frame"  Add a reference after this statement for those who want to see the derivation.
 One of the given references probably does already does this, so just pick one
+\end{itshape}
+
+An additional comment has been included in the problem formulation section with a reference to a source of the detailed derivation.
+\begin{correction}
+Following a straightforward application of the Euler-Lagrange equations, a more detailed derivation is provided in~\cite{szebehely1967}, results in the following equations of motion defined in the rotating reference frame
+\end{correction}
 

### 2017-06-30 manuscript.tex
Add reference to derivation of three body problem
Reviewer 1 comment 10

@@
 \end{equation}
-Following a straightforward application of the Euler-Lagrange equations results in the following equations of motion defined in the rotating reference frame
+Following a straightforward application of the Euler-Lagrange equations, a more detailed derivation is provided in~\cite{szebehely1967}, results in the following equations of motion defined in the rotating reference frame
 \begin{equation}\label{eqn:cont_dyn}

### 2017-06-30 Response.tex
Mention earlier that control is bounded
Reviewer 1 Comment 11

@@
 
+\item 
+    \begin{itshape}
 It appears from equation 5 and your control input definition that you control is unbounded?
@@
 5) that bounds will be added to the control.
+\end{itshape}
+
+An additional comment has been included mentioning that the control is assumed to be bounded in magnitude.
+
+\begin{correction}
+\ldots and the control inputs is defined as \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \) and assumed to be continuously variable but bounded in magnitude, i.e.
+\( \vecbf{u}^T \vecbf{u} \leq u_{max}^2 \).
+\end{correction}
 

### 2017-07-03 manuscript.tex
Fix citation to equation that no longer exists
Remove more material
Condense discussion about invariant manifolds and variational
integrators

@@
 In the vicinity of \( m_1\) or \(m_2\) there exists a potential well. 
-As the energy level increases there are five critical points of the effective potential~\cref{eq:eff_pot} where the slope is zero.
+As the energy level increases there are five critical points of the effective potential where the slope is zero.
 Three collinear saddle points on the \(\hat{e}_1\) axis and two equilateral points.
@@
 \end{figure}
-%\subsection{Periodic Orbits and Invariant Manifolds}
-%Used to generate a Periodic orbit about a lagrange point
-%Periodic orbits about the lagrange points
-%\subsection{\Poincare Section}
-%Intersection of states with a lower dimensional plane
-%Introduce invariant manifolds
-\subsubsection{Invariant Manifolds and \Poincare Map} 
-% discuss usefulness and basic generation of invariant manifolds
-
-%Planar problem has four dimensional phase space
-%associated with periodic solutions about lagrange equilibrium points
-%tubes, invariant manifolds are four dimenionsal surface which serve to separate classes of trajectories
-%Localized linear analysis, Monodromy matrix, is used to study and compute invariant manifold
-
-% intro to invariant manifolds
-Dynamics systems theory has been applied to design control-free maneuvers in the restricted three-body problem~\cite{koon2011}.
+
+\subsubsection{Invariant Manifolds and \Poincare Map}\label{sec:invariant_manifold}
+Dynamics systems theory has been applied to the design of control-free manuevers in the restricted three-body problem~\cite{koon2011}.
 As previously introduced in~\Cref{sec:jacobi}, there exist five equilibrium points in the equations of motion for the three-body problem.
@@
 These invariant manifolds only exist as a result of the dynamic formulation of the three-body problem in a rotating reference frame. 
-
-\subsection{Invariant Manifolds}\label{sec:invariant_manifold}
-% technical definition of manifold and some equations
 Invariant manifolds serve as a higher dimensional generalization of the concept of seperatrices from linear systems as applied to the case of nonlinear systems.
-To compute these manifolds first one must determine an appropriate periodic orbit. 
-The process of differential correction and numerical continuation are used to determine a periodic orbit with a desired energy level. 
-Let the trajectories of the equations of motion defined in~\cref{eqn:cont_dyn} be defined by the flow map \( \phi(t;\vecbf{x}_0) \) originating from \( \vecbf{x}(t_0) = \vecbf{x}_0 \) and we define a periodic orbit as \( \gamma\).
-The invariant manifolds are defined as : 
-\begin{align*}
-        W^s ( \vecbf{x}_0 ) &= \braces{ \vecbf{x} : \phi(t;\vecbf{x}_0) \to \gamma \text{ as } t \to \infty} , \\
-        W^u ( \vecbf{x}_0 ) &= \braces{ \vecbf{x} : \phi(t;\vecbf{x}_0) \to \gamma \text{ as } t \to -\infty} .
-\end{align*}
-The stable manifold, \( W^s(\vecbf{x}_0) \), is the set of trajectories which asymptotically approach the periodic orbit.
-Similarly, the unstable manifold, \( W^u(\vecbf{x}_0) \), is the set of trajectories which asymptotically depart the periodic orbit.
-For the nonlinear system, defined by~\cref{eqn:cont_dyn} or~\cref{eq:discrete_eoms}, we can determine the state transition matrix \( \Phi \) along the periodic orbit, from \( 0 \to T\), by numerical integration of the variational equations.
-In the neighborhood of the periodic orbit \( \gamma \), the tangent spaces to the stable and unstable manifolds are provided by the generalized eigenspaces \( E^s \) and \( E^u \) of the linearization \( A = \D f(\vecbf{x}_0) \), where \( \vecbf{x}_0 \) is on the periodic orbit.
-The eigenvectors of the monodromy matrix, \( M = \Phi(T) \), serve to locally approximate the directions of the stable and unstable manifolds. 
-A perturbation in the direction of the unstable eigenvector, with \( \norm{\lambda} > 1 \), is used to generate an initial condition on the unstable manifold. 
-Similarly, a perturbation in the direction of the stable eigenvector, with \( \norm{\lambda} < 1 \), is used to generate an initial condition on the stable manifold. 
-Each periodic orbit has two stable and unstable manifold sets associated with both a positive or negative perturbation along the respective eigenvector.
-This technique of determining the invariant manifolds is graphically presented in~\cref{fig:invariant_manifold}.
-\begin{figure}
-        \centering
-        \includegraphics[width=0.5\textwidth]{invariant_manifold}
-        \caption{Stable and unstable manifold of a periodic orbit\label{fig:invariant_manifold}}
-\end{figure}
-
-% introduce \Poincare section theory and math
+
 \Poincare maps are a useful tool in the analysis of the flow near periodic orbits in the three-body problem.
@@
 
-\subsection{Control-free Orbital Transfers}\label{sec:control-free_transfers}
-% intersect invariant manifolds on poincare section
-% find intersection between stable and unstable manifolds
-% define the sections used in Koon2011 (placed strategically near heteroclinic points) - rotation of manifolds
+% drawbacks of this approach on how we seek to improve upon it
 Combining invariant manifolds and an appropriate \Poincare section provides a conceptually simple manner to determine trajectories which connect wide regions of the phase space.
-\Cref{fig:manifold_transfer_example} demonstrates how a \Poincare map may be used to determine a connection between periodic orbits.
-The initial orbit is a planar periodic orbit about \( L_2 \) in the Earth-Moon system while the final orbit is a periodic orbit about \( L_1 \).
-Both periodic orbits have the same Jacobi energy constant and as a result it is possible to transfer between the orbits without any energy change.
-\Cref{fig:manifolds} shows the unstable manifold of the initial orbit and the stable manifold of the desired orbit. 
-Both invariant manifolds are globalized and propogated to the first intersection with the \Poincare surface of section located at the position of the Moon, \( x = 1 - \mu \).
-The \Poincare section in~\cref{fig:manifold_transfer_example} is defined as
-\begin{align*}
-        \Sigma = \braces{ ( y, \dot{y}) \, | \, x = 1- \mu, y > 0} , 
-\end{align*}
-and located to allow an intersection between both invariant manifolds.
-The configuration space of the invariant manifold is \( \vecbf{x} \in \R^4 \).
-The surface of section, and the resulting intersection of the four-dimensional space with the hyperplane \( \Sigma \), reduces the resulting intersection to a surface in \( \R^3\).
-Since the two orbits share the same Jacobi constant, the intersection phase space is further reduced to \( \R^2\).
-The intersection of the invariant manifolds on the \Poincare section is central to determining transfer connections between the initial and final orbits.
-
-% describe this specific poincare section and why it's placed and the axes
-The position, \( y \), and velocity, \( \dot{y} \), at the \Poincare section are plotted in~\cref{fig:poincare}.
-The surface of section is placed at a strategic location in order to allow for a cross section of the flow within the three-dimensional energy surface. 
-In a frame aligned with the rotating frame, \( ( \hat{e}_1, \hat{e}_2 ) \), but centered on the equilibrium point \( L_1 \) or \( L_2\), the unstable manifolds are locally moving in forward time in the second and forth quadrant.
-Similarly, the stable manifold branches are locally moving in backward time in the first and third quadrants as illustrated in~\cref{fig:invariant_manifold}. 
-As a result, the location of the section in~\cref{fig:poincare} is chosen to intersect both the stable manifolds of \( L_1 \) periodic orbits as well as unstable manifolds of \( L_2 \) periodic orbits.
-Additional \Poincare sections are located at other heteroclinic points in the phase space to enable design throughout the phase space~\cite{koon2011}.
-In this way, it is possible to design trajectories that intersect the invariant manifolds and connect widely separated regions of the state space. 
-This technique has been heavily investigated in~\cite{koon2011}, applied to operational missions in~\cite{koon1999}, and applied to potential multi-moon orbiter missions in~\cite{tanaka2011}.
-In addition, much recent work has focused on potential return missions to the Moon~\cite{zanzottera2012,campagnola2012,mingotti2011,ozimek2010a,mingotti2009}.
-\begin{figure}
-     \centering
-        \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\columnwidth]{U2_Manifolds}
-                \caption{Invariant manifolds }
-                \label{fig:manifolds}
-        \end{subfigure}%
-        ~
-        \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\columnwidth]{U2_poincare}
-                \caption{\(U_2\) \Poincare section}
-                \label{fig:poincare}
-        \end{subfigure}
-        \caption{Illustration of process to identify transfer trajectories}
-        \label{fig:manifold_transfer_example}
-\end{figure}
-
-% drawbacks of this approach on how we seek to improve upon it
 However, the results previously developed are highly case specific and difficult to generalize to arbitrary transfers.
@@
 
-\section{Variational Integrator}\label{sec:discrete_var}
-% description of variational integrator development
+\subsection{Variational Integrator}\label{sec:discrete_var}
 Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invaraint properties and symplecticity.
-Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations~\cite{marsden2001}.
+Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations.
 As a result, integrators developed in this manner have the desirable properties that they are symplectic and momentum preserving.
 In addition, they exhibit improved energy behavior over long integration periods.
-We present a short background on the variational principle for mechanical systems. 
-We then develop a discrete approximation of the action integral and construct a variational integrator for the PCRTBP.
-
-\subsection{Variational Principle}
-Consider a continuous mechanical system described by the Lagrangian, \( L( q, \dot{q} ) \), for the generalized position, \( q\), and velocity, \( \dot{q} \).
-In the standard approach of variational mechanics the action integral is formed by integrating the continuous Lagrangian along a path \( q(t) \) that the system follows from time \( t = 0 \) to \( t = T \)~\cite{greenwood1988}.
-In the continuous time the action integral is defined as
-\begin{align}\label{eq:action_integral}
-        S = \int_{0}^T L\left( q, \dot{q}\right) \, dt \, .
-\end{align}
-Hamilton's principle states that the actual path followed by a holonomic system results in a stationary action integral with respect to path variations for fixed endpoints.
-Taking the variation of~\cref{eq:action_integral} gives
-\begin{align}\label{eq:var_principle}
-        \delta S %&= \int_{0}^T \deriv{L}{q} \delta q + \deriv{L}{\dot{q}} \delta \dot{q} \, dt \\
-                %&= \int_{0}^T \deriv{L}{q} \delta q - \frac{d}{dt} \left( \deriv{L}{\dot{q}}\right) \delta q \, dt - \left. \left[ \deriv{L}{\dot{q}} \delta q\right] \right|_0^T \\
-        &= \int_{0}^T \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}}     \right) \, dt \, ,
-\end{align}
-where we have used integration by parts and the conditions \( \delta q(0) = 0 \) and \( \delta q(T) = 0\).
-For Hamilton's principle to be valid for all admissible variations \( \delta q \), the integrand of~\cref{eq:var_principle} must be zero for all \( t\), giving the continuous Euler-Lagrange Equations~\cite{lanczos1970}.
-\begin{align}\label{eq:euler-lagrange}
-        0 = \deriv{L}{q} - \frac{d}{dt} \left( \deriv{L}{\dot{q}} \right) \, .
-\end{align}
-Hamilton's equations are derivable through the use of the Legendre transformation which is a mapping \( \left( q, \dot{q},t\right) \rightarrow \left(q, p, t \right) \) where \( p_i\) is the generalized momenta.
-\begin{align}\label{eq:legendre_transform}
-        p_i = \deriv{L}{\dot{q}_i} \, .
-\end{align}
-In the continuous time case the Hamiltonian is defined as
-\begin{align}\label{eq:hamiltonian}
-        H &= \sum_{i = 1}^n p_i \dot{q}_i - L \left( q,\dot{q}, t \right) \, .
-\end{align}
-Applying~\cref{eq:legendre_transform} and taking the variation of~\cref{eq:hamiltonian} allows us to derive the equations of motion in Hamiltonian form
-\begin{subequations}\label{eq:hamilton_eq}
-\begin{align}
-        \dot{q}_i &= \deriv{H}{p_i} \, ,\\
-        \dot{p}_i &= - \deriv{H}{q_i} \, , \\
-        \deriv{L}{t} &= -\deriv{H}{t} \, .
-\end{align}
-\end{subequations}
-Both~\cref{eq:euler-lagrange,eq:hamilton_eq} result in equations of motion for the mechanical system and are equivalent via the Legendre transform.
-\Cref{eq:euler-lagrange} results in \( n \) second order differential equations while~\cref{eq:hamilton_eq} results in \( 2n \) first order differential equations.
-\subsection{Discrete Variational Mechanics}
-A discrete analogue of Hamilton's principle and the action integral is formed.
-Rather than taking a position, \( q \), and velocity, \( \dot{q} \), consider two positions \( q_0 \) and \( q_1 \) and a fixed time step \( h \in \R \).
-The two positions are points on the curve \( q(t) \) such that \( q_0 \approx q(0) \) and \( q_1 \approx q(h) \).
-A discrete time Lagrangian \( L_d( q_0, q_1) \) is formed which approximates the action integral between \( q_0 \) and \( q_1 \) as 
-\begin{align}\label{eq:discrete_lagrangian_gen}
-        L_d\left( q_0 , q_1 \right) \approx \int_{0}^{h} L \left( q , \dot{q} \right) \, dt \, .
-\end{align}
-Since~\cref{eq:discrete_lagrangian_gen} is calculated as a numerical integral, an appropriate quadrature rule is required.
-There are multiple possible methods one can use to approximate the integral in~\cref{eq:discrete_lagrangian_gen}.
-An appropriate approximation rule is determined based on the ease of implementation and accuracy desired.
-\begin{table}[htbp]
-\caption{Selected Quadrature Rules\label{tab:quadrature}}
-\begin{center}
-\begin{tabular}{l|l}Rectangle & \( L_d(q_0,q_1) =L(q_0,\frac{q_1-q_0}{h}) h \)  \\ \hline
-Midpoint & \( L_d(q_0,q_1) = L(\frac{q_0 + q_1}{2},\frac{q_1 - q_0}{h}) h \) \\ \hline
-Trapezoidal & \( L_d(q_0, q_1) = \frac{1}{2} \left[ L(q_0, \frac{q_1 - q_0}{h} ) + L(q_1, \frac{q_1 - q_0 }{h} )\right] h \)
-\end{tabular} 
-\end{center}
-\end{table}
-\Cref{tab:quadrature} shows several possible approximation rules that are typically applied.
-The rectangle rule is a first order accurate method and offers a straightforward implementation.
-The midpoint and trapezoidal rules are both second order accurate methods. 
-However, the midpoint rule results in an implicit form which adds further complexity to the equations of motion.
-%In this work, the trapezoidal approximation is applied to the PCRTBP.
-
-Once an appropriate discrete Lagrangian is formed a discrete action sum is formed as the discrete analogue of~\cref{eq:action_integral}
-\begin{align}\label{eq:action_sum}
-        S_d = \sum_{k=0}^{N-1} L_d(q_k, q_{k+1}) \, .
-\end{align}
-Once again a discrete version of Hamilton's principle is applied to~\cref{eq:action_sum}.
-This yields a discrete time counterpart to~\cref{eq:var_principle} as
-\begin{align}\label{eq:dis_var_principle}
-        \delta S_d %&= \sum_{k=0}^{N-1} \deriv{L_d(q_k, q_{k+1})}{q_k} \delta q_k + \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \delta q_{k+1} \\
-        &= \sum_{k=1}^{N-1} \bracket{ \deriv{L_d(q_k, q_{k+1})}{q_k} + \deriv{L_d(q_{k-1}, q_k)}{q_{k+1}}} \delta q_k \, .
-\end{align}
-
-For the discrete action sum to be stationary with respect to all admissible path variations, with fixed endpoints, the discrete Euler-Lagrange equations must be satisfied for \( k = 1, \cdots, N-1 \) resulting in
-\begin{align}\label{eq:discrete_euler-lagrange}
-        0 = \deriv{L_d(q_k, q_{k+1})}{q_k} + \deriv{L_d(q_{k-1}, q_k)}{q_{k+1}} \, .
-\end{align}
-A discrete version of the Legendre transformation, referred to as a discrete fiber derivative, results in the equivalent Hamiltonian form expression.
-The discrete fiber derivative is given as 
-\begin{subequations}\label{eq:discrete_legendre}
-\begin{align}
-        p_k &= \deriv{L_d(q_{k-1},q_k)}{q_{k+1}} = - \deriv{L_d(q_k, q_{k+1})}{q_k} \, ,\\
-        p_{k+1} &= \deriv{L_d(q_k, q_{k+1})}{q_{k+1}} \, .
-\end{align}
-\end{subequations}
-This yields a discrete Hamiltonian map \( (q_k, p_k) \to (q_{k+1}, p_{k+1}) \).
-A more extensive development of variational integrators can be found in~\cite{marsden2001}.
-
-\subsection{Discrete Equations of Motion}
+A thorough discussion of variational integrators is provided in~\cite{west2004,marsden2001}.
+We use this approach to construct a variational integrator for the PCRTBP with the inclusion of low-thrust propulsion.
+
 The discrete equations of motion for the PCRTBP are derived by choosing an appropriate quadrature rule to discretize the Lagrangian in~\cref{eq:lagrangian}. 
@@
 Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques.
-
 \section{Optimal Control Formulation for Reachability Set}

### 2017-07-03 manuscript.tex
Add lagrangian back into paper

@@
 \end{equation}
+In the rotating reference frame the Lagrangian is given by
+\begin{equation}\label{eq:lagrangian}
+    L = \frac{1}{2} \parenth{\parenth{\dot{x} - y}^2 + \parenth{\dot{y} + x}^2} + \frac{1 - \mu}{r_1} + \frac{\mu}{r_2},
+\end{equation}
+where the distances \( r_1, r_2 \in \R \) define the distance from the spacecraft to each primary and are defined as
+\begin{align}
+    r_1 &= \sqrt{\parenth{x + \mu}^2 + y^2} , \\
+    r_2 &= \sqrt{\parenth{x -1 + \mu}^2 + y^2} .
+\end{align}
 Following a straightforward application of the Euler-Lagrange equations, a more detailed derivation is provided in~\cite{szebehely1967}, results in the following equations of motion defined in the rotating reference frame

### 2017-07-03 manuscript.tex
Change title of section

@@
 
-\subsection{Variational Integrator}\label{sec:discrete_var}
+\subsection{Variational Integrator for PCRTBP}\label{sec:discrete_var}
 Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invaraint properties and symplecticity.

### 2017-07-03 manuscript.tex
Remove some equation references that don't matter

@@
 The discrete equations of motion for the PCRTBP are derived by choosing an appropriate quadrature rule to discretize the Lagrangian in~\cref{eq:lagrangian}.
-In this work, the trapezoidal approximation is applied.
-The trapezoid rule allows for an explicit second order accurate approximation.
+In this work, we apply the trapezoidal rule, which allows for an explicit second order accurate approximation of the continuous Lagrangian.
 The discrete Lagrangian is given by
@@
 Applying a discrete version of the Lagrange-d'Alembert principle allows for inclusion of an external control force on the system~\cite{marsden2001}.
-Using~\cref{eq:discrete_legendre,eq:discrete_lagrangian} and some manipulation, the equations of motion are given by
+Using~\cref{eq:discrete_lagrangian} and some manipulation, gives the discrete equations of motion as
 \begin{subequations}\label{eq:discrete_eoms}
@@
 \end{subequations}
-The discrete equations of motion are given in the Lagrangian form after applying the discrete fiber derivative from~\cref{eq:discrete_legendre} as \( p_{\xk} = \dot{x}_k - \yk \) and \( p_{\yk} = \dot{y}_k + \xk \).
+The discrete equations of motion are given in the Lagrangian form after applying the discrete fiber derivative as \( p_{\xk} = \dot{x}_k - \yk \) and \( p_{\yk} = \dot{y}_k + \xk \).
 The state is defined as \( \vecbf{x}_k = \begin{bmatrix} \xk & \yk & \dot{x}_k & \dot{y}_k \end{bmatrix}^T\) and the control input is \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \).

### 2017-07-05 Response.tex
Modify all the figures to use axis equal
All of the figures with position states now use axis equal

Reviewer 1 Comment 12
Reduce discussion about variational integrators
Reviewer 1 comment 13

@@
 
+\item 
+    \begin{itshape}
 Figure 1 should probably have "axis equal" applied to it.
@@
 If not, please regenerate those plots.
+\end{itshape}
 
+All of the figures showing position states have been modified to use \texttt{axis equal}.
+
+\item 
+    \begin{itshape}
 Section 3.1 has been discussed in many references, the reader may want to direct them to this material.
 (Choose one of Marsden or West's papers for example.)
+\end{itshape}
+
+The discussion on the variational integrators has been reduced. 
+Much of the background material on Lagrangian mechanics is removed and an additional paragraph is included to summarize the key ideas of variational integrators. 
+The additional paragraph is shown below.
+
+\begin{correction}
+Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invaraint properties and symplecticity.
+Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations.
+As a result, integrators developed in this manner have the desirable properties that they are symplectic and momentum preserving.
+In addition, they exhibit improved energy behavior over long integration periods.
+A thorough discussion of variational integrators is provided in~\cite{west2004,marsden2001}.
+We use this approach to construct a variational integrator for the PCRTBP with the inclusion of low-thrust propulsion.
+
+Consider the autonomous mechanical system described by the Lagrangian, \( L(q, \dot{q} ) \), for the generalized coordinates, \( q\) and velocities \( \dot{q} \).
+The integration of the continuous Lagrangian along a path, \( q (t) \), followed by the system must satisfy Hamilton's principle, which results in the well-known Euler-Lagrange equations~\cite{lanczos1970}.
+In the discrete time scenario, the continuous path \( q(t), \dot{q}(t) \), is replaced by a finite difference approximation over a fixed time step, e.g.
+\( q_0, \frac{q_1 - q_0}{h} \), which converges to the true velocity as \( h \) tends towards zero.
+In this manner, we define a discrete Lagrangian, \( L_d (q_0, q_1, h) \) which approximates the integral of the true Lagrangian over the time interval \( h \) between \( q_0 \) and \( q_1\)~\cite{marsden2001}.
+\end{correction}
 
@@
 It is very likely that the integration accuracy will introduce some errors in the accuracy of the partial derivatives and in the states.
-
 Section 3.2 - please cite a reference as to where the selected quadrature rules presented in Table 1 have been derived.

### 2017-07-05 manuscript.tex
Reduce discussion about variational integrators
Reviewer 1 comment 13

@@
 
-The discrete equations of motion for the PCRTBP are derived by choosing an appropriate quadrature rule to discretize the Lagrangian in~\cref{eq:lagrangian}.
-In this work, we apply the trapezoidal rule, which allows for an explicit second order accurate approximation of the continuous Lagrangian.
-The discrete Lagrangian is given by
+Consider the autonomous mechanical system described by the Lagrangian, \( L(q, \dot{q} ) \), for the generalized coordinates, \( q\) and velocities \( \dot{q} \).
+The integration of the continuous Lagrangian along a path, \( q (t) \), followed by the system must satisfy Hamilton's principle, which results in the well-known Euler-Lagrange equations~\cite{lanczos1970}.
+In the discrete time scenario, the continuous path \( q(t), \dot{q}(t) \), is replaced by a finite difference approximation over a fixed time step, e.g.
+\( q_0, \frac{q_1 - q_0}{h} \), which converges to the true velocity as \( h \) tends towards zero.
+In this manner, we define a discrete Lagrangian, \( L_d (q_0, q_1, h) \) which approximates the integral of the true Lagrangian over the time interval \( h \) between \( q_0 \) and \( q_1\)~\cite{marsden2001}.
+
+The discrete equations of motion for the PCRTBP are derived by choosing an appropriate quadrature rule to discretize the continuous Lagrangian in~\cref{eq:lagrangian}~\cite{marsden2001}.
+In this work, we apply the trapezoidal rule to approximate the action integral over a fixed time interval.
+The trapezoidal rule allows for a second order accurate approximation and alleviates the difficulties in applying implicit quadrature schemes.
+The discrete Lagrangian for the PCRTBP is given by
 \begin{align}\label{eq:discrete_lagrangian}

### 2017-07-07 Response.tex
Address some of the comments on variational integrators
Reviewer 1 Comments 14, 15

@@
 
+\item 
+    \begin{itshape}
 The second-order trapezoidal rule will give significantly poorer state vector accuracy with respect to a high-precision RK integrator, even if the state is guaranteed to lie on an energy surface.
@@
 It is very likely that the integration accuracy will introduce some errors in the accuracy of the partial derivatives and in the states.
+\end{itshape}
+
+We chose the trapezoidal rule as it provides a second order quadrature approximation.
+In addition, it provides an explicit rule that eases the complexity in using an implicit solution method.
+
+\item 
+    \begin{itshape}
 Section 3.2 - please cite a reference as to where the selected quadrature rules presented in Table 1 have been derived.
+\end{itshape}
+
+In the interest of conciseness and to avoid repeating material that has already been covered in the literature, the discussion on variational integrators has been substantially reduced.
+As a result, this table has been removed from the manuscript.
 
@@
 
+\item 
+    \begin{itshape}
 Comparison like that of Figure 3.4 are very common in the well-known literature concerning variational integrators.

### 2017-07-07 Response.tex
Start updating the figures for the variational integrator comparison
Reviewer 1 Comment 19

@@
 http://lagrange.mechse.illinois.edu/pubs/LeMaOrWe2004/LeMaOrWe2004.pdf
+\end{itshape}
+
+The comparison plot has been modified to follow the precedent in the literature.
+The plot shows the deviation of the Jacobi constant as the time step is varied. 
+In addition, an additional plot has been added, and some related discussion which shows the decrease in computation cost through the use of the variational integrator.
 

### 2017-07-07 manuscript.tex

@@
 
-\section*{Appendix B: Gauss Jordan Elimination}\label{sec:costate_gauss_jordan}
-The costate equations of motion are given by~\cref{eq:costate_eom} and repeated here as
-\begin{align}\label{eq:costate_eom_transpose}
-        \begin{bmatrix} 
-                \fonex & \ftwox & \fthreex & \ffourx \\
-                \foney & \ftwoy & \fthreey & \ffoury \\
-                \fonexd & \ftwoxd & \fthreexd & \ffourxd \\
-                \foneyd & \ftwoyd & \fthreeyd & \ffouryd
-        \end{bmatrix}
-        \begin{bmatrix} \lambda_{\xkp} \\ \lambda_{\ykp} \\ \lambda_{\xdotkp} \\ \lambda_{\ydotkp} \end{bmatrix}
-        =
-        \begin{bmatrix} \lambda_{\xk} \\ \lambda_{\yk} \\ \lambda_{\xdotk} \\ \lambda_{\ydotk} \end{bmatrix} \, .
-\end{align}
-To determine the discrete update map \( \lambda_k \to \lambda_{k+1}\) the inverse of the Jacobian matrix is required.
-In order to avoid the need of an explicit inversion a Gauss Jordan method is implemented.
-To begin, several terms are defined which are required to carry out the row operations and are defined as
-\begin{subequations}\label{eq:ref_scale}
-\begin{align}
-        a &= -\frac{\foney}{\fonex} \, ,\\
-        b &= -\frac{\fonexd}{\fonex} \, , \\
-        c &= -\frac{\foneyd}{\fonex} \, ,\\
-        e &= -\frac{\ftwoxd + b \ftwox}{\ftwoy + a \ftwox} \, ,\\
-        f &= -\frac{\ftwoyd + c \ftwox}{\ftwoy + a \ftwox} \, ,\\
-        g &= -\frac{\fthreeyd + c \fthreex + f \parenth{\fthreey + a \fthreex}}{\fthreexd + b \fthreex + e \parenth{\fthreey + a \fthreex}}\, .
-\end{align}
-\end{subequations}
-\Cref{eq:costate_eom_transpose} is transformed to row echelon form using elementary row operations and is defined as
-\begin{align}\label{eq:costate_ref}
-        \begin{bmatrix} 
-                \alpha_{11} & \alpha_{12} & \alpha_{13} & \alpha_{14} \\
-                0 & \alpha_{22} & \alpha_{23} & \alpha_{24} \\
-                0 & 0 & \alpha_{33} & \alpha_{34} \\
-                0 & 0 & 0 & \alpha_{44}
-        \end{bmatrix}
-        \begin{bmatrix} \lambda_{\xkp} \\ \lambda_{\ykp} \\ \lambda_{\xdotkp} \\ \lambda_{\ydotkp} \end{bmatrix}
-        =
-        \begin{bmatrix} \beta_1 \\ \beta_2 \\ \beta_3 \\ \beta_4 \end{bmatrix} \, ,
-\end{align}
-where the terms \( \alpha_{ij} \) and \( \beta_{i} \) are defined as follows
-\begin{subequations}
-\begin{align}
-        \alpha_{11} &= \fonex \, ,\\
-        \alpha_{12} &= \ftwox \, ,\\
-        \alpha_{13} &= \fthreex \, ,\\
-        \alpha_{14} &= \ffourx \, ,\\
-        \alpha_{22} &= \ftwoy + a \ftwox \, ,\\
-        \alpha_{23} &= \fthreey + a\fthreex \, ,\\
-        \alpha_{24} &= \ffoury + a \ffourx \, ,\\
-        \alpha_{33} &= \fthreexd + b \fthreex + e \parenth{\fthreey + a \fthreex}\, ,\\
-        \alpha_{34} &= \ffourxd + b \ffourx + e \parenth{\ffoury + a \ffourx} \, ,\\
-        \alpha_{44} &= \ffouryd + c \ffourx + f \parenth{\ffoury + a \ffourx} + g \parenth{\ffourxd + b \ffourx + e \parenth{\ffoury + a \ffourx}} \, ,\\
-        \beta_1 &= \lambda_{\xk} \, ,\\
-        \beta_2 &= \lambda_{\yk} + a \lambda_{\xk} \, ,\\
-        \beta_3 &= \lambda_{\xdotk} + b \lambda_{\xk} + e \parenth{\lambda_{\yk} + a \lambda_{\xk}} \, ,\\
-        \beta_4 &= \lambda_{\ydotk} + c \lambda_{\xk} + f \parenth{\lambda_{\yk}+a \lambda_{\xk}} + g \parenth{\lambda_{\xdotk} + b \lambda_{\xk} + e \parenth{\lambda_{\yk} + a \lambda_{\xk}}} \, .
-\end{align}
-\end{subequations}
-Finally, backsubstituion is used to determine explicit equations for the discrete update map \( \vecbf{\lambda}_k \to \vecbf{\lambda}_{k+1} \) which is defined as
-\begin{subequations}\label{eq:costate_update}
-\begin{align}
-        \lambda_{\ydotkp} &= \frac{\beta_4}{\alpha_{44}} \, ,\\
-        \lambda_{\xdotkp} &= \frac{\beta_3}{\alpha_{33}} - \frac{\alpha_{34}}{\alpha_{33}} \lambda_{\ydotkp} \, ,\\
-        \lambda_{\ykp} &= \frac{\beta_2}{\alpha_22} - \frac{\alpha_{33}}{\alpha_{22}}\lambda_{\xdotkp} - \frac{\alpha_{24}}{\alpha_{22}} \lambda_{\ydotkp} \, ,\\
-        \lambda_{\xkp} &= \frac{\beta_1}{\alpha_{11}} - \frac{\alpha_{12}}{\alpha_{11}} \lambda_{\ykp} - \frac{\alpha_{13}}{\alpha_{11}} \lambda_{\xdotkp} - \frac{\alpha_{14}}{\alpha_{11}} \lambda_{\ydotkp} \, .
-\end{align}
-\end{subequations}
-%Once again, these equations exhibit a cascade type relationship. 
-%As as result any errors will compound and tend to accumulate most prevalently in \( \lambda_{\xkp} \).
-
 %\begin{acknowledgements}

### 2017-07-07 Response.tex
Remove appendix B
remove the derivation of the costate variational update

Reviewer 1 Comment 17
Remove tricky monodromy matrix and periodic orbit discussion
Reviewer 1 Comment 18

@@
 
+\item 
+    \begin{itshape}
 I recommend removing appendix B.
 Giving the equations and stating that the Matrix inversion is from Gauss-Jordan elimination is probably sufficient, unless there is some compelling algebraic manipulation that is very unique to the problem.
+\end{itshape}
 
+Appendix B has been removed from the revised manuscript.
+
+\item 
+    \begin{itshape}
 In line 14 (page 12), you define the state transition matrix over one orbital period, and then on line 18 you start referring to it as the monodromy matrix.
@@
 Some of this section can be substantially reduced and other source material could be referenced anyway.
+\end{itshape}
+
+The discussion about the invariant manifolds and \Poincare sections has been reduced.
+This discussion has been removed from the revised manuscript and replaced with a general summary and some relevant citations.
 

### 2017-07-07 Response.tex
Add more detail on variational integrators
Reviewer 1 Comment 20, 21

@@
 
+\item 
+    \begin{itshape}
 It may also be generally worth acknowledging the benefits of Taylor's method within the context of variational integrators:
@@
 Online copy: https://arxiv.org/pdf/1703.06599.pdf 
+\end{itshape}
 
+This manuscript is not really focused on variational integrators but rather the design of transfer trajectories using reachability sets.
+Interested readers can find a wide variety of literature on the relative benefits of all manner of variational integrators in papers focused on that topic specifically.
+
+\item 
+    \begin{itshape}
 "Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques".
 I think you need to add "in which energy conservation is mandatory for characterizing the solution space."
+\end{itshape}
+
+The additional clarification has been added to the manuscript.
 

### 2017-07-07 manuscript.tex
Add more detail on variational integrators
Reviewer 1 Comment 20, 21

@@
 However, the conventional Runge-Kutta method demonstrates a clear energy drift of \num{4.2814e-8}. 
-Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques.
+Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
+
 \section{Optimal Control Formulation for Reachability Set}

### 2017-07-10 manuscript.tex
Update energy deviation plot
Modify plot and discussion of integrator comparison

@@
         \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{energy} 
-                \caption{Jacobi Integral comparison} \label{fig:compare_energy} 
+                \includegraphics[width=\textwidth]{figures/cputimevsE.pdf} 
+                \caption{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}} \label{fig:compare_energy} 
         \end{subfigure} 
@@
 A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
-The variational integrator uses a step size of \SI{47.22}{\second} while the Runge-Kutta method uses a variable step size implemented via ODE45 in Matlab.
+    The variational integrator uses a range of step sizes between \SIrange{4.72}{472}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
+    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
 \Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference frame for both integration schemes.
@@
 However, as time progresses the trajectories begin to diverge due to the differences in system energy.
-\Cref{fig:compare_energy} shows the evolution of the Jacobi integral.
-The variational integrator exhibits a bounded behavior about the initial energy with a mean variation of \num{4.2522e-20}.
-However, the conventional Runge-Kutta method demonstrates a clear energy drift of \num{4.2814e-8}. 
+\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
+For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the convential integration scheme.
 Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.

### 2017-07-11 manuscript.tex
Update trajectory plot in integrator comparison
Update discussion of comparision between conventional and variational integrator
Reviewer 1 Comment 19 and 21

@@
                 \includegraphics[width=\textwidth]{trajectory} 
-                \caption{Trajectory comparison} \label{fig:compare_trajectory} 
+                \caption{Earth-Moon three-body trajectory used for integrator comparision} \label{fig:compare_trajectory} 
         \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
@@
     The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
-\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference frame for both integration schemes.
+\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
 Both integration schemes result in trajectories that are initially nearly identical.
-The discrete equations of motion are an accurate approximation for the continuous dynamics as they closely match the solution of ODE45 over the initial portion of the simulation.
-However, as time progresses the trajectories begin to diverge due to the differences in system energy.
 \Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
 For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the convential integration scheme.
-Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
+Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
 

### 2017-07-11 Response.tex
Include additions to response
Include improved discussion about variational integrator

@@
 The comparison plot has been modified to follow the precedent in the literature.
-The plot shows the deviation of the Jacobi constant as the time step is varied. 
-In addition, an additional plot has been added, and some related discussion which shows the decrease in computation cost through the use of the variational integrator.
-
+The plot has been modified to show the energy deviation as a function of computational time for both the variational integrator and \texttt{ODE45}.
+The modified content is copied below.
+\begin{correction}
+\begin{figure} 
+        \centering 
+        \begin{subfigure}[h]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{trajectory} 
+                \caption{Earth-Moon three-body trajectory used for integrator comparision} \label{fig:compare_trajectory} 
+        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+%       \begin{subfigure}[htbp]{0.3\textwidth} 
+%               \includegraphics[width=\textwidth]{components} 
+%               \caption{State \( \parenth{x , y, \dot{x}, \dot{y}}\)} \label{fig:compare_components} 
+%       \end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+        \begin{subfigure}[htbp]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{figures/cputimevsE.pdf} 
+                \caption{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}} \label{fig:compare_energy} 
+        \end{subfigure} 
+        \caption{Integrator Comparison}
+        \label{fig:integrator_compare} 
+\end{figure}
+A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
+A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
+    The variational integrator uses a range of step sizes between \SIrange{4.72}{472}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
+    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
+\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
+Both integration schemes result in trajectories that are initially nearly identical.
+\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
+For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the convential integration scheme.
+Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
+\end{correction}
 \item 

### 2017-07-11 Response.tex
Add correction from paper

@@
 
-The additional clarification has been added to the manuscript.
+The additional clarification has been added to the manuscript and copied below.
+\begin{correction}
 
+Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
+\end{correction}
+
+\item 
+    \begin{itshape}
 Section 5 - This method, while very useful in its given context is highly dependent on the use of the PCRTB in the framework of the current analysis.
@@
 It can be done, but more subplots, dimensions, and data storing are necessary to view the subspace - save to say, it is a more challenging endeavor that must be acknowledged for general purpose low-thrust mission design.
-
+\end{itshape}
 Figures 10D and 13D - the authors need to present these plots in dimensional coordinates so that they may be reconciled with what existing technology can produce.

### 2017-07-11 Response.tex
Add detail on limitations of our approach to PCRTBP
Reviewer 1 Comment 22

@@
 \end{itshape}
+
+An additional paragraph has been added to the beginning of the section which discusses our reachability based transfer approach.
+This additional content is copied below.
+
+\begin{correction}
+The numerical examples presented in this section are designed in the context of the PCRTBP.
+The dynamic environment has a four dimensional state space and offers a convient integration constant in the form of the Jacobi integral.
+As a result, there are well defined methods to define and exploit \Poincare sections, which result in straightforward two-dimensional subspaces of the system.
+Our approach uses the \Poincare section to approximate the reachability set on this reduced subspace.
+As a result, this approach is more difficult to apply to three-dimensional transfers in the general three body problem.
+\Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
+However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories.
+\end{correction}
+
 Figures 10D and 13D - the authors need to present these plots in dimensional coordinates so that they may be reconciled with what existing technology can produce.

### 2017-07-11 manuscript.tex
Add detail on limitations of our approach to PCRTBP
Reviewer 1 Comment 22

@@
 The addition of continuous low thrust propulsion extends the control free design process developed previously and allows for a greater range of potential transfers with a reduced time of flight.
+
+The numerical examples presented in this section are designed in the context of the PCRTBP.
+The dynamic environment has a four dimensional state space and offers a convient integration constant in the form of the Jacobi integral.
+As a result, there are well defined methods to define and exploit \Poincare sections, which result in straightforward two-dimensional subspaces of the system.
+Our approach uses the \Poincare section to approximate the reachability set on this reduced subspace.
+As a result, this approach is more difficult to apply to three-dimensional transfers in the general three body problem.
+\Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
+However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories.
 

### 2017-07-11 Response.tex
Remove mention of appendix B
Reviewer 1 Comment 17

@@
 Appendix B has been removed from the revised manuscript.
+The manuscript has been modified to mention that Gauss Jordan elimination is used to avoid the use of the matrix inversion.
+
+\begin{correction}
+The derivation of~\cref{eq:costate_eom} is given in Appendix A.
+In addition, the computation of~\cref{eq:costate_eom} requires inversion of the Jacobian matrix.
+This is a computationally expensive operation that is prone to numerical error and instability.
+We use Gauss-Jordan elimination to avoid this inversion in~\cref{eq:costate_eom} and determine an explicit update map \( \vecbf{\lambda}_k \to \vecbf{\lambda}_{k+1} \).
+\end{correction}
 

### 2017-07-11 manuscript.tex
Remove mention of appendix B
Reviewer 1 Comment 17

@@
 This is a computationally expensive operation that is prone to numerical error and instability.
-A method is presented in Appendix B to avoid this inversion and determine an explicit update map \( \vecbf{\lambda}_k \to \vecbf{\lambda}_{k+1} \).
+We use Gauss-Jordan elimination to avoid this inversion in~\cref{eq:costate_eom} and determine an explicit update map \( \vecbf{\lambda}_k \to \vecbf{\lambda}_{k+1} \).
 

### 2017-07-11 Response.tex

@@
 
+\item 
+    \begin{itshape}
 Figure 2 - x-axis has no label but I think it's nondimensional time, please indicate.  Label all axes even if they are nondimensional units.
+\end{itshape}
+
+The plot has been removed and replaced with another comparision between the variational integrator and \texttt{ODE45}.
 

### 2017-07-12 manuscript.tex
Add some additional detail on the numerical examples

@@
 The reachability set on the \Poincare section allows for a straightforward method of determining transfer opportunities.
+The first example is a transfer from a periodic orbit of the \( L_1 \) Lagrange point to a fixed orbit of the moon.
+This example uses a single iteration of the reachable set computation in the design of the transfer.
+The second example is a transfer from geostationary orbit of the Earth to a period orbit of \( L_1 \). 
+This examples demonstrates the ability to extend the reachability process to multiple iterations, to allow for a much larger and more general transfer.
+With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare~sections.
 

### 2017-07-12 manuscript.tex
Remove explanation of periodic orbits in 3BP
Reviewer 1 Comment 27

@@
 \end{figure}
-\Cref{fig:moon_orbit} shows that the target set remain in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
+\Cref{fig:moon_orbit} shows that the target set remains in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
 This type of orbit would be useful for a variety of mission scenarios.
@@
 The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
-The initial set is a planar periodic orbit about \( L_1\).
-We generate the periodic orbit by using a linear approximation to first determine periodic solutions.
-This approximate periodic solution from the linear system is then implemented in the nonlinear dynamics through the method of differential correction, or single shooting~\cite{koon2011}.
-The initial \( \dot{y} \) is varied until the the periodic orbit makes a perpendicular crossing, \( \dot{x} = 0 \), after one half period.
-Due to the symmetry of the three-body problem, this ensures that the trajectory will return to the initial state after a single period.
+The initial set is a planar periodic orbit about \( L_1\), which is generated using the process of differential correction of a linear approximation~\cite{koon2011}.
 

### 2017-07-12 Response.tex
Add correct enumeration for reviewer 1 comments

@@
 I think that this statement needs to be quantified better because it gives a dishonest portrayal of high-precision Runge-Kutte integrators without numbers attached to it.
+\end{itshape}
 
+\item
+    \begin{itshape}
 "Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system [8].
@@
 
+\item
+    \begin{itshape}
 -"This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs."  Yes, but the order of the variational integrator still matters.
@@
 Characterizing the scope of the analysis in this way would be helpful to the audience.
+\end{itshape}
 
+\item 
+    \begin{itshape}
 "Optimal solutions are generally sensitive to small variations in the initial multipliers.
@@
 At best, you can change the statement to indicate that the accuracy in the gradients may improve, which is always helpful to gradient-based methods, but it's no guarantee that it eliminates major problem sensitivity.
+\end{itshape}
 
+\item
+    \begin{itshape}
 Authors may also want to point out the loss of accuracy in variable step integrators that can also degrade first-order gradient based methods.
@@
 Pellegrini, E.P., Russell, R.P., On the Accuracy of Trajectory State Transition Matrices, Paper AAS-15-785, AAS/AIAA Astrodynamics Specialist Conference, Vail, CO, Aug 2015
+\end{itshape}
 
+\item 
+    \begin{itshape}
 "This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods." I disagree that suboptimality is typical of direct methods.
@@
 For example, a hybrid direct method could use continuous optimal control theory control law, but directly minimize the cost function instead of solving the full TPBVP.
+\end{itshape}
 
@@
 
+\item
+    \begin{itshape}
 Figures 10D and 13D - the authors need to present these plots in dimensional coordinates so that they may be reconciled with what existing technology can produce.
@@
 It may be that the results are beyond the scope of current technology - that's fine - just say that they are examples and that it's ongoing work to formulate example problems that capture existing technology better.
+\end{itshape}
 
+\item
+    \begin{itshape}
 "The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches."  Did the authors attempt single shooting for their example problems before coming to this conclusion?
@@
 So, justification for its selection may be warranted.  You could always find a reference in which this trade space was really explored thoroughly to show why you made the choice.
+\end{itshape}
 
+\item
+    \begin{itshape}
 "In this work, we use the Matlab nonlinear solver fsolve to solve the system of nonlinear equations defined by the multiple shooting algorithm."  Can authors give some additional details?
@@
 For users of this method it's also useful to see the iteration history (more comments on this later).
+\end{itshape}
 
+\item
+    \begin{itshape}
 Did the authors experiment with the number of multiple shooting stages in their example problems?
@@
 If I missed it, I apologize, but I didn't see it.
+\end{itshape}
 
+\item 
+    \begin{itshape}
 Section 6.1 - I don't think you need to explain how you created the three-body orbits.
@@
 (Starting with line 49, page 18, up to line 21 page 19.)
+\end{itshape}
 
+\item
+    \begin{itshape}
 "We define a maximum magnitude of the thrust as umax = 0.75 and assume we can point the thrust in any direction within the plane." And  "we again assume a upper bound on the thrust magnitude of umax = 0.75."  What was the reason behind this?
@@
 In the context of what electric propulsion systems can deliver, it could be arguable that this value isn't really "low-thrust" but "finite-thrust".
+\end{itshape}
 
+\item 
+    \begin{itshape}
 Is this method exceptionally difficult to adapt to create a numerical example from LEO or GTO?
@@
 This should be called out in the conclusions or future work.
+\end{itshape}
 
+\item 
+    \begin{itshape}
 The authors should discuss some statistics of the convergence of their multiple shooting TPBVP - number of iterations, compute time, etc.
+\end{itshape}
 
+\item 
+    \begin{itshape}
 Can any intuition be gained from reporting the optimal converged cost function values in section 6.1 and 6.2?
 One could at least compare them to compare the reachability in the two different examples.
+\end{itshape}
 
+\item
+    \begin{itshape}
 "Low energy transfers from the Earth to the Lagrange points are necessary for future missions."  One must be careful not to confuse these with traditional "low-energy transfers" to the moon, which use Sun third-body perturbations to transfer into lunar orbit.
@@
 Online Copy: https://descanso.jpl.nasa.gov/monograph/series12/LunarTraj--Overall.pdf
+\end{itshape}
 
+\item 
+    \begin{itshape}
 For what it's worth, exploiting sun perturbations with low-thrust in the Earth-moon RTBP would be highly beneficial and represent a worthy avenue of future research (more on this below.)
+\end{itshape}
 
+\item 
+    \begin{itshape}
 The conclusions of the paper should acknowledge other potentially limiting factors of the presented analysis
@@
 -Thrust and mass are completely decoupled from the analysis, which is limiting in terms of finding suitable hardware that can deliver the controls needed, as currently presented.
+\end{itshape}
 

### 2017-07-12 Response.tex
Fix double citation in response letter

@@
                 \includegraphics[width=\textwidth]{trajectory} 
-                \caption{Earth-Moon three-body trajectory used for integrator comparision} \label{fig:compare_trajectory} 
+                \caption{Earth-Moon three-body trajectory used for integrator comparision}
         \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
@@
                 \includegraphics[width=\textwidth]{figures/cputimevsE.pdf} 
-                \caption{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}} \label{fig:compare_energy} 
+                \caption{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}} 
         \end{subfigure} 
         \caption{Integrator Comparison}
-        \label{fig:integrator_compare} 
 \end{figure}

### 2017-07-12 Response.tex
Add correction for excess detail on periodic orbits
Reviewer 1 Comment 31

@@
 
+The discussion of generating periodic orbits in the three body problem has been removed.
+It replaced with a small introduction mentioning the differential correction process and a citation to a relevant resource. 
+The revised paragraph is copied below.
+
+\begin{correction}
+The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
+The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
+The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
+\Cref{fig:moon_orbit} shows that the target set remains in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
+This type of orbit would be useful for a variety of mission scenarios.
+The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
+The initial set is a planar periodic orbit about \( L_1\), which is generated using the process of differential correction of a linear approximation~\cite{koon2011}.
+\end{correction}
+
 \item

### 2017-07-12 Response.tex
Add some justification for the multiple shooting approach
Reviewer 1 Comment 28

@@
 
+Our original attempt at solving the optimal control problem did rely on single shooting.
+However, during our investigation it was found that it was quite difficult for solutions to converge.
+As a result, we decided to implement the multiple shooting approach in hopes of generating more consistent and robust solutions.
+While it is possible to use single shooting, such as the provided reference, it is known that the single shooting approach will have greater sensitivity to initial costate errors as compared to a multiple shooting method.
+This can be alleviated, or entirely mitigated in the case of~\cite{ranieri2006}, it is highly case specific and dependent on a large variety of case specific techniques. 
+For example, the work in~\cite{ozimek2010a} used both a single shooting and multiple shooting approach to solve the optimal control problem.
+The large sensitivity of the single shooting approach was mitigated through the use of the multiple shooting method, at the expense of additional design variables.
+
+In the manuscript, additional detail has been added in this section which describes the difficulties of the single shooting approach and our decision to use the multiple shooting method.
+The modified portion of the manuscript is copied here.
+\begin{correction}
+Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs~\cite{stoer2013}.
+This multiple shooting approach incorporates additional interior constraints but reduces the sensitivity of the costates along each sub-arc.
+The use of the multiple shooting method reduces the sensitivity of changes in the initial costate at the expense of additional design variables, but has been shown to provide more stable and robust solutions~\cite{ozimek2010a}.
+\end{correction}
+
 \item

### 2017-07-12 manuscript.tex
Add some detail on the time units
Add some justification for the multiple shooting approach
Reviewer 1 Comment 28

@@
 Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs~\cite{stoer2013}.
-The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches.
+This multiple shooting approach incorporates additional interior constraints but reduces the sensitivity of the costates along each sub-arc.
+The use of the multiple shooting method reduces the sensitivity of changes in the initial costate at the expense of additional design variables, but has been shown to provide more stable and robust solutions~\cite{ozimek2010a}.
 
@@
 The numerical values associated with the green points denote the required time of flight along the invariant manifold in non-dimensional units.
-The required travel time for a transfer using the unstable invariant manifold is \( t_f \approx 3.1\).
+The required travel time for a transfer using the unstable invariant manifold is  approximately \( t_f \approx 3.1\) non-dimensional time units.
 \begin{figure} 

### 2017-07-12 manuscript.tex
Fix equation reference number

@@
 The state trajectory is intialized using the free trajectory of the periodic orbit. 
-Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propogated using the discrete equations of motion in~\cref{eq:costate_update}.
+Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propogated using the discrete equations of motion in~\cref{eq:costate_eom}.
 This results in the initial guess of the design vector \( \vecbf{\chi} = \begin{bmatrix} \vecbf{\lambda}_0 & \vecbf{x}_1^{-} & \vecbf{\lambda}_1^{-} & \vecbf{\beta} \end{bmatrix}^T\).

### 2017-07-12 Response.tex
Add detail on fsolve options from matlab
Reviewer 1 Comment 29

@@
 
+Additional detail has been incorporated which discusses the specific solver used in computing the solutions. 
+The description does not specifically mention all of the possible options in \texttt{fsolve} but rather gives the key components used in our solution.
+This approach seems to be consistent with the literature, e.g.~\cite{ozimek2010a,stuart2010}.
+The modified portion is copied below.
+
+\begin{correction}
+In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
+Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
+All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
+\end{correction}
+
 \item

### 2017-07-12 manuscript.tex
Add detail on fsolve options from matlab
Reviewer 1 Comment 29

@@
 In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
-
+Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
+All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
 % tpbvp is solved using a shooting method (fsolve in matlab)

### 2017-07-12 Response.tex
Add detail on choosing the number of shoooting stages
Reviewer 1 Comment 31

@@
 
+We did spend considerable time in determining the appropriate number of stages. 
+One key issue was that our implementation in software depended on dividing the total number of integration steps by the number of interior stages.
+For simplicity, this reduced to possible number of stages to an even number. 
+Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in covergence. 
+The solver would tend to ensure a continuous trajectory, by satisfying the interior stages, at the expense of the terminal stage. 
+Some additional detail has been provided that mentions our choice of the number of stages and has been copied below.
+
+\begin{correction}
+Based on experimentation, we use four stages in our multiple shooting method.
+This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
+\end{correction}
+
 \item 

### 2017-07-12 Response.tex
Additional detail on dimensional units of thrust for L1 reach
Reviewer 1 Comment 32 and 27

@@
 
+The maximum thrust magnitude was chosen to be approximately \SI{1}{\newton} for an assumed \SI{500}{\kilo\gram} spacecraft. 
+While larger than the NEXT thruster mentioned earlier, a cluster of similar engines could provide the required thrust (assuming adequate available power). 
+Additional detail has been included in the manuscript describing this thrust limit in dimensional units. 
+\begin{correction}
+Next, we determine the reachability set with the addition of a low thrust control input.
+We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
+This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
+We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
+This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
+Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
+Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
+The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
+This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit. 
+\end{correction}
+
 \item 

### 2017-07-12 manuscript.tex
Add detail on choosing the number of shoooting stages
Reviewer 1 Comment 31
Additional detail on dimensional units of thrust for L1 reach
Reviewer 1 Comment 32 and 27

@@
 Similarly, the constraint vector is augmented with the additional interior constraints defined in~\cref{eq:interior_constraints}.
+Based on experimentation, we use four stages in our multiple shooting method.
+This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
 The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
@@
 We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
+This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
+Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
+Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
 The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).

### 2017-07-12 Response.tex
Address lack of examples departing from LEO/GTO
Reviewer 1 Comment 33

@@
 
+This method is not difficult to adapt for other scenarios.
+However, there is a certain amount of experimentation to ensure that trajectories converge. 
+In addition, the details in this paper were designed in the context of the planar three body problem and as a result and extension to the nonplanar GTO is non trivial.
+The examples presented are primarily aimed at introducing and demonstrating the feasibility of this approach, rather than designing operational mission profiles for spacecraft.
+However, this is a valuable comment and a goal of future research. 
+Additional detail has been included in the conclusion discussing this topic.
+
 \item 

### 2017-07-13 Response.tex
Update some more responses to reviewer
Reviewer 1

@@
 
+\item
+    \begin{itshape}
 -It does not give a complete picture of the pros and cons of the proposed methodology.
@@
 The state of the art to design trajectories that depend on low-thrust engine hardware is either:
+
 -To produce constant thrust magnitude or no thrust magnitude (typical of nuclear electric propulsion, or very simple electric propulsion systems.)
+
 -To produce variable thrust magnitude associated with various throttle points at different power accessibility levels (solar electric propulsion).
@@
 a=T/m
+
 Thus - a MAJOR focus of future work should be the introduction of a seventh state (mass).
 This gives rise to a more challenging optimal control problem requiring the switching function, but it is the form that is generally needed in computing optimal open-loop trajectories that can actually be flown.
+\end{itshape}
 
@@
 The pros and cons of the PCRTBP
+
 -The PCRTB gives rise to exploiting Poincare sections in a systematic manner.
 2D solutions might be close enough for qualitative understanding of the solution space for many problems (as long as large plane changes aren't needed).
+
 -The PCRTB is limiting for many pragmatic departure orbits (e.g. LEO and GTO, the most common departure orbits for low-thrust transfers).
@@
 
+\item 
+    \begin{itshape}
 The pros and cons of variational integrators should really be acknowledged rather than completely dismissing the RK methods which are still the current state of the art in mission design (using higher-orders and suitable tolerances)
+
 -Pros = boundedness in the energy behavior.  Very long term integration will not deviate from the known dynamical structure.
+
 -Cons = you don't know where you are on the energy surface, just that you are on it somewhere.
@@
 Knowing the error is on an energy surface does not mean that you don't have error.
-
+\end{itshape}
 \end{itemize}
@@
 
+Most of the literature tends to not spend time discussing the low level details of the computational procedure, e.g.~\cite{grebow2011,stuart2010}.
+In addition, the number of iterations, computation time and even the eventual cost function value are highly dependent on the specific system and more specifically the implementation of the algorithm in software.
+There are a wide variety of works which are focused more exclusively on the computational implementation of various optimal control techniques which do spend significant time on the computational characteristics.
+However, our manuscript is primarily focused on demonstrating the feasbility of using reachability sets to design transfer trajectories.
+As a result, there is not much benefit to detailed statistics of the implementation of this methodology.
+
 \item 
@@
 \end{itshape}
+
+The cost function only parameterizes the distance between the controlled and uncontrolled terminal states on the \Poincare section. 
+In addition, each state which lies on the reachable set would typically have a different distance and therefore a different coverged cost function value. 
+Furthermore, it would be difficult to characterize a single parameter which characterizes the reachable set, in the form of the cost function, for all states of the reachable set. 
+Also, the cost function is highly dependent on the transit time of trajectory, i.e. a longer transit time corresponds to a larger reachable set and a larger magnitude of the cost function.
+However, this is an interesting suggestions for future research.
+For example, the cost function gives some indication of area/volume of the reachable set. 
+It may be possible to use this value as another way to characterize the subspace on the \Poincare section and is an intriguing avenue of future research.
 
@@
 
+This is a very useful and interesting suggestion and something that we will take into consideration for future research.
+
 \item 
@@
 The conclusions of the paper should acknowledge other potentially limiting factors of the presented analysis
+
 -Omits advantage of exploiting fourth body perturbation (Sun in Earth/moon RTBP)
+
 -Transfers are 2D only, method is more challenging to adapt to 3D
+
 -Presented control magnitudes are dimensionally large, and more work is needed to determine feasibility of approach for lower accelerations
+
 -Thrust and mass are completely decoupled from the analysis, which is limiting in terms of finding suitable hardware that can deliver the controls needed, as currently presented.

### 2017-07-13 Response.tex
Differentiate between low thrust and low energy
Reviewer 1 Comment 36

@@
 \end{itshape}
+This sentence has been modified to make it more clear that we're discussing low thrust propulsion rather than low-energy transfers.
+The modified sentence is copied below.
+
+\begin{correction}
+Transfers from the Earth to the Lagrange points, through the use of low-thrust electric propulsion, offer an additional and potentially shorter time of flight in comparison to the low-energy transfers which utilize solar perturbations.
+\end{correction}
 

### 2017-07-13 manuscript.tex
Differentiate between low thrust and low energy
Reviewer 1 Comment 36

@@
 Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
-Low energy transfers from the Earth to the Lagrange points are necessary for future missions.
+Transfers from the Earth to the Lagrange points, through the use of low-thrust electric propulsion, offer an additional and potentially shorter time of flight in comparison to the low-energy transfers which utilize solar perturbations.
 

### 2017-07-13 Response.tex
Update conclusion
Reviewer 1 Comment 38

@@
 
+The conclusion has been expanded with some discussion of the drawbacks of this research and avenues of future work.
+The modified conclusion is copied below.
+
+\begin{correction}
+In this paper, an optimal transfer process which combines concepts of reachability and \Poincare section is used to generate transfer between planar periodic orbits in the three-body problem.
+The \Poincare section allows for trajectory design on a lower dimensional phase space and simplifies the process.
+The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
+However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
+Furthermore, our approach relies heaviliy on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
+
+There is additional research to extend these results to more general transfer scenarios in future work.
+The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
+In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
+Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
+This will result in a more complicated optimal control problem and is a focus of future research.
+Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
+The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
+\end{correction}
 \end{enumerate}

### 2017-07-13 Response.tex
Update both plots with control units
Reviewer 1 Comment 27

@@
 
+The mentioned plots of the control input have been modified to show the data in dimensional units (Newtons). 
+Some additional detail has been provided in the manuscript which discusses the choice of \( u = 0.75 \) and it's realtionship to the current state of the art thruster systems.
+The modified plots are not included here but the additional content is copied below in response to a different but related comment.
+
 \item

### 2017-07-13 manuscript.tex

@@
 However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
-Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
-This analysis has also assumed perfect attitude control and the ability to orient thrust in any direction.
-The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability.
+Furthermore, our approach relies heaviliy on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
+
+There is additional research to extend these results to more general transfer scenarios in future work.
+The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
+In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
+Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
+This will result in a more complicated optimal control problem and is a focus of future research.
+Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
+The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
 

### 2017-07-13 Response.tex

@@
 
+The section on the variational integrator has been reduced and this specific sentence has been removed. 
+However, additional detail has been provided in a different section of the paper which discusses this drawback of the variational integrator.
+
+\begin{correction}
+Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
+In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
+This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
+However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
+\end{correction}
 \item 
@@
 The plot has been removed and replaced with another comparision between the variational integrator and \texttt{ODE45}.
+In addition, all of the plots now have axis labels.
 

### 2017-07-13 manuscript.tex

@@
 Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
+In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
+This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
+However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
 

### 2017-07-13 Response.tex
Add comment on optimal control with variational integrator
Reviewer 1 Comment 11
Add detail about variable step integration
Reviewer 1 Comment 12

@@
 
+Additional detail has been included in the manuscript discussing this potential drawback of variable step integrators.
+\begin{correction}
+In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
+The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.
+\end{correction}
 \item 

### 2017-07-13 manuscript.tex
Add detail about variable step integration
Reviewer 1 Comment 12

@@
 However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
+In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
+The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.
 
-\section{Optimal Control Formulation for Reachability Set}
+\section{Optimal Control Formulation for Reachability Set}\label{sec:optimal_control}
 % introduce method

### 2017-07-13 Response.tex
Remove optimal control sentence
Reviewer 1 comment 11
Mention some of the differences between indirect and direct optimal control
Reviewer 1 Comment 13

--- 2017-07-13 Response.tex
@@
 
+This sentence has been removed from the introduction.
+However, this is a valid comment and is noted for the future.
+
 \item
@@
 
+Additional detail has been included on the merits of indirect optimal control vs. direct optimal control methods.
+\begin{correction}
+This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on course or suboptimal control parameterizations.
+For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
+\end{correction}
 \item 

### 2017-07-13 manuscript.tex
Mention some of the differences between indirect and direct optimal control
Reviewer 1 Comment 13

--- 2017-07-13 manuscript.tex
@@
 We derive and solve the Euler-Lagrange equations to generate the necessary conditions for an optimal solution.
-This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods.
+This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on course or suboptimal control parameterizations.
+For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
 In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.

### 2017-07-13 Response.tex
Address general reviewer comments
Reviewer 1 General comments

@@
 
-    The paper is being reduced. 
-    The introduction has been reduced to avoid excessive repetition. 
-    In addition, much of the background material has been removed
+The manuscript has been reduced and much of the mathematical development has been removed.
 
@@
 \end{itshape}
+
+These pros/cons have been addressed in the conclusion.
 
@@
 \end{itshape}
+
+Additional detail has  been included in the manuscript under the section describing the variational integrator.
+These details are discussed further under the more specific comments below.
+
 \end{itemize}
@@
 \end{itshape}
+
+Additional detail has been included which quantifies the notion of long-term.
+
+\begin{correction}
+As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g.
+on the order of decades to centuries, which is typical of the three-body problem.
+\end{correction}
 

### 2017-07-13 Response.tex
Formatting for reviewer 2 comments

@@
 
-Reviewer \#2: PAPER SUMMARY:
+\begin{itshape}
+Reviewer \#2:
+
+PAPER SUMMARY:
 The contribution of the authors is the synthesis of reachability theory with traditional low-energy mission design approaches to ease the study of the low-thrust mission design space in the CR3BP.
@@
 ASSESSMENT:
+\begin{itemize}
+    \item
 The central premise of the paper - using optimization of Poincare distance as a guiding motivation to make low-thrust reachability analysis more tractable - is interesting, useful, and to my knowledge reasonably novel.
@@
 
+\item 
 First, in terms of composition, large portions of the paper are spent on lengthy development of foundational work that is well-known within the astrodynamics community and available in textbooks.
@@
 
+\item
 While there are the beginnings of a good article within this manuscript, it is not sufficiently close to publishable form to be accepted.
@@
 I hope that the authors will not find this verdict too discouraging, as I believe that stronger results are well within their reach.
+\end{itemize}
+\end{itshape}
 

### 2017-07-14 Response.tex
Organize reviewer 2 comments

@@
 ASSESSMENT:
-\begin{itemize}
-    \item
+
 The central premise of the paper - using optimization of Poincare distance as a guiding motivation to make low-thrust reachability analysis more tractable - is interesting, useful, and to my knowledge reasonably novel.
@@
 However, there are two major problems with this paper.
+\end{itshape}
+\begin{itemize}
 
 \item 
+    \begin{itshape}
 First, in terms of composition, large portions of the paper are spent on lengthy development of foundational work that is well-known within the astrodynamics community and available in textbooks. 
+\end{itshape}
+\item 
+    \begin{itshape}
         Secondly, the contributions and findings of the authors feel highly preliminary and do not seem to evaluate their method at significant depth.
+    \end{itshape}
 
 \item
+    \begin{itshape}
 While there are the beginnings of a good article within this manuscript, it is not sufficiently close to publishable form to be accepted.
@@
 The eventual publication should focus less on background and should provide more insight into the fundamental properties of their method, using more rigorously constructed and analyzed examples.
-
+\end{itshape}
+\item 
+    \begin{itshape}
 I can sympathize that there was quite a long road to get to the point of understanding this topic and generating results in it.
 I hope that the authors will not find this verdict too discouraging, as I believe that stronger results are well within their reach.
+\end{itshape}
 \end{itemize}
-\end{itshape}
 
 SPECIFIC FEEDBACK:
-\begin{itemize}
+\begin{enumerate}
     \item
@@
 The introduction is reduced and many of the paragraphs have been reworded and combined. 
-In addition, the extended discussion of the benefits of variational integrators is removeda.
-In the interest of space/clarity the modified introduction is not copied here.
+The mathematical background has been condensed into a single section with seperate subsections which discuss some of the mathematical details.
+In addition, the extended discussion of the benefits of variational integrators is removed from the introduction.
 
+\item
+    \begin{itshape}
 The development of the CR3BP, invariant manifolds and Poincare maps, all go on a bit too long.
@@
 Spending several pages on variational principles is way too much.
+\end{itshape}
 
+\item
+    \begin{itshape}
 Section 4 ends on page 14 with essentially a thesis statement ("In this paper, we...").
 Strive to push the boundary between background and new developments to a much earlier point in the paper; it shouldn't happen so late that you need to remind the audience what the paper is for.
+\end{itshape}
 
+\item 
+    \begin{itshape}
 A missing relevant reference: "On target for Venus — set oriented computation of energy efficient low thrust trajectories" (Dellnitz, Junge, Post, and Thiere, 2006)
+\end{itshape}
 
+\item
+    \begin{itshape}
 *** Variational Integrators
@@
 
+\end{itshape}
+
+\item
+    \begin{itshape}
 Apart from these claims, I question the need for a variational integrator in the first place.
@@
 Also, fixed step seems inappropriate for a path through diverse dynamic regimes, and it's not clear that the problem scope calls specifically for energy preservation (especially since the idea is to use low thrust to mitigate some time cost of exploiting the problem structure).
+\end{itshape}
 
+\item
+    \begin{itshape}
 *** Figures
@@
 
+\end{itshape}
+
+\item
+    \begin{itshape}
 Figure 4 does not show what I think you intend it to show; it looks like two periodic circular orbits.
+\end{itshape}
 
+\item
+    \begin{itshape}
 Figure 6 is also very unclear.  Is phi supposed to be \texttt{theta\_d}?  Is the bounding curve specifically circular, or should it be a more general shape?
+\end{itshape}
 
+\item
+    \begin{itshape}
 Consider plotting the intersections of manifolds with the Poincare section as continuous curves.  
+\end{itshape}
 
+\item
+    \begin{itshape}
 Also consider zooming/scaling as necessary to show the details you discuss.
+\end{itshape}
 
+\item
+    \begin{itshape}
 *** Method/Examples
@@
 More precise claims and assurances about solution quality (and therefore the validity of the computed reachable set boundary) are needed.
+\end{itshape}
 
+\item
+    \begin{itshape}
 How damaging is the assumption of a fixed travel time?  And how about the specification of a specific departure point on the initial orbit?
+\end{itshape}
 
+\item
+    \begin{itshape}
 In Fig 10b, it is hard to see, but appears that the low-thrust-reachable set is actually *centered* on part of the target set.
@@
 Perhaps beginning with the time horizon of the zero-thrust opportunity, and successively lowering it, would help.
+\end{itshape}
 
+\item
+    \begin{itshape}
 Perhaps there is also a better way to show what the control is doing, e.g., in some alternative parametrization?
 This could help provide assurances that the solution is doing something meaningful and is not just stuck in a weird local minimum.
+\end{itshape}
 
+\item
+    \begin{itshape}
 In the second example, what do you lose by chaining together multiple reachability-based steps?
@@
 And can you use knowledge about directionality of movement on the Poincare section to limit the scope of your reachability search?
+\end{itshape}
 
+\item
+    \begin{itshape}
 In general, please devote more depth to probing the strengths, weaknesses, and fundamental properties of the method.  Considering extreme cases can help with this.
-\end{itemize}
+\end{itshape}
+\end{enumerate}
 \bibliography{library}

### 2017-07-14 Response.tex
Address notion of suboptimal direct control techniques
Reviewer 1 Comment  13

@@
 For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
+\end{correction}
+
+In addition, the introduction has been modified to explicitly mention that direct methods which use suboptimal control parameterizations will result in suboptimal solutions.
+
+\begin{correction}
+The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which generally rely on suboptimal representations of the control.
 \end{correction}

### 2017-07-14 manuscript.tex
Address notion of suboptimal direct control techniques
Reviewer 1 Comment  13

@@
 Alternatively, indirect methods apply the calculus of variations to derive the necessary conditions for optimality.
-The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which result in sub-optimal solutions.
+The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which generally rely on suboptimal representations of the control.
 Second, in order to implement any optimization method a sufficiently accurate ``initial guess'' is required.

### 2017-07-14 manuscript.tex
Sentence grammar was off

@@
 We utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
-Through the use of a simple distance metric on a \Poincare~section greatly reduces the dimensionality of the reachability analysis and provides an approximation of the reachability set.
+The use of a simple distance metric on a \Poincare~section greatly reduces the dimensionality of the reachability analysis and provides an approximation of the reachability set.
 In addition, through the use of geometric integrators we more accurately capture the effects of low thrust on the system dynamics in the numerical simulation over extended periods which are typical of electric propulsion systems.

### 2017-07-14 Response.tex
Discuss reduction of introductory sections
Reviewer 2 Comment 2

@@
 
+The mathematical background has been reduced and condensed into a single section with multiple subsections.
+The discussion of the PCRTBP, invariant manifolds and \Poincare maps are combined into a single subsection.
+Furthermore, the variational integrator discussion has been reduced and the background material on the derivation of variational mechanics has been omitted.
+Instead, only the specific discussion of the development of the variational integrator is included.
+
 \item

### 2017-07-14 Response.tex
Add additional reference to paper that discusses reachability sets
Reviewer 2 Comment 4
Update figure captions to be more descriptive
Reviewer 2 Comment 7

@@
 \usepackage{jas_tikz_packages}
+\tdplotsetmaincoords{60}{125} % view angle in spherical coordinates
 \input{macros}
@@
 
+The reference has been noted in the discussion of reachability sets.
+\begin{correction}
+More recently, reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a,dellnitz2006}.
+\end{correction}
+
 \item
@@
 Altogether, the figures need more descriptive captions.  
-
-\end{itshape}
-
+\end{itshape}
+
+Thank you for the comment, and we agree that the figures will benefit from more descriptive captions.
+We've added additional descriptions for all of the figures.
+The figure captions are shown below.
+
+\begin{correction}
+\begin{figure}[htbp]
+        \centering
+        \includegraphics[width=0.5\textwidth]{energy_contours}
+        \caption{Contour plot of Jacobi integral: Zero velocity curves of constant Jacobi integral.
+A particle with fixed energy level cannot cross the contour lines and is therefore limited to specific regions of the phase space.
+}
+\end{figure}
+
+\begin{figure}
+        \centering
+        \begin{scaletikzpicturetowidth}{0.5\textwidth}
+    \begin{tikzpicture}[tdplot_main_coords,
+      poincare/.style={opacity=.2,very thick,fill=blue},
+      orbit/.style={very thick,black},
+      orbit hidden/.style={very thick,dashed},
+      grid/.style={very thin,gray!50},
+      axis/.style={->,blue,thick}, scale=\tikzscale]
+
+    % nodes for the poincare section
+    \node[label=above:\(\Sigma\)] (upper_right) at (0,5,5) {};
+    \node[] (upper_left) at (0,1,5) {};
+    \node[] (lower_left) at (0,1,0) {};
+    \node[] (lower_right) at (0,5,0) {};
+
+    % draw poincare section
+    \draw[poincare] (upper_right.center) -- (upper_left.center) -- (lower_left.center) -- (lower_right.center) -- (upper_right.center);
+
+    % draw a periodic orbit
+    \coordinate (center) at (0,0,2);
+    \node[below right] (x0) at (0,2,2) {\(\vecbf{q}_0\)};
+    \filldraw (0,2,2) circle (3pt);
+
+    \node[below right] (x1) at (0,3,2) {\(\vecbf{q}_1\)};
+    \filldraw (0,3,2) circle (3pt);
+
+    \tdplotdrawarc[orbit hidden]{(center)}{2}{90}{190}{}{};
+    \tdplotdrawarc[orbit,<-]{(center)}{2}{-170}{90}{}{};
+
+    \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{199}{}{};
+    \tdplotdrawarc[orbit,<-]{(center)}{3}{-161}{90}{}{};
+
+        \end{tikzpicture}
+    \end{scaletikzpicturetowidth}
+    \caption{Diagram of the \Poincare map: Periodic orbits will appear as fixed points on the \Poincare section \( \Sigma \).
+Stabiilty of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.}
+\end{figure}
+
+\begin{figure} 
+        \centering 
+        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparision: A stable trajectory is used to test the variational and conventional integrators. 
+        The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
+        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
+        \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}}
+        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
+\end{figure}
+\end{correction}
 \item

### 2017-07-14 manuscript.tex
Add additional reference to paper that discusses reachability sets
Reviewer 2 Comment 4
Update figure captions to be more descriptive
Reviewer 2 Comment 7

@@
         \includegraphics[width=0.5\textwidth]{energy_contours}
-        \caption{Contour plot of Jacobi integral}
+        \caption{Contour plot of Jacobi integral: Zero velocity curves of constant Jacobi integral.
+A particle with fixed energy level cannot cross the contour lines and is therefore limited to specific regions of the phase space.
+}
         \label{fig:energy_contour}
@@
     \end{scaletikzpicturetowidth}
-        \caption{Diagram of the \Poincare map\label{fig:poincare_map}}
+    \caption{Diagram of the \Poincare map: Periodic orbits will appear as fixed points on the \Poincare section \( \Sigma \).
+Stabiilty of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.\label{fig:poincare_map}}
 \end{figure}
@@
         \centering 
-        \begin{subfigure}[h]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{trajectory} 
-                \caption{Earth-Moon three-body trajectory used for integrator comparision} \label{fig:compare_trajectory} 
-        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-%       \begin{subfigure}[htbp]{0.3\textwidth} 
-%               \includegraphics[width=\textwidth]{components} 
-%               \caption{State \( \parenth{x , y, \dot{x}, \dot{y}}\)} \label{fig:compare_components} 
-%       \end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{figures/cputimevsE.pdf} 
-                \caption{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}} \label{fig:compare_energy} 
-        \end{subfigure} 
-        \caption{Integrator Comparison}
+        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparision: A stable trajectory is used to test the variational and conventional integrators. 
+        The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.\label{fig:compare_trajectory}}
+        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
+        \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}\label{fig:compare_energy}}
+        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
     \label{fig:integrator_compare} 
 \end{figure}
+
 A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
@@
 The theory formally supporting reachability has been extensively developed and is directly derivable from optimal control theory~\cite{varaiya2000,lygeros2002,lygeros2004}.
-More recently, reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a}.
+More recently, reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a,dellnitz2006}.
 Computation of the reachable set for a system involves solving the Hamilton-Jacobi partial differential equation or satisfying a dynamic programming principle.

### 2017-07-17 Response.tex
Add detail to Poincare map figure
Reviewer 2 Comment 8

@@
 
+We disagree as the figure is meant to demonstrate the purpose and use of the \Poincare section.
+The \Poincare section figure has been moved but it is meant to demonstrate in a pictoral form the benefit of this visualization technique.
+The periodic orbits intersect the section at fixed points, as a result it becomes simple to visualize and determine these fixed solutions by searching for fixed points on the section
+Conversely, stable/unstable trajectories will also appear on the section with a distinctive signature, and by tracking successive intersections of the trajectory with the \Poincare section.
+
+In order to improve the clarity and reduce confusion, we have added some additional detail to the discussion of the \Poincare map to the manuscript.
+This addition has been copied below.
+\begin{correction}
+\Poincare maps are a useful tool in the analysis of the flow near periodic orbits in the three-body problem.
+We let \( \Sigma \) define a hypersurface of section chosen such that all trajectories in the vicinity of a state \( \vecbf{q} \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
+A \Poincare map, \( P(\vecbf{q}) = \phi(T;\vecbf{q}) \), maps the state of a trajectory from one intersection to the next.
+Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
+In~\cref{fig:poincare_map} we show two examples of periodic trajectories intersecting the \Poincare section. 
+Periodic solutions will appear as fixed points on the section, such as \( q_0, q_1 \) in~\cref{fig:poincare_map}, while stable or unstable trajectories become clearly visible by viewing successive intersections of the section.
+This allows for greater insight into the stability and dynamics of periodic solutions of a dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories.
+\end{correction}
+
 \item

### 2017-07-17 manuscript.tex
Add detail to Poincare map figure
Reviewer 2 Comment 8

@@
 Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
+In~\cref{fig:poincare_map} we show two examples of periodic trajectories intersecting the \Poincare section. 
+Periodic solutions will appear as fixed points on the section, such as \( q_0, q_1 \) in~\cref{fig:poincare_map}, while stable or unstable trajectories become clearly visible by viewing successive intersections of the section.
 This allows for greater insight into the stability and dynamics of periodic solutions of a dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories.

### 2017-07-17 Response.tex
Correct reachability plot and add some detail
Reviewer 2 Comment 9

@@
 
+Thank you for the comment.
+The angle in the figure is meant to match the terminal constraint \( \theta_d\).
+In addition, the reachable set is only circular for ease of design in the figure.
+In reality the figure will be a general shape, rather than specifically circular.
+
+The manuscript has been modified to correct the figure and also incorporate some more discussion of the figure. 
+The additions are included below as well as the updated figure.
+\begin{correction}
+\Cref{fig:reachability_set} illustrates how, without any control input, trajectories will intersect with the \Poincare section at \( \vecbf{x}_n \). 
+However, the addition of low thrust propulsion allows the spacecraft to depart from the natural dynamics and intersect the \Poincare section at a different location.
+We use a cost function to define a distance metric on the \Poincare section from the control-free intersection to an intersection under the influence of the control input.
+Maximization of this distance along varying directions enables us to generate the largest reachability set under the bounded control input.
+In~\cref{fig:reachability_set} the reachable set is shown as a circular region on the \Poincare section.
+In practice, the reachable set will be of a general shape and also higher dimensional in the nonplanar case.
+\begin{figure}
+        \centering
+%       \includegraphics[width=0.5\textwidth]{reachability_set_hand}
+        \begin{scaletikzpicturetowidth}{0.5\textwidth}
+        \begin{tikzpicture}[tdplot_main_coords,
+          poincare/.style={opacity=.2,very thick,fill=blue},
+          orbit/.style={very thick,black},
+          orbit hidden/.style={very thick,dashed},
+          grid/.style={very thin,black},
+          axis/.style={->,blue,thick},
+          reachability/.style={thick,blue},scale=\tikzscale]
+
+        % nodes for the poincare section
+        \node[label=above:\(\Sigma\)] (upper_right) at (0,5,5) {};
+        \node[] (upper_left) at (0,1,5) {};
+        \node[] (lower_left) at (0,1,0) {};
+        \node[] (lower_right) at (0,5,0) {};
+
+        % draw poincare section
+        \draw[poincare] (upper_right.center) -- (upper_left.center) -- (lower_left.center) -- (lower_right.center) -- (upper_right.center);
+
+        % draw a periodic orbit
+        \coordinate (center) at (0,0,2);
+        \node[label=below:\(\vecbf{x}_n\)] (x0) at (0,3,2) {};
+        % \node[label=below:\(\vecbf{x}_n\)] at (x0) {};
+        \filldraw (x0) circle (3pt);
+
+        % \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{200}{}{};
+        \tdplotdrawarc[orbit,<-]{(center)}{3}{-160}{90}{}{};
+
+        % draw reachability set on the poincare section
+        \coordinate (reach) at (0,4.5,2);
+        \tdplotsetthetaplanecoords{90}
+
+        \draw[tdplot_rotated_coords,grid] (x0) -- (reach);
+        \draw[tdplot_rotated_coords,grid] (x0) -- ++(-45:1.5);
+
+        \tdplotdrawarc[tdplot_rotated_coords,grid]{(x0)}{0.5}{-45}{90}{above}{\(\theta_d\)};
+
+        % draw terminal state on reachability set
+        \node[tdplot_rotated_coords,label=above:\(\vecbf{x}_f\)] (xf) at ($ (x0)+(-45:1.5) $) {};
+        \filldraw (xf) circle (3pt);
+
+        \node[tdplot_rotated_coords,label=below:\(J\)] at (xf) {};
+
+        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
+        % place
+    \end{tikzpicture}
+        \end{scaletikzpicturetowidth}
+        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
+        }
+\end{figure}
+\end{correction}
 \item

### 2017-07-17 manuscript.tex
Correct reachability plot and add some detail
Reviewer 2 Comment 9

@@
 Maximization of this distance along varying directions enables us to generate the largest reachability set under the bounded control input.
+In~\cref{fig:reachability_set} the reachable set is shown as a circular region on the \Poincare section.
+In practice, the reachable set will be of a general shape and also higher dimensional in the nonplanar case.
 \begin{figure}
@@
 
-        \tdplotdrawarc[tdplot_rotated_coords,grid]{(x0)}{0.5}{-45}{90}{above}{\(\phi\)};
+        \tdplotdrawarc[tdplot_rotated_coords,grid]{(x0)}{0.5}{-45}{90}{above}{\(\theta_d\)};
 
@@
         \end{scaletikzpicturetowidth}
-        \caption{Reachability set on a \Poincare section\label{fig:reachability_set}}
+        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
+        \label{fig:reachability_set}}
 \end{figure}

### 2017-07-18 Response.tex
Not going to plot the poincare sections with continuous curves
Reviewer 2 Comment 10

@@
 
+Thank you for the comment.
+The manifold intersections with the \Poincare section are plotted as discrete points as the curves were generated in a discrete manner. 
+A series of trajectories were generated by perturbing the trajectory at various points along the periodic orbit. 
+As a result, we show that the intersection is a discrete approximation of the true intersection of the manifold with the section by using the discrete points rather than a continuous curve.
+
 \item

### 2017-07-18 Response.tex
Address comment on zooming in on figures
Reviewer 2 Comment 11

@@
 
+Thank you for the comment, however we feel that the figures adequately show the desired information.
+The figures are centered around the various intersections of the reachable set and the \Poincare section.
+Furthermore, in the case of the much larger geostationary transfer requires it becomes more difficult to zoom in on the fine details.
+However, a detailed view is presented which shows the various iterations of the reachability set computations.
+
 \item

### 2017-07-18 Response.tex
Address some comments about assumptions used in the manuscript
Reviewer 2 Comment 13

@@
 
+The fixed travel time is primarily used at this stage to reduce the complexity of the optimal control formulation. 
+Furthermore, the reachability set is defined in terms of a fixed terminal time.
+A variable terminal travel time would add an additional dimension, time, to the reachable space and add additional complexity to the reachability set visualization.
+The fixed travel time does not adversely affect the optimality of the computed trajectories on the reachable set, as the reachable set is defined in terms of a fixed travel time.
+
+The choice of starting from a fixed point on the initial periodic orbit is again mostly used for convience. 
+In this case of the periodic orbit the departing trajectories are all along the unstable invariant manifold associated with the periodic orbit. 
+The uncontrolled trajecotries associated with the invariant manifolds serve as the ``center'' of each reachability set computation.
+The low-thrust propulsion system is then used to enlarge the distance between the uncontrolled and controlled states. 
+Starting from a different location on the periodic orbit will not create a qualitatively different situation as the entire invariant manifold also intersects the \Poincare section.
+However, different locations along the periodic orbit may have vastly different travel times before intersecting the \Poincare section.
+This will require modification of the optimal control problem, with a larger terminal time, but will result in a similar behavior regardless of the location on the intial periodic orbit.
+
 \item

### 2017-07-18 Response.tex
Address comment about figures difficult to see
Reviewer 2 Comment 14

@@
 
+The reachability set is closely spaced in terms of both the position space as well as on the \Poincare section.
+This is due to the fact that there is a very low-thrust control input on the spacecraft.
+As a result, the trajectories associated with the reachability set are all closely spaced, but more widely seperated from the terminal state without any control input.
+In the figure, there is a fixed point labeled \( x_n \) which is the terminal state without any control input, which will also lie on the periodic orbit.
+The reachable set, in red, is widely separated from the uncontrolled state due the effect of the low thrust proplusion, however the subsequent spread of the entire reachability set is much more constrained.
+The low thrust expands the reachability set from a fixed point to the red elliptical region. 
+This is also explained in the manuscript with a similar description of the figure.
+
 \item

### 2017-07-18 Response.tex
Address comment about the control plots
Reviewer 2 Comment 15

@@
 
+Thank you for the comment.
+The control plots show the planar components of the control thrust. 
+Furthermore, the scale has been modified to show the thrust in terms of Newtons for an assumed \SI{500}{\kilo\gram} satellite.
+The control plots are meant to demonstrate the control input required by the trajectories which lie on the maximum reachability set. 
+As a result, the control input is determined through optimization to maximize the reachability set in a particular direction on the \Poincare section.
+This is in contrast to typical optimal control applications where the control input is used directly in the trajectory design process.
+Our method seeks to approximate the reachability set, and from this set trajectories are chosen in a systematic method which drives the spacecraft towards a target.
+
 \item

### 2017-07-18 Response.tex
Address comment on stringing multiple reachable set computations
Reviewer 2 Comment 16

@@
 
+While the trajectories do return to the energy level of the \Poincare section, there is a transfer of ``energy'' between the states of the vehicle.
+It is in this manner that the reachability set is enlarged to eventually include the target state.
+
+Conducting a single reachability analysis becomes more complicated as the larger time of flight increases the sensitivity of the terminal states to variations in the initial costates. 
+Furthermore, the larger time of flight creates a much larger and more chaotic reachable set for the analysis.
+Through this work it was found that shorter trajectories tended to have more stable and consistent solutions.
+
 \item

### 2017-07-19 Response.tex
Update more captions for the figures
Reviewer 2 Comment 7

@@
 \begin{correction}
-\begin{figure}[htbp]
+\begin{figure}[H]
         \centering
@@
 
-\begin{figure}
+\begin{figure}[H]
         \centering
@@
 
-\begin{figure} 
+\begin{figure}[H]
         \centering 
@@
         [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
+\end{figure}
+
+\begin{figure}[H]
+        \centering
+%       \includegraphics[width=0.5\textwidth]{reachability_set_hand}
+        \begin{scaletikzpicturetowidth}{0.5\textwidth}
+        \begin{tikzpicture}[tdplot_main_coords,
+          poincare/.style={opacity=.2,very thick,fill=blue},
+          orbit/.style={very thick,black},
+          orbit hidden/.style={very thick,dashed},
+          grid/.style={very thin,black},
+          axis/.style={->,blue,thick},
+          reachability/.style={thick,blue},scale=\tikzscale]
+
+        % nodes for the poincare section
+        \node[label=above:\(\Sigma\)] (upper_right) at (0,5,5) {};
+        \node[] (upper_left) at (0,1,5) {};
+        \node[] (lower_left) at (0,1,0) {};
+        \node[] (lower_right) at (0,5,0) {};
+
+        % draw poincare section
+        \draw[poincare] (upper_right.center) -- (upper_left.center) -- (lower_left.center) -- (lower_right.center) -- (upper_right.center);
+
+        % draw a periodic orbit
+        \coordinate (center) at (0,0,2);
+        \node[label=below:\(\vecbf{x}_n\)] (x0) at (0,3,2) {};
+        % \node[label=below:\(\vecbf{x}_n\)] at (x0) {};
+        \filldraw (x0) circle (3pt);
+
+        % \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{200}{}{};
+        \tdplotdrawarc[orbit,<-]{(center)}{3}{-160}{90}{}{};
+
+        % draw reachability set on the poincare section
+        \coordinate (reach) at (0,4.5,2);
+        \tdplotsetthetaplanecoords{90}
+
+        \draw[tdplot_rotated_coords,grid] (x0) -- (reach);
+        \draw[tdplot_rotated_coords,grid] (x0) -- ++(-45:1.5);
+
+        \tdplotdrawarc[tdplot_rotated_coords,grid]{(x0)}{0.5}{-45}{90}{above}{\(\theta_d\)};
+
+        % draw terminal state on reachability set
+        \node[tdplot_rotated_coords,label=above:\(\vecbf{x}_f\)] (xf) at ($ (x0)+(-45:1.5) $) {};
+        \filldraw (xf) circle (3pt);
+
+        \node[tdplot_rotated_coords,label=below:\(J\)] at (xf) {};
+
+        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
+        % place
+    \end{tikzpicture}
+        \end{scaletikzpicturetowidth}
+        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
+            The terminal state, \( x_n\), is the intersection without any control input. 
+            Adding a control input allows for the terminal state, \( x_f \), to be displaced by some distance/cost \( J \) as measured on the section.
+            We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
+            Computation of the maximum distance, or reacahbility, for a variety angles gives a discrete approximation of the reachability set.}
+\end{figure}
+
+\begin{figure}[H]
+        \centering
+        \begin{scaletikzpicturetowidth}{1\textwidth}
+                \begin{tikzpicture}[scale=\tikzscale]
+                    % \draw[help lines] (0,0) grid (20,5) ; %grid
+
+                    \node[label=west:\(\vecbf{x}\)] (x0m) at (0,5) {};
+                    \node[] (x1m) at (5,5) {};
+                    \node[] (x2m) at (10,5) {};
+                    \node[] (x3m) at (15,5) {};
+                    \node[] (xnm) at (20,5) {};
+
+                    \node[label=below:\(k_0\)] (t0) at (0,0) {};
+                    \node[label=below:\(k_1\)] (t1) at (5,0) {};
+                    \node[label=below:\(k_2\)] (t2) at (10,0) {};
+                    \node[label=below:\(k_{n-1}\)] (t3) at (15,0) {};
+                    \node[label=below:\(k_n\)] (tn) at (20,0) {};
+
+                    % nodes for each subsegment
+                    \node[label=left:\(\vecbf{x}_0\)] (x0) at (0,1) {\pgfuseplotmark{*}};
+                    \node[label=below left:\(\vecbf{\lambda}_0\)] at (x0) {};
+
+                    \node[label=right:\(\vecbf{x}_1^{-}\)] (x1minus) at (5,2) {\pgfuseplotmark{*}};
+                    \node[label=below right:\(\vecbf{\lambda}_1^{-}\)] at (x1minus) {};
+
+                    \node[label=above left:\(\vecbf{x}_1^{+}\)] (x1plus) at (5,3) {\pgfuseplotmark{o}};
+                    \node[label=left:\(\vecbf{\lambda}_1^{+}\)] at (x1plus) {};
+
+                    \node[label=right:\(\vecbf{x}_2^{-}\)] (x2minus) at (10,2) {\pgfuseplotmark{*}};
+                    \node[label=below right:\(\vecbf{\lambda}_2^{-}\)] at (x2minus) {};
+
+                    \node[label=above left:\(\vecbf{x}_2^{+}\)] (x2plus) at (10,3) {\pgfuseplotmark{o}};
+                    \node[label=left:\(\vecbf{\lambda}_2^{+}\)] at (x2plus) {};
+
+                    \node[label=right:\(\vecbf{x}_{n-1}^{-}\)] (x3minus) at (15,2) {\pgfuseplotmark{*}};
+                    \node[label=below right:\(\vecbf{\lambda}_{n-1}^{-}\)] at (x3minus) {};
+
+                    \node[label=above left:\(\vecbf{x}_{n-1}^{+}\)] (x3plus) at (15,3) {\pgfuseplotmark{o}};
+                    \node[label=left:\(\vecbf{\lambda}_{n-1}^{+}\)] at (x3plus) {};
+
+                    \node[label=right:\(\vecbf{x}_n\)] (xnminus) at (20,3) {\pgfuseplotmark{*}};
+                    \node[label=below right:\(\vecbf{\lambda}_n\)] at (xnminus) {};
+                    % draw axes
+                    \draw [<->,thick] (tn.center) -- (t0.center) -- (x0m.center);
+
+                    % draw segement dividers
+                    \draw [dashed,thick] (t1.center) -- (x1m.center);
+                    \draw [dashed,thick] (t2.center) -- (x2m.center);
+                    \draw [dashed,thick] (t3.center) -- (x3m.center);
+                    \draw [dashed,thick] (tn.center) -- (xnm.center);
+
+                    % draw markers at each subsegment
+
+                    % draw the subsegments
+                    \draw [->] (x0.center) to [bend left=30] (x1minus.center);
+                    \draw [->] (x1plus.center) to [bend left=30] (x2minus.center);
+                    % dotted lines here
+                    \draw[decorate sep={1mm}{5mm},fill] (x2plus.east) to [bend left=10] (x3minus.west);
+
+                    \draw [->] (x3plus.center) to [bend left=30] (xnminus.center);
+                \end{tikzpicture}
+        \end{scaletikzpicturetowidth}
+        \caption{Schematic diagram of the multiple shooting method:
+        The complete trajectory is split into a nubmer of sub-segments, and additional interior constraints are included to ensure state and costate continunity.
+    Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.}
+\end{figure}
+
+\begin{figure}[H]
+   \centering
+   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
+   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstraiting the initial and target orbit.
+   Without the low-thrust proplusion, the spacecraft is constrained to the initial periodic orbit. 
+   We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
+\end{figure}
+
+\begin{figure}[H]
+        \centering 
+        \begin{subfigure}[htbp]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{manifold_trajectory} 
+                \caption{Position space view of invariant manifold} \label{fig:manifold_trajectory} 
+        \end{subfigure}~ %
+        \begin{subfigure}[htbp]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{manifold_poincare} 
+                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} \label{fig:manifold_poincare} 
+        \end{subfigure} 
+        \caption{Invariant manifold transfer: An example transfer using the invariant manifolds is shown in both the position and \Poincare spaces.
+        The control free transfer from the initial periodic orbit to the target orbit result in a long time of flight. 
+    In addition, the manifold only intersects the target orbit on the descending or far side of the moon.}
+\end{figure}
+\begin{figure}[H]
+        \centering 
+        \begin{subfigure}[htbp]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{reach_trajectory} 
+                \caption{Position space view of reachability set}  
+        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+        \begin{subfigure}[htbp]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{poincare_compare} 
+                \caption{\Poincare section view of reachability set}  
+        \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
+
+        \begin{subfigure}[htbp]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{reach_transfer} 
+                \caption{Transfer trajectory selected from reachabiltiy set viewed in the position space}  
+        \end{subfigure}~ 
+        \begin{subfigure}[htbp]{0.5\textwidth} 
+                \includegraphics[width=\textwidth]{control_input_l1} 
+                \caption{Control input for the selected transfer trajectory}  
+        \end{subfigure}~
+        \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
+        The reachability set is shown in the upper two figures in both the position and \Poincare space.
+    From this reachabiilty set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
+The optimal control to achieve this transfer is shown in the lower right figure.} 
 \end{figure}

### 2017-07-19 manuscript.tex
Update more captions for the figures
Reviewer 2 Comment 7

@@
         \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
-        \label{fig:reachability_set}}
+            The terminal state, \( x_n\), is the intersection without any control input. 
+            Adding a control input allows for the terminal state, \( x_f \), to be displaced by some distance/cost \( J \) as measured on the section.
+            We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
+            Computation of the maximum distance, or reacahbility, for a variety angles gives a discrete approximation of the reachability set.\label{fig:reachability_set}}
 \end{figure}
@@
         \end{scaletikzpicturetowidth}
-        \caption{Multiple Shooting method\label{fig:multiple shooting}}
+        \caption{Schematic diagram of the multiple shooting method:
+        The complete trajectory is split into a nubmer of sub-segments, and additional interior constraints are included to ensure state and costate continunity.
+    Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.\label{fig:multiple shooting}}
 \end{figure}
@@
    \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
-   \caption{Target Orbit Region}
+   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstraiting the initial and target orbit.
+   Without the low-thrust proplusion, the spacecraft is constrained to the initial periodic orbit. 
+   We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
    \label{fig:moon_orbit}
@@
                 \includegraphics[width=\textwidth]{manifold_trajectory} 
-                \caption{Invariant Manifold} \label{fig:manifold_trajectory} 
+                \caption{Position space view of invariant manifold} \label{fig:manifold_trajectory} 
         \end{subfigure}~ %
@@
                 \includegraphics[width=\textwidth]{manifold_poincare} 
-                \caption{Poincar\`e Section} \label{fig:manifold_poincare} 
+                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} \label{fig:manifold_poincare} 
         \end{subfigure} 
-        \caption{Invariant manifold transfer}
+        \caption{Invariant manifold transfer: An example transfer using the invariant manifolds is shown in both the position and \Poincare spaces.
+        The control free transfer from the initial periodic orbit to the target orbit result in a long time of flight. 
+    In addition, the manifold only intersects the target orbit on the descending or far side of the moon.}
         \label{fig:invariant_manifold_transfer} 
@@
                 \includegraphics[width=\textwidth]{reach_trajectory} 
-                \caption{Reachability set trajectories} \label{fig:reach_trajectory} 
+                \caption{Position space view of reachability set} \label{fig:reach_trajectory} 
         \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
@@
                 \includegraphics[width=\textwidth]{poincare_compare} 
-                \caption{\Poincare section} \label{fig:poincare_compare} 
+                \caption{\Poincare section view of reachability set} \label{fig:poincare_compare} 
         \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
@@
                 \includegraphics[width=\textwidth]{reach_transfer} 
-                \caption{Final transfer trajectory} \label{fig:reach_transfer} 
+                \caption{Transfer trajectory selected from reachabiltiy set viewed in the position space} \label{fig:reach_transfer} 
         \end{subfigure}~ 
@@
                 \includegraphics[width=\textwidth]{control_input_l1} 
-                \caption{Control input} \label{fig:control_l1} 
+                \caption{Control input for the selected transfer trajectory} \label{fig:control_l1} 
         \end{subfigure}~
-        \caption{Reachability set transfer}
+        \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
+        The reachability set is shown in the upper two figures in both the position and \Poincare space.
+    From this reachabiilty set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
+The optimal control to achieve this transfer is shown in the lower right figure.}
         \label{fig:reachability_set_transfer} 

### 2017-07-20 manuscript.tex
Add more detail to start of geo transfer

@@
 There are many situations where a more complicated and extensive orbital transfer is desired. 
-For example, we present a simulation of transferring from a geostationary orbit to the a periodic orbit about the Moon.
+In this section, we present a simulation of transferring from a geostationary orbit to a periodic orbit about the \( L_1 \) Lagrange point.
+From this location, it is possible to transition to the Moon, as shown in~\cref{sec:periodic_orbit_transfer}, or beyond the Earth-Moon system through the use of invariant manifolds~\cite{koon2011}.
 This type of low energy transfer would be most suitable for unmanned spacecraft transitioning from the Earth to the Moon.
 The long time of flight would make such a transfer unsuitable for manned missions due to life support constraints.
-%This type of low energy transfer is best suited for autonomous spacecraft missions as the long time of flight is undesirable for manned spacecraft. 
 Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
@@
 
-The previous example demonstrated the capability of determining an orbital transfer after determining the intersection of the reachability set and a target orbit.
+\Cref{sec:periodic_orbit_transfer} demonstrated the capability of determining an orbital transfer after locating the intersection of the reachability set and a target orbit.
 However, it may not be possible to achieve an intersection on the \Poincare section after a single iteration. 
@@
 In this manner, it is possible to determine more complicated transfers by a simple selection of states from the reachability set.
+This affords a systematic method of determining general transfer trajectories in the three body problem.
 

### 2017-07-20 Response.tex
Update figure caption
Reviewer 2 Comment 7

@@
 \end{figure}
+
+\begin{figure}[H]
+   \centering
+   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
+   \caption{Geostationary to \( L_1 \) transfer: Representation of the initial and target orbits for the reachability transfer. 
+   Vehicle is assumed to begin on the geostationary orbit about \( m_1\) and will transfer to the periodic orbit about \( L_1\)}
+\end{figure}
 \end{correction}

### 2017-07-20 manuscript.tex
Update figure caption
Reviewer 2 Comment 7

@@
    \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
-   \caption{Transfer target}
+   \caption{Geostationary to \( L_1 \) transfer: Representation of the initial and target orbits for the reachability transfer. 
+   Vehicle is assumed to begin on the geostationary orbit about \( m_1\) and will transfer to the periodic orbit about \( L_1\)}
    \label{fig:geo_transfer_target}
 \end{figure}
-In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
-We seek to generate a transfer from the geostationary orbit onto the stable invariant manfiold.
+In~\cref{fig:geo_transfer_target}, the initial orbit is located in the center of the figure and centered about \( m_1 \) while the target periodic orbit is determined about the \( L_1 \) lagrange point.
+Instead of determining a transfer directly between the geostationary orbit and the periodic orbit, we seek to instead transfer to the stable manifold of the periodic orbit.
+This stable manifold will then asymptotically approach the target orbit without any further control input.
 Once on the manifold, the spacecraft will coast in an uncontrolled fashion and asymptotically arrive at the desired periodic orbit.

### 2017-07-20 manuscript.tex
Add more detailed figures for the geo transfer

@@
     \centering
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_trajectory_zoom.pdf} 
+        \caption{Stage 1 Trajectory~\label{fig:stage1_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_poincare.pdf} 
+        \caption{Stage 1 \Poincare~section \label{fig:stage1_poincare}} 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_trajectory_zoom.pdf} 
+        \caption{Stage 2 Trajectory~\label{fig:stage2_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_poincare.pdf} 
+        \caption{Stage 2 \Poincare~section \label{fig:stage2_poincare}} 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_trajectory_zoom.pdf} 
+        \caption{Stage 3 Trajectory~\label{fig:stage3_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_poincare.pdf} 
+        \caption{Stage 3 \Poincare~section \label{fig:stage3_poincare}} 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
+        \caption{Stage 4 Trajectory~\label{fig:stage4_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage4_poincare.pdf} 
+        \caption{Stage 4 \Poincare~section \label{fig:stage4_poincare}} 
+    \end{subfigure}   
+    \caption{Stage 1-4 reachability sets: The first four reachability sets visualized in both the position (left) and \Poincare space (right).
+        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
+    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit~\label{fig:stage1to4_reachability}}
+\end{figure}
+
+\begin{figure}[htbp]
+    \centering
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
+        \caption{Stage 5 Trajectory~\label{fig:stage5_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_poincare.pdf} 
+        \caption{Stage 5 \Poincare~section \label{fig:stage5_poincare}} 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_trajectory_zoom.pdf} 
+        \caption{Stage 6 Trajectory~\label{fig:stage6_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_poincare.pdf} 
+        \caption{Stage 6 \Poincare~section \label{fig:stage6_poincare}} 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_trajectory_zoom.pdf} 
+        \caption{Stage 7 Trajectory~\label{fig:stage7_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_poincare.pdf} 
+        \caption{Stage 7 \Poincare~section \label{fig:stage7_poincare}} 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_trajectory_zoom.pdf} 
+        \caption{Stage 8 Trajectory~\label{fig:stage8_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_poincare.pdf} 
+        \caption{Stage 8 \Poincare~section \label{fig:stage8_poincare}} 
+    \end{subfigure}   
+    \caption{Stage 5-8 reachability sets: The last four reachability sets visualized in both the position (left) and \Poincare space (right).
+        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
+    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit~\label{fig:stage1to4_reachability}}
+\end{figure}
+\begin{figure}[htbp]
+   \centering
    \includegraphics[width=0.5\textwidth]{third_reachability_set} % requires the graphicx package

### 2017-07-20 manuscript.tex
Working on geostationary orbit clarity

@@
 In addition, determining the intersections between various invariant manifolds is not trivial. 
-It requires an appropriate \Poincare section and the generation of several invariant manifolds.
+It requires an appropriate \Poincare~section and the generation of several invariant manifolds.
 There is no clear method of selecting the periodic orbits required based on a the type of transfer desired. 
@@
 % describe initial and target states for this example
-This numerical simulation demonstrates the ability of the proposed approach. 
+This numerical simulation demonstrates the ability of the our approach to determine more general transfer trajectories in the PCRTBP. 
 We use multiple iterations of the reachability set to achieve a more complicated transfer.
@@
 Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional Earth-Moon three-body system. 
+While the geostationary orbit is not a typical ``parking orbit'' for most spacecraft missions it serves as a convienent altitude to demonstrate our approach.
 The geostationary orbit about the Earth is transformed into the rotating reference frame of the Earth-Moon three body problem.

### 2017-07-21 manuscript.tex
Add some detail for the first stage

@@
 
+We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
+The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in the position space visualizations given in~\cref{fig:stage1to4_reachability,fig:stage5to8_reachability,fig:geo_transfer_full}.
+The stable manifold is denoted by the green trajecotries which span the majority of the phase space. 
+In addition, we plot the intersection of the stable manifold with the \Poincare section with green markers.
+The initial orbit and the stable invariant manifold do not intersect and the transfer objective is to enlarge the reachability set to include the stable manifold.
+
+Next, we compute the reachable sets originating on the geo-stationary orbit and subject to the maximum control constraint. 
+The multiple shooting approach is used to solve the optimal control problem with two segments.
+For the first reachability set computation, the initial geostationary orbit is used to initialize the multiple shooting algorithm.
+We use the same initial guess of \( \vecbf{\lambda}_0\) as in~\cref{sec:periodic_orbit_transfer} and propogate over the fixed time horizon of the period of the geostationary orbit to initialize \( \vecbf{\lambda}_i^{-}\).
+
 \begin{figure}[htbp]
@@
 \end{figure}
-
-\begin{figure}[htbp]
-    \centering
-    \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
-        \caption{Stage 5 Trajectory~\label{fig:stage5_trajecotry_zoom}} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_poincare.pdf} 
-        \caption{Stage 5 \Poincare~section \label{fig:stage5_poincare}} 
-    \end{subfigure}
-
-    \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_trajectory_zoom.pdf} 
-        \caption{Stage 6 Trajectory~\label{fig:stage6_trajecotry_zoom}} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_poincare.pdf} 
-        \caption{Stage 6 \Poincare~section \label{fig:stage6_poincare}} 
-    \end{subfigure}
-
-    \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_trajectory_zoom.pdf} 
-        \caption{Stage 7 Trajectory~\label{fig:stage7_trajecotry_zoom}} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_poincare.pdf} 
-        \caption{Stage 7 \Poincare~section \label{fig:stage7_poincare}} 
-    \end{subfigure}
-
-    \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_trajectory_zoom.pdf} 
-        \caption{Stage 8 Trajectory~\label{fig:stage8_trajecotry_zoom}} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_poincare.pdf} 
-        \caption{Stage 8 \Poincare~section \label{fig:stage8_poincare}} 
-    \end{subfigure}   
-    \caption{Stage 5-8 reachability sets: The last four reachability sets visualized in both the position (left) and \Poincare space (right).
-        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
-    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit~\label{fig:stage1to4_reachability}}
-\end{figure}
-\begin{figure}[htbp]
-   \centering
-   \includegraphics[width=0.5\textwidth]{third_reachability_set} % requires the graphicx package
-   \caption{Reachable set after three iterations}
-   \label{fig:third_reachability_set}
-\end{figure}
-Next, we generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
-The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
-Next, we compute the reachable sets originating on the geo-stationary orbit and subject to the maximum control constraint. 
-The multiple shooting approach is used to solve the optimal control problem with two segments.
-For the first reachability set computation, the initial geostationary orbit is used to initialize the multiple shooting algorithm.
-We use the same initial guess of \( \vecbf{\lambda}_0\) as in~\cref{sec:periodic_orbit_transfer} and propogate over the fixed time horizon of the period of the geostationary orbit to initialize \( \vecbf{\lambda}_i^{-}\).
-
 The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_{max} = 0.75 \).
@@
 This allows us to approximate the set of states that are achievable in the \( \parenth{x ,\dot{x}} \) space.
+The intersection of the first reachable set with the \Poincare section is shown in~\cref{fig:stage1_poincare} in red.
+While the reachable set does not intersect the stable manifold, it does reduce some of the distance in the \(\dot{x}\) dimension.
+From this reachable set, we chose a trajectory which minimizes the distance towards the stable manifold, which is shown in~\cref{fig:stage1_poincare} by the black marker.
+The minimum trajectory is also shown in~\cref{fig:stage1_trajecotry_zoom} and is used to initilize the following stage.
+Over the relatively short time horizon of the geostationary orbital period, the reachability set remains quite close to the initial orbit. 
+However, the addition of the control input has increased the velocity component of the trajectory.
+It is this velocity change that is subsequently exploited to generate the transfer trajectory.
+
 After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds as the function \( d \)
@@
 
+\begin{figure}[htbp]
+    \centering
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
+        \caption{Stage 5 Trajectory~\label{fig:stage5_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_poincare.pdf} 
+        \caption{Stage 5 \Poincare~section \label{fig:stage5_poincare}} 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_trajectory_zoom.pdf} 
+        \caption{Stage 6 Trajectory~\label{fig:stage6_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_poincare.pdf} 
+        \caption{Stage 6 \Poincare~section \label{fig:stage6_poincare}} 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_trajectory_zoom.pdf} 
+        \caption{Stage 7 Trajectory~\label{fig:stage7_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_poincare.pdf} 
+        \caption{Stage 7 \Poincare~section \label{fig:stage7_poincare}} 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_trajectory_zoom.pdf} 
+        \caption{Stage 8 Trajectory~\label{fig:stage8_trajecotry_zoom}} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.5\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_poincare.pdf} 
+        \caption{Stage 8 \Poincare~section \label{fig:stage8_poincare}} 
+    \end{subfigure}   
+    \caption{Stage 5-8 reachability sets: The last four reachability sets visualized in both the position (left) and \Poincare space (right).
+        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
+    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit~\label{fig:stage5to8_reachability}}
+\end{figure}
+\begin{figure}[htbp]
+   \centering
+   \includegraphics[width=0.5\textwidth]{third_reachability_set} % requires the graphicx package
+   \caption{Reachable set after three iterations}
+   \label{fig:third_reachability_set}
+\end{figure}
+
 With each reachable set we move the controlled trajectory closer to the target stable manifold.

### 2017-07-21 manuscript.tex
Add more detail to manuscript
More discussion about the geostationary transfer

@@
 
+The minimum trajectory from the first iteration of the reachabiilty set is used to initial the following stage.
+From the terminal state of the first iteration, the second reachability set is computed and displayed in~\cref{fig:stage2_trajecotry_zoom,fig:stage2_poincare}.
+The second reachability set is shown on the \Poincare section in~\cref{fig:stage2_poincare}.
+This iteration greatly decreases the distance in the \( x \) component between the trajectory and the target, at the expense of a small deviation in the \( \dot{x} \) component.
+In~\cref{fig:stage2_trajecotry_zoom}, the first stage trajectory is shown in red while the additional stage is shown in blue, which demonstrates the decrease in the \( x \) component.
+
 After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds as the function \( d \)
@@
 Once the reachability set intersects the stable manifold, we can achieve a complete transfer from the initial geostationary orbit to the stable manifold of \( L_1 \).
-%A state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
-For example,~\cref{fig:third_reachability_set} shows the reachable set after three iterations in red and the minimum distance to the stable manifold in green.
-The state and costate trajectories associated with this point on the rechable set is used to initialize the following iteration. 
-In this manner, the multiple shooting algorithm solves the optimal control problem and generates reachable sets which move away from the intial orbit and towards the desired target.
 
@@
 \end{figure}
-\begin{figure}[htbp]
-   \centering
-   \includegraphics[width=0.5\textwidth]{third_reachability_set} % requires the graphicx package
-   \caption{Reachable set after three iterations}
-   \label{fig:third_reachability_set}
-\end{figure}
 
 With each reachable set we move the controlled trajectory closer to the target stable manifold.
-\Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
+\Cref{fig:stage1to4_reachability,fig:stage5to8_reachability} shows all of the intermediate reachabiilty iterations to transfer between the initial orbit and the stable invariant manifold.
+The complete transfer trajectory, after eight iterations, is shown in~\Cref{fig:geo_transfer}. 
 Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
+Each iteration systematically moves the reachable set towards the stable manifold. 
+Furthermore, the optimal control formulation is simplified as each iteration is initialized using a simple distance metric on the \Poincare section.
 \Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory, with the final trajectory shown in red which ensures the intersection with the stable invariant manifold.
+\Cref{fig:geo_transfer_poincare} shows the \Poincare section with the minimum reachable state from each iteration.
+Each iteration seeks to decrease the distance to the stable manifold on the \Poincare section.
+Furthermore, these minimum states serve to initialize each subsequent stage of the transfer. 
+In this manner, there is a systematic and simple methodology to determine transfer trajectories.
 Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
 \Cref{fig:control_input_geo} shows the control input during the powered portion of the transfer. 
+We utilize the same spacecraft assumption of \SI{500}{\kilo\gram} from~\cref{sec:periodic_orbit_transfer} which gives a maximum thrust of approximately \SI{1}{\newton}.
 The spacecraft maintains a bounded control magnitude during the transfer to the stable manifold.
@@
         \end{subfigure} 
-        \caption{Transfer from geostationary orbit to the \( L_1 \) stable manifold using multiple iterations of the reachability set}
+        \caption{Geostationary to \( L_1 \) periodic orbit transfer: }
         \label{fig:geo_transfer} 

### 2017-07-21 manuscript.tex
Update figure caption
Use captionbox and longer more descriptive captions

@@
         \centering 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{geo_transfer_full} 
-                \caption{Completed transfer trajectory} \label{fig:geo_transfer_full} 
-        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{geo_transfer_zoom} 
-                \caption{Detailed view of transfer} \label{fig:geo_transfer_zoom} 
-        \end{subfigure}
+        \subcaptionbox{Transfer trajectory: Complete trajectory consisting of eight iterations (1-7 black and final in red) of the reachability sets with a final control-free coast (blue)  on the stable manifold (green).\label{fig:geo_transfer_full}}{\includegraphics[width=0.45\textwidth]{geo_transfer_full}}\hfill
+        \subcaptionbox{Detailed view of transfer: Centered at \(m_1\) the eight
+            iterations from the reachability analysis are shown in black.  The
+            final segment to the stable manifold is shown in red.  Each
+            iteration progressively approaches the stable manifold.  Once on
+        the stable manifold the vehicle can coast without thrust towards the
+    periodic orbit.\label{fig:geo_transfer_zoom}
+}{\includegraphics[width=0.45\textwidth]{geo_transfer_zoom}}
 

### 2017-07-21 manuscript.tex
Better captions for last figure

@@
 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{poincare} 
-                \caption{\Poincare section} \label{fig:geo_transfer_poincare} 
-        \end{subfigure}~%
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{control_input_geo} 
-                \caption{Control input} \label{fig:control_input_geo} 
-        \end{subfigure} 
-        \caption{Geostationary to \( L_1 \) periodic orbit transfer: }
-        \label{fig:geo_transfer} 
+\subcaptionbox{\Poincare section: Minimum distance states from each iteration of the reachability set analysis (red) with respect to the stable invariant manifold (green)\label{fig:geo_transfer_poincare}}{\includegraphics[width=0.45\textwidth]{poincare}}
+    \hfill
+    \subcaptionbox{Control Input: Combined control history for each stage of the reacability analysis. 
+    The control always remains within the maximum bound.\label{fig:control_input_geo}}{\includegraphics[width=0.45\textwidth]{figures/control_input_geo.pdf}}
+        \caption{Geostationary to \( L_1 \) periodic orbit transfer: Complete transfer from the geostationary orbit to the stable manifold.\label{fig:geo_transfer}}
 \end{figure}

### 2017-07-21 manuscript.tex
Update discussion of geo transfer

@@
 Furthermore, these minimum states serve to initialize each subsequent stage of the transfer. 
-In this manner, there is a systematic and simple methodology to determine transfer trajectories.
+This method provides a systematic and simple methodology to determine transfer trajectories.
+Instead of relying purely on numerical optimization to find an appropriate trajectory, our method instead utilizes the reachability set to determine suitable trajectories.
 Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
@@
 
-This approach demonstrates the ability to link several computations of the reachability set to enable a more general transfer.
-We use eight iterations in order to transfer from the geostationary orbit to the stable manifold.
+This numerical example demonstrates the ability to link several computations of the reachability set to enable a more general transfer.
+We use eight iterations of computing the reachability set in order to transfer from the geostationary orbit to the stable manifold.
 This approach allows for a larger class of potential transfers which leverage the capabilities of low-thrust propulsion systems.

### 2017-07-21 manuscript.tex
Update conclusion

@@
 \section{Conclusions}\label{sec:conclusion}
-In this paper, an optimal transfer process which combines concepts of reachability and \Poincare section is used to generate transfer between planar periodic orbits in the three-body problem.
-The \Poincare section allows for trajectory design on a lower dimensional phase space and simplifies the process.
+In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
+The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
 The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
+Direct optimal control techniques must rely entirely on the ability of the numerical optimzation routine to determine a feasible solution.
+Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
+Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
 However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.

### 2017-07-22 Response.tex
Add additional detail to response letter
Copying over detail and discussing the changes in more detail

@@
 
-The manuscript has been reduced and much of the mathematical development has been removed.
+    Thank you for the comment, and we agree that there is significant repetition of material previously covered in other sources.
+    The manuscript has been reduced and reorganized to enhance the clarity and clearly seperate previously material and the contributions of this paper.
+    The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
+    The second section has been retitiled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
+The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, followed by conclusions.
+The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.
 
@@
 
-These pros/cons have been addressed in the conclusion.
+Thank you for the comment and more focus has been included on the pros and cons of this proposed method.
+Additional detail has been added to the conclusion which discusses the advantages and the disadvantages, as well as the avenues for future research.
+The conclusion is copied below.
+\begin{correction}
+In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
+The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
+The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
+Direct optimal control techniques must rely entirely on the ability of the numerical optimzation routine to determine a feasible solution.
+Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
+Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
+However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
+Furthermore, our approach relies heaviliy on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
+
+There is additional research to extend these results to more general transfer scenarios in future work.
+The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
+In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
+Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
+This will result in a more complicated optimal control problem and is a focus of future research.
+Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
+The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
+\end{correction}
 
@@
 
+Thank you for the comment and these characteristics of our choice of the PCRTBP.
 Additional detail has been included in the problem formulation section.
@@
 
-Additional detail has  been included in the manuscript under the section describing the variational integrator.
-These details are discussed further under the more specific comments below.
-
+Thank you for the comment and the detail on the benefits/drawbacks of variatonal integrators.
+Additional discussion has been included in the paper in the subsection which introduces the variational integrator for the PCRTBP.
+Much of the discussion of variational integrators in the introduction has also been removed.
+The additional content, which discusses the relative benefits of variational integrators is copied below.
+
+\begin{correction}
+Both integration schemes result in trajectories that are initially nearly identical.
+\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
+For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the convential integration scheme.
+Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
+In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
+This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
+However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
+In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
+The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.
+\end{correction}
 \end{itemize}

### 2017-07-22 Response.tex
Add more detail to responses

@@
 
-        The introductory paragraphs have been reworded and reorganized for clarity. 
+        Thank you for the comment and the additional reference.
+        In addition to this change we have made other modifications to the introductory paragraphs to improve the organization and clarity.
         Specifically, the first paragraph has been divided, which motivates the use of electric propulsion and the advantages of small satellites which use electric propulsion.
@@
 
+Thank you for the comment.
 The sentence has been modified as follows.
@@
 
-Additional detail has been included which quantifies the notion of long-term.
-
+Thank you for the comment. 
+The claims about the variational integrator have been qualified with some additional detail.
+The correction is copied below.
 \begin{correction}

### 2017-07-22 Response.tex
Update response comments

@@
 
+Thank you for the comment.
+Evidently in some styles, such as \href{https://www.dlsweb.rmit.edu.au/lsu/content/1_studyskills/study_tuts/apa_ll/position.html}{APA}, it is acceptable to start with a reference.
+We agree that it is not the best with the numeric style citations.
 The sentences have been reworded to avoid starting with a reference number. 
@@
 
+Thank you for the clarification.
 The sentence has been modified as follows.
@@
 
+Thank you for the comment.
+The PCRTBP was primarily chosen as a simple dynamic system which still captures the primary structure of the three-body problem.
 The introduction has been modified with an additional sentence indicating our use of the planar three body problem
@@
 
+Thank you for the comment.
 Some additional detail has been included mentioning the difficulties in applying the \Poincare~section approach to the 3D case.
@@
 
+Thank you for the comment.
 The section on the variational integrator has been reduced and this specific sentence has been removed. 
-However, additional detail has been provided in a different section of the paper which discusses this drawback of the variational integrator.
+However, additional detail has been provided in a the variational integrator subsection which discusses this drawback.
 
@@
 
+Thank you for the comment.
 This sentence has been removed from the introduction.
-However, this is a valid comment and is noted for the future.
 
@@
 
+Thank you for the comment and helpful reference.
 Additional detail has been included in the manuscript discussing this potential drawback of variable step integrators.
+j
 \begin{correction}
@@
 \end{correction}
+
 \item 
@@
 
+Thank you for the comment.
+The suboptimal claims of direct optimization has been reduced and modified.
 Additional detail has been included on the merits of indirect optimal control vs. direct optimal control methods.
@@
 
+Thank you for the comment.
 An additional comment has been included in the problem formulation section with a reference to a source of the detailed derivation.
@@
 
+Thank you for the comment.
 An additional comment has been included mentioning that the control is assumed to be bounded in magnitude.
@@
 
+Thank you for the comment.
 All of the figures showing position states have been modified to use \texttt{axis equal}.
@@
 
+Thank you for the comment.
 The discussion on the variational integrators has been reduced. 
 Much of the background material on Lagrangian mechanics is removed and an additional paragraph is included to summarize the key ideas of variational integrators. 
+In addition, the variational integrator details are moved into a titled subsection.
 The additional paragraph is shown below.
@@
 We chose the trapezoidal rule as it provides a second order quadrature approximation.
-In addition, it provides an explicit rule that eases the complexity in using an implicit solution method.
+Furthermore, instead of using Taylor methods or higher order options, the trapezoidal rule provides an explicit second order accurate integration rule.
+This greatly reduces the complexity of the optimization routine.
+In addition, the background material on the variational integrators have been removed from the manuscript. 
+Only the specific details covering the development of the variational integrator for the PCRTBP are discussed.
 
@@
 In the interest of conciseness and to avoid repeating material that has already been covered in the literature, the discussion on variational integrators has been substantially reduced.
-As a result, this table has been removed from the manuscript.
+The background material on the variational integrators has been removed and this table has also been removed from the manuscript.
 
@@
 
-The plot has been removed and replaced with another comparision between the variational integrator and \texttt{ODE45}.
-In addition, all of the plots now have axis labels.
-
+Thank you for the comment. 
+All of the plots have been updated to include all axes labels.
+Furthermore, the specific plot in questions has been replaced with another plot more explicitly showing the benefits of variational integrators over conventional integration schemes.
+This modified plot is shown below.
+
+\begin{correction}
+    \begin{figure}[H] 
+        \centering 
+        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparision: A stable trajectory is used to test the variational and conventional integrators. 
+        The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
+        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
+        \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}}
+        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
+\end{figure}
+\end{correction}
 \item 

### 2017-07-23 Response.tex
Add control plots to response

@@
 
+Thank you for the comment.
 Appendix B has been removed from the revised manuscript.
@@
 
+Thank you for the comment.
 The additional clarification has been added to the manuscript and copied below.
@@
 
-An additional paragraph has been added to the beginning of the section which discusses our reachability based transfer approach.
+Thank you for the comment.
+An additional paragraph has been added to the beginning of the mathematical background section which discusses our reachability based transfer approach.
 This additional content is copied below.
@@
 
+Thank you for the comment.
 The mentioned plots of the control input have been modified to show the data in dimensional units (Newtons). 
@@
 The modified plots are not included here but the additional content is copied below in response to a different but related comment.
+
+\begin{correction}
+Next, we determine the reachability set with the addition of a low thrust control input.
+We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
+This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
+We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
+This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
+Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
+Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
+The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
+This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit. 
+\begin{figure}[H]
+    \centering
+    \subcaptionbox{Control input for the selected transfer trajectory}{\includegraphics[width=0.5\textwidth]{control_input_l1}}~
+    \subcaptionbox{Control Input: Combined control history for each stage of the reacability analysis. 
+    The control always remains within the maximum bound.}{\includegraphics[width=0.5\textwidth]{figures/control_input_geo.pdf}}
+\end{figure}
+\end{correction}
 

### 2017-07-23 Response.tex
Update all the responses for reviewer 1

@@
 
+Thank you for the comment.
 Additional detail has been incorporated which discusses the specific solver used in computing the solutions. 
@@
 One key issue was that our implementation in software depended on dividing the total number of integration steps by the number of interior stages.
-For simplicity, this reduced to possible number of stages to an even number. 
+For simplicity, this reduced the possible number of stages to an even number. 
 Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in covergence. 
@@
 
+Thank you for the comment.
 The discussion of generating periodic orbits in the three body problem has been removed.
-It replaced with a small introduction mentioning the differential correction process and a citation to a relevant resource. 
+It is replaced with a small introduction mentioning the differential correction process and a citation to a relevant resource. 
 The revised paragraph is copied below.
@@
 Additional detail has been included in the conclusion discussing this topic.
+The modified conclusion is copied below.
+
+\begin{correction}
+In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
+The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
+The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
+Direct optimal control techniques must rely entirely on the ability of the numerical optimzation routine to determine a feasible solution.
+Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
+Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
+However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
+Furthermore, our approach relies heaviliy on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
+
+There is additional research to extend these results to more general transfer scenarios in future work.
+The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
+In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
+Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
+This will result in a more complicated optimal control problem and is a focus of future research.
+Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
+The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
+\end{correction}
 
@@
 
+Thank you for the comment, howver it seems to be rare for the research community to give detailed description of the numerical optimization method.
 Most of the literature tends to not spend time discussing the low level details of the computational procedure, e.g.~\cite{grebow2011,stuart2010}.
@@
 However, our manuscript is primarily focused on demonstrating the feasbility of using reachability sets to design transfer trajectories.
-As a result, there is not much benefit to detailed statistics of the implementation of this methodology.
+As a result, there is not much benefit to detailed statistics of the implementation of this methodology, and we have omitted it from the manuscript.
 
@@
 \end{itshape}
+
+Thank you for the comment.
 This sentence has been modified to make it more clear that we're discussing low thrust propulsion rather than low-energy transfers.
@@
 
-This is a very useful and interesting suggestion and something that we will take into consideration for future research.
+Thank you for the comment.
+This is a very interesting suggestion and something that we will take into consideration for future research.
 
@@
 
+Thank you for the comment.
 The conclusion has been expanded with some discussion of the drawbacks of this research and avenues of future work.
@@
 \end{enumerate}
+
 \section*{Reviewer 2}

### 2017-07-23 Response.tex
Start updating response to reviewer 2

@@
 \end{itshape}
+
+    Thank you for the comment, and we agree that there is significant repetition of material previously covered in other sources.
+    The manuscript has been reduced and reorganized to enhance the clarity and clearly seperate previously material and the contributions of this paper.
+    The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
+    The second section has been retitiled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
+The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, followed by conclusions.
+The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.
+
 \item 
@@
     \end{itshape}
+
+    Thank you for the comment. 
+    The numerical examples are meant to demonstrate the use of this approach and the feasibility of using the reachability set on a \Poincare section to determine transfer trajecotries.
+    The examples give two demonstrations of finding a transfer trajectory in the PCRTBP and demonstrate combining several iterations of the reachability set for a transfer.
 
@@
 \end{itshape}
+
 \item 
@@
 
+Thank you for the comment.
 The introduction is reduced and many of the paragraphs have been reworded and combined. 
+In addition, the extended discussion of the benefits of variational integrators is removed from the introduction.
 The mathematical background has been condensed into a single section with seperate subsections which discuss some of the mathematical details.
-In addition, the extended discussion of the benefits of variational integrators is removed from the introduction.
+The background section is focused on reviewing the relevant material rather than reintroducing topics.
+In the interest of space the reorganized section is omitted in this response.
 
@@
 
+Thank you for the comment.
 The mathematical background has been reduced and condensed into a single section with multiple subsections.
@@
 \end{itshape}
+
+Thank you for the comment. 
+The introductory detail has been reduced and combined into a single section. 
+As a result, the third section begins with the new contributions which describe the optimal control formulation of the reachability set.
 

### 2017-07-23 Response.tex
Add red for comments that need some help

@@
 
+Thank you for the comment.
 The reference has been noted in the discussion of reachability sets.
@@
 \item
+    {\color{red}
     \begin{itshape}
@@
 \end{itshape}
-
-\item
+}
+
+\item
+    {\color{red}
     \begin{itshape}
@@
 \end{itshape}
+}
 
@@
 We've added additional descriptions for all of the figures.
-The figure captions are shown below.
+The figure captions are shown below, with the figures reduced in size to save some space.
 
@@
         \centering
-        \includegraphics[width=0.5\textwidth]{energy_contours}
+        \includegraphics[width=0.2\textwidth]{energy_contours}
         \caption{Contour plot of Jacobi integral: Zero velocity curves of constant Jacobi integral.
@@
         \centering
-        \begin{scaletikzpicturetowidth}{0.5\textwidth}
+        \begin{scaletikzpicturetowidth}{0.2\textwidth}
     \begin{tikzpicture}[tdplot_main_coords,
@@
         The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
-        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
+        [0.5\textwidth]{\includegraphics[width=0.2\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
         \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}}
-        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
+        [0.5\textwidth]{\includegraphics[width=0.2\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
 \end{figure}
@@
 %       \includegraphics[width=0.5\textwidth]{reachability_set_hand}
-        \begin{scaletikzpicturetowidth}{0.5\textwidth}
+        \begin{scaletikzpicturetowidth}{0.2\textwidth}
         \begin{tikzpicture}[tdplot_main_coords,
@@
         \centering
-        \begin{scaletikzpicturetowidth}{1\textwidth}
+        \begin{scaletikzpicturetowidth}{0.7\textwidth}
                 \begin{tikzpicture}[scale=\tikzscale]
@@
    \centering
-   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
+   \includegraphics[width=0.2\textwidth]{moon_orbit} % requires the graphicx package
    \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstraiting the initial and target orbit.
@@
         \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{manifold_trajectory} 
+                \includegraphics[width=0.5\textwidth]{manifold_trajectory} 
                 \caption{Position space view of invariant manifold} \label{fig:manifold_trajectory} 
@@
         \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{manifold_poincare} 
+                \includegraphics[width=0.5\textwidth]{manifold_poincare} 
                 \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} \label{fig:manifold_poincare} 
@@
         \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{reach_trajectory} 
+                \includegraphics[width=0.5\textwidth]{reach_trajectory} 
                 \caption{Position space view of reachability set}  
@@
         \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{poincare_compare} 
+                \includegraphics[width=0.5\textwidth]{poincare_compare} 
                 \caption{\Poincare section view of reachability set}  
@@
         \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{reach_transfer} 
+                \includegraphics[width=0.5\textwidth]{reach_transfer} 
                 \caption{Transfer trajectory selected from reachabiltiy set viewed in the position space}  
@@
         \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{control_input_l1} 
+                \includegraphics[width=0.5\textwidth]{control_input_l1} 
                 \caption{Control input for the selected transfer trajectory}  
@@
    \centering
-   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
+   \includegraphics[width=0.2\textwidth]{initial_final} % requires the graphicx package
    \caption{Geostationary to \( L_1 \) transfer: Representation of the initial and target orbits for the reachability transfer. 
@@
 \end{figure}
-\end{correction}
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_trajectory_zoom.pdf} 
+        \caption{Stage 1 Trajectory} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_poincare.pdf} 
+        \caption{Stage 1 \Poincare~section } 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_trajectory_zoom.pdf} 
+        \caption{Stage 2 Trajectory~} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_poincare.pdf} 
+        \caption{Stage 2 \Poincare~section } 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_trajectory_zoom.pdf} 
+        \caption{Stage 3 Trajectory} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_poincare.pdf} 
+        \caption{Stage 3 \Poincare~section } 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
+        \caption{Stage 4 Trajectory~} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage4_poincare.pdf} 
+        \caption{Stage 4 \Poincare~section } 
+    \end{subfigure}   
+    \caption{Stage 1-4 reachability sets: The first four reachability sets visualized in both the position (left) and \Poincare space (right).
+        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
+    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
+        \caption{Stage 5 Trajectory~} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_poincare.pdf} 
+        \caption{Stage 5 \Poincare~section } 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_trajectory_zoom.pdf} 
+        \caption{Stage 6 Trajectory} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_poincare.pdf} 
+        \caption{Stage 6 \Poincare~section } 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_trajectory_zoom.pdf} 
+        \caption{Stage 7 Trajectory} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_poincare.pdf} 
+        \caption{Stage 7 \Poincare~section } 
+    \end{subfigure}
+
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_trajectory_zoom.pdf} 
+        \caption{Stage 8 Trajectory} 
+    \end{subfigure}~
+    \begin{subfigure}[htbp]{0.2\textwidth} 
+        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_poincare.pdf} 
+        \caption{Stage 8 \Poincare~section } 
+    \end{subfigure}   
+    \caption{Stage 5-8 reachability sets: The last four reachability sets visualized in both the position (left) and \Poincare space (right).
+        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
+    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
+\end{figure}
+\end{correction}
+
 \item
@@
 
-We disagree as the figure is meant to demonstrate the purpose and use of the \Poincare section.
-The \Poincare section figure has been moved but it is meant to demonstrate in a pictoral form the benefit of this visualization technique.
+Thank you for the comment, however the figure is drawn to show a pictoral representation of a \Poincare section.
 The periodic orbits intersect the section at fixed points, as a result it becomes simple to visualize and determine these fixed solutions by searching for fixed points on the section
@@
 This addition has been copied below.
+
 \begin{correction}
@@
 In practice, the reachable set will be of a general shape and also higher dimensional in the nonplanar case.
-\begin{figure}
+\begin{figure}[H]
         \centering
 %       \includegraphics[width=0.5\textwidth]{reachability_set_hand}
-        \begin{scaletikzpicturetowidth}{0.5\textwidth}
+        \begin{scaletikzpicturetowidth}{0.3\textwidth}
         \begin{tikzpicture}[tdplot_main_coords,
@@
 \end{correction}
+
 \item
@@
 \end{itshape}
+}
 \end{enumerate}

### 2017-07-23 Response.tex
Commit at home

--- 2017-07-23 Response.tex
@@
 
+Thank you for the comment.
 The discussion about the invariant manifolds and \Poincare sections has been reduced.
 This discussion has been removed from the revised manuscript and replaced with a general summary and some relevant citations.
+This specific discussion no longer exists in the manuscript but rather cites several references with the details on determining the invariant manifolds.
 
@@
 
+Thank you for the comment.
 The comparison plot has been modified to follow the precedent in the literature.
@@
 
-This manuscript is not really focused on variational integrators but rather the design of transfer trajectories using reachability sets.
+Thank you for the comment.
+However, this manuscript is not really focused on variational integrators but rather the design of transfer trajectories using reachability sets.
+There is a wide range of topic under the umbrella of varaitional integrators that are not directly applicable to this manuscript and work.
 Interested readers can find a wide variety of literature on the relative benefits of all manner of variational integrators in papers focused on that topic specifically.
+As a result, we do not feel it is applicable to include additional discussion on the merits of taylor integration methods.
 
@@
 
-Thank you for the comment.
-An additional paragraph has been added to the beginning of the mathematical background section which discusses our reachability based transfer approach.
+An additional paragraph has been added to the beginning of the section which discusses our reachability based transfer approach.
 This additional content is copied below.
@@
 
-Thank you for the comment.
 The mentioned plots of the control input have been modified to show the data in dimensional units (Newtons). 
@@
 
+\item
+    \begin{itshape}
+"The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches."  Did the authors attempt single shooting for their example problems before coming to this conclusion?
+With proper scaling, and problem formulation, there are examples in the literature of single shooting yielding robust results.
+
+Example: 
+Ranieri, C. and Ocampo,C..  "Indirect Optimization of Spiral Trajectories", Journal of Guidance, Control, and Dynamics, Vol. 29, No. 6 (2006), pp. 1360-1366.
+
+So, justification for its selection may be warranted.  You could always find a reference in which this trade space was really explored thoroughly to show why you made the choice.
+\end{itshape}
+
+Our original attempt at solving the optimal control problem did rely on single shooting.
+However, during our investigation it was found that it was quite difficult for solutions to converge.
+As a result, we decided to implement the multiple shooting approach in hopes of generating more consistent and robust solutions.
+While it is possible to use single shooting, such as the provided reference, it is known that the single shooting approach will have greater sensitivity to initial costate errors as compared to a multiple shooting method.
+This can be alleviated, or entirely mitigated in the case of~\cite{ranieri2006}, it is highly case specific and dependent on a large variety of case specific techniques. 
+For example, the work in~\cite{ozimek2010a} used both a single shooting and multiple shooting approach to solve the optimal control problem.
+The large sensitivity of the single shooting approach was mitigated through the use of the multiple shooting method, at the expense of additional design variables.
+
+In the manuscript, additional detail has been added in this section which describes the difficulties of the single shooting approach and our decision to use the multiple shooting method.
+The modified portion of the manuscript is copied here.
+\begin{correction}
+Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs~\cite{stoer2013}.
+This multiple shooting approach incorporates additional interior constraints but reduces the sensitivity of the costates along each sub-arc.
+The use of the multiple shooting method reduces the sensitivity of changes in the initial costate at the expense of additional design variables, but has been shown to provide more stable and robust solutions~\cite{ozimek2010a}.
+\end{correction}
+
+\item
+    \begin{itshape}
+"In this work, we use the Matlab nonlinear solver fsolve to solve the system of nonlinear equations defined by the multiple shooting algorithm."  Can authors give some additional details?
+Which fsolve algorithm?
+trust region dogleg,  trust region, or Levenberg-Marquardt?
+Which additional options?
+E.g.
+various tolerances.
+For users of this method it's also useful to see the iteration history (more comments on this later).
+\end{itshape}
+
+Additional detail has been incorporated which discusses the specific solver used in computing the solutions. 
+The description does not specifically mention all of the possible options in \texttt{fsolve} but rather gives the key components used in our solution.
+This approach seems to be consistent with the literature, e.g.~\cite{ozimek2010a,stuart2010}.
+The modified portion is copied below.
+
+\begin{correction}
+In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
+Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
+All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
+\end{correction}
+
+\item
+    \begin{itshape}
+Did the authors experiment with the number of multiple shooting stages in their example problems?
+There must be some sort of "sweet spot" in number of stages for the example problems attempted.
+If I missed it, I apologize, but I didn't see it.
+\end{itshape}
+
+We did spend considerable time in determining the appropriate number of stages. 
+One key issue was that our implementation in software depended on dividing the total number of integration steps by the number of interior stages.
+For simplicity, this reduced to possible number of stages to an even number. 
+Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in covergence. 
+The solver would tend to ensure a continuous trajectory, by satisfying the interior stages, at the expense of the terminal stage. 
+Some additional detail has been provided that mentions our choice of the number of stages and has been copied below.
+
+\begin{correction}
+Based on experimentation, we use four stages in our multiple shooting method.
+This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
+\end{correction}
+
+\item 
+    \begin{itshape}
+Section 6.1 - I don't think you need to explain how you created the three-body orbits.
+This can be found in many references - You can just cite a reference and give the orbit specifications.
+(Starting with line 49, page 18, up to line 21 page 19.)
+\end{itshape}
+
+The discussion of generating periodic orbits in the three body problem has been removed.
+It replaced with a small introduction mentioning the differential correction process and a citation to a relevant resource. 
+The revised paragraph is copied below.
+
+\begin{correction}
+The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
+The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
+The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
+\Cref{fig:moon_orbit} shows that the target set remains in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
+This type of orbit would be useful for a variety of mission scenarios.
+The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
+The initial set is a planar periodic orbit about \( L_1\), which is generated using the process of differential correction of a linear approximation~\cite{koon2011}.
+\end{correction}
+
+\item
+    \begin{itshape}
+"We define a maximum magnitude of the thrust as umax = 0.75 and assume we can point the thrust in any direction within the plane." And  "we again assume a upper bound on the thrust magnitude of umax = 0.75."  What was the reason behind this?
+Isn't this a pretty large value of dimensional acceleration for your problem of interest?
+Even if it's chosen for demonstrative purposes, there should be some sort of explanation as to why.
+In the context of what electric propulsion systems can deliver, it could be arguable that this value isn't really "low-thrust" but "finite-thrust".
+\end{itshape}
+
+The maximum thrust magnitude was chosen to be approximately \SI{1}{\newton} for an assumed \SI{500}{\kilo\gram} spacecraft. 
+While larger than the NEXT thruster mentioned earlier, a cluster of similar engines could provide the required thrust (assuming adequate available power). 
+Additional detail has been included in the manuscript describing this thrust limit in dimensional units. 
 \begin{correction}
@@
 This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit. 
-\begin{figure}[H]
-    \centering
-    \subcaptionbox{Control input for the selected transfer trajectory}{\includegraphics[width=0.5\textwidth]{control_input_l1}}~
-    \subcaptionbox{Control Input: Combined control history for each stage of the reacability analysis. 
-    The control always remains within the maximum bound.}{\includegraphics[width=0.5\textwidth]{figures/control_input_geo.pdf}}
-\end{figure}
-\end{correction}
-
-\item
-    \begin{itshape}
-"The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches."  Did the authors attempt single shooting for their example problems before coming to this conclusion?
-With proper scaling, and problem formulation, there are examples in the literature of single shooting yielding robust results.
-
-Example: 
-Ranieri, C. and Ocampo,C..  "Indirect Optimization of Spiral Trajectories", Journal of Guidance, Control, and Dynamics, Vol. 29, No. 6 (2006), pp. 1360-1366.
-
-So, justification for its selection may be warranted.  You could always find a reference in which this trade space was really explored thoroughly to show why you made the choice.
-\end{itshape}
-
-Our original attempt at solving the optimal control problem did rely on single shooting.
-However, during our investigation it was found that it was quite difficult for solutions to converge.
-As a result, we decided to implement the multiple shooting approach in hopes of generating more consistent and robust solutions.
-While it is possible to use single shooting, such as the provided reference, it is known that the single shooting approach will have greater sensitivity to initial costate errors as compared to a multiple shooting method.
-This can be alleviated, or entirely mitigated in the case of~\cite{ranieri2006}, it is highly case specific and dependent on a large variety of case specific techniques. 
-For example, the work in~\cite{ozimek2010a} used both a single shooting and multiple shooting approach to solve the optimal control problem.
-The large sensitivity of the single shooting approach was mitigated through the use of the multiple shooting method, at the expense of additional design variables.
-
-In the manuscript, additional detail has been added in this section which describes the difficulties of the single shooting approach and our decision to use the multiple shooting method.
-The modified portion of the manuscript is copied here.
-\begin{correction}
-Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs~\cite{stoer2013}.
-This multiple shooting approach incorporates additional interior constraints but reduces the sensitivity of the costates along each sub-arc.
-The use of the multiple shooting method reduces the sensitivity of changes in the initial costate at the expense of additional design variables, but has been shown to provide more stable and robust solutions~\cite{ozimek2010a}.
-\end{correction}
-
-\item
-    \begin{itshape}
-"In this work, we use the Matlab nonlinear solver fsolve to solve the system of nonlinear equations defined by the multiple shooting algorithm."  Can authors give some additional details?
-Which fsolve algorithm?
-trust region dogleg,  trust region, or Levenberg-Marquardt?
-Which additional options?
-E.g.
-various tolerances.
-For users of this method it's also useful to see the iteration history (more comments on this later).
-\end{itshape}
-
-Thank you for the comment.
-Additional detail has been incorporated which discusses the specific solver used in computing the solutions. 
-The description does not specifically mention all of the possible options in \texttt{fsolve} but rather gives the key components used in our solution.
-This approach seems to be consistent with the literature, e.g.~\cite{ozimek2010a,stuart2010}.
-The modified portion is copied below.
-
-\begin{correction}
-In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
-Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
-All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
-\end{correction}
-
-\item
-    \begin{itshape}
-Did the authors experiment with the number of multiple shooting stages in their example problems?
-There must be some sort of "sweet spot" in number of stages for the example problems attempted.
-If I missed it, I apologize, but I didn't see it.
-\end{itshape}
-
-We did spend considerable time in determining the appropriate number of stages. 
-One key issue was that our implementation in software depended on dividing the total number of integration steps by the number of interior stages.
-For simplicity, this reduced the possible number of stages to an even number. 
-Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in covergence. 
-The solver would tend to ensure a continuous trajectory, by satisfying the interior stages, at the expense of the terminal stage. 
-Some additional detail has been provided that mentions our choice of the number of stages and has been copied below.
-
-\begin{correction}
-Based on experimentation, we use four stages in our multiple shooting method.
-This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
-\end{correction}
-
-\item 
-    \begin{itshape}
-Section 6.1 - I don't think you need to explain how you created the three-body orbits.
-This can be found in many references - You can just cite a reference and give the orbit specifications.
-(Starting with line 49, page 18, up to line 21 page 19.)
-\end{itshape}
-
-Thank you for the comment.
-The discussion of generating periodic orbits in the three body problem has been removed.
-It is replaced with a small introduction mentioning the differential correction process and a citation to a relevant resource. 
-The revised paragraph is copied below.
-
-\begin{correction}
-The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
-The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
-The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
-\Cref{fig:moon_orbit} shows that the target set remains in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
-This type of orbit would be useful for a variety of mission scenarios.
-The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
-The initial set is a planar periodic orbit about \( L_1\), which is generated using the process of differential correction of a linear approximation~\cite{koon2011}.
-\end{correction}
-
-\item
-    \begin{itshape}
-"We define a maximum magnitude of the thrust as umax = 0.75 and assume we can point the thrust in any direction within the plane." And  "we again assume a upper bound on the thrust magnitude of umax = 0.75."  What was the reason behind this?
-Isn't this a pretty large value of dimensional acceleration for your problem of interest?
-Even if it's chosen for demonstrative purposes, there should be some sort of explanation as to why.
-In the context of what electric propulsion systems can deliver, it could be arguable that this value isn't really "low-thrust" but "finite-thrust".
-\end{itshape}
-
-The maximum thrust magnitude was chosen to be approximately \SI{1}{\newton} for an assumed \SI{500}{\kilo\gram} spacecraft. 
-While larger than the NEXT thruster mentioned earlier, a cluster of similar engines could provide the required thrust (assuming adequate available power). 
-Additional detail has been included in the manuscript describing this thrust limit in dimensional units. 
-\begin{correction}
-Next, we determine the reachability set with the addition of a low thrust control input.
-We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
-This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
-We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
-This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
-Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
-Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
-The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
-This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit. 
 \end{correction}
@@
 Additional detail has been included in the conclusion discussing this topic.
+
+\item 
+    \begin{itshape}
+The authors should discuss some statistics of the convergence of their multiple shooting TPBVP - number of iterations, compute time, etc.
+\end{itshape}
+
+Most of the literature tends to not spend time discussing the low level details of the computational procedure, e.g.~\cite{grebow2011,stuart2010}.
+In addition, the number of iterations, computation time and even the eventual cost function value are highly dependent on the specific system and more specifically the implementation of the algorithm in software.
+There are a wide variety of works which are focused more exclusively on the computational implementation of various optimal control techniques which do spend significant time on the computational characteristics.
+However, our manuscript is primarily focused on demonstrating the feasbility of using reachability sets to design transfer trajectories.
+As a result, there is not much benefit to detailed statistics of the implementation of this methodology.
+
+\item 
+    \begin{itshape}
+Can any intuition be gained from reporting the optimal converged cost function values in section 6.1 and 6.2?
+One could at least compare them to compare the reachability in the two different examples.
+\end{itshape}
+
+The cost function only parameterizes the distance between the controlled and uncontrolled terminal states on the \Poincare section. 
+In addition, each state which lies on the reachable set would typically have a different distance and therefore a different coverged cost function value. 
+Furthermore, it would be difficult to characterize a single parameter which characterizes the reachable set, in the form of the cost function, for all states of the reachable set. 
+Also, the cost function is highly dependent on the transit time of trajectory, i.e. a longer transit time corresponds to a larger reachable set and a larger magnitude of the cost function.
+However, this is an interesting suggestions for future research.
+For example, the cost function gives some indication of area/volume of the reachable set. 
+It may be possible to use this value as another way to characterize the subspace on the \Poincare section and is an intriguing avenue of future research.
+
+\item
+    \begin{itshape}
+"Low energy transfers from the Earth to the Lagrange points are necessary for future missions."  One must be careful not to confuse these with traditional "low-energy transfers" to the moon, which use Sun third-body perturbations to transfer into lunar orbit.
+These are what many mission designers define as "low-energy" lunar transfers.
+This is actually a four-body problem (Sun, Earth, moon, spacecraft), and represents additional delta-V savings (as was done with NASA's GRAIL mission.)
+
+Example:
+
+Parker, J. and Anderson, R., "Low-Energy Lunar Trajectory Design," DEEP SPACE COMMUNICATIONS AND NAVIGATION SERIES, Jet Propulsion Laboratory Pasadena, California, July 2013.
+
+Online Copy: https://descanso.jpl.nasa.gov/monograph/series12/LunarTraj--Overall.pdf
+\end{itshape}
+This sentence has been modified to make it more clear that we're discussing low thrust propulsion rather than low-energy transfers.
+The modified sentence is copied below.
+
+\begin{correction}
+Transfers from the Earth to the Lagrange points, through the use of low-thrust electric propulsion, offer an additional and potentially shorter time of flight in comparison to the low-energy transfers which utilize solar perturbations.
+\end{correction}
+
+\item 
+    \begin{itshape}
+For what it's worth, exploiting sun perturbations with low-thrust in the Earth-moon RTBP would be highly beneficial and represent a worthy avenue of future research (more on this below.)
+\end{itshape}
+
+This is a very useful and interesting suggestion and something that we will take into consideration for future research.
+
+\item 
+    \begin{itshape}
+The conclusions of the paper should acknowledge other potentially limiting factors of the presented analysis
+
+-Omits advantage of exploiting fourth body perturbation (Sun in Earth/moon RTBP)
+
+-Transfers are 2D only, method is more challenging to adapt to 3D
+
+-Presented control magnitudes are dimensionally large, and more work is needed to determine feasibility of approach for lower accelerations
+
+-Thrust and mass are completely decoupled from the analysis, which is limiting in terms of finding suitable hardware that can deliver the controls needed, as currently presented.
+\end{itshape}
+
+The conclusion has been expanded with some discussion of the drawbacks of this research and avenues of future work.
 The modified conclusion is copied below.
@@
 \begin{correction}
-In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
-The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
+In this paper, an optimal transfer process which combines concepts of reachability and \Poincare section is used to generate transfer between planar periodic orbits in the three-body problem.
+The \Poincare section allows for trajectory design on a lower dimensional phase space and simplifies the process.
 The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
-Direct optimal control techniques must rely entirely on the ability of the numerical optimzation routine to determine a feasible solution.
-Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
-Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
 However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
@@
 \end{correction}
-
-\item 
-    \begin{itshape}
-The authors should discuss some statistics of the convergence of their multiple shooting TPBVP - number of iterations, compute time, etc.
-\end{itshape}
-
-Thank you for the comment, howver it seems to be rare for the research community to give detailed description of the numerical optimization method.
-Most of the literature tends to not spend time discussing the low level details of the computational procedure, e.g.~\cite{grebow2011,stuart2010}.
-In addition, the number of iterations, computation time and even the eventual cost function value are highly dependent on the specific system and more specifically the implementation of the algorithm in software.
-There are a wide variety of works which are focused more exclusively on the computational implementation of various optimal control techniques which do spend significant time on the computational characteristics.
-However, our manuscript is primarily focused on demonstrating the feasbility of using reachability sets to design transfer trajectories.
-As a result, there is not much benefit to detailed statistics of the implementation of this methodology, and we have omitted it from the manuscript.
-
-\item 
-    \begin{itshape}
-Can any intuition be gained from reporting the optimal converged cost function values in section 6.1 and 6.2?
-One could at least compare them to compare the reachability in the two different examples.
-\end{itshape}
-
-The cost function only parameterizes the distance between the controlled and uncontrolled terminal states on the \Poincare section. 
-In addition, each state which lies on the reachable set would typically have a different distance and therefore a different coverged cost function value. 
-Furthermore, it would be difficult to characterize a single parameter which characterizes the reachable set, in the form of the cost function, for all states of the reachable set. 
-Also, the cost function is highly dependent on the transit time of trajectory, i.e. a longer transit time corresponds to a larger reachable set and a larger magnitude of the cost function.
-However, this is an interesting suggestions for future research.
-For example, the cost function gives some indication of area/volume of the reachable set. 
-It may be possible to use this value as another way to characterize the subspace on the \Poincare section and is an intriguing avenue of future research.
-
-\item
-    \begin{itshape}
-"Low energy transfers from the Earth to the Lagrange points are necessary for future missions."  One must be careful not to confuse these with traditional "low-energy transfers" to the moon, which use Sun third-body perturbations to transfer into lunar orbit.
-These are what many mission designers define as "low-energy" lunar transfers.
-This is actually a four-body problem (Sun, Earth, moon, spacecraft), and represents additional delta-V savings (as was done with NASA's GRAIL mission.)
-
-Example:
-
-Parker, J. and Anderson, R., "Low-Energy Lunar Trajectory Design," DEEP SPACE COMMUNICATIONS AND NAVIGATION SERIES, Jet Propulsion Laboratory Pasadena, California, July 2013.
-
-Online Copy: https://descanso.jpl.nasa.gov/monograph/series12/LunarTraj--Overall.pdf
-\end{itshape}
-
-Thank you for the comment.
-This sentence has been modified to make it more clear that we're discussing low thrust propulsion rather than low-energy transfers.
-The modified sentence is copied below.
-
-\begin{correction}
-Transfers from the Earth to the Lagrange points, through the use of low-thrust electric propulsion, offer an additional and potentially shorter time of flight in comparison to the low-energy transfers which utilize solar perturbations.
-\end{correction}
-
-\item 
-    \begin{itshape}
-For what it's worth, exploiting sun perturbations with low-thrust in the Earth-moon RTBP would be highly beneficial and represent a worthy avenue of future research (more on this below.)
-\end{itshape}
-
-Thank you for the comment.
-This is a very interesting suggestion and something that we will take into consideration for future research.
-
-\item 
-    \begin{itshape}
-The conclusions of the paper should acknowledge other potentially limiting factors of the presented analysis
-
--Omits advantage of exploiting fourth body perturbation (Sun in Earth/moon RTBP)
-
--Transfers are 2D only, method is more challenging to adapt to 3D
-
--Presented control magnitudes are dimensionally large, and more work is needed to determine feasibility of approach for lower accelerations
-
--Thrust and mass are completely decoupled from the analysis, which is limiting in terms of finding suitable hardware that can deliver the controls needed, as currently presented.
-\end{itshape}
-
-Thank you for the comment.
-The conclusion has been expanded with some discussion of the drawbacks of this research and avenues of future work.
-The modified conclusion is copied below.
-
-\begin{correction}
-In this paper, an optimal transfer process which combines concepts of reachability and \Poincare section is used to generate transfer between planar periodic orbits in the three-body problem.
-The \Poincare section allows for trajectory design on a lower dimensional phase space and simplifies the process.
-The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
-However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
-Furthermore, our approach relies heaviliy on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
-
-There is additional research to extend these results to more general transfer scenarios in future work.
-The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
-In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
-Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
-This will result in a more complicated optimal control problem and is a focus of future research.
-Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
-The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
-\end{correction}
 \end{enumerate}
-
 \section*{Reviewer 2}

### 2017-07-24 manuscript.tex
Correct some typos

@@
 %Computational geometric mechanics allows for the development of variational integrators and a discrete optimal control problem.
-This is solved numerically as a discrete optimal control problem using a variational integrator, which preserve the geometric structure of the motion in the three-body problem.
+This is solved numerically as a discrete optimal control problem using a variational integrator, which preserves the geometric structure of the motion in the three-body problem.
 We demonstrate our approach with two numerical simulations of transfers in the Earth-Moon three-body system.
@@
 Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
-As a result numerical methods are often required to investigate solutions to the three body problem.
+As a result, numerical methods are often required to investigate solutions to the three body problem.
 Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
-As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g.
+Therefore, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g.
 on the order of decades to centuries, which is typical of the three-body problem.
@@
 Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system. 
-The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from teh free motion trajectory to allow for increased transfer opportunities.
+The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.
 
@@
 We defne an optimal control problem to determine the reachability set on a \Poincare~section.
-The reachable set, defined as the set of all attainable states subect to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
+The reachable set, defined as the set of all attainable states subject to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
 Defining the reachable set on a \Poincare~section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
@@
 Through the use of low-thrust control input, the reachable set on the \Poincare~section is enlarged and enables a larger space of potential transfers.
-By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare~section allows for the creation of general transfer trajectories.
+By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare~section, we generate general transfer trajectories.
 
@@
 In addition, this model allows for a systematic process to define and exploit \Poincare sections.
-Furthermore, the 2D solutions afforded by the PCRTBP are frequently use to gain a qualitative understanding of the trade space of transfers in the Earth-Moon system.
+Furthermore, the 2D solutions afforded by the PCRTBP are frequently used to gain a qualitative understanding of the trade space of transfers in the Earth-Moon system.
 The PCRTBP approach offers insight into the fundamental dynamical structure while capturing the major dynamic properties of the planar motion.
-As a result, this approach is best suited for trajectories which do not require large plane changes.
+As a result, this approach is best suited for preliminary trajectories which do not require large plane changes.
 
@@
 
-A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
+A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
 A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
@@
 This examples demonstrates the ability to extend the reachability process to multiple iterations, to allow for a much larger and more general transfer.
-With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare~sections.
+With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare sections.
 
@@
 This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
-Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
+Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of approximately \SI{1}{\newton}.
 Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
@@
 The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in the position space visualizations given in~\cref{fig:stage1to4_reachability,fig:stage5to8_reachability,fig:geo_transfer_full}.
-The stable manifold is denoted by the green trajecotries which span the majority of the phase space. 
+The stable manifold is denoted by the green trajectories which span the majority of the phase space. 
 In addition, we plot the intersection of the stable manifold with the \Poincare section with green markers.
@@
 From this reachable set, we chose a trajectory which minimizes the distance towards the stable manifold, which is shown in~\cref{fig:stage1_poincare} by the black marker.
+The distance on the \Poincare section between the reachable set and the stable manifolds is defined as the function \( d \),
+\begin{align*}
+        d(\vecbf{x}(N)) = \sqrt{\parenth{x - x_t}^2 + \parenth{\dot{x} - \dot{x}_t}^2}.
+\end{align*}
+From the reachable set, the trajectory which minimizes \( d \) is used to intialize the next iteration.
 The minimum trajectory is also shown in~\cref{fig:stage1_trajecotry_zoom} and is used to initilize the following stage.
@@
 In~\cref{fig:stage2_trajecotry_zoom}, the first stage trajectory is shown in red while the additional stage is shown in blue, which demonstrates the decrease in the \( x \) component.
-
-After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds as the function \( d \)
-\begin{align*}
-        d(\vecbf{x}(N)) = \sqrt{\parenth{x - x_t}^2 + \parenth{\dot{x} - \dot{x}_t}^2}.
-\end{align*}
-From the reachable set, the trajectory which minimizes \( d \) is used to intialize the next iteration.
-Once the reachability set intersects the stable manifold, we can achieve a complete transfer from the initial geostationary orbit to the stable manifold of \( L_1 \).
 
@@
 The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
-In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
+In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems and future research is aimed at investigating smaller magnitude control inputs.
 Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
@@
 Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
-The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
+The addition of attitude dynamics and realistic pointing constraints would also significantly improve the applicability of this work.
 

### 2017-07-27 manuscript.tex
Incorporate some suggestions from Dr. Lee

@@
 on the order of decades to centuries, which is typical of the three-body problem.
+Furthermore, the variational integrator can allow for reduced computational effort without any loss of numerical stability or energy drift.
 
@@
 We derive and solve the Euler-Lagrange equations to generate the necessary conditions for an optimal solution.
-This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on course or suboptimal control parameterizations.
+This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
 For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
@@
 \Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
-For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the convential integration scheme.
+For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
 Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
@@
 This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
-However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
+However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
 In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
-The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.
+The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
+The fixed step size of the presented variational integrator mitigates the artificial ``noise'' of variable step integration schemes in gradient computation.
+In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
+The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.
 
@@
 \Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
-However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories.
+However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories~\cite{kulumani2016d}.
 
@@
 With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare sections.
+The numerical examples presented in this section satisfy the necessary conditions for local optimality and obtaining a globally optimal solution is considered beyond the schope of this paper.
 

### 2017-07-28 manuscript.tex
Correct initial condition for variational integrator comparison

@@
 A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
-A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
-    The variational integrator uses a range of step sizes between \SIrange{4.72}{472}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
+A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
+    The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
     The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.

### 2017-07-28 Response.tex
Correct typos
Use vim spell to fix spelling mistakes

@@
 \usepackage{xr-hyper}
+\usepackage{xcite}
+\externalcitedocument{manuscript}
 % \usepackage{nameref,zref-xr}
@@
 
+\newcommand{\EditTL}[1]{{\color{red}\protect #1}}
+%\renewcommand{\EditTL}[1]{{\protect #1}}
+
 \begin{document}
@@
 
-    Thank you for the comment, and we agree that there is significant repetition of material previously covered in other sources.
-    The manuscript has been reduced and reorganized to enhance the clarity and clearly seperate previously material and the contributions of this paper.
+    Thank you for the comment, and we agree that there is repetition of material previously covered in other publications.
+    The challenge is that the proposed approach utilizes several tools and concepts integrated together to design low-thrust orbital maneuvers in a systematic way, and in order to describe the key idea in a self-contained manner, each of prior approach should be described properly.
+
+    To address this, the manuscript has been reduced and reorganized to enhance the clarity and clearly separate previously material and the contributions of this paper.
     The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
-    The second section has been retitiled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
-The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, followed by conclusions.
+    The second section has been reorganized to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
 The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.
+The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, which are the main contributions of this paper.
+
+    In short, the manuscript has been completely reorganized such that all of the background materials are summarized in Section 2.
+The remaining sections dealing with the proposed approach are expanded with more detailed descriptions and numerical results.
 
@@
 The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
-Direct optimal control techniques must rely entirely on the ability of the numerical optimzation routine to determine a feasible solution.
+Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.
 Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
@@
 However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
-Furthermore, our approach relies heaviliy on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
+Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
 
@@
 Thank you for the comment and these characteristics of our choice of the PCRTBP.
-Additional detail has been included in the problem formulation section.
-This includes some discussion on the choice of the PCRTBP and the relative strengths/weaknesses of this approach.
-The additional paragraph has been copied below.
+The authors would like to point out that the key idea of designing low-thrust orbits through the reachability analysis on the \Poincare section may be applied to other dynamic system.
+This manuscript illustrates such idea through PCRTB.
+
+However, according to the reviewers suggestion, additional details have been stated in the problem formulation section, including discussions on the choice of the PCRTBP and the relative strengths/weaknesses of this approach as copied below.
 
@@
 
-Thank you for the comment and the detail on the benefits/drawbacks of variatonal integrators.
-Additional discussion has been included in the paper in the subsection which introduces the variational integrator for the PCRTBP.
+Thank you for the comment and the detail on the benefits/drawbacks of variational integrators.
+The authors agree that the general-purpose numerical integrators such as higher-order RK methods are suitable in many cases.
+The proposed approach utilizes the variational integrator, particularly due to its numerical structural stability of capturing long-term behaviors of small effects accurately and efficiently, which is critical in low-thrust orbital maneuvers.
+
+While the structural stability of geometric numerical integrators is often illustrated by the conserved quantities such as energy or momentum, it has been shown that variational integrators are also effective in computing the rate of energy change for non-conservative systems~\cite{marsden2001}.
+
+To clarify these, additional discussion has been included in the paper in the subsection which introduces the variational integrator for the PCRTBP.
 Much of the discussion of variational integrators in the introduction has also been removed.
@@
 In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
-The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.
+The fixed step size of the presented variational integrator mitigates the artificial ``noise'' of variable step integration schemes in gradient computation.
+In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
+The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.
 \end{correction}
 \end{itemize}
-From here on out comments specific to various sections of the paper follow:
+From here on out, comments specific to various sections of the paper follow:
 
@@
             Electric propulsion systems offer a much greater specific impulse than chemical systems.
-            As a result, the greatly increased effeciency allows for greater payload mass or extended duration missions.
-            However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital manuevers have a much longer time of flight.
+            As a result, the greatly increased efficiency allows for greater payload mass or extended duration missions.
+            However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital maneuvers have a much longer time of flight.
             In spite of this drawback, a wide variety of missions, such as communication and deep space probes, have utilized the unique benefits of low thrust electric propulsion to great effect~\cite{choueiri2009}.
@@
             Recent developments in miniature electric propulsion offer the potential for new research opportunities for small spacecraft~\cite{haque2013}.
-            The advancements in minature electric propulsion allows for greater flexibility in the deployment of cubesats as secondary payloads.
+            The advancements in miniature electric propulsion allows for greater flexibility in the deployment of cubesats as secondary payloads.
             Without the explicit requirement of a dedicated launch trajectory, cubesats are able to exploit a much wider range of possible mission designs~\cite{folta2015}.
@@
 \begin{correction}
-
 Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
@@
 Thank you for the comment.
-The claims about the variational integrator have been qualified with some additional detail.
+Those numerical dissipations would depend on several parameters.
+But, for the particular numerical comparison illustrated by \Cref{fig:compare_energy}, such difference in conservation properties becomes noticeable in the simulation for 15 years.
+As suggested, the claims about the variational integrator have been qualified with some additional detail.
 The correction is copied below.
@@
 As a result, the long term effects of low-thrust on the spacecraft trajectory are not accurately captured."  Again, I think situationally I would disagree.
-Transfers on the order of 100's of days can maintain accuracies of 1e-10 (nondimensional units) in the Hamiltonian, which is generally considered a "small" amount of energy drift.
-\end{itshape}
-
-These sentences have been modified and combined to enhance clarity and reduce some of the unecessary bloat. 
+Transfers on the order of 100's of days can maintain accuracies of 1e-10 (non-dimensional units) in the Hamiltonian, which is generally considered a "small" amount of energy drift.
+\end{itshape}
+
+These sentences have been modified and combined to enhance clarity and reduce some of the unnecessary bloat. 
 In addition, some additional detail has been added to explicitly define the notion of ``long-term'' in the context of numerical integration.
+As shown by \Cref{fig:compare_energy}, the variational integrator exhibit smaller conservation errors than ODE45 at the similar computational efforts, for the particular numerical case considered.
 The modifications are copied below.
@@
 
-While the geometric integrator will not exactly characterize the terminal energy or state of a dynamic system, it will better preserve the total energy variation as compared to convential integration schemes.
-In addition, the total energy deviation will remain bounded and less than than of convential schemes which experience unbounded energy deviation. 
+While the geometric integrator will not exactly characterize the terminal energy or state of a dynamic system, it will better preserve the total energy variation as compared to conventional integration schemes.
+In addition, the total energy deviation will remain bounded and less than than of conventional schemes which experience unbounded energy deviation. 
 As a result, the terminal state/energy will be more accurately characterized using a variational integrator as compared to conventional schemes. 
@@
     \begin{itshape}
--"The Poincare section reduces the dimensionality of the system dyanmics to the study of a related discrete update map." It does, but at the expense if a large up-front computational run that must be run well in advance to capture the problem characteristics.
+-"The Poincare section reduces the dimensionality of the system dynamics to the study of a related discrete update map." It does, but at the expense if a large up-front computational run that must be run well in advance to capture the problem characteristics.
 Also, it is more challenging to execute in 3D, which should be noted.
@@
 The section on the variational integrator has been reduced and this specific sentence has been removed. 
-However, additional detail has been provided in a the variational integrator subsection which discusses this drawback.
-
-\begin{correction}
+However, additional detail has been provided in the variational integrator subsection which discusses this drawback.
+
+\begin{correction}
+
 Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
@@
 This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
-However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
+However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
+
 \end{correction}
@@
 
-Thank you for the comment.
-This sentence has been removed from the introduction.
+It has been shown that the structure-preserving properties of geometric numerical integrators are also useful to numerical optimizations for improved computational efficiency and stability.
+For example, see the following reference.
+
+\begin{itemize}
+\item[-] Ober-Blöbaum, Sina, Oliver Junge, and Jerrold E. Marsden. "Discrete mechanics and optimal control: an analysis." ESAIM: Control, Optimisation and Calculus of Variations 17.2 (2011): 322-352.
+\end{itemize}
+
+The proposed approach utilizes desirable numerical properties of the variational integrator.
+However, the main contribution is systematic design of low-thrust orbital maneuvers through reachability set analysis on the \Poincare section.
+The above sentences may be confusing, and the authors decided to remove them from the introduction of the revised manuscript.
 
@@
 Thank you for the comment and helpful reference.
+
+It is well known that variable step size may break conservational properties of a geometric numerical integrator.
+Variational integrators are no exception, and there has been a multisymplectic framework to address it.
+See
+
+\begin{itemize}
+\item[-] Lew, Adrian, et al. "Asynchronous variational integrators." Archive for Rational Mechanics and Analysis 167.2 (2003): 85-146.
+\end{itemize}
+
+However, the above formulation for structure-preserving, adaptive time-step variational integrators is beyond the scope of this paper, and we use a fixed time step throughout the presented numerical results.
+
 Additional detail has been included in the manuscript discussing this potential drawback of variable step integrators.
-j
-\begin{correction}
+
+\begin{correction}
+
 In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
-The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.
+The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
+
 \end{correction}
@@
 \begin{correction}
-This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on course or suboptimal control parameterizations.
+This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
 For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
@@
 \begin{correction}
-Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invaraint properties and symplecticity.
+Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invariant properties and symplecticity.
 Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations.
@@
     \begin{itshape}
-Figure 2 - x-axis has no label but I think it's nondimensional time, please indicate.  Label all axes even if they are nondimensional units.
+Figure 2 - x-axis has no label but I think it's non-dimensional time, please indicate.  Label all axes even if they are non-dimensional units.
 \end{itshape}
@@
         \centering 
-        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparision: A stable trajectory is used to test the variational and conventional integrators. 
+        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparison: A stable trajectory is used to test the variational and conventional integrators. 
         The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
@@
                 \includegraphics[width=\textwidth]{trajectory} 
-                \caption{Earth-Moon three-body trajectory used for integrator comparision}
+                \caption{Earth-Moon three-body trajectory used for integrator comparison}
         \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
@@
 A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
-A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
-    The variational integrator uses a range of step sizes between \SIrange{4.72}{472}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
+A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
+    The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
     The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
@@
 However, this manuscript is not really focused on variational integrators but rather the design of transfer trajectories using reachability sets.
-There is a wide range of topic under the umbrella of varaitional integrators that are not directly applicable to this manuscript and work.
+There is a wide range of topic under the umbrella of variational integrators that are not directly applicable to this manuscript and work.
 Interested readers can find a wide variety of literature on the relative benefits of all manner of variational integrators in papers focused on that topic specifically.
@@
 \begin{correction}
-
 Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
@@
 The numerical examples presented in this section are designed in the context of the PCRTBP.
-The dynamic environment has a four dimensional state space and offers a convient integration constant in the form of the Jacobi integral.
+The dynamic environment has a four dimensional state space and offers a convenient integration constant in the form of the Jacobi integral.
 As a result, there are well defined methods to define and exploit \Poincare sections, which result in straightforward two-dimensional subspaces of the system.
@@
 \Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
-However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories.
+However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories~\cite{kulumani2016d}.
 \end{correction}
@@
 The mentioned plots of the control input have been modified to show the data in dimensional units (Newtons). 
-Some additional detail has been provided in the manuscript which discusses the choice of \( u = 0.75 \) and it's realtionship to the current state of the art thruster systems.
+Some additional detail has been provided in the manuscript which discusses the choice of \( u = 0.75 \) and it's relationship to the current state of the art thruster systems.
 The modified plots are not included here but the additional content is copied below in response to a different but related comment.
@@
     \subcaptionbox{Control input for the selected transfer trajectory}{\includegraphics[width=0.5\textwidth]{control_input_l1}}~
-    \subcaptionbox{Control Input: Combined control history for each stage of the reacability analysis. 
+    \subcaptionbox{Control Input: Combined control history for each stage of the reachability analysis. 
     The control always remains within the maximum bound.}{\includegraphics[width=0.5\textwidth]{figures/control_input_geo.pdf}}
@@
 For simplicity, this reduced the possible number of stages to an even number. 
-Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in covergence. 
+Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in convergence. 
 The solver would tend to ensure a continuous trajectory, by satisfying the interior stages, at the expense of the terminal stage. 
@@
 
-Thank you for the comment, howver it seems to be rare for the research community to give detailed description of the numerical optimization method.
+Thank you for the comment, however it seems to be rare for the research community to give detailed description of the numerical optimization method.
 Most of the literature tends to not spend time discussing the low level details of the computational procedure, e.g.~\cite{grebow2011,stuart2010}.
@@
 There are a wide variety of works which are focused more exclusively on the computational implementation of various optimal control techniques which do spend significant time on the computational characteristics.
-However, our manuscript is primarily focused on demonstrating the feasbility of using reachability sets to design transfer trajectories.
+However, our manuscript is primarily focused on demonstrating the feasibility of using reachability sets to design transfer trajectories.
 As a result, there is not much benefit to detailed statistics of the implementation of this methodology, and we have omitted it from the manuscript.
+However, we have added some rough detail on the computation time and computer hardware used in the numerical optimization.
+This additional detail is copied below.
+
+\begin{correction}
+Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
+Each solution on the reachability set is computed in approximately \SI{30}{\second} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
+\end{correction}
 
@@
 The cost function only parameterizes the distance between the controlled and uncontrolled terminal states on the \Poincare section. 
-In addition, each state which lies on the reachable set would typically have a different distance and therefore a different coverged cost function value. 
+In addition, each state which lies on the reachable set would typically have a different distance and therefore a different converged cost function value. 
 Furthermore, it would be difficult to characterize a single parameter which characterizes the reachable set, in the form of the cost function, for all states of the reachable set. 
@@
     Thank you for the comment, and we agree that there is significant repetition of material previously covered in other sources.
-    The manuscript has been reduced and reorganized to enhance the clarity and clearly seperate previously material and the contributions of this paper.
+    The manuscript has been reduced and reorganized to enhance the clarity and clearly separate previously material and the contributions of this paper.
     The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
-    The second section has been retitiled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
+    The second section has been retitled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
 The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, followed by conclusions.
@@
     Thank you for the comment. 
-    The numerical examples are meant to demonstrate the use of this approach and the feasibility of using the reachability set on a \Poincare section to determine transfer trajecotries.
+    The numerical examples are meant to demonstrate the use of this approach and the feasibility of using the reachability set on a \Poincare section to determine transfer trajectories.
     The examples give two demonstrations of finding a transfer trajectory in the PCRTBP and demonstrate combining several iterations of the reachability set for a transfer.
@@
 \end{itshape}
+
+The authors believe that the given two nontrivial numerical examples illustrate the efficacy of the proposed approach in designing low-thrust maneuvers through reachability set analysis on the \Poincare section.
+Furthermore, the manuscript has been completely reorganized to emphasize the contributions and clearly separate the background material. 
+In particular,~\cref{sec:simulation} includes additional paragraphs and figures for the geostationary to periodic orbit transfer.
+These additional figures demonstrate the ability to chain together multiple iterations of the reachability set computation to generate a nontrivial transfer trajectory.
+Combined together, the numerical examples demonstrate the ability to depart from the vicinity of the Earth to the vicinity of the Moon using our proposed method of coupling low-thrust maneuvers with reachability set analysis.
 
@@
 In addition, the extended discussion of the benefits of variational integrators is removed from the introduction.
-The mathematical background has been condensed into a single section with seperate subsections which discuss some of the mathematical details.
+The mathematical background has been condensed into a single section with separate subsections which discuss some of the mathematical details.
 The background section is focused on reviewing the relevant material rather than reintroducing topics.
 In the interest of space the reorganized section is omitted in this response. 
+In short, the manuscript is completely reorganized and all of the background materials are summarized in~\cref{sec:background}.
 
@@
     \begin{itshape}
-A missing relevant reference: "On target for Venus — set oriented computation of energy efficient low thrust trajectories" (Dellnitz, Junge, Post, and Thiere, 2006)
+A missing relevant reference: "On target for Venus a set oriented computation of energy efficient low thrust trajectories" (Dellnitz, Junge, Post, and Thiere, 2006)
 \end{itshape}
@@
 \item
-    {\color{red}
     \begin{itshape}
@@
 \end{itshape}
-}
-
-\item
-    {\color{red}
+
+The main contribution is systematic design of low-thrust orbital maneuvers through reachability set analysis on the \Poincare section.
+The authors decided to utilize the variational integrator as it is more computationally efficient in simulating a complex dynamic system accurately.
+
+More explicitly, while the structural stability of geometric numerical integrators is often illustrated by the conserved quantities such as energy or momentum, it has been shown that variational integrators are also effective in computing the rate of energy change for non-conservative systems accurately~\cite{marsden2001}.
+This is particularly useful for low-thrust orbital maneuvers where the energy dissipate or accumulates slowly.
+Furthermore, it has been shown that the structure-preserving properties of geometric numerical integrators are also useful to numerical optimizations for improved computational efficiency and stability.
+For example, see the following reference.
+
+\begin{itemize}
+\item[-] Ober-Blöbaum, Sina, Oliver Junge, and Jerrold E. Marsden. "Discrete mechanics and optimal control: an analysis." ESAIM: Control, Optimisation and Calculus of Variations 17.2 (2011): 322-352.
+\end{itemize}
+
+In the revised manuscript, all of the comments for the variational integrator are re-examined and reorganized to Section 2.2, as also discussed in the response to the first reviewer.
+
+\item
     \begin{itshape}
@@
 \end{itshape}
-}
+
+In the revised manuscript, the computational properties of the variational integrator is explicitly compared with a general purpose integrator in~\Cref{fig:compare_energy}.
+According to the reviewers suggestion, the error in energy conservation is illustrated for several computational costs.
+It is shown that the variational integrator exhibits lower conservation error  against a Runge-Kutta method for the similar level of CPU time.
+
+With regards to the comments on the adaptive step size, the variational integrator can be generalized according to the multisymplectic framework, as in the following reference. 
+\begin{itemize}
+\item[-] Lew, Adrian, et al. "Asynchronous variational integrators." Archive for Rational Mechanics and Analysis 167.2 (2003): 85-146.
+\end{itemize}
+However, the above formulation for structure-preserving, adaptive time-step variational integrators is beyond the scope of this paper, and we use a fixed time step throughout the presented numerical results.
+
+More importantly, it has been shown variational integrators are also effective in computing the rate of energy change for non-conservative systems accurately~\cite{marsden2001}, even for the fixed-step sizes.
 
@@
     \caption{Diagram of the \Poincare map: Periodic orbits will appear as fixed points on the \Poincare section \( \Sigma \).
-Stabiilty of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.}
+Stability of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.}
 \end{figure}
@@
         \end{scaletikzpicturetowidth}
-        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
+        \caption{Reachability set on a \Poincare section: Pictorial representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
             The terminal state, \( x_n\), is the intersection without any control input. 
@@
             We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
-            Computation of the maximum distance, or reacahbility, for a variety angles gives a discrete approximation of the reachability set.}
+            Computation of the maximum distance, or reachability, for a variety angles gives a discrete approximation of the reachability set.}
 \end{figure}
@@
         \caption{Schematic diagram of the multiple shooting method:
-        The complete trajectory is split into a nubmer of sub-segments, and additional interior constraints are included to ensure state and costate continunity.
+        The complete trajectory is split into a number of sub-segments, and additional interior constraints are included to ensure state and costate continuity.
     Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.}
@@
    \includegraphics[width=0.2\textwidth]{moon_orbit} % requires the graphicx package
-   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstraiting the initial and target orbit.
-   Without the low-thrust proplusion, the spacecraft is constrained to the initial periodic orbit. 
+   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstrating the initial and target orbit.
+   Without the low-thrust propulsion, the spacecraft is constrained to the initial periodic orbit. 
    We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
@@
                 \includegraphics[width=0.5\textwidth]{reach_transfer} 
-                \caption{Transfer trajectory selected from reachabiltiy set viewed in the position space}  
+                \caption{Transfer trajectory selected from reachability set viewed in the position space}  
         \end{subfigure}~ 
@@
         The reachability set is shown in the upper two figures in both the position and \Poincare space.
-    From this reachabiilty set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
+    From this reachability set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
 The optimal control to achieve this transfer is shown in the lower right figure.} 
@@
     \caption{Stage 1-4 reachability sets: The first four reachability sets visualized in both the position (left) and \Poincare space (right).
-        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
+        The minimum trajectories from the preceding stages are shown in red, while the next stage is shown in blue.
     The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
@@
     \caption{Stage 5-8 reachability sets: The last four reachability sets visualized in both the position (left) and \Poincare space (right).
-        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
+        The minimum trajectories from the preceding stages are shown in red, while the next stage is shown in blue.
     The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
@@
 
-Thank you for the comment, however the figure is drawn to show a pictoral representation of a \Poincare section.
+Thank you for the comment, however the figure is drawn to show a pictorial representation of a \Poincare section.
 The periodic orbits intersect the section at fixed points, as a result it becomes simple to visualize and determine these fixed solutions by searching for fixed points on the section
@@
         \end{scaletikzpicturetowidth}
-        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
+        \caption{Reachability set on a \Poincare section: Pictorial representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
         }
@@
 
-Thank you for the comment, however we feel that the figures adequately show the desired information.
+Thank you for the comment. The figure have been revised to adequately show the desired information in the revised manuscript. 
 The figures are centered around the various intersections of the reachable set and the \Poincare section.
@@
 \end{itshape}
-}
+
+The presented numerical results satisfy the necessary conditions for optimality.
+As common in gradient-based numerical optimizations for non-convex systems, it guarantees local optimality.
+Applying various techniques to obtain the globally optimal solution is beyond the scope of this paper that is focused on systematic design of low-thrust orbital maneuvers thorough reachability set analysis on the \Poincare section.
+
+To clarify these, the following comment is added at the first paragraph of~\cref{sec:simulation}.
+
+\begin{correction}
+The numerical examples presented in this section satisfy necessary conditions for local optimality, and obtaining a globally optimal solution is considered beyond the scope of this paper.
+\end{correction}
 
@@
 
-The choice of starting from a fixed point on the initial periodic orbit is again mostly used for convience. 
+The choice of starting from a fixed point on the initial periodic orbit is again mostly used for convenience. 
 In this case of the periodic orbit the departing trajectories are all along the unstable invariant manifold associated with the periodic orbit. 
-The uncontrolled trajecotries associated with the invariant manifolds serve as the ``center'' of each reachability set computation.
+The uncontrolled trajectories associated with the invariant manifolds serve as the ``center'' of each reachability set computation.
 The low-thrust propulsion system is then used to enlarge the distance between the uncontrolled and controlled states. 
@@
 However, different locations along the periodic orbit may have vastly different travel times before intersecting the \Poincare section.
-This will require modification of the optimal control problem, with a larger terminal time, but will result in a similar behavior regardless of the location on the intial periodic orbit.
+This will require modification of the optimal control problem, with a larger terminal time, but will result in a similar behavior regardless of the location on the initial periodic orbit.
 
@@
 This is due to the fact that there is a very low-thrust control input on the spacecraft.
-As a result, the trajectories associated with the reachability set are all closely spaced, but more widely seperated from the terminal state without any control input.
+As a result, the trajectories associated with the reachability set are all closely spaced, but more widely separated from the terminal state without any control input.
 In the figure, there is a fixed point labeled \( x_n \) which is the terminal state without any control input, which will also lie on the periodic orbit.
-The reachable set, in red, is widely separated from the uncontrolled state due the effect of the low thrust proplusion, however the subsequent spread of the entire reachability set is much more constrained.
+The reachable set, in red, is widely separated from the uncontrolled state due the effect of the low thrust propulsion, however the subsequent spread of the entire reachability set is much more constrained.
 The low thrust expands the reachability set from a fixed point to the red elliptical region. 
@@
 \end{itshape}
-}
+
+The authors believe that the given two nontrivial numerical examples illustrate the efficacy of the proposed approach in designing low-thrust maneuvers through reachability set analysis on the \Poincare section.
+The examples demonstrate the ability to generate transfer trajectories in the PCRTBP and can be easily extended to non-planar transfers~\cite{kulumani2016d}.
+The proposed approach provides for a systematic design process for low-thrust orbital transfers, in contrast to other methods which must rely entirely on numerical optimization to determine trajectories.
+This design process leverages the reachability set on a \Poincare section to greatly reduce the computational complexity of an unmotivated search of the reachable set.
+Through the use of the reachability set, our proposed approach provides a methodical process to determine transfers, using single or multiple iterations.
+
+There are several drawbacks of this approach.
+First, it relies heavily on the PCRTBP, which limits the applicability to planar trajectories.
+Furthermore, this reduces it's benefit to operational missions which are often non-planar, e.g. LEO to GTO missions.
+Second, the method does not consider the effect of variable mass on the system dynamics. 
+This is a focus of future work to incorporate the effect of variable mass on the optimal control problem.
+Finally, the variational integrator provides for a bounded energy behavior but does not ensure exact energy over the simulation. 
+As a result, there is still error in the variational integrator with respect to energy deviation.
+
+The manuscript has been completely reorganized to emphasize the contributions and clearly separate the background material. 
+In particular,~\cref{sec:simulation} includes additional paragraphs and figures for the geostationary to periodic orbit transfer.
+These additional figures demonstrate the ability to chain together multiple iterations of the reachability set computation to generate a nontrivial transfer trajectory.
+Combined together, the numerical examples demonstrate the ability to depart from the vicinity of the Earth to the vicinity of the Moon using our proposed method of coupling low-thrust maneuvers with reachability set analysis.
+
+In addition, the conclusion has been modified to better demonstrate the strengths and weaknesses of our approach as well as avenues for future research.
+The conclusion is copied below.
+
+\begin{correction}
+
+In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
+The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
+The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
+Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.
+Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
+Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
+However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
+Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
+
+There is additional research to extend these results to more general transfer scenarios in future work.
+The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
+In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems and future research is aimed at investigating smaller magnitude control inputs.
+Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
+This will result in a more complicated optimal control problem and is a focus of future research.
+Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
+The addition of attitude dynamics and realistic pointing constraints would also significantly improve the applicability of this work.
+\end{correction}
+
 \end{enumerate}
-\bibliography{library}
-\bibliographystyle{spmpsci}
+% \bibliography{library}
+% \bibliographystyle{spmpsci}
 \end{document}

### 2017-07-28 manuscript.tex
Add some detail on computation time required
Correct typos
Use vim spell to fix spelling mistakes

@@
 Electric propulsion systems offer a much greater specific impulse than chemical systems.
-As a result, the greatly increased effeciency allows for greater payload mass or extended duration missions.
-However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital manuevers have a much longer time of flight.
+As a result, the greatly increased efficiency allows for greater payload mass or extended duration missions.
+However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital maneuvers have a much longer time of flight.
 In spite of this drawback, a wide variety of missions, such as communication and deep space probes, have utilized the unique benefits of low thrust electric propulsion to great effect~\cite{choueiri2009}.
@@
 Recent developments in miniature electric propulsion offer the potential for new research opportunities for small spacecraft~\cite{haque2013}.
-The advancements in minature electric propulsion allows for greater flexibility in the deployment of cubesats as secondary payloads.
+The advancements in miniature electric propulsion allows for greater flexibility in the deployment of cubesats as secondary payloads.
 Without the explicit requirement of a dedicated launch trajectory, cubesats are able to exploit a much wider range of possible mission designs~\cite{folta2015}.
@@
 There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}. 
-Inspite of the long history in astrodynamics, a number of difficulties arise when implenting optimal control techniques.
+In spite of the long history in astrodynamics, a number of difficulties arise when implementing optimal control techniques.
 First, the optimal control problem is typically solved using direct methods, which approximate the continuous problem as a parameter optimization problem.
@@
 
-% control-free trajectories in teh three-body problem 
+% control-free trajectories in the three-body problem 
 There exists a rich structure in the three-body problem~\cite{koon2011,ross2006}.
@@
 
-We defne an optimal control problem to determine the reachability set on a \Poincare~section.
+We define an optimal control problem to determine the reachability set on a \Poincare~section.
 The reachable set, defined as the set of all attainable states subject to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
@@
 \subsubsection{Invariant Manifolds and \Poincare Map}\label{sec:invariant_manifold}
-Dynamics systems theory has been applied to the design of control-free manuevers in the restricted three-body problem~\cite{koon2011}.
+Dynamics systems theory has been applied to the design of control-free maneuvers in the restricted three-body problem~\cite{koon2011}.
 As previously introduced in~\Cref{sec:jacobi}, there exist five equilibrium points in the equations of motion for the three-body problem.
@@
     \caption{Diagram of the \Poincare map: Periodic orbits will appear as fixed points on the \Poincare section \( \Sigma \).
-Stabiilty of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.\label{fig:poincare_map}}
+Stability of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.\label{fig:poincare_map}}
 \end{figure}
@@
 \subsection{Variational Integrator for PCRTBP}\label{sec:discrete_var}
-Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invaraint properties and symplecticity.
+Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invariant properties and symplecticity.
 Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations.
@@
         \centering 
-        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparision: A stable trajectory is used to test the variational and conventional integrators. 
+        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparison: A stable trajectory is used to test the variational and conventional integrators. 
         The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.\label{fig:compare_trajectory}}
@@
 In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
-This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
+This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge-Kutta methods. 
 However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
@@
             We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
-            Computation of the maximum distance, or reacahbility, for a variety angles gives a discrete approximation of the reachability set.\label{fig:reachability_set}}
+            Computation of the maximum distance, or reachability, for a variety angles gives a discrete approximation of the reachability set.\label{fig:reachability_set}}
 \end{figure}
@@
 Consider the vector of initial conditions, \( \vecbf{\chi} = \braces{\vecbf{x}_0, \vecbf{\lambda}_0}\), which is varied to satisfy some terminal constraints of the form \( \vecbf{G}(\vecbf{\chi}) = \braces{\vecbf{x}_t - \vecbf{x}_n} = 0 \).
-The free variables at the terminal time are computed by propogation of \( \vecbf{\chi} \) over the selected time horizon. 
+The free variables at the terminal time are computed by propagation of \( \vecbf{\chi} \) over the selected time horizon. 
 At the terminal time, the constraint vector is calculated and if not satisfied \( \vecbf{\chi}\) is varied.
@@
         \caption{Schematic diagram of the multiple shooting method:
-        The complete trajectory is split into a nubmer of sub-segments, and additional interior constraints are included to ensure state and costate continunity.
+        The complete trajectory is split into a number of sub-segments, and additional interior constraints are included to ensure state and costate continuity.
     Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.\label{fig:multiple shooting}}
@@
 With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare sections.
-The numerical examples presented in this section satisfy the necessary conditions for local optimality and obtaining a globally optimal solution is considered beyond the schope of this paper.
+The numerical examples presented in this section satisfy the necessary conditions for local optimality and obtaining a globally optimal solution is considered beyond the scope of this paper.
 
@@
    \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
-   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstraiting the initial and target orbit.
-   Without the low-thrust proplusion, the spacecraft is constrained to the initial periodic orbit. 
+   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstrating the initial and target orbit.
+   Without the low-thrust propulsion, the spacecraft is constrained to the initial periodic orbit. 
    We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
@@
 From the initial state on the periodic orbit, a series of optimal trajectories are generated to determine the reachable set.
-The multiple shooting approach is implemented to sovle the optimal control problem. 
+The multiple shooting approach is implemented to solve the optimal control problem. 
 We divide the time horizon into two equal length segments. 
-The state trajectory is intialized using the free trajectory of the periodic orbit. 
-Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propogated using the discrete equations of motion in~\cref{eq:costate_eom}.
+The state trajectory is initialized using the free trajectory of the periodic orbit. 
+Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propagated using the discrete equations of motion in~\cref{eq:costate_eom}.
 This results in the initial guess of the design vector \( \vecbf{\chi} = \begin{bmatrix} \vecbf{\lambda}_0 & \vecbf{x}_1^{-} & \vecbf{\lambda}_1^{-} & \vecbf{\beta} \end{bmatrix}^T\).
@@
 Choosing a new angle \( \theta_d \) corresponds to a different direction as well as a new optimal control problem which is again solved using the multiple shooting approach laid out previously.
+Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
+Each solution on the reachability set is computed in approximately \SI{30}{\second} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
 The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:invariant_manifold_transfer}.
@@
                 \includegraphics[width=\textwidth]{reach_transfer} 
-                \caption{Transfer trajectory selected from reachabiltiy set viewed in the position space} \label{fig:reach_transfer} 
+                \caption{Transfer trajectory selected from reachability set viewed in the position space} \label{fig:reach_transfer} 
         \end{subfigure}~ 
@@
         The reachability set is shown in the upper two figures in both the position and \Poincare space.
-    From this reachabiilty set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
+    From this reachability set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
 The optimal control to achieve this transfer is shown in the lower right figure.}
@@
 In addition, by generating the reachability set we are able to compute the required control input to exactly intersect the target orbit.
-This avoids having to compute and accomplish a secondary impulsive manuever to transition from the invariant manifold to the target orbit.
+This avoids having to compute and accomplish a secondary impulsive maneuver to transition from the invariant manifold to the target orbit.
 %Since The intersection on the \Poincare section only shows that the \( x \text{ and } \dot{x} \) components intersect.
@@
 It requires an appropriate \Poincare~section and the generation of several invariant manifolds.
-There is no clear method of selecting the periodic orbits required based on a the type of transfer desired. 
+There is no clear method of selecting the periodic orbits required based on the type of transfer desired. 
 Determining an intersection between these invariant manifolds generally requires extended flight times, involving several orbits of the primaries, before an appropriate intersection is found.
@@
 Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional Earth-Moon three-body system. 
-While the geostationary orbit is not a typical ``parking orbit'' for most spacecraft missions it serves as a convienent altitude to demonstrate our approach.
+While the geostationary orbit is not a typical ``parking orbit'' for most spacecraft missions it serves as a convenient altitude to demonstrate our approach.
 The geostationary orbit about the Earth is transformed into the rotating reference frame of the Earth-Moon three body problem.
-In addition, we nondimensionalize the initial state to find \( \vecbf{x}_0 = \begin{bmatrix} 0.0972 & 0 & 0 & 3.0010\end{bmatrix}^T\) as the initial condition of the spacecraft on the geostationary orbit.
+In addition, we non-dimensionalize the initial state to find \( \vecbf{x}_0 = \begin{bmatrix} 0.0972 & 0 & 0 & 3.0010\end{bmatrix}^T\) as the initial condition of the spacecraft on the geostationary orbit.
 It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
@@
 \end{figure}
-In~\cref{fig:geo_transfer_target}, the initial orbit is located in the center of the figure and centered about \( m_1 \) while the target periodic orbit is determined about the \( L_1 \) lagrange point.
+In~\cref{fig:geo_transfer_target}, the initial orbit is located in the center of the figure and centered about \( m_1 \) while the target periodic orbit is determined about the \( L_1 \) Lagrange point.
 Instead of determining a transfer directly between the geostationary orbit and the periodic orbit, we seek to instead transfer to the stable manifold of the periodic orbit.
@@
 For the first reachability set computation, the initial geostationary orbit is used to initialize the multiple shooting algorithm.
-We use the same initial guess of \( \vecbf{\lambda}_0\) as in~\cref{sec:periodic_orbit_transfer} and propogate over the fixed time horizon of the period of the geostationary orbit to initialize \( \vecbf{\lambda}_i^{-}\).
+We use the same initial guess of \( \vecbf{\lambda}_0\) as in~\cref{sec:periodic_orbit_transfer} and propagate over the fixed time horizon of the period of the geostationary orbit to initialize \( \vecbf{\lambda}_i^{-}\).
 
@@
     \caption{Stage 1-4 reachability sets: The first four reachability sets visualized in both the position (left) and \Poincare space (right).
-        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
+        The minimum trajectories from the preceding stages are shown in red, while the next stage is shown in blue.
     The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit~\label{fig:stage1to4_reachability}}
@@
 \end{align*}
-From the reachable set, the trajectory which minimizes \( d \) is used to intialize the next iteration.
-The minimum trajectory is also shown in~\cref{fig:stage1_trajecotry_zoom} and is used to initilize the following stage.
+From the reachable set, the trajectory which minimizes \( d \) is used to initialize the next iteration.
+The minimum trajectory is also shown in~\cref{fig:stage1_trajecotry_zoom} and is used to initialize the following stage.
 Over the relatively short time horizon of the geostationary orbital period, the reachability set remains quite close to the initial orbit. 
@@
 
-The minimum trajectory from the first iteration of the reachabiilty set is used to initial the following stage.
+The minimum trajectory from the first iteration of the reachability set is used to initial the following stage.
 From the terminal state of the first iteration, the second reachability set is computed and displayed in~\cref{fig:stage2_trajecotry_zoom,fig:stage2_poincare}.
@@
     \caption{Stage 5-8 reachability sets: The last four reachability sets visualized in both the position (left) and \Poincare space (right).
-        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
+        The minimum trajectories from the preceding stages are shown in red, while the next stage is shown in blue.
     The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit~\label{fig:stage5to8_reachability}}
@@
 With each reachable set we move the controlled trajectory closer to the target stable manifold.
-\Cref{fig:stage1to4_reachability,fig:stage5to8_reachability} shows all of the intermediate reachabiilty iterations to transfer between the initial orbit and the stable invariant manifold.
+\Cref{fig:stage1to4_reachability,fig:stage5to8_reachability} shows all of the intermediate reachability iterations to transfer between the initial orbit and the stable invariant manifold.
 The complete transfer trajectory, after eight iterations, is shown in~\Cref{fig:geo_transfer}. 
@@
     \hfill
-    \subcaptionbox{Control Input: Combined control history for each stage of the reacability analysis. 
+    \subcaptionbox{Control Input: Combined control history for each stage of the reachability analysis. 
     The control always remains within the maximum bound.\label{fig:control_input_geo}}{\includegraphics[width=0.45\textwidth]{figures/control_input_geo.pdf}}

### 2017-07-28 Response.tex
Remove duplicate labels

@@
                 \includegraphics[width=0.5\textwidth]{manifold_trajectory} 
-                \caption{Position space view of invariant manifold} \label{fig:manifold_trajectory} 
+                \caption{Position space view of invariant manifold} 
         \end{subfigure}~ %
@@
                 \includegraphics[width=0.5\textwidth]{manifold_poincare} 
-                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} \label{fig:manifold_poincare} 
+                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} 
         \end{subfigure} 

### 2017-07-30 Response.tex
Remove gradient comment and update timing of computation

@@
 In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
-The fixed step size of the presented variational integrator mitigates the artificial ``noise'' of variable step integration schemes in gradient computation.
 In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
@@
 Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
-Each solution on the reachability set is computed in approximately \SI{30}{\second} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
+Each solution on the reachability set is computed in approximately \SI{2}{\minute} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
 \end{correction}

### 2017-07-30 manuscript.tex
Remove gradient comment and update timing of computation

@@
 The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
-The fixed step size of the presented variational integrator mitigates the artificial ``noise'' of variable step integration schemes in gradient computation.
 In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
@@
 Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
-Each solution on the reachability set is computed in approximately \SI{30}{\second} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
+Each solution on the reachability set is computed in approximately \SI{2}{\minute} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
 The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:invariant_manifold_transfer}.

### 2017-07-30 manuscript.tex
Correct some more typos

@@
 We demonstrate our approach with two numerical simulations of transfers in the Earth-Moon three-body system.
-\keywords{First keyword \and Second keyword \and More}
+\keywords{three body problem \and reachability \and optimal control}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 The use of a simple distance metric on a \Poincare~section greatly reduces the dimensionality of the reachability analysis and provides an approximation of the reachability set.
-In addition, through the use of geometric integrators we more accurately capture the effects of low thrust on the system dynamics in the numerical simulation over extended periods which are typical of electric propulsion systems.
+In addition, through the use of geometric integrators, we more accurately capture the effects of low thrust on the system dynamics in the numerical simulation over extended periods, which are typical of electric propulsion systems.
 With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low thrust propulsion systems.
@@
 The reachable set, defined as the set of all attainable states subject to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
-Defining the reachable set on a \Poincare~section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
-While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
-Instead of relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on the \Poincare~section.
-Through the use of low-thrust control input, the reachable set on the \Poincare~section is enlarged and enables a larger space of potential transfers.
-By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare~section, we generate general transfer trajectories.
+Defining the reachable set on a \Poincare section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
+While the \Poincare section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
+Instead of relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on the \Poincare section.
+Through the use of low-thrust control input, the reachable set on the \Poincare section is enlarged and enables a larger space of potential transfers.
+By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare section, we generate general transfer trajectories.
 
@@
 \end{align}
-and the control inputs is defined as \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \) and assumed to be continuously variable but bounded in magnitude, i.e.
+and the control input is defined as \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \) and assumed to be continuously variable but bounded in magnitude, i.e.
 \( \vecbf{u}^T \vecbf{u} \leq u_{max}^2 \).
@@
     The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
-    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
+    The step size of the variational integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
 \Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
@@
         \vecbf{\lambda}_2^{-} - \vecbf{\lambda}_2^{+} &= 0 , \\
-        &\vdots \\
+        &\vdots \nonumber \\
         \vecbf{x}_{n-1}^{-} - \vecbf{x}_{n-1}^{+} &= 0 , \\ 
@@
         The control free transfer from the initial periodic orbit to the target orbit result in a long time of flight. 
-    In addition, the manifold only intersects the target orbit on the descending or far side of the moon.}
+    In addition, the manifold only intersects the target orbit on the ascending or far side of the moon.}
         \label{fig:invariant_manifold_transfer} 
@@
 
-The minimum trajectory from the first iteration of the reachability set is used to initial the following stage.
+The minimum trajectory from the first iteration of the reachability set is used to initialize the following stage.
 From the terminal state of the first iteration, the second reachability set is computed and displayed in~\cref{fig:stage2_trajecotry_zoom,fig:stage2_poincare}.

### 2017-07-30 Response.tex

@@
 \usepackage{xr-hyper}
-\usepackage{xcite}
-\externalcitedocument{manuscript}
 % \usepackage{nameref,zref-xr}
@@
 
+% \usepackage{xcite}
+% \externalcitedocument{manuscript}
 % edit cref citations for IEEE format
@@
     The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
-    The second section has been reorganized to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
+    The second section has been reorganized to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the and variational integrator.
 The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.
@@
 While the geometric integrator will not exactly characterize the terminal energy or state of a dynamic system, it will better preserve the total energy variation as compared to conventional integration schemes.
-In addition, the total energy deviation will remain bounded and less than than of conventional schemes which experience unbounded energy deviation. 
+In addition, the total energy deviation will remain bounded and less than that of conventional schemes which experience unbounded energy deviation. 
 As a result, the terminal state/energy will be more accurately characterized using a variational integrator as compared to conventional schemes. 

### 2018-05-08 utilities/setup_repo.sh
Add travis setup files

@@ -1,0 +1,20 @@
+#!/bin/bash
+
+# add the correct remote repositories
+bb_repo="git@bitbucket.org:shankarkulumani/2017_JAS_manuscript.git"
+git_repo="git@github.com:skulumani/2017_JAS_manuscript.git"
+fdcl_repo="git@github.com:fdcl-gwu/2017_JAS_manuscript.git"
+
+printf "Setting the origin remote to point to Github and Bitbucket:\n\n"
+printf "Github: $git_repo\n"
+printf "Bitbucket: $bb_repo\n"
+printf "FDCL-GWU: $fdcl_repo\n\n"
+
+git remote set-url origin --push --add $bb_repo
+git remote set-url origin --push --add $fdcl_repo
+git remote set-url origin --push --add $git_repo
+
+printf "Finished! The new remotes are listed.\n"
+printf "\n"
+
+git remote -v

### 2018-05-08 Response.tex
Add new reviewer comments

@@
 
+\section*{Second round of responses to the reviewers' comments for JASS-D-17-00005}
+
+We thank the reviewers for their comments and aid in imporoving the quality of our manuscript.
+
+\subsection*{Reviewer 1}
+\begin{itshape}
+This is a vastly improved iteration of the paper and I thank the authors for taking the time to improve it.
+I commend the authors for responding to all of the feedback, especially modifying figures such as Figure 2, which I know takes a lot of time.
+I was also happy to see the intentional reductions in the paper to make it more focused, and the fact that the authors reconciled their low-thrust transfers with existing technology.
+I still have a few issues that persist based on the current submission, but the paper is far closer to being acceptable for publication.
+\end{itshape}
+
+\begin{itemize}
+    \item 
+        \begin{itshape}
+            Section 2.3
+
+            "This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods."
+
+            "Potentially lower state accuracy" is still overselling.
+Just because your second order method is bounded on an energy surface does not guarantee that the state is accurate.
+It is likely that a 2nd order integrator in the PCRTBP contains more state error than a high-order RK which is the standard for modern trajectory planning in flight operations.
+I think it's best to focus on the primary advantage that the variational integrator provides reduced computational demands.
+        \end{itshape}
+
+    \item \begin{itshape}
+            Response to feedback item 34:
+
+            I disagree that it is abnormal to cite convergence statistics, especially for a new optimization approach or methodology.
+Yes, there are some dependencies to the specific example and use-case, but this is a paper in which researchers may wish to reproduce your results as a learning exercise to build on this research.
+Some expectation as to what they can expect for the results is therefore of pragmatic interest.
+
+            See for example, Table 2 in
+
+            Ellison, D., et al., M., "Application and Analysis of Bounded-Impulse Trajectory Models with Analytic Gradients," Journal of Guidance, Control, and Dynamics, 2018.
+
+            https://arc.aiaa.org/doi/pdf/10.2514/1.G003078
+
+            There are many other paper examples that I could cite that do this, but this is just one.
+        \end{itshape}
+
+    \item \begin{itshape}
+Conclusions:
+
+"The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints."
+
+I don't agree with the statement.
+Often, indirect methods are exceptionally difficult to append path constraints onto.
+In contrast, direct methods such as collocation approaches, are very amenable to the formulation of path constraints.
+    \end{itshape}
+
+\end{itemize}
+\subsection*{Reviewer 2}
+
+\begin{itshape}
+I thank the authors for their thorough and organized rebuttal.
+The revised manuscript itself feels greatly improved in terms of its structure and coverage of background material.
+Nonetheless I have some remaining qualms about the introduction (ISSUE 1), and some more serious concerns regarding the analysis/illustration of the developed method (ISSUE 2).
+\end{itshape}
+
+\begin{itshape}
+ISSUE 1: introduction (requires minor revisions): 
+Despite some reduction, the intro is still long-winded.  For example,
+\end{itshape}
+
+\begin{itemize}
+    \item
+        \begin{itshape}
+            It's superfluous to explain at such length that spacecraft trajectory design is a problem of interest, that cubesats are a thing, etc.
+Broad points like these could potentially be condensed into single, citation-heavy sentences.
+        \end{itshape}
+    \item 
+        \begin{itshape}
+            There is also still significant repetition. For instance, all the paragraph-to-paragraph transitions on page 3 feel very similar.
+        \end{itshape}
+    \item 
+        \begin{itshape}
+            Some paragraphs are far too long
+        \end{itshape}
+    \item 
+        \begin{itshape}
+            I would like to reiterate my previous suggestion to organize the intro into subsections.
+This will help you weed out repetition and will make the paper more reader-friendly instead of a wall of text.
+        \end{itshape}
+    \item 
+        \begin{itshape}
+            The instability of RK for decades-to-centuries propagations seems irrelevant or misleading given the time scales used in this paper.
+        \end{itshape}
+    \item 
+        \begin{itshape}
+            The point brought up about "suboptimal direct optimization" versus the authors' approach isn't really translated into anything meaningful in the body of the paper.
+Along with the RK vs variational point, I feel like the route you've chosen has some nice properties but shouldn't necessarily be sold as a crucial enabler; the key of the paper lies elsewhere.
+        \end{itshape}
+\end{itemize}
+
+\begin{itshape}
+ISSUE 2: analysis/illustration (requires major revisions): 
+My other highlighted area of concern with the original manuscript was the degree of (in)completeness in illustrating the method and its pros and cons.
+I do not feel that the authors have done very much to improve this aspect of the paper beyond adding some caveats that feel a little half-hearted rather than properly contextualized discussions.
+This aspect of the research remains a major shortcoming that must be addressed before publication.
+
+While my request for improvements was itself not very specific, I had hoped that some of my other questions would reveal some areas of consideration for the authors.
+Addressal/discussion of those questions furthermore needs to appear adequately in the paper, not just the rebuttal.
+Below I will try to sketch how these points of inquiry can lead to a more demonstrative analysis.
+
+\end{itshape}
+
+\begin{itemize}
+    \item 
+        \begin{itshape}
+            The assumption of fixed travel time.
+The authors respond that this choice is made to simplify the optimization problem.
+Of course this is the case, and is conventional, but this choice is made in a rather arbitrary fashion rather than properly analyzed (especially considering that a central purpose of the problem setup is to enable faster travel times).
+Imagine plotting low-thrust reachability curves as in Fig8b for each of a sequence of travel time choices.
+This would reveal the influence of the arbitrarily handled parameter, and show that there's some critical point at which the reachablility curve intersects the targeted manifold.
+Maybe exhaustive exploration of this dimension is not appropriate in every application, but the relationship is worth illustrating.
+        \end{itshape}
+
+    \item 
+        \begin{itshape}
+            The specification of a departure point from the periodic orbit.
+In response to this question, the authors describe how the use of low-thrust propulsion enlarges the reachable set about some "center" defined by one branch of the uncontrolled manifold.
+This is a key way to think about the approach you've posed but it is not illustrated in the paper.
+Why not generate some curves under different thrust bounds for a specific departure point/manifold branch?
+Again you'd be using a sidelined parameter to illustrate something fundamental about the approach.
+And just maybe there would be a point at which, with high thrust, you start to see "bang-bang" type control and establish more connection to intuition.
+        \end{itshape}
+
+    \item 
+        \begin{itshape}
+            Actually, the influence of the starting point selection *is* somewhat illustrated by the green unstable manifold of Fig7b/8b.
+The way a mission designer might think about applying this tool could be to start with this information as a baseline and then use principles revealed by the preceding two types of illustration to make some informed way of selecting a combination of [departure point, travel time, max thrust] to get the sets to intersect without oversizing their system.
+That said, it's a little confusing how the red and green curves of Fig8b don't line up with this description to fit the comment of P18L47; additional consideration/analysis may be needed with regards to the number of revs/crossings used to define a section for comparison.
+        \end{itshape}
+
+    \item
+        \begin{itshape}
+            I previously mentioned my dissatisfaction with the interpretability of the control solution (which is plotted in planar components).
+On further consideration, I think a time series plot of the Jacobi energy would be beneficial to help vet whether the method is working reasonably.
+From intuition, one might expect an efficient solution to exhibit a span of monotonic energy increase, a relatively constant coasting span, and then a monotonic decrease back to the Poincare section value.
+As is, the two roughly constant thrusting phases visible in the plot suggest that this might be happening, but it's hard to be sure.
+Less consistent behavior might indicate, for example, that a longer-than-necessary transfer time was used.
+        \end{itshape}
+
+    \item 
+        \begin{itshape}
+            Following that thought, I have serious concerns about chaining this Poincare-based approach, as done in the second problem and as I previously asked about.
+The authors' responded: "While the trajectories do return to the energy level of the Poincare section, there is a transfer of "energy" between the states of the vehicle." Frankly, this statement doesn't make sense to me.
+I think that you are actually getting a very inefficient solution by forcing these repeated (and at fixed intervals) returns to the same energy level (again, plot the Jacobi time series and see if you can square it with your intuition).
+Furthermore posing the problem as a transfer to the stable manifold kind of hamstrings the motivation of reducing transfer times.
+Honestly, I think this second example should simply be axed.
+If the first one is expanded enough to illustrate the problem fundamentals, it will be sufficient.
+Conducting the same kind of analysis at a couple different rev counts would
+            reveal the dimension that this second example added, and help reveal the limits of the approach under certain scenarios and thrust capabilities.
+        \end{itshape}
+
+    \item 
+        \begin{itshape}
+            To summarize my thoughts on ISSUE 2:
+            As they exist, the examples illustrate that solutions can be found but they do not enlighten the reader as to the characteristics of the problem formulation/tool that should be understood and kept in mind when applying it or when planning related research efforts.
+Furthermore, I am very suspicious that Example 2 is not actually a good solution or a compelling use case for the approach.
+I do not necessarily ask or expect the authors to follow my above outline precisely, but I insist that some changes occur to make the analysis/illustrations better convey the general nature of the tool, not just that it gives you an answer.
+
+            A more collected and focused discussion about how the parameter choices, assumptions, and limitations of the tool should inform its use is also warranted; currently remarks to this effect feel like scattered afterthoughts.
+One organized way to think about this is to distinguish all the dimensions of the problem and identify what is being solved rigorously, what is being set manually, what is being numerically explored, etc.
+        \end{itshape}
+\end{itemize}
 \section*{Response to the Reviewers' Comments for JASS-D-17-00005}

### 2018-05-08 setup_repo.sh
Remove setup repo script

@@ -1,20 +1,0 @@
-#!/bin/bash
-
-# add the correct remote repositories
-git_repo="git@github.com:skulumani/2017_JAS_manuscript.git"
-bb_repo="git@bitbucket.org:shankarkulumani/2017_jas_manuscript.git"
-fdcl_repo="git@github.com:fdcl-gwu/2017_JAS_manuscript.git"
-
-printf "Setting the origin remote to point to Github and Bitbucket:\n\n"
-printf "Github: $git_repo\n"
-printf "Bitbucket: $bb_repo\n\n"
-
-
-git remote set-url origin --push --add $bb_repo
-git remote set-url origin --push --add $git_repo
-git remote add fdcl $fdcl_repo
-
-printf "Finished! The new remotes are listed.\n"
-printf "\n"
-
-git remote -v

### 2018-05-09 Response.tex
Reviewer 1 comment 1
Incorporate changes for first comment. Overselling the integrator

@@
 
+    This is a very good point. 
+    The manuscript has been modified to remove the about the benefits of state accuracy.
+    The section has been modified to highlight the reduced computational load of the varaitional integrator.
+    In addition, an additional sentence has been added specifically highlighting the fact that high order Runge-Kutta methods will probably provide better state accuracy at the expense of additional computations and poorer energy behavior.
+
+    \begin{correction}
+        This work uses a second order integrator which has improved energy behavior in comparision to first order integration techniques without a significant increase in computational demand.
+        The variational integrator provides improved energy behavior over long integration times but does not guarantee state accuracy.
+        High order Runge-Kutta methods will likely provide a more accurate state trajectory, but at the expense of additional computational demand and total energy deviations.
+
+    \end{correction}
     \item \begin{itshape}

### 2018-05-09 manuscript.tex
Fix indent
Reviewer 1 comment 1
Incorporate changes for first comment. Overselling the integrator

@@
 In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
-This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge-Kutta methods. 
+This work uses a second order integrator which has improved energy behavior in comparision to first order integration techniques without a significant increase in computational demand.
+The variational integrator provides improved energy behavior over long integration times but does not guarantee state accuracy.
+High order Runge-Kutta methods will likely provide a more accurate state trajectory, but at the expense of additional computational demand and total energy deviations.
 However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.

### 2018-05-11 manuscript.tex
Modify sentence to reduce selling of indirect control
Reviewer 1 #3

@@
 \section{Conclusions}\label{sec:conclusion}
-In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
+In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~sections is used to generate a transfer between planar periodic orbits in the three-body problem.
 The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
-The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
+The indirect optimal control formulation direclty solves for the state and costate vectors.
 Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.

### 2018-05-11 Response.tex
Modify sentence to reduce selling of indirect control
Reviewer 1 #3
Remove previous version of reviisoin comments

@@
 
-    This is a very good point. 
     The manuscript has been modified to remove the about the benefits of state accuracy.
@@
 
+    This sentence has been modified.
+    \begin{correction}
+The indirect optimal control formulation direclty solves for the state and costate vectors.
+    \end{correction}
 \end{itemize}
@@
 \end{itemize}
-\section*{Response to the Reviewers' Comments for JASS-D-17-00005}
-
-I would like to thank the reviewers for their thoughtful comments, which are aimed towards improving the quality of the manuscript and the clarity of the results. 
-In accordance with the comments and suggestions, the manuscript has been revised, and the answers to all comments are addressed as follows.
-
-(In the revised manuscript, the citation numbers for equations, assumptions, propositions, and references are changed. This answer is written according to the new item numbers.)
-
-\subsection*{Reviewer 1} 
-\begin{itshape}
-Reviewer \#1: I have some issues with this paper that, in my opinion necessitate a major revision.
-These issues are:
-\end{itshape}
-
-\begin{itemize}
-    \item 
-        \begin{itshape}
--It extensively re-covers issues that have been covered in many papers, making it difficult to disseminate what the original contributions are.
-For example, I don't think the addition of the thrust term into the dynamics necessitates re-doing every piece of analysis that has been done with various resources in the literature about variational integrators (including comparisons with RK methods).
-Things that can be found in other research should be cited, and summary comments can be used to point the readers to fill in this knowledge.
-The content of this paper easily be reduced considerably to maintain focus on the original contributions.
-    \end{itshape}
-
-    Thank you for the comment, and we agree that there is repetition of material previously covered in other publications.
-    The challenge is that the proposed approach utilizes several tools and concepts integrated together to design low-thrust orbital maneuvers in a systematic way, and in order to describe the key idea in a self-contained manner, each of prior approach should be described properly.
-
-    To address this, the manuscript has been reduced and reorganized to enhance the clarity and clearly separate previously material and the contributions of this paper.
-    The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
-    The second section has been reorganized to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the and variational integrator.
-The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.
-The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, which are the main contributions of this paper.
-
-    In short, the manuscript has been completely reorganized such that all of the background materials are summarized in Section 2.
-The remaining sections dealing with the proposed approach are expanded with more detailed descriptions and numerical results.
-
-\item
-    \begin{itshape}
--It does not give a complete picture of the pros and cons of the proposed methodology.
-This method is full of both advantages and disadvantages.
-I don't think the authors need to be concerned with "selling" the work to the community, but rather showing the full and honest picture given the current state of the research.
-In the conclusions, the potential cons to this research can be mitigated by setting a path forward for future research to further improve the methodology.
-
-These pros/cons include:
-
--The formulation of the problem with acceleration magnitude is a major technical challenge for mission design practitioners.
-The state of the art to design trajectories that depend on low-thrust engine hardware is either:
-
--To produce constant thrust magnitude or no thrust magnitude (typical of nuclear electric propulsion, or very simple electric propulsion systems.)
-
--To produce variable thrust magnitude associated with various throttle points at different power accessibility levels (solar electric propulsion).
-That is not to say that useful solutions cannot be extracted from the problem, as the authors have shown.
-However, a great many solutions in this approach will be unusable due to the fact that acceleration and thrust magnitude are different, i.e.
-a=T/m
-
-Thus - a MAJOR focus of future work should be the introduction of a seventh state (mass).
-This gives rise to a more challenging optimal control problem requiring the switching function, but it is the form that is generally needed in computing optimal open-loop trajectories that can actually be flown.
-\end{itshape}
-
-Thank you for the comment and more focus has been included on the pros and cons of this proposed method.
-Additional detail has been added to the conclusion which discusses the advantages and the disadvantages, as well as the avenues for future research.
-The conclusion is copied below.
-\begin{correction}
-In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
-The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
-The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
-Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.
-Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
-Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
-However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
-Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
-
-There is additional research to extend these results to more general transfer scenarios in future work.
-The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
-In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
-Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
-This will result in a more complicated optimal control problem and is a focus of future research.
-Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
-The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
-\end{correction}
-
-\item 
-    \begin{itshape}
-The pros and cons of the PCRTBP
-
--The PCRTB gives rise to exploiting Poincare sections in a systematic manner.
-2D solutions might be close enough for qualitative understanding of the solution space for many problems (as long as large plane changes aren't needed).
-
--The PCRTB is limiting for many pragmatic departure orbits (e.g. LEO and GTO, the most common departure orbits for low-thrust transfers).
-\end{itshape}
-
-Thank you for the comment and these characteristics of our choice of the PCRTBP.
-The authors would like to point out that the key idea of designing low-thrust orbits through the reachability analysis on the \Poincare section may be applied to other dynamic system.
-This manuscript illustrates such idea through PCRTB.
-
-However, according to the reviewers suggestion, additional details have been stated in the problem formulation section, including discussions on the choice of the PCRTBP and the relative strengths/weaknesses of this approach as copied below.
-
-\begin{correction}
-We utilize the planar circular restricted three body problem~(PCRTBP) as the basis of our system definition. 
-It is a popular model in the preliminary analysis of multibody spacecraft trajectories. 
-In the context of Earth based mission, the PCRTBP affords a relatively simple dynamic model while still capturing the major third body perturbation of the Moon.
-In addition, this model allows for a systematic process to define and exploit \Poincare sections.
-Furthermore, the 2D solutions afforded by the PCRTBP are frequently use to gain a qualitative understanding of the trade space of transfers in the Earth-Moon system.
-The PCRTBP approach offers insight into the fundamental dynamical structure while capturing the major dynamic properties of the planar motion.
-As a result, this approach is best suited for trajectories which do not require large plane changes.
-\end{correction}
-
-\item 
-    \begin{itshape}
-The pros and cons of variational integrators should really be acknowledged rather than completely dismissing the RK methods which are still the current state of the art in mission design (using higher-orders and suitable tolerances)
-
--Pros = boundedness in the energy behavior.  Very long term integration will not deviate from the known dynamical structure.
-
--Cons = you don't know where you are on the energy surface, just that you are on it somewhere.
-This is still a problem in actual trajectory design.
-Knowing the error is on an energy surface does not mean that you don't have error.
-\end{itshape}
-
-Thank you for the comment and the detail on the benefits/drawbacks of variational integrators.
-The authors agree that the general-purpose numerical integrators such as higher-order RK methods are suitable in many cases.
-The proposed approach utilizes the variational integrator, particularly due to its numerical structural stability of capturing long-term behaviors of small effects accurately and efficiently, which is critical in low-thrust orbital maneuvers.
-
-While the structural stability of geometric numerical integrators is often illustrated by the conserved quantities such as energy or momentum, it has been shown that variational integrators are also effective in computing the rate of energy change for non-conservative systems~\cite{marsden2001}.
-
-To clarify these, additional discussion has been included in the paper in the subsection which introduces the variational integrator for the PCRTBP.
-Much of the discussion of variational integrators in the introduction has also been removed.
-The additional content, which discusses the relative benefits of variational integrators is copied below.
-
-\begin{correction}
-Both integration schemes result in trajectories that are initially nearly identical.
-\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
-For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
-Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
-In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
-This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
-However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
-In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
-In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
-The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.
-\end{correction}
-\end{itemize}
-From here on out, comments specific to various sections of the paper follow:
-
-\begin{enumerate}
-    \item
-        \begin{itshape}
-            In the statement "In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential
-            missions at a reduced energy cost [15]" - you may want to reference, for example, the Lunar Ice Cube mission paper which absolutely requires low-thrust propulsion and non-keplerian dynamics to satisfy its mission objectives.
-
-            Folta, D., Dichmann, D., Clark, P., Haapala, A., and Howell, K., "LunarCube Transfer Trajectory Options," AAS/AIAA Space Flight Mechanics Meeting, Williamsburg, Virginia, January 11-15, 2015.
-        \end{itshape}
-
-        Thank you for the comment and the additional reference.
-        In addition to this change we have made other modifications to the introductory paragraphs to improve the organization and clarity.
-        Specifically, the first paragraph has been divided, which motivates the use of electric propulsion and the advantages of small satellites which use electric propulsion.
-        In addition, the suggested reference has been included in the paper.
-        The modified paragraphs are duplicated below.
-
-        \begin{correction}
-            Designing spacecraft trajectories is a classic and ongoing topic of research.
-            There has been significant research into the design of orbital transfers for space vehicles.
-            Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
-            Electric propulsion systems offer a much greater specific impulse than chemical systems.
-            As a result, the greatly increased efficiency allows for greater payload mass or extended duration missions.
-            However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital maneuvers have a much longer time of flight.
-            In spite of this drawback, a wide variety of missions, such as communication and deep space probes, have utilized the unique benefits of low thrust electric propulsion to great effect~\cite{choueiri2009}.
-
-            With reduced development intervals and decreased launch costs, small satellites have gained increased attention as a cost effective means of scientific and technologic development. 
-            Furthermore, these cubesats are easily launched as secondary payloads rather than requiring a dedicated launch.
-            The merger of small satellites with miniature electric propulsion enables inexpensive and responsive missions requiring large changes in orbital energy or extended mission lifespan.
-            Recent developments in miniature electric propulsion offer the potential for new research opportunities for small spacecraft~\cite{haque2013}.
-            The advancements in miniature electric propulsion allows for greater flexibility in the deployment of cubesats as secondary payloads.
-            Without the explicit requirement of a dedicated launch trajectory, cubesats are able to exploit a much wider range of possible mission designs~\cite{folta2015}.
-            With the potential for more demanding missions, even greater importance is placed on the mission design to ensure that optimal trajectories satisfy mission requirements. 
-            In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{folta2015,koon2011}.
-            Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional patched-conic approach~\cite{ross2006,gomez2001}.
-            Optimally combining the structure of the dynamic environment with low-thrust propulsion systems is vital for future mission success.
-        \end{correction}
-
-\item 
-    \begin{itshape}
-"Since there is an insufficient number of analytical constants, or integral of motion, numerical methods must be used to investigate solutions to the three-body problem."  This isn't completely true as there are also approximations to the (uncontrolled) three-body motion, such such as the Richardson-Cary approximation.
-Saying numerical methods are required may be a bit strong.
-May want to say "often required".
-\end{itshape}
-
-Thank you for the comment.
-The sentence has been modified as follows.
-
-\begin{correction}
-Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
-As a result numerical methods are often required to investigate solutions to the three body problem.
-\end{correction}
-
-\item 
-    \begin{itshape}
-"These techniques suffer from numerical instability and energy drift behaviors which make them ill-suited for long-term propagation.
-These dissipative effects are even more detrimental with the addition of low-thrust propulsion to the dynamic equations of motion."  It really depends on what you mean by "long term".
-If transfers are on the order of hundreds of days, then high precision runge-kutte integrator, for example a Runge-Kutte Verner 8(9), can maintain suitable accuracy such that the energy drift in the Hamiltonian is very small, as can be seen in some of the references that the authors have cited.
-If, on the other hand you mean 10-100's of years or more, i.e., when you create your Poincare maps, then this drift is more apparent.
-I think that this statement needs to be quantified better because it gives a dishonest portrayal of high-precision Runge-Kutte integrators without numbers attached to it.
-\end{itshape}
-
-Thank you for the comment.
-Those numerical dissipations would depend on several parameters.
-But, for the particular numerical comparison illustrated by \Cref{fig:compare_energy}, such difference in conservation properties becomes noticeable in the simulation for 15 years.
-As suggested, the claims about the variational integrator have been qualified with some additional detail.
-The correction is copied below.
-\begin{correction}
-As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g.
-on the order of decades to centuries, which is typical of the three-body problem.
-\end{correction}
-
-\item
-    \begin{itshape}
-"Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system [8].
-As a result, the long term effects of low-thrust on the spacecraft trajectory are not accurately captured."  Again, I think situationally I would disagree.
-Transfers on the order of 100's of days can maintain accuracies of 1e-10 (non-dimensional units) in the Hamiltonian, which is generally considered a "small" amount of energy drift.
-\end{itshape}
-
-These sentences have been modified and combined to enhance clarity and reduce some of the unnecessary bloat. 
-In addition, some additional detail has been added to explicitly define the notion of ``long-term'' in the context of numerical integration.
-As shown by \Cref{fig:compare_energy}, the variational integrator exhibit smaller conservation errors than ODE45 at the similar computational efforts, for the particular numerical case considered.
-The modifications are copied below.
-\begin{correction}
-    Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
-    As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g.
-on the order of decades to centuries, which is typical of the three-body problem.
-\end{correction}
-
-\item 
-\begin{itshape}
-- "[15,25] have illustrated the rich structure that exists in the three-body problem." 
-- "[23,6] implement the solutions using conventional Runge-Kutta integration techniques."
-I don't think you can grammatically begin a sentence with reference numbers.  Please fix all instances of this.
-\end{itshape}
-
-Thank you for the comment.
-Evidently in some styles, such as \href{https://www.dlsweb.rmit.edu.au/lsu/content/1_studyskills/study_tuts/apa_ll/position.html}{APA}, it is acceptable to start with a reference.
-We agree that it is not the best with the numeric style citations.
-The sentences have been reworded to avoid starting with a reference number. 
-The corrections are shown below.
-
-\begin{correction}
-    \begin{itemize}
-        \item Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
-        \item There exists a rich structure in the three-body problem~\cite{koon2011,ross2006}.
-    \end{itemize}
-\end{correction}
-
-\item 
-    \begin{itshape}
-"It has been shown that there exist multi-dimensional tubes, or invariant manifolds, of constant energy trajectories that span the state space."  The "tube" statement is really only true for some libration point orbits at certain Jacobi Constants.
-For some orbits these structures are more chaotic and don't resemble tubes.
-May want to soften this language or change "tubes" to "structures".
-\end{itshape}
-
-Thank you for the clarification.
-The sentence has been modified as follows.
-
-\begin{correction}
-It has been shown that there exist multi-dimensional structures, termed invariant manifolds, of constant energy trajectories that span the state space. 
-\end{correction}
-
-\item 
-    \begin{itshape}
--"In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem." Why planar?
-For simplicity?
-Almost all parking orbits departing the Earth are likely going to have some inclination with respect to the rotating reference plane, and will likely require some three-dimensional thrusting to target the manifold.
-You may want to indicate that this is a first step in the research to exploit the Poincare sections more easily, which is why I'm assuming it was done.
-It doesn't hurt to be straightforward about this, as it is a first-step towards characterizing the design space.
-\end{itshape}
-
-Thank you for the comment.
-The PCRTBP was primarily chosen as a simple dynamic system which still captures the primary structure of the three-body problem.
-The introduction has been modified with an additional sentence indicating our use of the planar three body problem
-
-\begin{correction}
-Utilizing the planar assumption greatly simplifies the problem and allows for the use of established methods in determining periodic orbits and \Poincare sections.
-\end{correction}
-
-\item 
-    \begin{itshape}
--"In addition, through the use of geometric integrators we accurately capture the effects of low thrust on the system dynamics in the numerical simulation."  A variational integrator will accurately capture the qualitative effects of the transfer (e.g.
-propellant use), however it will not necessarily accurately capture accuracy in the final state vector - it is guaranteed to be on an energy surface, but it is not guaranteed to be at exactly the right location on the energy surface.
-I would soften this claim or reword.
-\end{itshape}
-
-While the geometric integrator will not exactly characterize the terminal energy or state of a dynamic system, it will better preserve the total energy variation as compared to conventional integration schemes.
-In addition, the total energy deviation will remain bounded and less than that of conventional schemes which experience unbounded energy deviation. 
-As a result, the terminal state/energy will be more accurately characterized using a variational integrator as compared to conventional schemes. 
-Also this difference becomes more evident over increasing time horizons.
-Some additional detail has been added stating that the increased accuracy is most evident over long time periods.
-
-\begin{correction}
-In addition, through the use of geometric integrators we more accurately capture the effects of low thrust on the system dynamics in the numerical simulation over extended periods which are typical of electric propulsion systems.
-\end{correction}
-
-\item 
-    \begin{itshape}
--"The Poincare section reduces the dimensionality of the system dynamics to the study of a related discrete update map." It does, but at the expense if a large up-front computational run that must be run well in advance to capture the problem characteristics.
-Also, it is more challenging to execute in 3D, which should be noted.
-\end{itshape}
-
-Thank you for the comment.
-Some additional detail has been included mentioning the difficulties in applying the \Poincare~section approach to the 3D case.
-
-\begin{correction}
-While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
-\end{correction}
-
-\item
-    \begin{itshape}
--"This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs."  Yes, but the order of the variational integrator still matters.
-The drift in the state vector accuracy is still important to a mission designer, even if it is guaranteed to lie on an energy surface.
-I think you should indicate that this is more for preliminary analysis to characterize the transfers, when this drift in state vector accuracy is tolerable.
-I.E., it's useful for the introductory phases of mission design when trying to globally characterize the design space.
-Characterizing the scope of the analysis in this way would be helpful to the audience.
-\end{itshape}
-
-Thank you for the comment.
-The section on the variational integrator has been reduced and this specific sentence has been removed. 
-However, additional detail has been provided in the variational integrator subsection which discusses this drawback.
-
-\begin{correction}
-
-Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
-In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
-This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
-However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
-
-\end{correction}
-\item 
-    \begin{itshape}
-"Optimal solutions are generally sensitive to small variations in the initial multipliers.
-As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance."  I don't agree that this implies that variational integrators are better for gradient-based methods.
-The primary issue is that gradient-based methods that are solving optimal control problems are using first-order approximations to iterate on the problem.
-The first-order approximation is what is poor, not necessarily the gradients.
-Even if these gradients are of exact accuracy (and I agree that they should always be as accurate as possible), the first-order methods in the chaotic problem are what give the difficulties.
-One could use a variational or standard integrator in a first-order gradient based method and nothing will change the fact that the method will be highly sensitive, save for partitioning the sensitivities as you did with your multiple shooting approach.
-At best, you can change the statement to indicate that the accuracy in the gradients may improve, which is always helpful to gradient-based methods, but it's no guarantee that it eliminates major problem sensitivity.
-\end{itshape}
-
-It has been shown that the structure-preserving properties of geometric numerical integrators are also useful to numerical optimizations for improved computational efficiency and stability.
-For example, see the following reference.
-
-\begin{itemize}
-\item[-] Ober-Blöbaum, Sina, Oliver Junge, and Jerrold E. Marsden. "Discrete mechanics and optimal control: an analysis." ESAIM: Control, Optimisation and Calculus of Variations 17.2 (2011): 322-352.
-\end{itemize}
-
-The proposed approach utilizes desirable numerical properties of the variational integrator.
-However, the main contribution is systematic design of low-thrust orbital maneuvers through reachability set analysis on the \Poincare section.
-The above sentences may be confusing, and the authors decided to remove them from the introduction of the revised manuscript.
-
-\item
-    \begin{itshape}
-Authors may also want to point out the loss of accuracy in variable step integrators that can also degrade first-order gradient based methods.
-I didn't see this anywhere and it's something that should be acknowledged.
-Basically, variable step integrators will introduce artificial "noise" into the gradients:
-
-Pellegrini, E.P., Russell, R.P., On the Accuracy of Trajectory State Transition Matrices, Paper AAS-15-785, AAS/AIAA Astrodynamics Specialist Conference, Vail, CO, Aug 2015
-\end{itshape}
-
-Thank you for the comment and helpful reference.
-
-It is well known that variable step size may break conservational properties of a geometric numerical integrator.
-Variational integrators are no exception, and there has been a multisymplectic framework to address it.
-See
-
-\begin{itemize}
-\item[-] Lew, Adrian, et al. "Asynchronous variational integrators." Archive for Rational Mechanics and Analysis 167.2 (2003): 85-146.
-\end{itemize}
-
-However, the above formulation for structure-preserving, adaptive time-step variational integrators is beyond the scope of this paper, and we use a fixed time step throughout the presented numerical results.
-
-Additional detail has been included in the manuscript discussing this potential drawback of variable step integrators.
-
-\begin{correction}
-
-In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
-The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
-
-\end{correction}
-
-\item 
-    \begin{itshape}
-"This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods." I disagree that suboptimality is typical of direct methods.
-A direct method will typically solve the KKT conditions, which can be directly transcribed into the costates in the optimal control problem.
-A direct method fully satisfying the KKT conditions can map to the indirect problem.
-If you want to strengthen this argument, then you should argue that it applies to direct methods which rely on coarse or suboptimal representations of the control.
-For example, a hybrid direct method could use continuous optimal control theory control law, but directly minimize the cost function instead of solving the full TPBVP.
-\end{itshape}
-
-Thank you for the comment.
-The suboptimal claims of direct optimization has been reduced and modified.
-Additional detail has been included on the merits of indirect optimal control vs. direct optimal control methods.
-\begin{correction}
-This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
-For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
-\end{correction}
-
-In addition, the introduction has been modified to explicitly mention that direct methods which use suboptimal control parameterizations will result in suboptimal solutions.
-
-\begin{correction}
-The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which generally rely on suboptimal representations of the control.
-\end{correction}
-
-\item 
-    \begin{itshape}
-"Application of the Euler-Lagrange equations results in following equations of motion defined in the rotating reference frame"  Add a reference after this statement for those who want to see the derivation.
-One of the given references probably does already does this, so just pick one
-\end{itshape}
-
-Thank you for the comment.
-An additional comment has been included in the problem formulation section with a reference to a source of the detailed derivation.
-\begin{correction}
-Following a straightforward application of the Euler-Lagrange equations, a more detailed derivation is provided in~\cite{szebehely1967}, results in the following equations of motion defined in the rotating reference frame
-\end{correction}
-
-\item 
-    \begin{itshape}
-It appears from equation 5 and your control input definition that you control is unbounded?
-Later on in reading the paper I found that you do add bounds on the control.
-You may want to indicate earlier on in the paper (closer to eq.
-5) that bounds will be added to the control.
-\end{itshape}
-
-Thank you for the comment.
-An additional comment has been included mentioning that the control is assumed to be bounded in magnitude.
-
-\begin{correction}
-\ldots and the control inputs is defined as \( \vecbf{u} = \begin{bmatrix} u_x & u_y \end{bmatrix}^T \in \R^{2\times1} \) and assumed to be continuously variable but bounded in magnitude, i.e.
-\( \vecbf{u}^T \vecbf{u} \leq u_{max}^2 \).
-\end{correction}
-
-\item 
-    \begin{itshape}
-Figure 1 should probably have "axis equal" applied to it.
-For that matter, please check all trajectory plots (those which plot the position state), and ensure that all plots are using "axis equal".
-If not, please regenerate those plots.
-\end{itshape}
-
-Thank you for the comment.
-All of the figures showing position states have been modified to use \texttt{axis equal}.
-
-\item 
-    \begin{itshape}
-Section 3.1 has been discussed in many references, the reader may want to direct them to this material.
-(Choose one of Marsden or West's papers for example.)
-\end{itshape}
-
-Thank you for the comment.
-The discussion on the variational integrators has been reduced. 
-Much of the background material on Lagrangian mechanics is removed and an additional paragraph is included to summarize the key ideas of variational integrators. 
-In addition, the variational integrator details are moved into a titled subsection.
-The additional paragraph is shown below.
-
-\begin{correction}
-Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invariant properties and symplecticity.
-Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations.
-As a result, integrators developed in this manner have the desirable properties that they are symplectic and momentum preserving.
-In addition, they exhibit improved energy behavior over long integration periods.
-A thorough discussion of variational integrators is provided in~\cite{west2004,marsden2001}.
-We use this approach to construct a variational integrator for the PCRTBP with the inclusion of low-thrust propulsion.
-
-Consider the autonomous mechanical system described by the Lagrangian, \( L(q, \dot{q} ) \), for the generalized coordinates, \( q\) and velocities \( \dot{q} \).
-The integration of the continuous Lagrangian along a path, \( q (t) \), followed by the system must satisfy Hamilton's principle, which results in the well-known Euler-Lagrange equations~\cite{lanczos1970}.
-In the discrete time scenario, the continuous path \( q(t), \dot{q}(t) \), is replaced by a finite difference approximation over a fixed time step, e.g.
-\( q_0, \frac{q_1 - q_0}{h} \), which converges to the true velocity as \( h \) tends towards zero.
-In this manner, we define a discrete Lagrangian, \( L_d (q_0, q_1, h) \) which approximates the integral of the true Lagrangian over the time interval \( h \) between \( q_0 \) and \( q_1\)~\cite{marsden2001}.
-\end{correction}
-
-\item 
-    \begin{itshape}
-The second-order trapezoidal rule will give significantly poorer state vector accuracy with respect to a high-precision RK integrator, even if the state is guaranteed to lie on an energy surface.
-Why didn't the authors consider going to a higher-order variational integrator, or even a Taylor Integrator?
-It is very likely that the integration accuracy will introduce some errors in the accuracy of the partial derivatives and in the states.
-\end{itshape}
-
-We chose the trapezoidal rule as it provides a second order quadrature approximation.
-Furthermore, instead of using Taylor methods or higher order options, the trapezoidal rule provides an explicit second order accurate integration rule.
-This greatly reduces the complexity of the optimization routine.
-In addition, the background material on the variational integrators have been removed from the manuscript. 
-Only the specific details covering the development of the variational integrator for the PCRTBP are discussed.
-
-\item 
-    \begin{itshape}
-Section 3.2 - please cite a reference as to where the selected quadrature rules presented in Table 1 have been derived.
-\end{itshape}
-
-In the interest of conciseness and to avoid repeating material that has already been covered in the literature, the discussion on variational integrators has been substantially reduced.
-The background material on the variational integrators has been removed and this table has also been removed from the manuscript.
-
-\item 
-    \begin{itshape}
-Figure 2 - x-axis has no label but I think it's non-dimensional time, please indicate.  Label all axes even if they are non-dimensional units.
-\end{itshape}
-
-Thank you for the comment. 
-All of the plots have been updated to include all axes labels.
-Furthermore, the specific plot in questions has been replaced with another plot more explicitly showing the benefits of variational integrators over conventional integration schemes.
-This modified plot is shown below.
-
-\begin{correction}
-    \begin{figure}[H] 
-        \centering 
-        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparison: A stable trajectory is used to test the variational and conventional integrators. 
-        The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
-        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
-        \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}}
-        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
-\end{figure}
-\end{correction}
-\item 
-    \begin{itshape}
-I recommend removing appendix B.
-Giving the equations and stating that the Matrix inversion is from Gauss-Jordan elimination is probably sufficient, unless there is some compelling algebraic manipulation that is very unique to the problem.
-\end{itshape}
-
-Thank you for the comment.
-Appendix B has been removed from the revised manuscript.
-The manuscript has been modified to mention that Gauss Jordan elimination is used to avoid the use of the matrix inversion.
-
-\begin{correction}
-The derivation of~\cref{eq:costate_eom} is given in Appendix A.
-In addition, the computation of~\cref{eq:costate_eom} requires inversion of the Jacobian matrix.
-This is a computationally expensive operation that is prone to numerical error and instability.
-We use Gauss-Jordan elimination to avoid this inversion in~\cref{eq:costate_eom} and determine an explicit update map \( \vecbf{\lambda}_k \to \vecbf{\lambda}_{k+1} \).
-\end{correction}
-
-\item 
-    \begin{itshape}
-In line 14 (page 12), you define the state transition matrix over one orbital period, and then on line 18 you start referring to it as the monodromy matrix.
-I understand what you're doing, but you may want to make the definition more explicit.
-Some of this section can be substantially reduced and other source material could be referenced anyway.
-\end{itshape}
-
-Thank you for the comment.
-The discussion about the invariant manifolds and \Poincare sections has been reduced.
-This discussion has been removed from the revised manuscript and replaced with a general summary and some relevant citations.
-This specific discussion no longer exists in the manuscript but rather cites several references with the details on determining the invariant manifolds.
-
-\item 
-    \begin{itshape}
-Comparison like that of Figure 3.4 are very common in the well-known literature concerning variational integrators.
-Moreover, the proper way to present such a comparison is with different delta-T's in the integration step, showing when the Runge-Kutta methods actually break down in terms of maintaining an accurate portrayal of the Energy/Hamiltonian conservation.
-
-For example -
-
-Figure 2.4
-Lew, A., Marsden, J., Ortiz, M., West, M., "An Overview of Variational Integrators," FINITE ELEMENT METHODS: 1970's AND BEYOND L.P. Franca (Ed.) c CIMNE, Barcelona, Spain 2003.
-
-Online copy:
-http://lagrange.mechse.illinois.edu/pubs/LeMaOrWe2004/LeMaOrWe2004.pdf
-\end{itshape}
-
-Thank you for the comment.
-The comparison plot has been modified to follow the precedent in the literature.
-The plot has been modified to show the energy deviation as a function of computational time for both the variational integrator and \texttt{ODE45}.
-The modified content is copied below.
-\begin{correction}
-\begin{figure} 
-        \centering 
-        \begin{subfigure}[h]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{trajectory} 
-                \caption{Earth-Moon three-body trajectory used for integrator comparison}
-        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-%       \begin{subfigure}[htbp]{0.3\textwidth} 
-%               \includegraphics[width=\textwidth]{components} 
-%               \caption{State \( \parenth{x , y, \dot{x}, \dot{y}}\)} \label{fig:compare_components} 
-%       \end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=\textwidth]{figures/cputimevsE.pdf} 
-                \caption{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}} 
-        \end{subfigure} 
-        \caption{Integrator Comparison}
-\end{figure}
-A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
-A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
-    The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
-    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
-\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
-Both integration schemes result in trajectories that are initially nearly identical.
-\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
-For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
-Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
-\end{correction}
-
-\item 
-    \begin{itshape}
-It may also be generally worth acknowledging the benefits of Taylor's method within the context of variational integrators:
-Schmitt, J., Shingel, T., Leok, M., "LAGRANGIAN AND HAMILTONIAN TAYLOR VARIATIONAL INTEGRATORS," eprint arXiv:1703.06599, 2017
-
-Online copy: https://arxiv.org/pdf/1703.06599.pdf 
-\end{itshape}
-
-Thank you for the comment.
-However, this manuscript is not really focused on variational integrators but rather the design of transfer trajectories using reachability sets.
-There is a wide range of topic under the umbrella of variational integrators that are not directly applicable to this manuscript and work.
-Interested readers can find a wide variety of literature on the relative benefits of all manner of variational integrators in papers focused on that topic specifically.
-As a result, we do not feel it is applicable to include additional discussion on the merits of taylor integration methods.
-
-\item 
-    \begin{itshape}
-"Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques".
-I think you need to add "in which energy conservation is mandatory for characterizing the solution space."
-\end{itshape}
-
-Thank you for the comment.
-The additional clarification has been added to the manuscript and copied below.
-\begin{correction}
-Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
-\end{correction}
-
-\item 
-    \begin{itshape}
-Section 5 - This method, while very useful in its given context is highly dependent on the use of the PCRTB in the framework of the current analysis.
-This needs to somehow be conveyed.
-3D transfers are significantly more challenging when it comes to exploiting Poincare sections.
-It can be done, but more subplots, dimensions, and data storing are necessary to view the subspace - save to say, it is a more challenging endeavor that must be acknowledged for general purpose low-thrust mission design.
-\end{itshape}
-
-Thank you for the comment.
-An additional paragraph has been added to the beginning of the mathematical background section which discusses our reachability based transfer approach.
-This additional content is copied below.
-
-\begin{correction}
-The numerical examples presented in this section are designed in the context of the PCRTBP.
-The dynamic environment has a four dimensional state space and offers a convenient integration constant in the form of the Jacobi integral.
-As a result, there are well defined methods to define and exploit \Poincare sections, which result in straightforward two-dimensional subspaces of the system.
-Our approach uses the \Poincare section to approximate the reachability set on this reduced subspace.
-As a result, this approach is more difficult to apply to three-dimensional transfers in the general three body problem.
-\Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
-However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories~\cite{kulumani2016d}.
-\end{correction}
-
-\item
-    \begin{itshape}
-Figures 10D and 13D - the authors need to present these plots in dimensional coordinates so that they may be reconciled with what existing technology can produce.
-I recommend Newtons or milliNewtons.
-For example, the NEXT (NASA evolutionary xenon thruster) can produce 0.25 Newtons of thrust at maximum power (~7 kW).
-The XR5 Hall thruster can produce comparable thrust at 4.5 kW of power.
-These are currently the highest-power single engines other than the ARM Hall thruster which is not flight ready yet.
-For these examples to be reconciled against the capability of existing technology, we need to see the required forces that are being demanded from the optimal control problem.
-It may be that the results are beyond the scope of current technology - that's fine - just say that they are examples and that it's ongoing work to formulate example problems that capture existing technology better.
-\end{itshape}
-
-Thank you for the comment.
-The mentioned plots of the control input have been modified to show the data in dimensional units (Newtons). 
-Some additional detail has been provided in the manuscript which discusses the choice of \( u = 0.75 \) and it's relationship to the current state of the art thruster systems.
-The modified plots are not included here but the additional content is copied below in response to a different but related comment.
-
-\begin{correction}
-Next, we determine the reachability set with the addition of a low thrust control input.
-We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
-This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
-We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
-This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
-Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
-Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
-The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
-This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit. 
-\begin{figure}[H]
-    \centering
-    \subcaptionbox{Control input for the selected transfer trajectory}{\includegraphics[width=0.5\textwidth]{control_input_l1}}~
-    \subcaptionbox{Control Input: Combined control history for each stage of the reachability analysis. 
-    The control always remains within the maximum bound.}{\includegraphics[width=0.5\textwidth]{figures/control_input_geo.pdf}}
-\end{figure}
-\end{correction}
-
-\item
-    \begin{itshape}
-"The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches."  Did the authors attempt single shooting for their example problems before coming to this conclusion?
-With proper scaling, and problem formulation, there are examples in the literature of single shooting yielding robust results.
-
-Example: 
-Ranieri, C. and Ocampo,C..  "Indirect Optimization of Spiral Trajectories", Journal of Guidance, Control, and Dynamics, Vol. 29, No. 6 (2006), pp. 1360-1366.
-
-So, justification for its selection may be warranted.  You could always find a reference in which this trade space was really explored thoroughly to show why you made the choice.
-\end{itshape}
-
-Our original attempt at solving the optimal control problem did rely on single shooting.
-However, during our investigation it was found that it was quite difficult for solutions to converge.
-As a result, we decided to implement the multiple shooting approach in hopes of generating more consistent and robust solutions.
-While it is possible to use single shooting, such as the provided reference, it is known that the single shooting approach will have greater sensitivity to initial costate errors as compared to a multiple shooting method.
-This can be alleviated, or entirely mitigated in the case of~\cite{ranieri2006}, it is highly case specific and dependent on a large variety of case specific techniques. 
-For example, the work in~\cite{ozimek2010a} used both a single shooting and multiple shooting approach to solve the optimal control problem.
-The large sensitivity of the single shooting approach was mitigated through the use of the multiple shooting method, at the expense of additional design variables.
-
-In the manuscript, additional detail has been added in this section which describes the difficulties of the single shooting approach and our decision to use the multiple shooting method.
-The modified portion of the manuscript is copied here.
-\begin{correction}
-Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs~\cite{stoer2013}.
-This multiple shooting approach incorporates additional interior constraints but reduces the sensitivity of the costates along each sub-arc.
-The use of the multiple shooting method reduces the sensitivity of changes in the initial costate at the expense of additional design variables, but has been shown to provide more stable and robust solutions~\cite{ozimek2010a}.
-\end{correction}
-
-\item
-    \begin{itshape}
-"In this work, we use the Matlab nonlinear solver fsolve to solve the system of nonlinear equations defined by the multiple shooting algorithm."  Can authors give some additional details?
-Which fsolve algorithm?
-trust region dogleg,  trust region, or Levenberg-Marquardt?
-Which additional options?
-E.g.
-various tolerances.
-For users of this method it's also useful to see the iteration history (more comments on this later).
-\end{itshape}
-
-Thank you for the comment.
-Additional detail has been incorporated which discusses the specific solver used in computing the solutions. 
-The description does not specifically mention all of the possible options in \texttt{fsolve} but rather gives the key components used in our solution.
-This approach seems to be consistent with the literature, e.g.~\cite{ozimek2010a,stuart2010}.
-The modified portion is copied below.
-
-\begin{correction}
-In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
-Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
-All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
-\end{correction}
-
-\item
-    \begin{itshape}
-Did the authors experiment with the number of multiple shooting stages in their example problems?
-There must be some sort of "sweet spot" in number of stages for the example problems attempted.
-If I missed it, I apologize, but I didn't see it.
-\end{itshape}
-
-We did spend considerable time in determining the appropriate number of stages. 
-One key issue was that our implementation in software depended on dividing the total number of integration steps by the number of interior stages.
-For simplicity, this reduced the possible number of stages to an even number. 
-Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in convergence. 
-The solver would tend to ensure a continuous trajectory, by satisfying the interior stages, at the expense of the terminal stage. 
-Some additional detail has been provided that mentions our choice of the number of stages and has been copied below.
-
-\begin{correction}
-Based on experimentation, we use four stages in our multiple shooting method.
-This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
-\end{correction}
-
-\item 
-    \begin{itshape}
-Section 6.1 - I don't think you need to explain how you created the three-body orbits.
-This can be found in many references - You can just cite a reference and give the orbit specifications.
-(Starting with line 49, page 18, up to line 21 page 19.)
-\end{itshape}
-
-Thank you for the comment.
-The discussion of generating periodic orbits in the three body problem has been removed.
-It is replaced with a small introduction mentioning the differential correction process and a citation to a relevant resource. 
-The revised paragraph is copied below.
-
-\begin{correction}
-The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
-The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
-The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
-\Cref{fig:moon_orbit} shows that the target set remains in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
-This type of orbit would be useful for a variety of mission scenarios.
-The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
-The initial set is a planar periodic orbit about \( L_1\), which is generated using the process of differential correction of a linear approximation~\cite{koon2011}.
-\end{correction}
-
-\item
-    \begin{itshape}
-"We define a maximum magnitude of the thrust as umax = 0.75 and assume we can point the thrust in any direction within the plane." And  "we again assume a upper bound on the thrust magnitude of umax = 0.75."  What was the reason behind this?
-Isn't this a pretty large value of dimensional acceleration for your problem of interest?
-Even if it's chosen for demonstrative purposes, there should be some sort of explanation as to why.
-In the context of what electric propulsion systems can deliver, it could be arguable that this value isn't really "low-thrust" but "finite-thrust".
-\end{itshape}
-
-The maximum thrust magnitude was chosen to be approximately \SI{1}{\newton} for an assumed \SI{500}{\kilo\gram} spacecraft. 
-While larger than the NEXT thruster mentioned earlier, a cluster of similar engines could provide the required thrust (assuming adequate available power). 
-Additional detail has been included in the manuscript describing this thrust limit in dimensional units. 
-\begin{correction}
-Next, we determine the reachability set with the addition of a low thrust control input.
-We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
-This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
-We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
-This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
-Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
-Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
-The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
-This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit. 
-\end{correction}
-
-\item 
-    \begin{itshape}
-Is this method exceptionally difficult to adapt to create a numerical example from LEO or GTO?
-These would be the most pragmatically interesting departure orbits to the mission design community.
-Section 6.2 presents an example from GEO, but it is unlikely that GEO represents a "typical" parking orbit, since there is no reason to insert a spacecraft into GEO before it transfers to the destination orbit.
-This should be called out in the conclusions or future work.
-\end{itshape}
-
-This method is not difficult to adapt for other scenarios.
-However, there is a certain amount of experimentation to ensure that trajectories converge. 
-In addition, the details in this paper were designed in the context of the planar three body problem and as a result and extension to the nonplanar GTO is non trivial.
-The examples presented are primarily aimed at introducing and demonstrating the feasibility of this approach, rather than designing operational mission profiles for spacecraft.
-However, this is a valuable comment and a goal of future research. 
-Additional detail has been included in the conclusion discussing this topic.
-The modified conclusion is copied below.
-
-\begin{correction}
-In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
-The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
-The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
-Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.
-Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
-Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
-However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
-Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
-
-There is additional research to extend these results to more general transfer scenarios in future work.
-The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
-In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
-Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
-This will result in a more complicated optimal control problem and is a focus of future research.
-Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
-The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
-\end{correction}
-
-\item 
-    \begin{itshape}
-The authors should discuss some statistics of the convergence of their multiple shooting TPBVP - number of iterations, compute time, etc.
-\end{itshape}
-
-Thank you for the comment, however it seems to be rare for the research community to give detailed description of the numerical optimization method.
-Most of the literature tends to not spend time discussing the low level details of the computational procedure, e.g.~\cite{grebow2011,stuart2010}.
-In addition, the number of iterations, computation time and even the eventual cost function value are highly dependent on the specific system and more specifically the implementation of the algorithm in software.
-There are a wide variety of works which are focused more exclusively on the computational implementation of various optimal control techniques which do spend significant time on the computational characteristics.
-However, our manuscript is primarily focused on demonstrating the feasibility of using reachability sets to design transfer trajectories.
-As a result, there is not much benefit to detailed statistics of the implementation of this methodology, and we have omitted it from the manuscript.
-However, we have added some rough detail on the computation time and computer hardware used in the numerical optimization.
-This additional detail is copied below.
-
-\begin{correction}
-Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
-Each solution on the reachability set is computed in approximately \SI{2}{\minute} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
-\end{correction}
-
-\item 
-    \begin{itshape}
-Can any intuition be gained from reporting the optimal converged cost function values in section 6.1 and 6.2?
-One could at least compare them to compare the reachability in the two different examples.
-\end{itshape}
-
-The cost function only parameterizes the distance between the controlled and uncontrolled terminal states on the \Poincare section. 
-In addition, each state which lies on the reachable set would typically have a different distance and therefore a different converged cost function value. 
-Furthermore, it would be difficult to characterize a single parameter which characterizes the reachable set, in the form of the cost function, for all states of the reachable set. 
-Also, the cost function is highly dependent on the transit time of trajectory, i.e. a longer transit time corresponds to a larger reachable set and a larger magnitude of the cost function.
-However, this is an interesting suggestions for future research.
-For example, the cost function gives some indication of area/volume of the reachable set. 
-It may be possible to use this value as another way to characterize the subspace on the \Poincare section and is an intriguing avenue of future research.
-
-\item
-    \begin{itshape}
-"Low energy transfers from the Earth to the Lagrange points are necessary for future missions."  One must be careful not to confuse these with traditional "low-energy transfers" to the moon, which use Sun third-body perturbations to transfer into lunar orbit.
-These are what many mission designers define as "low-energy" lunar transfers.
-This is actually a four-body problem (Sun, Earth, moon, spacecraft), and represents additional delta-V savings (as was done with NASA's GRAIL mission.)
-
-Example:
-
-Parker, J. and Anderson, R., "Low-Energy Lunar Trajectory Design," DEEP SPACE COMMUNICATIONS AND NAVIGATION SERIES, Jet Propulsion Laboratory Pasadena, California, July 2013.
-
-Online Copy: https://descanso.jpl.nasa.gov/monograph/series12/LunarTraj--Overall.pdf
-\end{itshape}
-
-Thank you for the comment.
-This sentence has been modified to make it more clear that we're discussing low thrust propulsion rather than low-energy transfers.
-The modified sentence is copied below.
-
-\begin{correction}
-Transfers from the Earth to the Lagrange points, through the use of low-thrust electric propulsion, offer an additional and potentially shorter time of flight in comparison to the low-energy transfers which utilize solar perturbations.
-\end{correction}
-
-\item 
-    \begin{itshape}
-For what it's worth, exploiting sun perturbations with low-thrust in the Earth-moon RTBP would be highly beneficial and represent a worthy avenue of future research (more on this below.)
-\end{itshape}
-
-Thank you for the comment.
-This is a very interesting suggestion and something that we will take into consideration for future research.
-
-\item 
-    \begin{itshape}
-The conclusions of the paper should acknowledge other potentially limiting factors of the presented analysis
-
--Omits advantage of exploiting fourth body perturbation (Sun in Earth/moon RTBP)
-
--Transfers are 2D only, method is more challenging to adapt to 3D
-
--Presented control magnitudes are dimensionally large, and more work is needed to determine feasibility of approach for lower accelerations
-
--Thrust and mass are completely decoupled from the analysis, which is limiting in terms of finding suitable hardware that can deliver the controls needed, as currently presented.
-\end{itshape}
-
-Thank you for the comment.
-The conclusion has been expanded with some discussion of the drawbacks of this research and avenues of future work.
-The modified conclusion is copied below.
-
-\begin{correction}
-In this paper, an optimal transfer process which combines concepts of reachability and \Poincare section is used to generate transfer between planar periodic orbits in the three-body problem.
-The \Poincare section allows for trajectory design on a lower dimensional phase space and simplifies the process.
-The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
-However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
-Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
-
-There is additional research to extend these results to more general transfer scenarios in future work.
-The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
-In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
-Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
-This will result in a more complicated optimal control problem and is a focus of future research.
-Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
-The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
-\end{correction}
-\end{enumerate}
-
-\section*{Reviewer 2}
-
-\begin{itshape}
-Reviewer \#2:
-
-PAPER SUMMARY:
-The contribution of the authors is the synthesis of reachability theory with traditional low-energy mission design approaches to ease the study of the low-thrust mission design space in the CR3BP.
-A crucial point is that reachability analysis in non-integrable systems is typically limited by the dimensionality of the control parameterization, which can be extremely large for continuous-thrust problems, depending on the time discretization used.
-The authors numerically probe the boundaries of the reachable space by optimizing a distance metric on the Poincare section, avoiding the extreme cost of unmotivated exploration.
-Numerical experiments show how the method could be applied in two CR3BP orbit transfer scenarios.
-For a multi-rev solution, the method is applied once per rev and the reachable end state with minimal distance to the target manifold is used as the basis for the next rev.
-
-ASSESSMENT:
-
-The central premise of the paper - using optimization of Poincare distance as a guiding motivation to make low-thrust reachability analysis more tractable - is interesting, useful, and to my knowledge reasonably novel.
-Furthermore, the quality of the writing is good (at local scales) and there is an appropriate degree of literature review.
-However, there are two major problems with this paper.
-\end{itshape}
-\begin{itemize}
-
-\item 
-    \begin{itshape}
-First, in terms of composition, large portions of the paper are spent on lengthy development of foundational work that is well-known within the astrodynamics community and available in textbooks. 
-\end{itshape}
-
-    Thank you for the comment, and we agree that there is significant repetition of material previously covered in other sources.
-    The manuscript has been reduced and reorganized to enhance the clarity and clearly separate previously material and the contributions of this paper.
-    The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
-    The second section has been retitled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
-The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, followed by conclusions.
-The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.
-
-\item 
-    \begin{itshape}
-        Secondly, the contributions and findings of the authors feel highly preliminary and do not seem to evaluate their method at significant depth.
-    \end{itshape}
-
-    Thank you for the comment. 
-    The numerical examples are meant to demonstrate the use of this approach and the feasibility of using the reachability set on a \Poincare section to determine transfer trajectories.
-    The examples give two demonstrations of finding a transfer trajectory in the PCRTBP and demonstrate combining several iterations of the reachability set for a transfer.
-
-\item
-    \begin{itshape}
-While there are the beginnings of a good article within this manuscript, it is not sufficiently close to publishable form to be accepted.
-I recommend that the paper be rejected and that the authors push their investigation a bit further.
-The eventual publication should focus less on background and should provide more insight into the fundamental properties of their method, using more rigorously constructed and analyzed examples.
-\end{itshape}
-
-The authors believe that the given two nontrivial numerical examples illustrate the efficacy of the proposed approach in designing low-thrust maneuvers through reachability set analysis on the \Poincare section.
-Furthermore, the manuscript has been completely reorganized to emphasize the contributions and clearly separate the background material. 
-In particular,~\cref{sec:simulation} includes additional paragraphs and figures for the geostationary to periodic orbit transfer.
-These additional figures demonstrate the ability to chain together multiple iterations of the reachability set computation to generate a nontrivial transfer trajectory.
-Combined together, the numerical examples demonstrate the ability to depart from the vicinity of the Earth to the vicinity of the Moon using our proposed method of coupling low-thrust maneuvers with reachability set analysis.
-
-\item 
-    \begin{itshape}
-I can sympathize that there was quite a long road to get to the point of understanding this topic and generating results in it.
-I hope that the authors will not find this verdict too discouraging, as I believe that stronger results are well within their reach.
-\end{itshape}
-\end{itemize}
-
-SPECIFIC FEEDBACK:
-\begin{enumerate}
-    \item
-    \begin{itshape}
-*** Bloat
-In the introduction, the exposition of the problem could be considerably condensed, given the venue of publication.
-Additionally, the somewhat sprawling discussion jumps back and forth, and many key points seem to be repeated more than necessary.
-It needs to be organized into titled subsections with specific scope.
-\end{itshape}
-
-Thank you for the comment.
-The introduction is reduced and many of the paragraphs have been reworded and combined. 
-In addition, the extended discussion of the benefits of variational integrators is removed from the introduction.
-The mathematical background has been condensed into a single section with separate subsections which discuss some of the mathematical details.
-The background section is focused on reviewing the relevant material rather than reintroducing topics.
-In the interest of space the reorganized section is omitted in this response. 
-In short, the manuscript is completely reorganized and all of the background materials are summarized in~\cref{sec:background}.
-
-\item
-    \begin{itshape}
-The development of the CR3BP, invariant manifolds and Poincare maps, all go on a bit too long.
-These should be refreshers to the audience that expose the specific equations and concepts that you'll be leveraging, not full-on introductions to well-established areas.
-Spending several pages on variational principles is way too much.
-\end{itshape}
-
-Thank you for the comment.
-The mathematical background has been reduced and condensed into a single section with multiple subsections.
-The discussion of the PCRTBP, invariant manifolds and \Poincare maps are combined into a single subsection.
-Furthermore, the variational integrator discussion has been reduced and the background material on the derivation of variational mechanics has been omitted.
-Instead, only the specific discussion of the development of the variational integrator is included.
-
-\item
-    \begin{itshape}
-Section 4 ends on page 14 with essentially a thesis statement ("In this paper, we...").
-Strive to push the boundary between background and new developments to a much earlier point in the paper; it shouldn't happen so late that you need to remind the audience what the paper is for.
-\end{itshape}
-
-Thank you for the comment. 
-The introductory detail has been reduced and combined into a single section. 
-As a result, the third section begins with the new contributions which describe the optimal control formulation of the reachability set.
-
-\item 
-    \begin{itshape}
-A missing relevant reference: "On target for Venus a set oriented computation of energy efficient low thrust trajectories" (Dellnitz, Junge, Post, and Thiere, 2006)
-\end{itshape}
-
-Thank you for the comment.
-The reference has been noted in the discussion of reachability sets.
-\begin{correction}
-More recently, reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a,dellnitz2006}.
-\end{correction}
-
-\item
-    \begin{itshape}
-*** Variational Integrators
-Many of the specific claims regarding the use of a variational integrator are unsubstantiated and, I believe, incorrect.
-Why is energy drift "even more detrimental" with regards to "the long term effects of low-thrust"?
-I would need to be convinced that a non-geometric integrator's energy drift relative to a constant energy baseline is significantly different from its drift relative to a gradually shifting baseline corresponding to low thrust.
-The specific claims about the numerical stability of derivatives also don't seem related to energy preservation.
-
-\end{itshape}
-
-The main contribution is systematic design of low-thrust orbital maneuvers through reachability set analysis on the \Poincare section.
-The authors decided to utilize the variational integrator as it is more computationally efficient in simulating a complex dynamic system accurately.
-
-More explicitly, while the structural stability of geometric numerical integrators is often illustrated by the conserved quantities such as energy or momentum, it has been shown that variational integrators are also effective in computing the rate of energy change for non-conservative systems accurately~\cite{marsden2001}.
-This is particularly useful for low-thrust orbital maneuvers where the energy dissipate or accumulates slowly.
-Furthermore, it has been shown that the structure-preserving properties of geometric numerical integrators are also useful to numerical optimizations for improved computational efficiency and stability.
-For example, see the following reference.
-
-\begin{itemize}
-\item[-] Ober-Blöbaum, Sina, Oliver Junge, and Jerrold E. Marsden. "Discrete mechanics and optimal control: an analysis." ESAIM: Control, Optimisation and Calculus of Variations 17.2 (2011): 322-352.
-\end{itemize}
-
-In the revised manuscript, all of the comments for the variational integrator are re-examined and reorganized to Section 2.2, as also discussed in the response to the first reviewer.
-
-\item
-    \begin{itshape}
-Apart from these claims, I question the need for a variational integrator in the first place.
-The example in section 3.4 is not convincing because there is no apples-to-apples basis for the comparison; a variable-step integrator with unspecified tolerance is compared to a fixed-step integrator with an arbitrary step size.
-Integrator comparisons should consider error in proportion to compute cost, which can be tuned by step size or tolerance.
-Also, fixed step seems inappropriate for a path through diverse dynamic regimes, and it's not clear that the problem scope calls specifically for energy preservation (especially since the idea is to use low thrust to mitigate some time cost of exploiting the problem structure).
-\end{itshape}
-
-In the revised manuscript, the computational properties of the variational integrator is explicitly compared with a general purpose integrator in~\Cref{fig:compare_energy}.
-According to the reviewers suggestion, the error in energy conservation is illustrated for several computational costs.
-It is shown that the variational integrator exhibits lower conservation error  against a Runge-Kutta method for the similar level of CPU time.
-
-With regards to the comments on the adaptive step size, the variational integrator can be generalized according to the multisymplectic framework, as in the following reference. 
-\begin{itemize}
-\item[-] Lew, Adrian, et al. "Asynchronous variational integrators." Archive for Rational Mechanics and Analysis 167.2 (2003): 85-146.
-\end{itemize}
-However, the above formulation for structure-preserving, adaptive time-step variational integrators is beyond the scope of this paper, and we use a fixed time step throughout the presented numerical results.
-
-More importantly, it has been shown variational integrators are also effective in computing the rate of energy change for non-conservative systems accurately~\cite{marsden2001}, even for the fixed-step sizes.
-
-\item
-    \begin{itshape}
-*** Figures
-Altogether, the figures need more descriptive captions.  
-\end{itshape}
-
-Thank you for the comment, and we agree that the figures will benefit from more descriptive captions.
-We've added additional descriptions for all of the figures.
-The figure captions are shown below, with the figures reduced in size to save some space.
-
-\begin{correction}
-\begin{figure}[H]
-        \centering
-        \includegraphics[width=0.2\textwidth]{energy_contours}
-        \caption{Contour plot of Jacobi integral: Zero velocity curves of constant Jacobi integral.
-A particle with fixed energy level cannot cross the contour lines and is therefore limited to specific regions of the phase space.
-}
-\end{figure}
-
-\begin{figure}[H]
-        \centering
-        \begin{scaletikzpicturetowidth}{0.2\textwidth}
-    \begin{tikzpicture}[tdplot_main_coords,
-      poincare/.style={opacity=.2,very thick,fill=blue},
-      orbit/.style={very thick,black},
-      orbit hidden/.style={very thick,dashed},
-      grid/.style={very thin,gray!50},
-      axis/.style={->,blue,thick}, scale=\tikzscale]
-
-    % nodes for the poincare section
-    \node[label=above:\(\Sigma\)] (upper_right) at (0,5,5) {};
-    \node[] (upper_left) at (0,1,5) {};
-    \node[] (lower_left) at (0,1,0) {};
-    \node[] (lower_right) at (0,5,0) {};
-
-    % draw poincare section
-    \draw[poincare] (upper_right.center) -- (upper_left.center) -- (lower_left.center) -- (lower_right.center) -- (upper_right.center);
-
-    % draw a periodic orbit
-    \coordinate (center) at (0,0,2);
-    \node[below right] (x0) at (0,2,2) {\(\vecbf{q}_0\)};
-    \filldraw (0,2,2) circle (3pt);
-
-    \node[below right] (x1) at (0,3,2) {\(\vecbf{q}_1\)};
-    \filldraw (0,3,2) circle (3pt);
-
-    \tdplotdrawarc[orbit hidden]{(center)}{2}{90}{190}{}{};
-    \tdplotdrawarc[orbit,<-]{(center)}{2}{-170}{90}{}{};
-
-    \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{199}{}{};
-    \tdplotdrawarc[orbit,<-]{(center)}{3}{-161}{90}{}{};
-
-        \end{tikzpicture}
-    \end{scaletikzpicturetowidth}
-    \caption{Diagram of the \Poincare map: Periodic orbits will appear as fixed points on the \Poincare section \( \Sigma \).
-Stability of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.}
-\end{figure}
-
-\begin{figure}[H]
-        \centering 
-        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparison: A stable trajectory is used to test the variational and conventional integrators. 
-        The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
-        [0.5\textwidth]{\includegraphics[width=0.2\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
-        \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}}
-        [0.5\textwidth]{\includegraphics[width=0.2\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
-\end{figure}
-
-\begin{figure}[H]
-        \centering
-%       \includegraphics[width=0.5\textwidth]{reachability_set_hand}
-        \begin{scaletikzpicturetowidth}{0.2\textwidth}
-        \begin{tikzpicture}[tdplot_main_coords,
-          poincare/.style={opacity=.2,very thick,fill=blue},
-          orbit/.style={very thick,black},
-          orbit hidden/.style={very thick,dashed},
-          grid/.style={very thin,black},
-          axis/.style={->,blue,thick},
-          reachability/.style={thick,blue},scale=\tikzscale]
-
-        % nodes for the poincare section
-        \node[label=above:\(\Sigma\)] (upper_right) at (0,5,5) {};
-        \node[] (upper_left) at (0,1,5) {};
-        \node[] (lower_left) at (0,1,0) {};
-        \node[] (lower_right) at (0,5,0) {};
-
-        % draw poincare section
-        \draw[poincare] (upper_right.center) -- (upper_left.center) -- (lower_left.center) -- (lower_right.center) -- (upper_right.center);
-
-        % draw a periodic orbit
-        \coordinate (center) at (0,0,2);
-        \node[label=below:\(\vecbf{x}_n\)] (x0) at (0,3,2) {};
-        % \node[label=below:\(\vecbf{x}_n\)] at (x0) {};
-        \filldraw (x0) circle (3pt);
-
-        % \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{200}{}{};
-        \tdplotdrawarc[orbit,<-]{(center)}{3}{-160}{90}{}{};
-
-        % draw reachability set on the poincare section
-        \coordinate (reach) at (0,4.5,2);
-        \tdplotsetthetaplanecoords{90}
-
-        \draw[tdplot_rotated_coords,grid] (x0) -- (reach);
-        \draw[tdplot_rotated_coords,grid] (x0) -- ++(-45:1.5);
-
-        \tdplotdrawarc[tdplot_rotated_coords,grid]{(x0)}{0.5}{-45}{90}{above}{\(\theta_d\)};
-
-        % draw terminal state on reachability set
-        \node[tdplot_rotated_coords,label=above:\(\vecbf{x}_f\)] (xf) at ($ (x0)+(-45:1.5) $) {};
-        \filldraw (xf) circle (3pt);
-
-        \node[tdplot_rotated_coords,label=below:\(J\)] at (xf) {};
-
-        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
-        % place
-    \end{tikzpicture}
-        \end{scaletikzpicturetowidth}
-        \caption{Reachability set on a \Poincare section: Pictorial representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
-            The terminal state, \( x_n\), is the intersection without any control input. 
-            Adding a control input allows for the terminal state, \( x_f \), to be displaced by some distance/cost \( J \) as measured on the section.
-            We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
-            Computation of the maximum distance, or reachability, for a variety angles gives a discrete approximation of the reachability set.}
-\end{figure}
-
-\begin{figure}[H]
-        \centering
-        \begin{scaletikzpicturetowidth}{0.7\textwidth}
-                \begin{tikzpicture}[scale=\tikzscale]
-                    % \draw[help lines] (0,0) grid (20,5) ; %grid
-
-                    \node[label=west:\(\vecbf{x}\)] (x0m) at (0,5) {};
-                    \node[] (x1m) at (5,5) {};
-                    \node[] (x2m) at (10,5) {};
-                    \node[] (x3m) at (15,5) {};
-                    \node[] (xnm) at (20,5) {};
-
-                    \node[label=below:\(k_0\)] (t0) at (0,0) {};
-                    \node[label=below:\(k_1\)] (t1) at (5,0) {};
-                    \node[label=below:\(k_2\)] (t2) at (10,0) {};
-                    \node[label=below:\(k_{n-1}\)] (t3) at (15,0) {};
-                    \node[label=below:\(k_n\)] (tn) at (20,0) {};
-
-                    % nodes for each subsegment
-                    \node[label=left:\(\vecbf{x}_0\)] (x0) at (0,1) {\pgfuseplotmark{*}};
-                    \node[label=below left:\(\vecbf{\lambda}_0\)] at (x0) {};
-
-                    \node[label=right:\(\vecbf{x}_1^{-}\)] (x1minus) at (5,2) {\pgfuseplotmark{*}};
-                    \node[label=below right:\(\vecbf{\lambda}_1^{-}\)] at (x1minus) {};
-
-                    \node[label=above left:\(\vecbf{x}_1^{+}\)] (x1plus) at (5,3) {\pgfuseplotmark{o}};
-                    \node[label=left:\(\vecbf{\lambda}_1^{+}\)] at (x1plus) {};
-
-                    \node[label=right:\(\vecbf{x}_2^{-}\)] (x2minus) at (10,2) {\pgfuseplotmark{*}};
-                    \node[label=below right:\(\vecbf{\lambda}_2^{-}\)] at (x2minus) {};
-
-                    \node[label=above left:\(\vecbf{x}_2^{+}\)] (x2plus) at (10,3) {\pgfuseplotmark{o}};
-                    \node[label=left:\(\vecbf{\lambda}_2^{+}\)] at (x2plus) {};
-
-                    \node[label=right:\(\vecbf{x}_{n-1}^{-}\)] (x3minus) at (15,2) {\pgfuseplotmark{*}};
-                    \node[label=below right:\(\vecbf{\lambda}_{n-1}^{-}\)] at (x3minus) {};
-
-                    \node[label=above left:\(\vecbf{x}_{n-1}^{+}\)] (x3plus) at (15,3) {\pgfuseplotmark{o}};
-                    \node[label=left:\(\vecbf{\lambda}_{n-1}^{+}\)] at (x3plus) {};
-
-                    \node[label=right:\(\vecbf{x}_n\)] (xnminus) at (20,3) {\pgfuseplotmark{*}};
-                    \node[label=below right:\(\vecbf{\lambda}_n\)] at (xnminus) {};
-                    % draw axes
-                    \draw [<->,thick] (tn.center) -- (t0.center) -- (x0m.center);
-
-                    % draw segement dividers
-                    \draw [dashed,thick] (t1.center) -- (x1m.center);
-                    \draw [dashed,thick] (t2.center) -- (x2m.center);
-                    \draw [dashed,thick] (t3.center) -- (x3m.center);
-                    \draw [dashed,thick] (tn.center) -- (xnm.center);
-
-                    % draw markers at each subsegment
-
-                    % draw the subsegments
-                    \draw [->] (x0.center) to [bend left=30] (x1minus.center);
-                    \draw [->] (x1plus.center) to [bend left=30] (x2minus.center);
-                    % dotted lines here
-                    \draw[decorate sep={1mm}{5mm},fill] (x2plus.east) to [bend left=10] (x3minus.west);
-
-                    \draw [->] (x3plus.center) to [bend left=30] (xnminus.center);
-                \end{tikzpicture}
-        \end{scaletikzpicturetowidth}
-        \caption{Schematic diagram of the multiple shooting method:
-        The complete trajectory is split into a number of sub-segments, and additional interior constraints are included to ensure state and costate continuity.
-    Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.}
-\end{figure}
-
-\begin{figure}[H]
-   \centering
-   \includegraphics[width=0.2\textwidth]{moon_orbit} % requires the graphicx package
-   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstrating the initial and target orbit.
-   Without the low-thrust propulsion, the spacecraft is constrained to the initial periodic orbit. 
-   We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
-\end{figure}
-
-\begin{figure}[H]
-        \centering 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=0.5\textwidth]{manifold_trajectory} 
-                \caption{Position space view of invariant manifold} 
-        \end{subfigure}~ %
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=0.5\textwidth]{manifold_poincare} 
-                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} 
-        \end{subfigure} 
-        \caption{Invariant manifold transfer: An example transfer using the invariant manifolds is shown in both the position and \Poincare spaces.
-        The control free transfer from the initial periodic orbit to the target orbit result in a long time of flight. 
-    In addition, the manifold only intersects the target orbit on the descending or far side of the moon.}
-\end{figure}
-\begin{figure}[H]
-        \centering 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=0.5\textwidth]{reach_trajectory} 
-                \caption{Position space view of reachability set}  
-        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=0.5\textwidth]{poincare_compare} 
-                \caption{\Poincare section view of reachability set}  
-        \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
-
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=0.5\textwidth]{reach_transfer} 
-                \caption{Transfer trajectory selected from reachability set viewed in the position space}  
-        \end{subfigure}~ 
-        \begin{subfigure}[htbp]{0.5\textwidth} 
-                \includegraphics[width=0.5\textwidth]{control_input_l1} 
-                \caption{Control input for the selected transfer trajectory}  
-        \end{subfigure}~
-        \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
-        The reachability set is shown in the upper two figures in both the position and \Poincare space.
-    From this reachability set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
-The optimal control to achieve this transfer is shown in the lower right figure.} 
-\end{figure}
-
-\begin{figure}[H]
-   \centering
-   \includegraphics[width=0.2\textwidth]{initial_final} % requires the graphicx package
-   \caption{Geostationary to \( L_1 \) transfer: Representation of the initial and target orbits for the reachability transfer. 
-   Vehicle is assumed to begin on the geostationary orbit about \( m_1\) and will transfer to the periodic orbit about \( L_1\)}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_trajectory_zoom.pdf} 
-        \caption{Stage 1 Trajectory} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_poincare.pdf} 
-        \caption{Stage 1 \Poincare~section } 
-    \end{subfigure}
-
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_trajectory_zoom.pdf} 
-        \caption{Stage 2 Trajectory~} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_poincare.pdf} 
-        \caption{Stage 2 \Poincare~section } 
-    \end{subfigure}
-
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_trajectory_zoom.pdf} 
-        \caption{Stage 3 Trajectory} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_poincare.pdf} 
-        \caption{Stage 3 \Poincare~section } 
-    \end{subfigure}
-
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
-        \caption{Stage 4 Trajectory~} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage4_poincare.pdf} 
-        \caption{Stage 4 \Poincare~section } 
-    \end{subfigure}   
-    \caption{Stage 1-4 reachability sets: The first four reachability sets visualized in both the position (left) and \Poincare space (right).
-        The minimum trajectories from the preceding stages are shown in red, while the next stage is shown in blue.
-    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
-        \caption{Stage 5 Trajectory~} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_poincare.pdf} 
-        \caption{Stage 5 \Poincare~section } 
-    \end{subfigure}
-
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_trajectory_zoom.pdf} 
-        \caption{Stage 6 Trajectory} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_poincare.pdf} 
-        \caption{Stage 6 \Poincare~section } 
-    \end{subfigure}
-
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_trajectory_zoom.pdf} 
-        \caption{Stage 7 Trajectory} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_poincare.pdf} 
-        \caption{Stage 7 \Poincare~section } 
-    \end{subfigure}
-
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_trajectory_zoom.pdf} 
-        \caption{Stage 8 Trajectory} 
-    \end{subfigure}~
-    \begin{subfigure}[htbp]{0.2\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_poincare.pdf} 
-        \caption{Stage 8 \Poincare~section } 
-    \end{subfigure}   
-    \caption{Stage 5-8 reachability sets: The last four reachability sets visualized in both the position (left) and \Poincare space (right).
-        The minimum trajectories from the preceding stages are shown in red, while the next stage is shown in blue.
-    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
-\end{figure}
-\end{correction}
-
-\item
-    \begin{itshape}
-Figure 4 does not show what I think you intend it to show; it looks like two periodic circular orbits.
-\end{itshape}
-
-Thank you for the comment, however the figure is drawn to show a pictorial representation of a \Poincare section.
-The periodic orbits intersect the section at fixed points, as a result it becomes simple to visualize and determine these fixed solutions by searching for fixed points on the section
-Conversely, stable/unstable trajectories will also appear on the section with a distinctive signature, and by tracking successive intersections of the trajectory with the \Poincare section.
-
-In order to improve the clarity and reduce confusion, we have added some additional detail to the discussion of the \Poincare map to the manuscript.
-This addition has been copied below.
-
-\begin{correction}
-\Poincare maps are a useful tool in the analysis of the flow near periodic orbits in the three-body problem.
-We let \( \Sigma \) define a hypersurface of section chosen such that all trajectories in the vicinity of a state \( \vecbf{q} \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
-A \Poincare map, \( P(\vecbf{q}) = \phi(T;\vecbf{q}) \), maps the state of a trajectory from one intersection to the next.
-Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
-In~\cref{fig:poincare_map} we show two examples of periodic trajectories intersecting the \Poincare section. 
-Periodic solutions will appear as fixed points on the section, such as \( q_0, q_1 \) in~\cref{fig:poincare_map}, while stable or unstable trajectories become clearly visible by viewing successive intersections of the section.
-This allows for greater insight into the stability and dynamics of periodic solutions of a dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories.
-\end{correction}
-
-\item
-    \begin{itshape}
-Figure 6 is also very unclear.  Is phi supposed to be \texttt{theta\_d}?  Is the bounding curve specifically circular, or should it be a more general shape?
-\end{itshape}
-
-Thank you for the comment.
-The angle in the figure is meant to match the terminal constraint \( \theta_d\).
-In addition, the reachable set is only circular for ease of design in the figure.
-In reality the figure will be a general shape, rather than specifically circular.
-
-The manuscript has been modified to correct the figure and also incorporate some more discussion of the figure. 
-The additions are included below as well as the updated figure.
-\begin{correction}
-\Cref{fig:reachability_set} illustrates how, without any control input, trajectories will intersect with the \Poincare section at \( \vecbf{x}_n \). 
-However, the addition of low thrust propulsion allows the spacecraft to depart from the natural dynamics and intersect the \Poincare section at a different location.
-We use a cost function to define a distance metric on the \Poincare section from the control-free intersection to an intersection under the influence of the control input.
-Maximization of this distance along varying directions enables us to generate the largest reachability set under the bounded control input.
-In~\cref{fig:reachability_set} the reachable set is shown as a circular region on the \Poincare section.
-In practice, the reachable set will be of a general shape and also higher dimensional in the nonplanar case.
-\begin{figure}[H]
-        \centering
-%       \includegraphics[width=0.5\textwidth]{reachability_set_hand}
-        \begin{scaletikzpicturetowidth}{0.3\textwidth}
-        \begin{tikzpicture}[tdplot_main_coords,
-          poincare/.style={opacity=.2,very thick,fill=blue},
-          orbit/.style={very thick,black},
-          orbit hidden/.style={very thick,dashed},
-          grid/.style={very thin,black},
-          axis/.style={->,blue,thick},
-          reachability/.style={thick,blue},scale=\tikzscale]
-
-        % nodes for the poincare section
-        \node[label=above:\(\Sigma\)] (upper_right) at (0,5,5) {};
-        \node[] (upper_left) at (0,1,5) {};
-        \node[] (lower_left) at (0,1,0) {};
-        \node[] (lower_right) at (0,5,0) {};
-
-        % draw poincare section
-        \draw[poincare] (upper_right.center) -- (upper_left.center) -- (lower_left.center) -- (lower_right.center) -- (upper_right.center);
-
-        % draw a periodic orbit
-        \coordinate (center) at (0,0,2);
-        \node[label=below:\(\vecbf{x}_n\)] (x0) at (0,3,2) {};
-        % \node[label=below:\(\vecbf{x}_n\)] at (x0) {};
-        \filldraw (x0) circle (3pt);
-
-        % \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{200}{}{};
-        \tdplotdrawarc[orbit,<-]{(center)}{3}{-160}{90}{}{};
-
-        % draw reachability set on the poincare section
-        \coordinate (reach) at (0,4.5,2);
-        \tdplotsetthetaplanecoords{90}
-
-        \draw[tdplot_rotated_coords,grid] (x0) -- (reach);
-        \draw[tdplot_rotated_coords,grid] (x0) -- ++(-45:1.5);
-
-        \tdplotdrawarc[tdplot_rotated_coords,grid]{(x0)}{0.5}{-45}{90}{above}{\(\theta_d\)};
-
-        % draw terminal state on reachability set
-        \node[tdplot_rotated_coords,label=above:\(\vecbf{x}_f\)] (xf) at ($ (x0)+(-45:1.5) $) {};
-        \filldraw (xf) circle (3pt);
-
-        \node[tdplot_rotated_coords,label=below:\(J\)] at (xf) {};
-
-        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
-        % place
-    \end{tikzpicture}
-        \end{scaletikzpicturetowidth}
-        \caption{Reachability set on a \Poincare section: Pictorial representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
-        }
-\end{figure}
-\end{correction}
-
-\item
-    \begin{itshape}
-Consider plotting the intersections of manifolds with the Poincare section as continuous curves.  
-\end{itshape}
-
-Thank you for the comment.
-The manifold intersections with the \Poincare section are plotted as discrete points as the curves were generated in a discrete manner. 
-A series of trajectories were generated by perturbing the trajectory at various points along the periodic orbit. 
-As a result, we show that the intersection is a discrete approximation of the true intersection of the manifold with the section by using the discrete points rather than a continuous curve.
-
-\item
-    \begin{itshape}
-Also consider zooming/scaling as necessary to show the details you discuss.
-\end{itshape}
-
-Thank you for the comment. The figure have been revised to adequately show the desired information in the revised manuscript. 
-The figures are centered around the various intersections of the reachable set and the \Poincare section.
-Furthermore, in the case of the much larger geostationary transfer requires it becomes more difficult to zoom in on the fine details.
-However, a detailed view is presented which shows the various iterations of the reachability set computations.
-
-\item
-    \begin{itshape}
-*** Method/Examples
-I urge a revisiting of the claims with regards to optimality.
-The optimization method is being initialized with *something*, and stationary points are *local* extrema.
-More precise claims and assurances about solution quality (and therefore the validity of the computed reachable set boundary) are needed.
-\end{itshape}
-
-The presented numerical results satisfy the necessary conditions for optimality.
-As common in gradient-based numerical optimizations for non-convex systems, it guarantees local optimality.
-Applying various techniques to obtain the globally optimal solution is beyond the scope of this paper that is focused on systematic design of low-thrust orbital maneuvers thorough reachability set analysis on the \Poincare section.
-
-To clarify these, the following comment is added at the first paragraph of~\cref{sec:simulation}.
-
-\begin{correction}
-The numerical examples presented in this section satisfy necessary conditions for local optimality, and obtaining a globally optimal solution is considered beyond the scope of this paper.
-\end{correction}
-
-\item
-    \begin{itshape}
-How damaging is the assumption of a fixed travel time?  And how about the specification of a specific departure point on the initial orbit?
-\end{itshape}
-
-The fixed travel time is primarily used at this stage to reduce the complexity of the optimal control formulation. 
-Furthermore, the reachability set is defined in terms of a fixed terminal time.
-A variable terminal travel time would add an additional dimension, time, to the reachable space and add additional complexity to the reachability set visualization.
-The fixed travel time does not adversely affect the optimality of the computed trajectories on the reachable set, as the reachable set is defined in terms of a fixed travel time.
-
-The choice of starting from a fixed point on the initial periodic orbit is again mostly used for convenience. 
-In this case of the periodic orbit the departing trajectories are all along the unstable invariant manifold associated with the periodic orbit. 
-The uncontrolled trajectories associated with the invariant manifolds serve as the ``center'' of each reachability set computation.
-The low-thrust propulsion system is then used to enlarge the distance between the uncontrolled and controlled states. 
-Starting from a different location on the periodic orbit will not create a qualitatively different situation as the entire invariant manifold also intersects the \Poincare section.
-However, different locations along the periodic orbit may have vastly different travel times before intersecting the \Poincare section.
-This will require modification of the optimal control problem, with a larger terminal time, but will result in a similar behavior regardless of the location on the initial periodic orbit.
-
-\item
-    \begin{itshape}
-In Fig 10b, it is hard to see, but appears that the low-thrust-reachable set is actually *centered* on part of the target set.
-Is this related to the existence of the uncontrolled transfer opportunity?
-There must be some more clear way to show how the low thrust set expands upon the zero thrust set.
-Perhaps beginning with the time horizon of the zero-thrust opportunity, and successively lowering it, would help.
-\end{itshape}
-
-The reachability set is closely spaced in terms of both the position space as well as on the \Poincare section.
-This is due to the fact that there is a very low-thrust control input on the spacecraft.
-As a result, the trajectories associated with the reachability set are all closely spaced, but more widely separated from the terminal state without any control input.
-In the figure, there is a fixed point labeled \( x_n \) which is the terminal state without any control input, which will also lie on the periodic orbit.
-The reachable set, in red, is widely separated from the uncontrolled state due the effect of the low thrust propulsion, however the subsequent spread of the entire reachability set is much more constrained.
-The low thrust expands the reachability set from a fixed point to the red elliptical region. 
-This is also explained in the manuscript with a similar description of the figure.
-
-\item
-    \begin{itshape}
-Perhaps there is also a better way to show what the control is doing, e.g., in some alternative parametrization?
-This could help provide assurances that the solution is doing something meaningful and is not just stuck in a weird local minimum.
-\end{itshape}
-
-Thank you for the comment.
-The control plots show the planar components of the control thrust. 
-Furthermore, the scale has been modified to show the thrust in terms of Newtons for an assumed \SI{500}{\kilo\gram} satellite.
-The control plots are meant to demonstrate the control input required by the trajectories which lie on the maximum reachability set. 
-As a result, the control input is determined through optimization to maximize the reachability set in a particular direction on the \Poincare section.
-This is in contrast to typical optimal control applications where the control input is used directly in the trajectory design process.
-Our method seeks to approximate the reachability set, and from this set trajectories are chosen in a systematic method which drives the spacecraft towards a target.
-
-\item
-    \begin{itshape}
-In the second example, what do you lose by chaining together multiple reachability-based steps?
-Aren't you imposing that the trajectory routinely returns to the energy level of the Poincare section?
-How feasible would it be to conduct a single reachability analysis covering multiple revolutions?
-And can you use knowledge about directionality of movement on the Poincare section to limit the scope of your reachability search?
-\end{itshape}
-
-While the trajectories do return to the energy level of the \Poincare section, there is a transfer of ``energy'' between the states of the vehicle.
-It is in this manner that the reachability set is enlarged to eventually include the target state.
-
-Conducting a single reachability analysis becomes more complicated as the larger time of flight increases the sensitivity of the terminal states to variations in the initial costates. 
-Furthermore, the larger time of flight creates a much larger and more chaotic reachable set for the analysis.
-Through this work it was found that shorter trajectories tended to have more stable and consistent solutions.
-
-\item
-    \begin{itshape}
-In general, please devote more depth to probing the strengths, weaknesses, and fundamental properties of the method.  Considering extreme cases can help with this.
-\end{itshape}
-
-The authors believe that the given two nontrivial numerical examples illustrate the efficacy of the proposed approach in designing low-thrust maneuvers through reachability set analysis on the \Poincare section.
-The examples demonstrate the ability to generate transfer trajectories in the PCRTBP and can be easily extended to non-planar transfers~\cite{kulumani2016d}.
-The proposed approach provides for a systematic design process for low-thrust orbital transfers, in contrast to other methods which must rely entirely on numerical optimization to determine trajectories.
-This design process leverages the reachability set on a \Poincare section to greatly reduce the computational complexity of an unmotivated search of the reachable set.
-Through the use of the reachability set, our proposed approach provides a methodical process to determine transfers, using single or multiple iterations.
-
-There are several drawbacks of this approach.
-First, it relies heavily on the PCRTBP, which limits the applicability to planar trajectories.
-Furthermore, this reduces it's benefit to operational missions which are often non-planar, e.g. LEO to GTO missions.
-Second, the method does not consider the effect of variable mass on the system dynamics. 
-This is a focus of future work to incorporate the effect of variable mass on the optimal control problem.
-Finally, the variational integrator provides for a bounded energy behavior but does not ensure exact energy over the simulation. 
-As a result, there is still error in the variational integrator with respect to energy deviation.
-
-The manuscript has been completely reorganized to emphasize the contributions and clearly separate the background material. 
-In particular,~\cref{sec:simulation} includes additional paragraphs and figures for the geostationary to periodic orbit transfer.
-These additional figures demonstrate the ability to chain together multiple iterations of the reachability set computation to generate a nontrivial transfer trajectory.
-Combined together, the numerical examples demonstrate the ability to depart from the vicinity of the Earth to the vicinity of the Moon using our proposed method of coupling low-thrust maneuvers with reachability set analysis.
-
-In addition, the conclusion has been modified to better demonstrate the strengths and weaknesses of our approach as well as avenues for future research.
-The conclusion is copied below.
-
-\begin{correction}
-
-In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
-The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
-The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
-Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.
-Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
-Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
-However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
-Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
-
-There is additional research to extend these results to more general transfer scenarios in future work.
-The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
-In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems and future research is aimed at investigating smaller magnitude control inputs.
-Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
-This will result in a more complicated optimal control problem and is a focus of future research.
-Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
-The addition of attitude dynamics and realistic pointing constraints would also significantly improve the applicability of this work.
-\end{correction}
-
-\end{enumerate}
 % \bibliography{library}

### 2018-05-11 Response.tex
Remove previous version of reviisoin comments

@@
     \begin{correction}
-The indirect optimal control formulation direclty solves for the state and costate vectors.
+        The indirect optimal control formulation direclty solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
+        Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution and typically results in a much larger parameter set.
     \end{correction}

### 2018-05-11 manuscript.tex
Remove previous version of reviisoin comments

@@
 The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
-The indirect optimal control formulation direclty solves for the state and costate vectors.
-Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.
+The indirect optimal control formulation direclty solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
+Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution and typically results in a much larger parameter set.
 Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.

### 2018-05-11 Response.tex
Add table for statistics for both examples
Reviewer 1 Comment 2

@@
 
+        Some additional statistics have been included in the manuscript for both examples.
+        \begin{correction}
+            Convergence statistics associated for this transfer are shown in~\cref{tab:l1_transfer_stats}.
+            \begin{table}
+                \centering
+                \begin{tabular}{llr}  
+                    \toprule
+                    Metric    & Value \\
+                    \midrule
+                    \texttt{fsolve} objective      & \num{6.03e-15}      \\
+                    \texttt{fsolve} major iterations       & \num{9}      \\
+                    \texttt{fsolve} first order optimality & \num{1.88e-13} \\
+                    Optimal cost       & \num{2.09e-31}      \\
+                    Executation time & \SI{1.49}{\second}       \\
+                    \bottomrule
+
+                \end{tabular}
+                \caption{Convergence statistics for the periodic orbit transfer\label{tab:l1_transfer_stats}}
+            \end{table}
+
+        \end{correction}
+
+        And also additional statistics are provided for the geo-stationary transfer.
+        \begin{correction}
+            The optimization statistics for the final transfer are shown in~\cref{tab:geo_transfer}.
+            \begin{table}[h]
+                \centering
+                \begin{tabular}{llr}  
+                    \toprule
+                    Metric    & Value \\
+                    \midrule
+                    \texttt{fsolve} objective      & \num{1.42e-11}      \\
+                    \texttt{fsolve} major iterations       & \num{18}      \\
+                    \texttt{fsolve} first order optimality & \num{7.2e-11} \\
+                    Optimal cost       & \num{4.30e-25}      \\
+                    Executation time & \SI{2.62}{\second}       \\
+                    \bottomrule
+                \end{tabular}
+                \caption{Convergence statistics for the geostationary orbit transfer\label{tab:geo_transfer}}
+            \end{table}
+
+        \end{correction}
+
+        In addition, some additional detail was included to highlight the fact that exact results will vary based on the choice of nonlinear solver.
+
+        \begin{correction}
+            In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm with a convergence tolerance of \num{1e-5}.
+            Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
+            All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
+            The numerical results are specific to the choice of nonlinear solver, and different tolerances or software tools may result in slight changes.
+        \end{correction}
     \item \begin{itshape}

### 2018-05-11 manuscript.tex
Add table for statistics for both examples
Reviewer 1 Comment 2

@@
 The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
-In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
+In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm with a convergence tolerance of \num{1e-5}.
 Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
 All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
+The numerical results are specific to the choice of nonlinear solver, and different tolerances or software tools may result in slight changes.
 % tpbvp is solved using a shooting method (fsolve in matlab)
@@
 The spacecraft achieves the desired target while satisfying the bounded control input.
+Convergence statistics associated for this transfer are shown in~\cref{tab:l1_transfer_stats}.
+\begin{table}
+    \centering
+    \begin{tabular}{llr}  
+        \toprule
+        Metric    & Value \\
+        \midrule
+        \texttt{fsolve} objective      & \num{6.03e-15}      \\
+        \texttt{fsolve} major iterations       & \num{9}      \\
+        \texttt{fsolve} first order optimality & \num{1.88e-13} \\
+        Optimal cost       & \num{2.09e-31}      \\
+        Executation time & \SI{1.49}{\second}       \\
+        \bottomrule
+
+    \end{tabular}
+    \caption{Convergence statistics for the periodic orbit transfer\label{tab:l1_transfer_stats}}
+\end{table}
 
@@
 It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
-The periodic orbit is defined by the initial condition \(\begin{bmatrix} 0.8057 & 0 & 0 & 0.2982 \end{bmatrix} \).
+The periodic orbit is defined by the initial condition \(\begin{bmatrix} 0.8057 & 0 & 0 & 0.2982 \end{bmatrix}^T \).
 The initial and target orbits are illustrated in~\cref{fig:geo_transfer_target}.
@@
 \Cref{fig:stage1to4_reachability,fig:stage5to8_reachability} shows all of the intermediate reachability iterations to transfer between the initial orbit and the stable invariant manifold.
+The final reachable set intersects the stable manifold of the periodic orbit. 
+A final fixed terminal time and terminal state optimal transfer is finally used to arrive at the stable manifold.
+The optimization statistics for the final transfer are shown in~\cref{tab:geo_transfer}.
+\begin{table}[h]
+    \centering
+    \begin{tabular}{llr}  
+        \toprule
+        Metric    & Value \\
+        \midrule
+        \texttt{fsolve} objective      & \num{1.42e-11}      \\
+        \texttt{fsolve} major iterations       & \num{18}      \\
+        \texttt{fsolve} first order optimality & \num{7.2e-11} \\
+        Optimal cost       & \num{4.30e-25}      \\
+        Executation time & \SI{2.62}{\second}       \\
+        \bottomrule
+    \end{tabular}
+    \caption{Convergence statistics for the geostationary orbit transfer\label{tab:geo_transfer}}
+\end{table}
 The complete transfer trajectory, after eight iterations, is shown in~\Cref{fig:geo_transfer}. 

### 2018-05-11 manuscript.tex
Rewording the introduction

@@
 % utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
-With reduced development intervals and decreased launch costs, small satellites have gained increased attention as a cost effective means of scientific and technologic development. 
-Furthermore, these cubesats are easily launched as secondary payloads rather than requiring a dedicated launch.
-The merger of small satellites with miniature electric propulsion enables inexpensive and responsive missions requiring large changes in orbital energy or extended mission lifespan.
-Recent developments in miniature electric propulsion offer the potential for new research opportunities for small spacecraft~\cite{haque2013}.
-The advancements in miniature electric propulsion allows for greater flexibility in the deployment of cubesats as secondary payloads.
-Without the explicit requirement of a dedicated launch trajectory, cubesats are able to exploit a much wider range of possible mission designs~\cite{folta2015}.
-With the potential for more demanding missions, even greater importance is placed on the mission design to ensure that optimal trajectories satisfy mission requirements. 
-In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{folta2015,koon2011}.
-Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional patched-conic approach~\cite{ross2006,gomez2001}.
+Recent developments in miniature electric propulsion and small satellites now offer the potential for new research oppportunities~\cite{folta2015,haque2013}.
+The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}.
 Optimally combining the structure of the dynamic environment with low-thrust propulsion systems is vital for future mission success.
-
-% --------------Issues/Challenges of previous work
-
-% use of direct optimal control
-% chaotic systems and initial guesses for optimization
-% numerical integration accuracy
-% preserving dynamical structure
 
@@
 Furthermore, the variational integrator can allow for reduced computational effort without any loss of numerical stability or energy drift.
+
+% --------------Issues/Challenges of previous work
+
+% use of direct optimal control
+% chaotic systems and initial guesses for optimization
+% numerical integration accuracy
+% preserving dynamical structure
 

### 2018-05-11 manuscript.tex
Rewording the introduction

@@
 % utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
-Recent developments in miniature electric propulsion and small satellites now offer the potential for new research oppportunities~\cite{folta2015,haque2013}.
+Recent developments in miniature electric propulsion and small satellites now offer the potential of new research oppportunities~\cite{folta2015,haque2013}.
 The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}.

### 2018-05-13 manuscript.tex
Combining some paragraphs

--- 2018-05-11 manuscript.tex
@@
 The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}.
-Optimally combining the structure of the dynamic environment with low-thrust propulsion systems is vital for future mission success.
-
 There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}. 
-In spite of the long history in astrodynamics, a number of difficulties arise when implementing optimal control techniques.
-First, the optimal control problem is typically solved using direct methods, which approximate the continuous problem as a parameter optimization problem.
-Alternatively, indirect methods apply the calculus of variations to derive the necessary conditions for optimality.
-The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which generally rely on suboptimal representations of the control.
-Second, in order to implement any optimization method a sufficiently accurate ``initial guess'' is required.
-Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system.
-Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
-As a result, numerical methods are often required to investigate solutions to the three body problem.
+Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
 Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
-Therefore, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g.
-on the order of decades to centuries, which is typical of the three-body problem.
-Furthermore, the variational integrator can allow for reduced computational effort without any loss of numerical stability or energy drift.
+Variational integrators can allow for reduced computational effort without any loss of numerical stability or energy drift which exist in conventional integration schemes.
 

### 2018-05-13 manuscript.tex
Rearranging introduction

@@
 The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}.
-There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}. 
-Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
+There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011,koon2011,ross2006}. 
+Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
+The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.
+Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
 Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
@@
 % control-free trajectories in the three-body problem 
-There exists a rich structure in the three-body problem~\cite{koon2011,ross2006}.
-Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
-It has been shown that there exist multi-dimensional structures, termed invariant manifolds, of constant energy trajectories that span the state space. 
-These invariant manifolds, which are associated with the periodic solutions of the three-body dynamics, allow for the spacecraft to traverse vast expanses of the state space with zero energy change. 
-However, the results presented are highly case specific and difficult to generalize to arbitrary transfers.
-Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system. 
-The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.
-
 % objectives of this work
@@
 In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem.
-Utilizing the planar assumption greatly simplifies the problem and allows for the use of established methods in determining periodic orbits and \Poincare~sections.
 We utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
-The use of a simple distance metric on a \Poincare~section greatly reduces the dimensionality of the reachability analysis and provides an approximation of the reachability set.
-In addition, through the use of geometric integrators, we more accurately capture the effects of low thrust on the system dynamics in the numerical simulation over extended periods, which are typical of electric propulsion systems.
-With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low thrust propulsion systems.
-
-We define an optimal control problem to determine the reachability set on a \Poincare~section.
 The reachable set, defined as the set of all attainable states subject to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
 Defining the reachable set on a \Poincare section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
-While the \Poincare section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
-Instead of relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on the \Poincare section.
 Through the use of low-thrust control input, the reachable set on the \Poincare section is enlarged and enables a larger space of potential transfers.
 By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare section, we generate general transfer trajectories.
+With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low thrust propulsion systems.
 

### 2018-05-13 manuscript.tex
Reduce teh introduction

@@
 In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
-Previous results in the design of optimal transfers have relied on suboptimal direct optimization methods and conventional integration techniques.
 This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
-We derive and solve the Euler-Lagrange equations to generate the necessary conditions for an optimal solution.
-This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
-For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
+% This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
+% For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
 In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.

### 2018-05-13 Response.tex
Update introduction
Reviewer 2 Comment 1

@@
 
+The introduction has been reduced in accordance with the recommendations below.
+Several paragraphs have been removed or refactored to reduce repetition and extraneous detail.
+The updated introduction is included below and specific comments for each comment is also provided.
+
 \begin{itemize}
+
     \item
@@
         \end{itshape}
+
+        Several introductory paragraphs have been combined to reduce the extraneous detail.
+
     \item 
@@
         \end{itshape}
+
+        Several paragraphs have been combined or removed entirely. 
+        The transitions are also updated to reduce repetition.
+
     \item 
@@
         \end{itshape}
+
+        The paragraphs in the introduction have been reworked to reduce length, repetition and superfluous information.
+
     \item 
@@
         \end{itshape}
+
+        The introduction has been reduced and divided into a subsection.
+
     \item 
@@
         \end{itshape}
+
+        This comment has been removed.
+
     \item 
@@
         \end{itshape}
+
+        The sentences have been removed from the manuscript.
+        The updated introduction is included below.
+
+        \begin{correction}
+            \section{Introduction}
+
+            %%-------------- Motivation for low thrust orbital transfers
+
+            % enables long duration missions with minimized propellant usage since Isp is so high
+
+            % Enables different mission scenarios which are not possible through the use of conventional chemical propulsion (high thrust) systems
+            Designing spacecraft trajectories is a classic and ongoing topic of research.
+            There has been significant research into the design of orbital transfers for space vehicles.
+            Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
+            Electric propulsion systems offer a much greater specific impulse than chemical systems.
+            As a result, the greatly increased efficiency allows for greater payload mass or extended duration missions.
+            However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital maneuvers have a much longer time of flight.
+            In spite of this drawback, a wide variety of missions, such as communication and deep space probes, have utilized the unique benefits of low thrust electric propulsion to great effect~\cite{choueiri2009}.
+
+            % utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
+            Recent developments in miniature electric propulsion and small satellites now offer the potential for new research oppportunities~\cite{folta2015,haque2013}.
+            The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}.
+            There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011,koon2011,ross2006}. 
+            Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
+            The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.
+            Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
+            Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
+            Variational integrators can allow for reduced computational effort without any loss of numerical stability or energy drift which exist in conventional integration schemes.
+
+            % --------------Issues/Challenges of previous work
+
+            % use of direct optimal control
+            % chaotic systems and initial guesses for optimization
+            % numerical integration accuracy
+            % preserving dynamical structure
+
+            % control-free trajectories in the three-body problem 
+            % objectives of this work
+
+            % description of our work - our approach
+
+            % concise summary of contributions
+
+            %-------------------- OBJECTIVES OF THIS WORK-----------------------------%
+            % SYSTEMATIC DESIGN TO AVOID DIFFICULTIES IN CHOOSING A GOOD INITIAL GUESS
+            % CAPTURE LONG TERM EFFECTS OF LOW THRUST ACCURATELY IN NUMERICAL SIMULATION
+            \subsection{Objective and Contribution}
+            In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem.
+            We utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
+            The reachable set, defined as the set of all attainable states subject to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
+            Defining the reachable set on a \Poincare section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
+            Through the use of low-thrust control input, the reachable set on the \Poincare section is enlarged and enables a larger space of potential transfers.
+            By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare section, we generate general transfer trajectories.
+            With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low thrust propulsion systems.
+
+            In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
+            This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
+            % This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
+            % For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
+            In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
+            Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
+            We avoid the issues inherent in selecting a valid initial condition for optimization.
+            Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
+            We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.
+
+        \end{correction}
 \end{itemize}

### 2018-05-13 manuscript.tex
Update introduction
Reviewer 2 Comment 1

@@
 % CAPTURE LONG TERM EFFECTS OF LOW THRUST ACCURATELY IN NUMERICAL SIMULATION
+\subsection{Objective and Contribution}
 In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem.

### 2018-05-14 manuscript.tex
Add jacobi integral plot for first example

@@
         \end{subfigure}~
+
+        \begin{subfigure}[htbp]{0.5\textwidth} 
+            \includegraphics[width=\textwidth, keepaspectratio]{figures/l1_transfer/jacobi.pdf} 
+            \caption{Jacobi Energy over transfer \label{fig:jacobi_l1}} 
+        \end{subfigure}
+
         \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
@@
 The spacecraft achieves the desired target while satisfying the bounded control input.
+The Jacobi energy integral, computed using~\cref{eq:jacobi}, is shown in~\cref{fig:jacobi_l1}.
+\Cref{fig:jacobi_l1} shows that the vehicle begins with an energy level equal to the periodic orbit and arrives at the target orbit with the appropriate energy.
+The first half of the transfer is associated with and increase in energy as the control is used to transition towards the target orbit which is followed by a energy decrease to the target orbit.
+This roughly corresponds with the expected optimal solution of a bang-coast-bang type orbital transfer~\cite{kirk2012}.
 Convergence statistics associated for this transfer are shown in~\cref{tab:l1_transfer_stats}.

### 2018-05-14 Response.tex
Add some discussion about the  jacobi integral
Response to comments on energy levels and use of example 2

Reviewer 2 Comment 5 and 6

@@
 
+        Additional plots have been included in the manuscript which shows the evolution of the Jacobi energy integral over the transfer horizon.
+
+        The additional plot and some additional discussion for the first periodic orbit transfer is presented below.
+        \begin{correction}
+            \begin{figure} 
+                \centering 
+                \begin{subfigure}[htbp]{0.4\textwidth} 
+                    \includegraphics[width=\textwidth]{reach_trajectory} 
+                    \caption{Position space view of reachability set} \label{fig:reach_trajectory} 
+                \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
+                \begin{subfigure}[htbp]{0.4\textwidth} 
+                    \includegraphics[width=\textwidth]{poincare_compare} 
+                    \caption{\Poincare section view of reachability set} \label{fig:poincare_compare} 
+                \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
+
+                \begin{subfigure}[htbp]{0.4\textwidth} 
+                    \includegraphics[width=\textwidth]{reach_transfer} 
+                    \caption{Transfer trajectory selected from reachability set viewed in the position space} \label{fig:reach_transfer} 
+                \end{subfigure}~ 
+                \begin{subfigure}[htbp]{0.4\textwidth} 
+                    \includegraphics[width=\textwidth]{control_input_l1} 
+                    \caption{Control input for the selected transfer trajectory} \label{fig:control_l1} 
+                \end{subfigure}~
+
+                \begin{subfigure}[htbp]{0.4\textwidth} 
+                    \includegraphics[width=\textwidth, keepaspectratio]{figures/l1_transfer/jacobi.pdf} 
+                    \caption{Jacobi Energy over transfer \label{fig:jacobi_l1}} 
+                \end{subfigure}
+
+                \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
+                    The reachability set is shown in the upper two figures in both the position and \Poincare space.
+                    From this reachability set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
+                The optimal control to achieve this transfer is shown in the lower right figure.}
+                \label{fig:reachability_set_transfer} 
+            \end{figure}
+
+            The Jacobi energy integral, computed using~\cref{eq:jacobi}, is shown in~\cref{fig:jacobi_l1}.
+            \Cref{fig:jacobi_l1} shows that the vehicle begins with an energy level equal to the periodic orbit and arrives at the target orbit with the appropriate energy.
+            The first half of the transfer is associated with and increase in energy as the control is used to transition towards the target orbit which is followed by a energy decrease to the target orbit.
+            This roughly corresponds with the expected optimal solution of a bang-coast-bang type orbital transfer~\cite{kirk2012}.
+
+        \end{correction}
+        The additional plot for the geostationary orbit transfer is presented below.
+        \begin{correction}
+
+Each stage of the transfer serves to raise the energy level of the vehicle.
+After eight stages the reachability set intersects the stable manifold and a demonstrates that a transfer is achievable.
+The final optimal control drives the vehicle towards the target manifold with the appropriate energy level.
+\begin{figure} 
+        \centering 
+        \subcaptionbox{Transfer trajectory: Complete trajectory consisting of eight iterations (1-7 black and final in red) of the reachability sets with a final control-free coast (blue)  on the stable manifold (green).\label{fig:geo_transfer_full}}{\includegraphics[width=0.4\textwidth]{geo_transfer_full}}\hfill
+        \subcaptionbox{Detailed view of transfer: Centered at \(m_1\) the eight
+            iterations from the reachability analysis are shown in black.  The
+            final segment to the stable manifold is shown in red.  Each
+            iteration progressively approaches the stable manifold.  Once on
+        the stable manifold the vehicle can coast without thrust towards the
+    periodic orbit.\label{fig:geo_transfer_zoom}
+}{\includegraphics[width=0.4\textwidth]{geo_transfer_zoom}}
+
+\subcaptionbox{\Poincare section: Minimum distance states from each iteration of the reachability set analysis (red) with respect to the stable invariant manifold (green)\label{fig:geo_transfer_poincare}}{\includegraphics[width=0.4\textwidth]{poincare}}
+    \hfill
+    \subcaptionbox{Control Input: Combined control history for each stage of the reachability analysis. 
+    The control always remains within the maximum bound.\label{fig:control_input_geo}}{\includegraphics[width=0.4\textwidth]{figures/control_input_geo.pdf}}
+
+    \subcaptionbox{Jacobi Energy: Energy history throughout the transfer.
+    Each stage of the transfer serves to increase the energy towards the stable manifold. 
+The final reachability set intersects the stable manifold and enables a large final manuever to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.4\textwidth]{figures/geo_transfer/jacobi_energy.pdf}}
+
+        \caption{Geostationary to \( L_1 \) periodic orbit transfer: Complete transfer from the geostationary orbit to the stable manifold.\label{fig:geo_transfer}}
+\end{figure}
+        \end{correction}
     \item 
@@
         \end{itshape}
+
+        The combination of several reachability sets does not force the system to return to a prescribed energy level.
+        Each reachability set computation serves to maximize a distance metric, defined in a reduced phase space corresponding to the \Poincare section.
+        The cost function is computed as the distance between the terminal states of the controlled and uncontrolled trajectories.
+        The additional plots shown in both examples highlight the fact that the Jacobi energy is continuously changing as the control input is used to maneuver the vehicle towards the desired target states.
+
+        The transfer to the stable manifold is presented as an example of the combination of existing control free  transfer methodologies with low thrust control inputs.
+        This example demonstrates one possible use of low-thrust control to enable a larger region of feasibility for the use of invariant manifold transfers in the three body problem.
 

### 2018-05-14 manuscript.tex
Add some discussion about the  jacobi integral
Response to comments on energy levels and use of example 2

Reviewer 2 Comment 5 and 6

@@
 The spacecraft maintains a bounded control magnitude during the transfer to the stable manifold.
+\Cref{fig:jacobi_geo} shows the evolution of the Jacobi energy over the transfer.
+Each stage of the transfer serves to raise the energy level of the vehicle.
+After eight stages the reachability set intersects the stable manifold and a demonstrates that a transfer is achievable.
+The final optimal control drives the vehicle towards the target manifold with the appropriate energy level.
 \begin{figure} 
@@
     The control always remains within the maximum bound.\label{fig:control_input_geo}}{\includegraphics[width=0.45\textwidth]{figures/control_input_geo.pdf}}
+
+    \subcaptionbox{Jacobi Energy: Energy history throughout the transfer.
+    Each stage of the transfer serves to increase the energy towards the stable manifold. 
+The final reachability set intersects the stable manifold and enables a large final manuever to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.5\textwidth]{figures/geo_transfer/jacobi_energy.pdf}}
+
         \caption{Geostationary to \( L_1 \) periodic orbit transfer: Complete transfer from the geostationary orbit to the stable manifold.\label{fig:geo_transfer}}

### 2018-05-15 Response.tex
Modifying response letter formatting

@@
 Below I will try to sketch how these points of inquiry can lead to a more demonstrative analysis.
-
 \end{itshape}
+\newline\newline
+Thank you for the comments and the additional suggestions to improve the manuscript.
+The revised manuscript incorporates many of the suggested changes and additions.
+We include several new figures to further illustrate the change in Jacobi energy.
+In addition, we add some additional examples demonstrating the effect of variations in the control bound and terminal time on the reachability set.
 
@@
             The first half of the transfer is associated with and increase in energy as the control is used to transition towards the target orbit which is followed by a energy decrease to the target orbit.
-            This roughly corresponds with the expected optimal solution of a bang-coast-bang type orbital transfer~\cite{kirk2012}.
+            This roughly corresponds with the expected optimal solution of a bang-coast-bang type orbital transfer.
 
@@
 I do not necessarily ask or expect the authors to follow my above outline precisely, but I insist that some changes occur to make the analysis/illustrations better convey the general nature of the tool, not just that it gives you an answer.
+        \end{itshape}
 
+    \item
+        \begin{itshape}
             A more collected and focused discussion about how the parameter choices, assumptions, and limitations of the tool should inform its use is also warranted; currently remarks to this effect feel like scattered afterthoughts.

### 2018-05-17 manuscript.tex
Adding discussion on variations of u_m and tf

@@
 
+Next, we determine the reachability set with addition of a low-thrust control input over a fixed time horizon.
+In~\cref{sssec:periodic_orbit_varying_tf}, we demonstrate the effect of variations in the choice of variations of maximum control bound and terminal time.
+In~\cref{sssec:periodic_orbit_transfer} a transfer is demonstrated using the intersection of the reachability set on the \Poincare section.
+A notional thrust bound is selected to match emulate currect spacecraft electric propulsion technology.
+
+\subsubsection{Reachability Set on the \Poincare section}~\label{sssec:periodic_orbit_varying_tf}
+From the initial state on the periodic orbit, a series of optimal trajectories are generated to determine the reachable set.
+The multiple shooting approach is implemented to solve the optimal control problem. 
+We divide the time horizon into two equal length segments. 
+The state trajectory is initialized using the free trajectory of the periodic orbit. 
+Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propagated using the discrete equations of motion in~\cref{eq:costate_eom}.
+This results in the initial guess of the design vector \( \vecbf{\chi} = \begin{bmatrix} \vecbf{\lambda}_0 & \vecbf{x}_1^{-} & \vecbf{\lambda}_1^{-} & \vecbf{\beta} \end{bmatrix}^T\).
+This design vector is then varied to ensure that the necessary conditions of optimality and the interior point constraints are satisfied. 
+\Cref{tab:varying_tf} shows the range of terminal times, \( t_f \), and maximum control bound, \( u_m \), are used to investigate their effect on the subsequent reachable set on the \Poincare section.
+\begin{table}
+    \centering
+    \begin{tabular}{ll}  
+        \toprule
+        \(t_f\) & \( u_m \) \\
+        \midrule
+        1.24 & 0.05 \\
+        1.30 & 0.25 \\
+        1.37 & 0.5 \\
+        1.44 & \\
+        \bottomrule
+    \end{tabular}
+    \caption{Combinations for \( t_f \) and \( u_m\) used to generate the reachability set~\label{tab:varying_tf}}
+\end{table}
+
+The angle \( \theta_d\) in~\cref{eq:constraints} is varied to select a different direction along the \Poincare section to maximize.
+We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
+Choosing a new angle \( \theta_d \) corresponds to a different direction as well as a new optimal control problem which is again solved using the multiple shooting approach laid out previously.
+Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
+Each solution on the reachability set is computed in approximately \SI{2}{\minute} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
+The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:varying_tf_reachability_sets}.
+
+\begin{figure}
+    \subcaptionbox{Position space view of reachability sets\label{fig:varying_tf_trajectory}}{\includegraphics[width=0.5\textwidth]{figures/l1_transfer/trajectory_varying_um.pdf}}~
+    \subcaptionbox{\Poincare section view of reachability sets\label{fig:varying_tf_poincare}}{\includegraphics[width=0.5\textwidth]{figures/l1_transfer/poincare_section_varying_um.pdf}}
+    \caption{Variation of \( t_f \) and \( u_m\) on the reachability set. 
+    Colors, \braces{red, blue, green, yellow}, are used note increasing \( t_f\) while markers, \braces{circle, square, cross}, are used to denote increasing \(u_m\).
+    Increasing the maximum control bound has a large effect and enables the reachability set to intersect the target manifold. 
+    Increases in \(t_f\) are less critical and have minimal impact on the distribution of the reachability set on the \Poincare section.~\label{fig:varying_tf_reachability_sets}}
+\end{figure}
+
+\Cref{fig:varying_tf_reachability_sets} shows the reachabilty set for each combination of \(t_f\) and \( u_m\) listed in~\cref{tab:varying_tf}.
+Assuming a small control bound of \( u_m = \num{0.05} \) is shown in ~\cref{fig:varying_tf_reachability_sets} using square markers with the various terminal times shown using different colors.
+With a small maximum control bound, \( u_m = \num{0.05} \), the reachable set is not dramatically changed from that of the no control solution.
+The reachable set is shown using square markers on the left most portion of \cref{fig:varying_tf_poincare}.
+Variations of \( t_f\) are indicated using different colors and also demonstrate that this parameters has a smaller impact than changes in \( u_m \).
+Increasing the control bound to \( u_m = \num{0.25} \) and \( u_m = \num{0.5} \) shows that the reachable set is progressively approaching the target set and there are several intersections.
+
+The reachable sets presented in~\cref{fig:varying_tf_reachability_sets} are highly dependent on the initial condition, terminal time, and maximum control bound. 
+This example demonstrates that increasing the \( u_m \) results in a larger displacement between the controlled and uncontrolled trajectories.
+The reachable set is 
+\subsubsection{Transfer to Periodic Orbit}~\label{sssec:periodic_orbit_transfer}
 % reachable set transfer

### 2018-05-17 manuscript.tex
reorder some paragraphs

@@
 In~\cref{sssec:periodic_orbit_varying_tf}, we demonstrate the effect of variations in the choice of variations of maximum control bound and terminal time.
-In~\cref{sssec:periodic_orbit_transfer} a transfer is demonstrated using the intersection of the reachability set on the \Poincare section.
-A notional thrust bound is selected to match emulate currect spacecraft electric propulsion technology.
+While~\cref{sssec:periodic_orbit_transfer} show a specific example of a transfer design process using the intersection of the reachability set on the \Poincare section.
+The analysis presented in the following sections define a maximum magnitude of the thrust as \( u_{m} \) and assume that the trust can be directed arbitrarily within the plane. 
+This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
+Assuming a fully actuated spacecraft model allows us to decouple the translation and rotational dynamics of the spacecraft.
 
@@
 This design vector is then varied to ensure that the necessary conditions of optimality and the interior point constraints are satisfied. 
-\Cref{tab:varying_tf} shows the range of terminal times, \( t_f \), and maximum control bound, \( u_m \), are used to investigate their effect on the subsequent reachable set on the \Poincare section.
+\Cref{tab:varying_tf} shows the range of terminal times, \( t_f \), and maximum control bound, \( u_m \), which are used to investigate their effect on the subsequent reachable set on the \Poincare section.
 \begin{table}
@@
 This example demonstrates that increasing the \( u_m \) results in a larger displacement between the controlled and uncontrolled trajectories.
-The reachable set is 
+The reachable set is enlarged from a single point, as shown in~\cref{fig:manifold_poincare} by the black point, to a larger region as shown in~\cref{fig:varying_tf_poincare}.
+The choice of \( u_m\), \( t_f \) and initial condition all combine to change the resulting reachability set.
+
 \subsubsection{Transfer to Periodic Orbit}~\label{sssec:periodic_orbit_transfer}
 % reachable set transfer
-Next, we determine the reachability set with the addition of a low thrust control input.
-We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
-This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
-We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
-This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
+Here we use the results of the preceeding section to demonstrate one specific example of a transfer designed using the reachability set.
+The acceleration limit is chosen as  \( u_{max} = 0.75 \approx \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
 Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of approximately \SI{1}{\newton}.
@@
 This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit.
-
-From the initial state on the periodic orbit, a series of optimal trajectories are generated to determine the reachable set.
-The multiple shooting approach is implemented to solve the optimal control problem. 
-We divide the time horizon into two equal length segments. 
-The state trajectory is initialized using the free trajectory of the periodic orbit. 
-Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propagated using the discrete equations of motion in~\cref{eq:costate_eom}.
-This results in the initial guess of the design vector \( \vecbf{\chi} = \begin{bmatrix} \vecbf{\lambda}_0 & \vecbf{x}_1^{-} & \vecbf{\lambda}_1^{-} & \vecbf{\beta} \end{bmatrix}^T\).
-This design vector is then varied to ensure that the necessary conditions of optimality and the interior point constraints are satisfied.
-
-By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
-We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
-Choosing a new angle \( \theta_d \) corresponds to a different direction as well as a new optimal control problem which is again solved using the multiple shooting approach laid out previously.
-Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
-Each solution on the reachability set is computed in approximately \SI{2}{\minute} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
-The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:invariant_manifold_transfer}.
 

### 2018-05-17 Response.tex
Update response with additions to manuscript

@@
 
+        Thank you for the comment. 
+        We agree that some further discussion of the effect of these ``tunable'' parameters would help to improve the manuscript. 
+        An additional section has been included which highlights some of the effects of variations in terminal time, and control magnitude (next comment).
+        The results presented are not an exhaustive description of the search space (in terms of travel time and/or control magnitude) but do highlight some of the important characteristics of these parameters on the resulting reachability set.
+        The additional discussion and plots are shown below.
+
+        \begin{correction}
+            Next, we determine the reachability set with addition of a low-thrust control input over a fixed time horizon.
+            In~\cref{sssec:periodic_orbit_varying_tf}, we demonstrate the effect of variations in the choice of variations of maximum control bound and terminal time.
+            While~\cref{sssec:periodic_orbit_transfer} show a specific example of a transfer design process using the intersection of the reachability set on the \Poincare section.
+            The analysis presented in the following sections define a maximum magnitude of the thrust as \( u_{m} \) and assume that the trust can be directed arbitrarily within the plane. 
+            This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
+            Assuming a fully actuated spacecraft model allows us to decouple the translation and rotational dynamics of the spacecraft.
+
+            \subsubsection{Reachability Set on the \Poincare section}
+            From the initial state on the periodic orbit, a series of optimal trajectories are generated to determine the reachable set.
+            The multiple shooting approach is implemented to solve the optimal control problem. 
+            We divide the time horizon into two equal length segments. 
+            The state trajectory is initialized using the free trajectory of the periodic orbit. 
+            Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propagated using the discrete equations of motion in~\cref{eq:costate_eom}.
+            This results in the initial guess of the design vector \( \vecbf{\chi} = \begin{bmatrix} \vecbf{\lambda}_0 & \vecbf{x}_1^{-} & \vecbf{\lambda}_1^{-} & \vecbf{\beta} \end{bmatrix}^T\).
+            This design vector is then varied to ensure that the necessary conditions of optimality and the interior point constraints are satisfied. 
+            \Cref{tab:varying_tf} shows the range of terminal times, \( t_f \), and maximum control bound, \( u_m \), which are used to investigate their effect on the subsequent reachable set on the \Poincare section.
+            \begin{table}
+                \centering
+                \begin{tabular}{ll}  
+                    \toprule
+                    \(t_f\) & \( u_m \) \\
+                    \midrule
+                    1.24 & 0.05 \\
+                    1.30 & 0.25 \\
+                    1.37 & 0.5 \\
+                    1.44 & \\
+                    \bottomrule
+                \end{tabular}
+                \caption{Combinations for \( t_f \) and \( u_m\) used to generate the reachability sets in~\cref{fig:varying_tf_reachability_sets}}
+            \end{table}
+
+            The angle \( \theta_d\) in~\cref{eq:constraints} is varied to select a different direction along the \Poincare section to maximize.
+            We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
+            Choosing a new angle \( \theta_d \) corresponds to a different direction as well as a new optimal control problem which is again solved using the multiple shooting approach laid out previously.
+            Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
+            Each solution on the reachability set is computed in approximately \SI{2}{\minute} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
+            The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:varying_tf_reachability_sets}.
+
+            \begin{figure}
+                \subcaptionbox{Position space view of reachability sets}{\includegraphics[width=0.5\textwidth]{figures/l1_transfer/trajectory_varying_um.pdf}}~
+                \subcaptionbox{\Poincare section view of reachability sets}{\includegraphics[width=0.5\textwidth]{figures/l1_transfer/poincare_section_varying_um.pdf}}
+                \caption{Variation of \( t_f \) and \( u_m\) on the reachability set. 
+                    Colors, \braces{red, blue, green, yellow}, are used note increasing \( t_f\) while markers, \braces{circle, square, cross}, are used to denote increasing \(u_m\).
+                    Increasing the maximum control bound has a large effect and enables the reachability set to intersect the target manifold. 
+                Increases in \(t_f\) are less critical and have minimal impact on the distribution of the reachability set on the \Poincare section.}
+            \end{figure}
+
+            \Cref{fig:varying_tf_reachability_sets} shows the reachabilty set for each combination of \(t_f\) and \( u_m\) listed in~\cref{tab:varying_tf}.
+            Assuming a small control bound of \( u_m = \num{0.05} \) is shown in ~\cref{fig:varying_tf_reachability_sets} using square markers with the various terminal times shown using different colors.
+            With a small maximum control bound, \( u_m = \num{0.05} \), the reachable set is not dramatically changed from that of the no control solution.
+            The reachable set is shown using square markers on the left most portion of \cref{fig:varying_tf_poincare}.
+            Variations of \( t_f\) are indicated using different colors and also demonstrate that this parameters has a smaller impact than changes in \( u_m \).
+            Increasing the control bound to \( u_m = \num{0.25} \) and \( u_m = \num{0.5} \) shows that the reachable set is progressively approaching the target set and there are several intersections.
+
+            The reachable sets presented in~\cref{fig:varying_tf_reachability_sets} are highly dependent on the initial condition, terminal time, and maximum control bound. 
+            This example demonstrates that increasing the \( u_m \) results in a larger displacement between the controlled and uncontrolled trajectories.
+            The reachable set is enlarged from a single point, as shown in~\cref{fig:manifold_poincare} by the black point, to a larger region as shown in~\cref{fig:varying_tf_poincare}.
+            The choice of \( u_m\), \( t_f \) and initial condition all combine to change the resulting reachability set.
+
+        \end{correction}
     \item 
@@
         \end{itshape}
+
+        Thank you for the comment.
+        We have included an additional section describing the effect of varaitions in \( u_m \) and terminal time on the reachable set.
+        As discussed above, the results are not an exhaustive exploration of the parameter space but illustrate the key effect of the control bound and terminal time.
+        In the example presented, increasing the magnitude of the control has the effect of enlarging and moving the reachable set. 
+        At a sufficiently large value of \( u_m \) the reachable set is able to intersect the target orbit.
 

### 2018-05-17 Response.tex
Finish responding to all the comments.

@@
 
+        Thank you for the comment.
+        As stated in the previous comments we have included additional discussion on the effect of variations of terminal time and control bound.
+
+        \begin{itshape}
+That said, it's a little confusing how the red and green curves of Fig8b don't line up with this description to fit the comment of P18L47; additional consideration/analysis may be needed with regards to the number of revs/crossings used to define a section for comparison.
+        \end{itshape}
+
+        Thank you for the question. 
+        The green points are the intersection of the unstable manifold with the \Poincare section while the red points are the reachability set.
+        Without any control input the spacecraft would remain on the periodic orbit and intersect the \Poincare seciton at \( x_n\). 
+        The addition of the control input allows for an expansion of the reachable set from a single point to the larger red region.
+        An additional clarification has been added to the manuscript to aid in the description of the figure.
+
+        \begin{correction}
+            In contrast, the low thrust control input we are able to enlarge the reachability set from a single point, \( x_n\) associated with the periodic orbit, to a larger ellipsoidal region shown in red in~\cref{fig:poincare_compare}.
+        \end{correction}
+
     \item
@@
 
+        Thank you for the comments.
+        The manuscript has been editied, as described previously, to incorporate several changes to illustrate the specific characteristics of this approach.
+        An additional section is presented which highlights the effect of terminal time and control magnitude on the reachability set.
+
+        The second example presents a larger and more involved example of combining several iterations of the reachable set to achieve a larger and more complex transfer. 
+        This aspect of the proposed method, more specifically of combining several reachability sets to approach a target, is not demonstrated in the first example.
+        As a result, we feel that the second example does add some additional insight into the use of this method for a more complicated transfer.
+
     \item
@@
         \end{itshape}
+
+        Thank you for the helpful comments to improve the manuscript.
+        The manuscript has been edited to reduce the length and repetition of the introduction.
+        In addition, an additional section is presented highlighting the effect of the variations of terminal time and control magnitude on the reachable set. 
+        Furthermore, the presented examples also incldue additional plots showing the change of the Jacobi energy during the transfer. 
 \end{itemize}

### 2018-05-17 manuscript.tex
Update response with additions to manuscript
Finish responding to all the comments.

@@
     \end{tabular}
-    \caption{Combinations for \( t_f \) and \( u_m\) used to generate the reachability set~\label{tab:varying_tf}}
+    \caption{Combinations for \( t_f \) and \( u_m\) used to generate the reachability sets in~\cref{fig:varying_tf_reachability_sets}~\label{tab:varying_tf}}
 \end{table}
@@
 In addition, only a small portion of the invariant manifold intersects with the target Moon orbit.
-In contrast, the low thrust control input we are able to enlarge the reachability set from a single point to a larger ellipsoidal region in~\cref{fig:poincare_compare}.
+In contrast, the low thrust control input we are able to enlarge the reachability set from a single point, \( x_n\) associated with the periodic orbit, to a larger ellipsoidal region shown in red in~\cref{fig:poincare_compare}.
 This achieves an intersection with the target orbit with a much lower time of flight as compared to the invariant manifold method.

### 2018-05-21 Response.tex
tylee edit

@@
 
-We thank the reviewers for their comments and aid in imporoving the quality of our manuscript.
+We thank the reviewers for their comments and aid in improving the quality of our manuscript. 
+In accordance with the comments and suggestions, the manuscript has been revised, and the answers to all comments are addressed as follow.
 
@@
 
-            "This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods."
+            "This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge-Kutta methods."
 
@@
     The manuscript has been modified to remove the about the benefits of state accuracy.
-    The section has been modified to highlight the reduced computational load of the varaitional integrator.
+    The section has been modified to highlight the reduced computational load of the variational integrator.
     In addition, an additional sentence has been added specifically highlighting the fact that high order Runge-Kutta methods will probably provide better state accuracy at the expense of additional computations and poorer energy behavior.
@@
     \begin{correction}
-        This work uses a second order integrator which has improved energy behavior in comparision to first order integration techniques without a significant increase in computational demand.
-        The variational integrator provides improved energy behavior over long integration times but does not guarantee state accuracy.
-        High order Runge-Kutta methods will likely provide a more accurate state trajectory, but at the expense of additional computational demand and total energy deviations.
+        This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
+        The variational integrator is constructed according to discrete Hamilton's principle, specifically for the given dynamic system, and it provides a long-term structural stability in capturing the effects of low-thrust propulsion systems.
+        General purpose numerical integrators, such as adaptive stepsize Runge-Kutta methods may yield a more accurate state trajectory depending on the choice of integration parameters, but at the expense of additional computational demand and total energy deviations.
 
@@
 
-    This sentence has been modified.
+    Thank you for the comments. This sentence has been modified.
     \begin{correction}
-        The indirect optimal control formulation direclty solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
+        The indirect optimal control formulation directly solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
         Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution and typically results in a much larger parameter set.
@@
         \begin{itshape}
-            It's superfluous to explain at such length that spacecraft trajectory design is a problem of interest, that cubesats are a thing, etc.
+            It's superfluous to explain at such length that spacecraft trajectory design is a problem of interest, that CubeSats are a thing, etc.
 Broad points like these could potentially be condensed into single, citation-heavy sentences.
@@
             % utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
-            Recent developments in miniature electric propulsion and small satellites now offer the potential for new research oppportunities~\cite{folta2015,haque2013}.
+            Recent developments in miniature electric propulsion and small satellites now offer the potential for new research opportunities~\cite{folta2015,haque2013}.
             The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}.
@@
             Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
-            Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
+            Conventional general-purpose Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, may fail to preserve geometric properties of the dynamic system numerically.
             Variational integrators can allow for reduced computational effort without any loss of numerical stability or energy drift which exist in conventional integration schemes.
@@
 Thank you for the comments and the additional suggestions to improve the manuscript. 
-The revised manuscript incorporates many of the suggested changes and additions.
+The main objective of the submitted manuscript is to construct a systematic approach to design low-thrust orbital maneuvers. 
+It is motivated by the specific challenges that electric propulsion systems generate minuscule level of thrust, and low-thrust orbital maneuvers rely on those effects accumulated over a long time period.
+It is quite cumbersome and challenging to apply numerical optimization tools or boundary value problem solvers successfully to low-thrust orbital design problems. 
+We aim to provide more organized tools to mitigate such difficulties by explicitly considering the characteristics of the natural dynamics and utilizing geometric integrators exhibiting long-term numerical stability.
+The former is achieved by incorporating the control-free maneuvers obtained by the Poincare section in the invariant manifolds, and the latter is achieved by variational integrators. 
+However, the proposed scheme does not guarantee any optimality in the constructed maneuvers. 
+It is heuristically implied that the constructed orbital maneuver would require no-excessive control input, as it is essentially constructed by adjusting the control-free trajectory. 
+In short,
+\begin{itemize}
+	\item[Pros] The proposed scheme utilizes the characteristics of the inherent three-body problem, in constructing low-thrust orbital maneuvers in a systematic way. 
+		This reduces the cumbersome process of guessing a \textit{good} set of initial parameters to make the direct application of optimization become successful. 
+	\item[Pros] The use of geometric integrators is particularly useful in capturing the long-term effects of low thrust computationally effectively. 
+	\item[Cons] While the proposed scheme generates orbital trajectories to the target with prescribed constraints, it does not guarantee any optimality. 
+\end{itemize}
+The conclusions section has been revised to summarize the above pros and cons of the proposed approach. 
+\begin{itshape}
+	Put the revised conclusions here.
+\end{itshape}
+
+Also, the revised manuscript incorporates many of the suggested changes and additions.
 We include several new figures to further illustrate the change in Jacobi energy.
@@
         We agree that some further discussion of the effect of these ``tunable'' parameters would help to improve the manuscript. 
+		According to the comments, we considered twelve additional cases for every combination of varying terminal time $t_f$ and the maximum control $u_m$ listed at Table \ref{fig:varying_tf_reachability_sets}.
         An additional section has been included which highlights some of the effects of variations in terminal time, and control magnitude (next comment).
@@
 
-        The combination of several reachability sets does not force the system to return to a prescribed energy level.
+		As pointed out by the reviewer, our respond stating `..the trajectories do return to the energy level of the Poincare section..' is not correct. 
+        Instead, the combination of several reachability sets does not force the system to return to a prescribed energy level.
         Each reachability set computation serves to maximize a distance metric, defined in a reduced phase space corresponding to the \Poincare section.
@@
 
+		With regards to the comments on the second example, the authors decided to keep them in the manuscript, as it illustrates the idea of multiple stages to construct orbital maneuvers with more challenging scenarios.
+		Nonetheless, the revised manuscript includes extended numerical results and discussions for the first example, for example to illustrate the effects of varying $t_f$ and $u_m$ and the evolution of the Jacobi constant.
     \item 
@@
         As a result, we feel that the second example does add some additional insight into the use of this method for a more complicated transfer.
+		Furthermore, the revised conclusions section summarizes the objective of the submitted manuscript, and the pros and the cons of the proposed approach.
 
@@
         In addition, an additional section is presented highlighting the effect of the variations of terminal time and control magnitude on the reachable set. 
-        Furthermore, the presented examples also incldue additional plots showing the change of the Jacobi energy during the transfer. 
+        Furthermore, the presented examples also include additional plots showing the change of the Jacobi energy during the transfer. 
 \end{itemize}

### 2018-05-21 Response.tex
Merge changes from Dr. Lee to manuscript

@@
         This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
-        The variational integrator is constructed according to discrete Hamilton's principle, specifically for the given dynamic system, and it provides a long-term structural stability in capturing the effects of low-thrust propulsion systems.
+        The variational integrator is constructed according to a discrete version of Hamilton's principle, specifically for the given dynamic system, and it provides long-term structural stability in capturing the effects of low-thrust propulsion systems.
         General purpose numerical integrators, such as adaptive stepsize Runge-Kutta methods may yield a more accurate state trajectory depending on the choice of integration parameters, but at the expense of additional computational demand and total energy deviations.
@@
 The main objective of the submitted manuscript is to construct a systematic approach to design low-thrust orbital maneuvers. 
-It is motivated by the specific challenges that electric propulsion systems generate minuscule level of thrust, and low-thrust orbital maneuvers rely on those effects accumulated over a long time period.
+It is motivated by the specific challenges that electric propulsion systems generate a minuscule level of thrust, and low-thrust orbital maneuvers rely on those effects accumulated over a long time period.
 It is quite cumbersome and challenging to apply numerical optimization tools or boundary value problem solvers successfully to low-thrust orbital design problems. 
@@
 \begin{itemize}
-	\item[Pros] The proposed scheme utilizes the characteristics of the inherent three-body problem, in constructing low-thrust orbital maneuvers in a systematic way. 
-		This reduces the cumbersome process of guessing a \textit{good} set of initial parameters to make the direct application of optimization become successful. 
-	\item[Pros] The use of geometric integrators is particularly useful in capturing the long-term effects of low thrust computationally effectively. 
+	\item[Pros] The proposed scheme utilizes the characteristics inherent in the three-body problem in constructing low-thrust orbital maneuvers in a systematic way. 
+		This reduces the cumbersome process of guessing a \textit{good} set of initial parameters to make the direct application of optimization successful. 
+	\item[Pros] The use of geometric integrators is particularly useful in capturing the long-term effects of low thrust in a computationaly efficient manner. 
 	\item[Cons] While the proposed scheme generates orbital trajectories to the target with prescribed constraints, it does not guarantee any optimality. 

### 2018-05-21 manuscript.tex
Merge changes from Dr. Lee to manuscript

@@
 % utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
-Recent developments in miniature electric propulsion and small satellites now offer the potential for new research oppportunities~\cite{folta2015,haque2013}.
+Recent developments in miniature electric propulsion and small satellites now offer the potential for new research opportunities~\cite{folta2015,haque2013}.
 The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}.
@@
 Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
-Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
+Conventional general-purpose Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, may fail to preserve geometric properties of the dynamic system numerically.
 Variational integrators can allow for reduced computational effort without any loss of numerical stability or energy drift which exist in conventional integration schemes.
@@
 In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
-This work uses a second order integrator which has improved energy behavior in comparision to first order integration techniques without a significant increase in computational demand.
-The variational integrator provides improved energy behavior over long integration times but does not guarantee state accuracy.
-High order Runge-Kutta methods will likely provide a more accurate state trajectory, but at the expense of additional computational demand and total energy deviations.
+This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
+The variational integrator is constructed according to a discrete version of Hamilton's principle, specifically for the given dynamic system, and it provides long-term structural stability in capturing the effects of low-thrust propulsion systems.
+General purpose numerical integrators, such as adaptive stepsize Runge-Kutta methods may yield a more accurate state trajectory depending on the choice of integration parameters, but at the expense of additional computational demand and total energy deviations.
 However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
@@
 The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
-The indirect optimal control formulation direclty solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
+The indirect optimal control formulation directly solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
 Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution and typically results in a much larger parameter set.
@@
 However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
+
 Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.

### 2018-05-21 Response.tex
Update response
Revise the conclusion

@@
 The conclusions section has been revised to summarize the above pros and cons of the proposed approach. 
-\begin{itshape}
-	Put the revised conclusions here.
-\end{itshape}
+\begin{correction}
+In this paper, we construct a systematic approach which combines the concepts of reachability sets and \Poincare~sections to generate orbital transfers between planar periodic orbits in the three-body problem using low-thrust propulsion systems.
+The \Poincare~section allows for trajectory design on a lower dimensional phase space thereby reducing the complexity and search space.
+The approximation of the reachability set enables a simple methodology for the selection of \textit{good} initial parameters which will allow for convergence of the optimal control formulation.
+The combination of these two concepts allows the engineer to capture the characteristics inherent in the three-body problem.
+Additionally, we utilize geometric integrators to efficiently capture the long-term effects of low-thrust on the system dynamics.
+Our approach enables the computation of orbital transfers within the three-body problem but does not guarantee optimality of the resulting transfer.
+Instead, the resulting transfers will require no excessive control inputs as it is formulated as a deviation from the control-free trajectory.
+The presented approach allows for a systematic method to determine and generate orbital transfers in the three-body problem using low-thrust proplusion.
+
+There is additional research to extend these results to more general transfer scenarios in future work.
+The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
+In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems and future research is aimed at investigating smaller magnitude control inputs.
+Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
+This will result in a more complicated optimal control problem and is a focus of future research.
+Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
+The addition of attitude dynamics and realistic pointing constraints would also significantly improve the applicability of this work.
+
+\end{correction}
 
@@
 
-		As pointed out by the reviewer, our respond stating `..the trajectories do return to the energy level of the Poincare section..' is not correct. 
+		As pointed out by the reviewer, our response stating `..the trajectories do return to the energy level of the Poincare section..' is not correct. 
         Instead, the combination of several reachability sets does not force the system to return to a prescribed energy level.

### 2018-05-21 manuscript.tex
Revise the conclusion

@@
 \section{Conclusions}\label{sec:conclusion}
-In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~sections is used to generate a transfer between planar periodic orbits in the three-body problem.
-The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
-The indirect optimal control formulation directly solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
-Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution and typically results in a much larger parameter set.
-Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
-Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
-However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
-
-Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
+In this paper, we construct a systematic approach which combines the concepts of reachability sets and \Poincare~sections to generate orbital transfers between planar periodic orbits in the three-body problem using low-thrust propulsion systems.
+The \Poincare~section allows for trajectory design on a lower dimensional phase space thereby reducing the complexity and search space.
+The approximation of the reachability set enables a simple methodology for the selection of \textit{good} initial parameters which will allow for convergence of the optimal control formulation.
+The combination of these two concepts allows the engineer to capture the characteristics inherent in the three-body problem.
+Additionally, we utilize geometric integrators to efficiently capture the long-term effects of low-thrust on the system dynamics.
+Our approach enables the computation of orbital transfers within the three-body problem but does not guarantee optimality of the resulting transfer.
+Instead, the resulting transfers will require no excessive control inputs as it is formulated as a deviation from the control-free trajectory.
+The presented approach allows for a systematic method to determine and generate orbital transfers in the three-body problem using low-thrust proplusion.
 

### 2018-05-23 Response.tex
tylee last revision

@@
 It is motivated by the specific challenges that electric propulsion systems generate a minuscule level of thrust, and low-thrust orbital maneuvers rely on those effects accumulated over a long time period.
-It is quite cumbersome and challenging to apply numerical optimization tools or boundary value problem solvers successfully to low-thrust orbital design problems. 
-We aim to provide more organized tools to mitigate such difficulties by explicitly considering the characteristics of the natural dynamics and utilizing geometric integrators exhibiting long-term numerical stability.
-The former is achieved by incorporating the control-free maneuvers obtained by the Poincare section in the invariant manifolds, and the latter is achieved by variational integrators. 
-However, the proposed scheme does not guarantee any optimality in the constructed maneuvers. 
+Consequently, it is quite cumbersome and challenging to apply numerical optimization tools or boundary value problem solvers successfully to low-thrust orbital design problems. 
+For example, it is difficult to initialize the optimization parameters such that the resulting numerical iteration converges successfully. 
+Conventional numerical integration schemes have shown to be possibly erroneous particularly in computing long-term energy behaviors correctly.
+
+In the submitted manuscript, we aim to provide more organized tools to mitigate such difficulties by explicitly considering the characteristics of the natural dynamics and utilizing geometric integrators exhibiting long-term numerical stability.
+The consideration of the natural, control-free dynamics is reasonable as low-thrust trajectories won't be drastically different from those over a limited time period. 
+Here, the former is achieved by incorporating the control-free maneuvers characterized by the Poincare section in the invariant manifolds, and the latter is achieved by variational integrators.
+
+However, the proposed scheme is focused on constructing low-thrust trajectories satisfying the prescribed boundary conditions, and it does not guarantee any optimality in the constructed maneuvers. 
 It is heuristically implied that the constructed orbital maneuver would require no-excessive control input, as it is essentially constructed by adjusting the control-free trajectory. 
+As such, it is rather focused on constructing a reasonable low-thrust trajectory that satisfies the given constraints on the maximum thrust bound and the boundary conditions.
+
 In short,
@@
 		This reduces the cumbersome process of guessing a \textit{good} set of initial parameters to make the direct application of optimization successful. 
-	\item[Pros] The use of geometric integrators is particularly useful in capturing the long-term effects of low thrust in a computationaly efficient manner. 
+	\item[Pros] The use of geometric integrators is particularly useful in capturing the long-term effects of low thrust in a computationally efficient manner. 
 	\item[Cons] While the proposed scheme generates orbital trajectories to the target with prescribed constraints, it does not guarantee any optimality. 
 \end{itemize}
-The conclusions section has been revised to summarize the above pros and cons of the proposed approach. 
+
+In the revised manuscript, the conclusions section has been rewritten to summarize the above pros and cons of the proposed approach. 
 \begin{correction}
@@
 Instead, the resulting transfers will require no excessive control inputs as it is formulated as a deviation from the control-free trajectory.
-The presented approach allows for a systematic method to determine and generate orbital transfers in the three-body problem using low-thrust proplusion.
+The presented approach allows for a systematic method to determine and generate orbital transfers in the three-body problem using low-thrust propulsion.
 
@@
 
-Also, the revised manuscript incorporates many of the suggested changes and additions.
+Also, the revised manuscript incorporates the suggested changes and additions.
 We include several new figures to further illustrate the change in Jacobi energy.
 In addition, we add some additional examples demonstrating the effect of variations in the control bound and terminal time on the reachability set.
+These are itemized below.
 
@@
         The results presented are not an exhaustive description of the search space (in terms of travel time and/or control magnitude) but do highlight some of the important characteristics of these parameters on the resulting reachability set.
-        The additional discussion and plots are shown below.
+		According to the new numerical results, the manuscript has been revised to discuss the effects of varying $t_f$ and $u_m$ as copied below.
 
@@
         Thank you for the comment.
-        We have included an additional section describing the effect of varaitions in \( u_m \) and terminal time on the reachable set.
+        We have included an additional section describing the effect of variations in \( u_m \) and terminal time on the reachable set.
         As discussed above, the results are not an exhaustive exploration of the parameter space but illustrate the key effect of the control bound and terminal time.
@@
         The green points are the intersection of the unstable manifold with the \Poincare section while the red points are the reachability set.
-        Without any control input the spacecraft would remain on the periodic orbit and intersect the \Poincare seciton at \( x_n\). 
+        Without any control input the spacecraft would remain on the periodic orbit and intersect the \Poincare section at \( x_n\). 
         The addition of the control input allows for an expansion of the reachable set from a single point to the larger red region.
@@
 
-		As pointed out by the reviewer, our response stating `..the trajectories do return to the energy level of the Poincare section..' is not correct. 
+		As pointed out by the reviewer, our prior response stating `..the trajectories do return to the energy level of the Poincare section..' is erroneous and misleading. 
         Instead, the combination of several reachability sets does not force the system to return to a prescribed energy level.
@@
         Thank you for the comments.
-        The manuscript has been editied, as described previously, to incorporate several changes to illustrate the specific characteristics of this approach.
-        An additional section is presented which highlights the effect of terminal time and control magnitude on the reachability set.
+        The manuscript has been edited, as described previously, to incorporate several changes to illustrate the specific characteristics of this approach.
+        Also, an additional section is presented which highlights the effect of terminal time and control magnitude on the reachability set.
 

### 2018-05-23 manuscript.tex

@@
 \end{equation}
-where the matrix \( A \) and psuedo gravitational potential gradient \( \nabla U\) are
+where the matrix \( A \) and pseudo gravitational potential gradient \( \nabla U\) are
 \begin{align}
@@
 % reachable set transfer
-Here we use the results of the preceeding section to demonstrate one specific example of a transfer designed using the reachability set.
+Here we use the results of the preceding section to demonstrate one specific example of a transfer designed using the reachability set.
 The acceleration limit is chosen as  \( u_{max} = 0.75 \approx \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
@@
         Optimal cost       & \num{2.09e-31}      \\
-        Executation time & \SI{1.49}{\second}       \\
+        Execution time & \SI{1.49}{\second}       \\
         \bottomrule
@@
         Optimal cost       & \num{4.30e-25}      \\
-        Executation time & \SI{2.62}{\second}       \\
+        Execution time & \SI{2.62}{\second}       \\
         \bottomrule
@@
     Each stage of the transfer serves to increase the energy towards the stable manifold. 
-The final reachability set intersects the stable manifold and enables a large final manuever to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.5\textwidth]{figures/geo_transfer/jacobi_energy.pdf}}
+The final reachability set intersects the stable manifold and enables a large final maneuver to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.5\textwidth]{figures/geo_transfer/jacobi_energy.pdf}}
 

### 2018-05-23 Response.tex
Fix some typos in response

@@
 
-    The manuscript has been modified to remove the about the benefits of state accuracy.
+    The manuscript has been modified to accurately discuss the drawbacks of variational integrators.
     The section has been modified to highlight the reduced computational load of the variational integrator.
@@
 Several paragraphs have been removed or refactored to reduce repetition and extraneous detail.
-The updated introduction is included below and specific comments for each comment is also provided.
+The updated introduction is included below and specific responses for each comment are also provided.
 
@@
 For example, it is difficult to initialize the optimization parameters such that the resulting numerical iteration converges successfully. 
-Conventional numerical integration schemes have shown to be possibly erroneous particularly in computing long-term energy behaviors correctly.
+Conventional numerical integration schemes have shown to be possibly erroneous particularly in correctly computing long-term energy behaviors.
 
@@
 The consideration of the natural, control-free dynamics is reasonable as low-thrust trajectories won't be drastically different from those over a limited time period. 
-Here, the former is achieved by incorporating the control-free maneuvers characterized by the Poincare section in the invariant manifolds, and the latter is achieved by variational integrators.
+Here, the former is achieved by incorporating the control-free maneuvers characterized by the \Poincare section in the invariant manifolds, and the latter is achieved by variational integrators.
 

### 2018-05-23 manuscript.tex
Reword for clarity

@@
 Next, we determine the reachability set with addition of a low-thrust control input over a fixed time horizon.
-In~\cref{sssec:periodic_orbit_varying_tf}, we demonstrate the effect of variations in the choice of variations of maximum control bound and terminal time.
+In~\cref{sssec:periodic_orbit_varying_tf}, we demonstrate the effect of variations in the choice of maximum control bound and terminal time.
 While~\cref{sssec:periodic_orbit_transfer} show a specific example of a transfer design process using the intersection of the reachability set on the \Poincare section.
@@
 
-\Cref{fig:varying_tf_reachability_sets} shows the reachabilty set for each combination of \(t_f\) and \( u_m\) listed in~\cref{tab:varying_tf}.
-Assuming a small control bound of \( u_m = \num{0.05} \) is shown in ~\cref{fig:varying_tf_reachability_sets} using square markers with the various terminal times shown using different colors.
-With a small maximum control bound, \( u_m = \num{0.05} \), the reachable set is not dramatically changed from that of the no control solution.
-The reachable set is shown using square markers on the left most portion of \cref{fig:varying_tf_poincare}.
+\Cref{fig:varying_tf_reachability_sets} shows the reachabilty set for twelve combinations of \(t_f\) and \( u_m\) listed in~\cref{tab:varying_tf}.
+With a small maximum control bound, \( u_m = \num{0.05} \) shown using square markers, the reachable set is not dramatically changed from that of the no control solution.
+The reachable set is denoted using square markers on the left most portion of \cref{fig:varying_tf_poincare}.
 Variations of \( t_f\) are indicated using different colors and also demonstrate that this parameters has a smaller impact than changes in \( u_m \).

### 2018-05-23 manuscript.tex
reword and fix figure caption

@@
     \caption{Variation of \( t_f \) and \( u_m\) on the reachability set. 
-    Colors, \braces{red, blue, green, yellow}, are used note increasing \( t_f\) while markers, \braces{circle, square, cross}, are used to denote increasing \(u_m\).
+        Colors, \braces{\text{red}, \text{blue}, \text{green}, \text{yellow}}, are used note increasing \( t_f\) while markers, \braces{\text{circle}, \text{square}, \text{cross}}, are used to denote increasing \(u_m\).
     Increasing the maximum control bound has a large effect and enables the reachability set to intersect the target manifold. 
@@
 Variations of \( t_f\) are indicated using different colors and also demonstrate that this parameters has a smaller impact than changes in \( u_m \).
-Increasing the control bound to \( u_m = \num{0.25} \) and \( u_m = \num{0.5} \) shows that the reachable set is progressively approaching the target set and there are several intersections.
+Increasing the control bound to \( u_m = \num{0.25} \) and \( u_m = \num{0.5} \) shows that the reachable set progressively approaches the target set.
 

### 2018-05-24 manuscript.tex

@@
         \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
-        The reachability set is shown in the upper two figures in both the position and \Poincare space.
-    From this reachability set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
-The optimal control to achieve this transfer is shown in the lower right figure.}
+            The reachability set is shown in~\cref{fig:reach_trajectory,fig:poincare_compare} in both the position and \Poincare space respectively.
+            From this reachability set we chose a trajectory which intersects the target orbit and it is shown in~\cref{fig:reach_transfer}.
+        The optimal control to achieve this transfer is shown in~\cref{fig:control_l1}.}
         \label{fig:reachability_set_transfer} 
@@
 This intersection generates a partial target state of \( x_t \text{ and } \dot{x}_t \).
-Using the energy level of the target region, defined by~\cref{eq:jacobi}, and the intersection state we can calculate the final component \( \dot{y} \) is calculated. 
+Using the energy level of the target region, defined by~\cref{eq:jacobi}, and the intersection state we can calculate the final component \( \dot{y} \). 
 This results in a complete target state \( \vecbf{x}_t \) which lies on the reachable set and on the target orbit. 

### 2018-05-24 Response.tex
Images in top level directory

@@
             \begin{figure}
-                \subcaptionbox{Position space view of reachability sets}{\includegraphics[width=0.5\textwidth]{figures/l1_transfer/trajectory_varying_um.pdf}}~
-                \subcaptionbox{\Poincare section view of reachability sets}{\includegraphics[width=0.5\textwidth]{figures/l1_transfer/poincare_section_varying_um.pdf}}
+                \subcaptionbox{Position space view of reachability sets}{\includegraphics[width=0.5\textwidth]{trajectory_varying_um.pdf}}~
+                \subcaptionbox{\Poincare section view of reachability sets}{\includegraphics[width=0.5\textwidth]{poincare_section_varying_um.pdf}}
                 \caption{Variation of \( t_f \) and \( u_m\) on the reachability set. 
@@
                 \begin{subfigure}[htbp]{0.4\textwidth} 
-                    \includegraphics[width=\textwidth, keepaspectratio]{figures/l1_transfer/jacobi.pdf} 
+                    \includegraphics[width=\textwidth, keepaspectratio]{jacobi.pdf} 
                     \caption{Jacobi Energy over transfer \label{fig:jacobi_l1}} 
@@
     \subcaptionbox{Control Input: Combined control history for each stage of the reachability analysis. 
-    The control always remains within the maximum bound.\label{fig:control_input_geo}}{\includegraphics[width=0.4\textwidth]{figures/control_input_geo.pdf}}
+    The control always remains within the maximum bound.\label{fig:control_input_geo}}{\includegraphics[width=0.4\textwidth]{control_input_geo.pdf}}
 
@@
     Each stage of the transfer serves to increase the energy towards the stable manifold. 
-The final reachability set intersects the stable manifold and enables a large final manuever to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.4\textwidth]{figures/geo_transfer/jacobi_energy.pdf}}
+The final reachability set intersects the stable manifold and enables a large final manuever to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.4\textwidth]{jacobi_energy.pdf}}
 

### 2018-05-24 manuscript.tex
Images in top level directory

@@
         \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}\label{fig:compare_energy}}
-        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
+        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{cputimevsE.pdf}}~
     \label{fig:integrator_compare} 
@@
 \begin{figure}
-    \subcaptionbox{Position space view of reachability sets\label{fig:varying_tf_trajectory}}{\includegraphics[width=0.5\textwidth]{figures/l1_transfer/trajectory_varying_um.pdf}}~
-    \subcaptionbox{\Poincare section view of reachability sets\label{fig:varying_tf_poincare}}{\includegraphics[width=0.5\textwidth]{figures/l1_transfer/poincare_section_varying_um.pdf}}
+    \subcaptionbox{Position space view of reachability sets\label{fig:varying_tf_trajectory}}{\includegraphics[width=0.5\textwidth]{trajectory_varying_um.pdf}}~
+    \subcaptionbox{\Poincare section view of reachability sets\label{fig:varying_tf_poincare}}{\includegraphics[width=0.5\textwidth]{poincare_section_varying_um.pdf}}
     \caption{Variation of \( t_f \) and \( u_m\) on the reachability set. 
@@
         \begin{subfigure}[htbp]{0.5\textwidth} 
-            \includegraphics[width=\textwidth, keepaspectratio]{figures/l1_transfer/jacobi.pdf} 
+            \includegraphics[width=\textwidth, keepaspectratio]{jacobi.pdf} 
             \caption{Jacobi Energy over transfer \label{fig:jacobi_l1}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_trajectory_zoom.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage1_trajectory_zoom.pdf} 
         \caption{Stage 1 Trajectory~\label{fig:stage1_trajecotry_zoom}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_poincare.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage1_poincare.pdf} 
         \caption{Stage 1 \Poincare~section \label{fig:stage1_poincare}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_trajectory_zoom.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage2_trajectory_zoom.pdf} 
         \caption{Stage 2 Trajectory~\label{fig:stage2_trajecotry_zoom}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_poincare.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage2_poincare.pdf} 
         \caption{Stage 2 \Poincare~section \label{fig:stage2_poincare}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_trajectory_zoom.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage3_trajectory_zoom.pdf} 
         \caption{Stage 3 Trajectory~\label{fig:stage3_trajecotry_zoom}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_poincare.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage3_poincare.pdf} 
         \caption{Stage 3 \Poincare~section \label{fig:stage3_poincare}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage5_trajectory_zoom.pdf} 
         \caption{Stage 4 Trajectory~\label{fig:stage4_trajecotry_zoom}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage4_poincare.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage4_poincare.pdf} 
         \caption{Stage 4 \Poincare~section \label{fig:stage4_poincare}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage5_trajectory_zoom.pdf} 
         \caption{Stage 5 Trajectory~\label{fig:stage5_trajecotry_zoom}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_poincare.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage5_poincare.pdf} 
         \caption{Stage 5 \Poincare~section \label{fig:stage5_poincare}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_trajectory_zoom.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage6_trajectory_zoom.pdf} 
         \caption{Stage 6 Trajectory~\label{fig:stage6_trajecotry_zoom}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_poincare.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage6_poincare.pdf} 
         \caption{Stage 6 \Poincare~section \label{fig:stage6_poincare}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_trajectory_zoom.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage7_trajectory_zoom.pdf} 
         \caption{Stage 7 Trajectory~\label{fig:stage7_trajecotry_zoom}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_poincare.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage7_poincare.pdf} 
         \caption{Stage 7 \Poincare~section \label{fig:stage7_poincare}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_trajectory_zoom.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage8_trajectory_zoom.pdf} 
         \caption{Stage 8 Trajectory~\label{fig:stage8_trajecotry_zoom}} 
@@
     \begin{subfigure}[htbp]{0.5\textwidth} 
-        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_poincare.pdf} 
+        \includegraphics[width=\textwidth, keepaspectratio]{stage8_poincare.pdf} 
         \caption{Stage 8 \Poincare~section \label{fig:stage8_poincare}} 
@@
     \subcaptionbox{Control Input: Combined control history for each stage of the reachability analysis. 
-    The control always remains within the maximum bound.\label{fig:control_input_geo}}{\includegraphics[width=0.45\textwidth]{figures/control_input_geo.pdf}}
+    The control always remains within the maximum bound.\label{fig:control_input_geo}}{\includegraphics[width=0.45\textwidth]{control_input_geo.pdf}}
 
@@
     Each stage of the transfer serves to increase the energy towards the stable manifold. 
-The final reachability set intersects the stable manifold and enables a large final maneuver to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.5\textwidth]{figures/geo_transfer/jacobi_energy.pdf}}
+The final reachability set intersects the stable manifold and enables a large final maneuver to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.5\textwidth]{jacobi_energy.pdf}}
 

### 2018-08-26 Response.tex

@@
 
-\section*{Second round of responses to the reviewers' comments for JASS-D-17-00005}
+\section*{Third round of responses to the reviewers' comments for JASS-D-17-00005}
 
@@
 \begin{itshape}
-This is a vastly improved iteration of the paper and I thank the authors for taking the time to improve it.
-I commend the authors for responding to all of the feedback, especially modifying figures such as Figure 2, which I know takes a lot of time.
-I was also happy to see the intentional reductions in the paper to make it more focused, and the fact that the authors reconciled their low-thrust transfers with existing technology.
-I still have a few issues that persist based on the current submission, but the paper is far closer to being acceptable for publication.
+    I appreciate all of the authors' hard work in organizing two very detailed revisions for this journal submission.
+I've reviewed the second round of revisions and I believe that all of my concerns have been addressed.
+My opinion is that this paper is worthy of publication by JAS.
+Thank you for the interesting read.
 \end{itshape}
+
+Thank you for the comments.
+
+\subsection*{Reviewer 2}
+\begin{itshape}
+    I thank the authors for their significant additional improvements relative
+    to the previous version of the manuscript, as well as for the organized
+    structure of their response.  Many of my previous points have been
+    addressed to a satisfactory degree.  The expositional sections of the paper
+    are in much better shape and there are a number of clarifications and
+    better acknowledgment of assumptions, pros/cons etc.  The parametric
+    analysis in the first example is perhaps a bit more limited than I hoped
+    for, but is still a good addition.
+\end{itshape}
+
+Thank you for the comments.
 
@@
         \begin{itshape}
-            Section 2.3
+            The primary remaining issue for me stems from my previous remark that
+            appears at the end of page 10 of the most recent response document.  The
+            author response on page 13 of that document has not resolved my concerns,
+            but rather it has made the nature of the persisting issue more clear to me
+            (e.g. through the clarification that the Jacobi energy is not identical for
+            all points on the (x,xdot) plots.)
 
-            "This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge-Kutta methods."
-
-            "Potentially lower state accuracy" is still overselling.
-Just because your second order method is bounded on an energy surface does not guarantee that the state is accurate.
-It is likely that a 2nd order integrator in the PCRTBP contains more state error than a high-order RK which is the standard for modern trajectory planning in flight operations.
-I think it's best to focus on the primary advantage that the variational integrator provides reduced computational demands.
+            I'll first discuss this in terms of merely minor issues for Sec 4.1.  This
+            uses a single stage of reachability analysis, and the initial and final
+            orbits have different energy levels.  However, the corresponding discussion
+            in the middle paragraph of page 15 remarks upon an "intersection" between
+            the unstable manifold of the initial orbit (E=-1.58 from Fig 9e) and the
+            final orbit (E=-1.62 from Fig 9e).  The middle of page 18 then contains a
+            correct description that acknowledges the Jacobi differences, which are
+            handled by the boundary values of the optimization problem.  The following
+            paragraph then again describes that "a small portion of the invariant
+            manifold intersects with the target Moon orbit" and alludes that a
+            "secondary impulsive maneuver" would be necessary to complete the transfer.
+            This is the first place where the implications of (x,xdot) intersections at
+            different energy levels are actively discussed, and it seems conceptually
+            odd to consider targeting impulsive transfers in this way since they do not
+            require xdot to be matched as a precondition.  So, Sec 4.1 would benefit
+            from a more clear treatment of the ydot/Jacobi dimension, but the results
+            are still valid, and the Jacobi plot implies a sensible solution (first
+            raise energy to travel faster, then lower energy to match the target).
         \end{itshape}
 
-    The manuscript has been modified to accurately discuss the drawbacks of variational integrators.
-    The section has been modified to highlight the reduced computational load of the variational integrator.
-    In addition, an additional sentence has been added specifically highlighting the fact that high order Runge-Kutta methods will probably provide better state accuracy at the expense of additional computations and poorer energy behavior.
-
+        The authors appreciate these comments. 
+        As suggested Section 4.1 is completely revised to clarity the following properties of the proposed approach.
+        \begin{itemize}
+            \item Due to the change of the Jacobi integral under low-thrust propulsion, the intersection between the rechability set and the stable manifold to the desired periodic orbit is computed in a lower-dimensional space, specifically excluding $\dot y$.
+            \item Once the intersection is found in the lower dimensional Poincare section, the missing value of $\dot y$ is computed from the Jacobi integral values of the desired periodic orbit. 
+                This specifies the complete target state among the desired period orbit.
+            \item Finally, an optimal control problem is solved with the target state defined in the previous step.
+        \end{itemize}
+        More specifically, the statements pointed out in the review are revised, and the following paragraph is added. 
         \begin{correction}
-        This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
-        The variational integrator is constructed according to a discrete version of Hamilton's principle, specifically for the given dynamic system, and it provides long-term structural stability in capturing the effects of low-thrust propulsion systems.
-        General purpose numerical integrators, such as adaptive stepsize Runge-Kutta methods may yield a more accurate state trajectory depending on the choice of integration parameters, but at the expense of additional computational demand and total energy deviations.
-
-    \end{correction}
-    \item \begin{itshape}
-            Response to feedback item 34:
-
-            I disagree that it is abnormal to cite convergence statistics, especially for a new optimization approach or methodology.
-Yes, there are some dependencies to the specific example and use-case, but this is a paper in which researchers may wish to reproduce your results as a learning exercise to build on this research.
-Some expectation as to what they can expect for the results is therefore of pragmatic interest.
-
-            See for example, Table 2 in
-
-            Ellison, D., et al., M., "Application and Analysis of Bounded-Impulse Trajectory Models with Analytic Gradients," Journal of Guidance, Control, and Dynamics, 2018.
-
-            https://arc.aiaa.org/doi/pdf/10.2514/1.G003078
-
-            There are many other paper examples that I could cite that do this, but this is just one.
-        \end{itshape}
-
-        Some additional statistics have been included in the manuscript for both examples.
-        \begin{correction}
-            Convergence statistics associated for this transfer are shown in~\cref{tab:l1_transfer_stats}.
-            \begin{table}
-                \centering
-                \begin{tabular}{llr}  
-                    \toprule
-                    Metric    & Value \\
-                    \midrule
-                    \texttt{fsolve} objective      & \num{6.03e-15}      \\
-                    \texttt{fsolve} major iterations       & \num{9}      \\
-                    \texttt{fsolve} first order optimality & \num{1.88e-13} \\
-                    Optimal cost       & \num{2.09e-31}      \\
-                    Executation time & \SI{1.49}{\second}       \\
-                    \bottomrule
-
-                \end{tabular}
-                \caption{Convergence statistics for the periodic orbit transfer\label{tab:l1_transfer_stats}}
-            \end{table}
-
-        \end{correction}
-
-        And also additional statistics are provided for the geo-stationary transfer.
-        \begin{correction}
-            The optimization statistics for the final transfer are shown in~\cref{tab:geo_transfer}.
-            \begin{table}[h]
-                \centering
-                \begin{tabular}{llr}  
-                    \toprule
-                    Metric    & Value \\
-                    \midrule
-                    \texttt{fsolve} objective      & \num{1.42e-11}      \\
-                    \texttt{fsolve} major iterations       & \num{18}      \\
-                    \texttt{fsolve} first order optimality & \num{7.2e-11} \\
-                    Optimal cost       & \num{4.30e-25}      \\
-                    Executation time & \SI{2.62}{\second}       \\
-                    \bottomrule
-                \end{tabular}
-                \caption{Convergence statistics for the geostationary orbit transfer\label{tab:geo_transfer}}
-            \end{table}
-
-        \end{correction}
-
-        In addition, some additional detail was included to highlight the fact that exact results will vary based on the choice of nonlinear solver.
-
-        \begin{correction}
-            In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm with a convergence tolerance of \num{1e-5}.
-            Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
-            All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
-            The numerical results are specific to the choice of nonlinear solver, and different tolerances or software tools may result in slight changes.
-        \end{correction}
-    \item \begin{itshape}
-Conclusions:
-
-"The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints."
-
-I don't agree with the statement.
-Often, indirect methods are exceptionally difficult to append path constraints onto.
-In contrast, direct methods such as collocation approaches, are very amenable to the formulation of path constraints.
-    \end{itshape}
-
-    Thank you for the comments. This sentence has been modified.
-    \begin{correction}
-        The indirect optimal control formulation directly solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
-        Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution and typically results in a much larger parameter set.
-    \end{correction}
-\end{itemize}
-\subsection*{Reviewer 2}
-
-\begin{itshape}
-I thank the authors for their thorough and organized rebuttal.
-The revised manuscript itself feels greatly improved in terms of its structure and coverage of background material.
-Nonetheless I have some remaining qualms about the introduction (ISSUE 1), and some more serious concerns regarding the analysis/illustration of the developed method (ISSUE 2).
-\end{itshape}
-
-\begin{itshape}
-ISSUE 1: introduction (requires minor revisions): 
-Despite some reduction, the intro is still long-winded.  For example,
-\end{itshape}
-
-The introduction has been reduced in accordance with the recommendations below.
-Several paragraphs have been removed or refactored to reduce repetition and extraneous detail.
-The updated introduction is included below and specific responses for each comment are also provided.
-
-\begin{itemize}
-
-    \item
-        \begin{itshape}
-            It's superfluous to explain at such length that spacecraft trajectory design is a problem of interest, that CubeSats are a thing, etc.
-Broad points like these could potentially be condensed into single, citation-heavy sentences.
-        \end{itshape}
-
-        Several introductory paragraphs have been combined to reduce the extraneous detail.
-
-    \item 
-        \begin{itshape}
-            There is also still significant repetition. For instance, all the paragraph-to-paragraph transitions on page 3 feel very similar.
-        \end{itshape}
-
-        Several paragraphs have been combined or removed entirely. 
-        The transitions are also updated to reduce repetition.
-
-    \item 
-        \begin{itshape}
-            Some paragraphs are far too long
-        \end{itshape}
-
-        The paragraphs in the introduction have been reworked to reduce length, repetition and superfluous information.
-
-    \item 
-        \begin{itshape}
-            I would like to reiterate my previous suggestion to organize the intro into subsections.
-This will help you weed out repetition and will make the paper more reader-friendly instead of a wall of text.
-        \end{itshape}
-
-        The introduction has been reduced and divided into a subsection.
-
-    \item 
-        \begin{itshape}
-            The instability of RK for decades-to-centuries propagations seems irrelevant or misleading given the time scales used in this paper.
-        \end{itshape}
-
-        This comment has been removed.
-
-    \item 
-        \begin{itshape}
-            The point brought up about "suboptimal direct optimization" versus the authors' approach isn't really translated into anything meaningful in the body of the paper.
-Along with the RK vs variational point, I feel like the route you've chosen has some nice properties but shouldn't necessarily be sold as a crucial enabler; the key of the paper lies elsewhere.
-        \end{itshape}
-
-        The sentences have been removed from the manuscript.
-        The updated introduction is included below.
-
-        \begin{correction}
-            \section{Introduction}
-
-            %%-------------- Motivation for low thrust orbital transfers
-
-            % enables long duration missions with minimized propellant usage since Isp is so high
-
-            % Enables different mission scenarios which are not possible through the use of conventional chemical propulsion (high thrust) systems
-            Designing spacecraft trajectories is a classic and ongoing topic of research.
-            There has been significant research into the design of orbital transfers for space vehicles.
-            Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
-            Electric propulsion systems offer a much greater specific impulse than chemical systems.
-            As a result, the greatly increased efficiency allows for greater payload mass or extended duration missions.
-            However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital maneuvers have a much longer time of flight.
-            In spite of this drawback, a wide variety of missions, such as communication and deep space probes, have utilized the unique benefits of low thrust electric propulsion to great effect~\cite{choueiri2009}.
-
-            % utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
-            Recent developments in miniature electric propulsion and small satellites now offer the potential for new research opportunities~\cite{folta2015,haque2013}.
-            The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}.
-            There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011,koon2011,ross2006}. 
-            Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
-            The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.
-            Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
-            Conventional general-purpose Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, may fail to preserve geometric properties of the dynamic system numerically.
-            Variational integrators can allow for reduced computational effort without any loss of numerical stability or energy drift which exist in conventional integration schemes.
-
-            % --------------Issues/Challenges of previous work
-
-            % use of direct optimal control
-            % chaotic systems and initial guesses for optimization
-            % numerical integration accuracy
-            % preserving dynamical structure
-
-            % control-free trajectories in the three-body problem 
-            % objectives of this work
-
-            % description of our work - our approach
-
-            % concise summary of contributions
-
-            %-------------------- OBJECTIVES OF THIS WORK-----------------------------%
-            % SYSTEMATIC DESIGN TO AVOID DIFFICULTIES IN CHOOSING A GOOD INITIAL GUESS
-            % CAPTURE LONG TERM EFFECTS OF LOW THRUST ACCURATELY IN NUMERICAL SIMULATION
-            \subsection{Objective and Contribution}
-            In this paper, we propose a systematic design method which enables low-thrust transfers in the planar circular restricted three-body problem.
-            We utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers. 
-            The reachable set, defined as the set of all attainable states subject to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
-            Defining the reachable set on a \Poincare section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
-            Through the use of low-thrust control input, the reachable set on the \Poincare section is enlarged and enables a larger space of potential transfers.
-            By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare section, we generate general transfer trajectories.
-            With this proposed method, the previous research on control-free trajectories will be generalized with the addition of low thrust propulsion systems.
-
-            In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
-            This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
-            % This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
-            % For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
-            In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
-            Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
-            We avoid the issues inherent in selecting a valid initial condition for optimization.
-            Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
-            We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.
-
-        \end{correction}
-\end{itemize}
-
-\begin{itshape}
-ISSUE 2: analysis/illustration (requires major revisions): 
-My other highlighted area of concern with the original manuscript was the degree of (in)completeness in illustrating the method and its pros and cons.
-I do not feel that the authors have done very much to improve this aspect of the paper beyond adding some caveats that feel a little half-hearted rather than properly contextualized discussions.
-This aspect of the research remains a major shortcoming that must be addressed before publication.
-
-While my request for improvements was itself not very specific, I had hoped that some of my other questions would reveal some areas of consideration for the authors.
-Addressal/discussion of those questions furthermore needs to appear adequately in the paper, not just the rebuttal.
-Below I will try to sketch how these points of inquiry can lead to a more demonstrative analysis.
-\end{itshape}
-\newline\newline
-Thank you for the comments and the additional suggestions to improve the manuscript. 
-The main objective of the submitted manuscript is to construct a systematic approach to design low-thrust orbital maneuvers. 
-It is motivated by the specific challenges that electric propulsion systems generate a minuscule level of thrust, and low-thrust orbital maneuvers rely on those effects accumulated over a long time period.
-Consequently, it is quite cumbersome and challenging to apply numerical optimization tools or boundary value problem solvers successfully to low-thrust orbital design problems. 
-For example, it is difficult to initialize the optimization parameters such that the resulting numerical iteration converges successfully. 
-Conventional numerical integration schemes have shown to be possibly erroneous particularly in correctly computing long-term energy behaviors.
-
-In the submitted manuscript, we aim to provide more organized tools to mitigate such difficulties by explicitly considering the characteristics of the natural dynamics and utilizing geometric integrators exhibiting long-term numerical stability.
-The consideration of the natural, control-free dynamics is reasonable as low-thrust trajectories won't be drastically different from those over a limited time period. 
-Here, the former is achieved by incorporating the control-free maneuvers characterized by the \Poincare section in the invariant manifolds, and the latter is achieved by variational integrators.
-
-However, the proposed scheme is focused on constructing low-thrust trajectories satisfying the prescribed boundary conditions, and it does not guarantee any optimality in the constructed maneuvers. 
-It is heuristically implied that the constructed orbital maneuver would require no-excessive control input, as it is essentially constructed by adjusting the control-free trajectory. 
-As such, it is rather focused on constructing a reasonable low-thrust trajectory that satisfies the given constraints on the maximum thrust bound and the boundary conditions.
-
-In short,
-\begin{itemize}
-	\item[Pros] The proposed scheme utilizes the characteristics inherent in the three-body problem in constructing low-thrust orbital maneuvers in a systematic way. 
-		This reduces the cumbersome process of guessing a \textit{good} set of initial parameters to make the direct application of optimization successful. 
-	\item[Pros] The use of geometric integrators is particularly useful in capturing the long-term effects of low thrust in a computationally efficient manner. 
-	\item[Cons] While the proposed scheme generates orbital trajectories to the target with prescribed constraints, it does not guarantee any optimality. 
-\end{itemize}
-
-In the revised manuscript, the conclusions section has been rewritten to summarize the above pros and cons of the proposed approach. 
-\begin{correction}
-In this paper, we construct a systematic approach which combines the concepts of reachability sets and \Poincare~sections to generate orbital transfers between planar periodic orbits in the three-body problem using low-thrust propulsion systems.
-The \Poincare~section allows for trajectory design on a lower dimensional phase space thereby reducing the complexity and search space.
-The approximation of the reachability set enables a simple methodology for the selection of \textit{good} initial parameters which will allow for convergence of the optimal control formulation.
-The combination of these two concepts allows the engineer to capture the characteristics inherent in the three-body problem.
-Additionally, we utilize geometric integrators to efficiently capture the long-term effects of low-thrust on the system dynamics.
-Our approach enables the computation of orbital transfers within the three-body problem but does not guarantee optimality of the resulting transfer.
-Instead, the resulting transfers will require no excessive control inputs as it is formulated as a deviation from the control-free trajectory.
-The presented approach allows for a systematic method to determine and generate orbital transfers in the three-body problem using low-thrust propulsion.
-
-There is additional research to extend these results to more general transfer scenarios in future work.
-The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
-In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems and future research is aimed at investigating smaller magnitude control inputs.
-Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
-This will result in a more complicated optimal control problem and is a focus of future research.
-Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
-The addition of attitude dynamics and realistic pointing constraints would also significantly improve the applicability of this work.
-
-\end{correction}
-
-Also, the revised manuscript incorporates the suggested changes and additions.
-We include several new figures to further illustrate the change in Jacobi energy.
-In addition, we add some additional examples demonstrating the effect of variations in the control bound and terminal time on the reachability set.
-These are itemized below.
-
-\begin{itemize}
-    \item 
-        \begin{itshape}
-            The assumption of fixed travel time.
-The authors respond that this choice is made to simplify the optimization problem.
-Of course this is the case, and is conventional, but this choice is made in a rather arbitrary fashion rather than properly analyzed (especially considering that a central purpose of the problem setup is to enable faster travel times).
-Imagine plotting low-thrust reachability curves as in Fig8b for each of a sequence of travel time choices.
-This would reveal the influence of the arbitrarily handled parameter, and show that there's some critical point at which the reachablility curve intersects the targeted manifold.
-Maybe exhaustive exploration of this dimension is not appropriate in every application, but the relationship is worth illustrating.
-        \end{itshape}
-
-        Thank you for the comment. 
-        We agree that some further discussion of the effect of these ``tunable'' parameters would help to improve the manuscript. 
-		According to the comments, we considered twelve additional cases for every combination of varying terminal time $t_f$ and the maximum control $u_m$ listed at Table \ref{fig:varying_tf_reachability_sets}.
-        An additional section has been included which highlights some of the effects of variations in terminal time, and control magnitude (next comment).
-        The results presented are not an exhaustive description of the search space (in terms of travel time and/or control magnitude) but do highlight some of the important characteristics of these parameters on the resulting reachability set.
-		According to the new numerical results, the manuscript has been revised to discuss the effects of varying $t_f$ and $u_m$ as copied below.
-
-        \begin{correction}
-            Next, we determine the reachability set with addition of a low-thrust control input over a fixed time horizon.
-            In~\cref{sssec:periodic_orbit_varying_tf}, we demonstrate the effect of variations in the choice of variations of maximum control bound and terminal time.
-            While~\cref{sssec:periodic_orbit_transfer} show a specific example of a transfer design process using the intersection of the reachability set on the \Poincare section.
-            The analysis presented in the following sections define a maximum magnitude of the thrust as \( u_{m} \) and assume that the trust can be directed arbitrarily within the plane. 
-            This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
-            Assuming a fully actuated spacecraft model allows us to decouple the translation and rotational dynamics of the spacecraft.
-
-            \subsubsection{Reachability Set on the \Poincare section}
-            From the initial state on the periodic orbit, a series of optimal trajectories are generated to determine the reachable set.
-            The multiple shooting approach is implemented to solve the optimal control problem. 
-            We divide the time horizon into two equal length segments. 
-            The state trajectory is initialized using the free trajectory of the periodic orbit. 
-            Similarly, the costate trajectory is initialized from an initial guess of \( \vecbf{\lambda}_0 = \begin{bmatrix} -1 & -1 & -1 & -1\end{bmatrix}^T\) and propagated using the discrete equations of motion in~\cref{eq:costate_eom}.
-            This results in the initial guess of the design vector \( \vecbf{\chi} = \begin{bmatrix} \vecbf{\lambda}_0 & \vecbf{x}_1^{-} & \vecbf{\lambda}_1^{-} & \vecbf{\beta} \end{bmatrix}^T\).
-            This design vector is then varied to ensure that the necessary conditions of optimality and the interior point constraints are satisfied. 
-            \Cref{tab:varying_tf} shows the range of terminal times, \( t_f \), and maximum control bound, \( u_m \), which are used to investigate their effect on the subsequent reachable set on the \Poincare section.
-            \begin{table}
-                \centering
-                \begin{tabular}{ll}  
-                    \toprule
-                    \(t_f\) & \( u_m \) \\
-                    \midrule
-                    1.24 & 0.05 \\
-                    1.30 & 0.25 \\
-                    1.37 & 0.5 \\
-                    1.44 & \\
-                    \bottomrule
-                \end{tabular}
-                \caption{Combinations for \( t_f \) and \( u_m\) used to generate the reachability sets in~\cref{fig:varying_tf_reachability_sets}}
-            \end{table}
-
-            The angle \( \theta_d\) in~\cref{eq:constraints} is varied to select a different direction along the \Poincare section to maximize.
-            We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
-            Choosing a new angle \( \theta_d \) corresponds to a different direction as well as a new optimal control problem which is again solved using the multiple shooting approach laid out previously.
-            Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
-            Each solution on the reachability set is computed in approximately \SI{2}{\minute} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
-            The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:varying_tf_reachability_sets}.
-
-            \begin{figure}
-                \subcaptionbox{Position space view of reachability sets}{\includegraphics[width=0.5\textwidth]{trajectory_varying_um.pdf}}~
-                \subcaptionbox{\Poincare section view of reachability sets}{\includegraphics[width=0.5\textwidth]{poincare_section_varying_um.pdf}}
-                \caption{Variation of \( t_f \) and \( u_m\) on the reachability set. 
-                    Colors, \braces{red, blue, green, yellow}, are used note increasing \( t_f\) while markers, \braces{circle, square, cross}, are used to denote increasing \(u_m\).
-                    Increasing the maximum control bound has a large effect and enables the reachability set to intersect the target manifold. 
-                Increases in \(t_f\) are less critical and have minimal impact on the distribution of the reachability set on the \Poincare section.}
-            \end{figure}
-
-            \Cref{fig:varying_tf_reachability_sets} shows the reachabilty set for each combination of \(t_f\) and \( u_m\) listed in~\cref{tab:varying_tf}.
-            Assuming a small control bound of \( u_m = \num{0.05} \) is shown in ~\cref{fig:varying_tf_reachability_sets} using square markers with the various terminal times shown using different colors.
-            With a small maximum control bound, \( u_m = \num{0.05} \), the reachable set is not dramatically changed from that of the no control solution.
-            The reachable set is shown using square markers on the left most portion of \cref{fig:varying_tf_poincare}.
-            Variations of \( t_f\) are indicated using different colors and also demonstrate that this parameters has a smaller impact than changes in \( u_m \).
-            Increasing the control bound to \( u_m = \num{0.25} \) and \( u_m = \num{0.5} \) shows that the reachable set is progressively approaching the target set and there are several intersections.
-
-            The reachable sets presented in~\cref{fig:varying_tf_reachability_sets} are highly dependent on the initial condition, terminal time, and maximum control bound. 
-            This example demonstrates that increasing the \( u_m \) results in a larger displacement between the controlled and uncontrolled trajectories.
-            The reachable set is enlarged from a single point, as shown in~\cref{fig:manifold_poincare} by the black point, to a larger region as shown in~\cref{fig:varying_tf_poincare}.
-            The choice of \( u_m\), \( t_f \) and initial condition all combine to change the resulting reachability set.
-
-        \end{correction}
-    \item 
-        \begin{itshape}
-            The specification of a departure point from the periodic orbit.
-In response to this question, the authors describe how the use of low-thrust propulsion enlarges the reachable set about some "center" defined by one branch of the uncontrolled manifold.
-This is a key way to think about the approach you've posed but it is not illustrated in the paper.
-Why not generate some curves under different thrust bounds for a specific departure point/manifold branch?
-Again you'd be using a sidelined parameter to illustrate something fundamental about the approach.
-And just maybe there would be a point at which, with high thrust, you start to see "bang-bang" type control and establish more connection to intuition.
-        \end{itshape}
-
-        Thank you for the comment.
-        We have included an additional section describing the effect of variations in \( u_m \) and terminal time on the reachable set.
-        As discussed above, the results are not an exhaustive exploration of the parameter space but illustrate the key effect of the control bound and terminal time.
-        In the example presented, increasing the magnitude of the control has the effect of enlarging and moving the reachable set. 
-        At a sufficiently large value of \( u_m \) the reachable set is able to intersect the target orbit.
-
-    \item 
-        \begin{itshape}
-            Actually, the influence of the starting point selection *is* somewhat illustrated by the green unstable manifold of Fig7b/8b.
-The way a mission designer might think about applying this tool could be to start with this information as a baseline and then use principles revealed by the preceding two types of illustration to make some informed way of selecting a combination of [departure point, travel time, max thrust] to get the sets to intersect without oversizing their system.
-That said, it's a little confusing how the red and green curves of Fig8b don't line up with this description to fit the comment of P18L47; additional consideration/analysis may be needed with regards to the number of revs/crossings used to define a section for comparison.
-        \end{itshape}
-
-        Thank you for the comment.
-        As stated in the previous comments we have included additional discussion on the effect of variations of terminal time and control bound.
-
-        \begin{itshape}
-That said, it's a little confusing how the red and green curves of Fig8b don't line up with this description to fit the comment of P18L47; additional consideration/analysis may be needed with regards to the number of revs/crossings used to define a section for comparison.
-        \end{itshape}
-
-        Thank you for the question. 
-        The green points are the intersection of the unstable manifold with the \Poincare section while the red points are the reachability set.
-        Without any control input the spacecraft would remain on the periodic orbit and intersect the \Poincare section at \( x_n\). 
-        The addition of the control input allows for an expansion of the reachable set from a single point to the larger red region.
-        An additional clarification has been added to the manuscript to aid in the description of the figure.
-
-        \begin{correction}
-            In contrast, the low thrust control input we are able to enlarge the reachability set from a single point, \( x_n\) associated with the periodic orbit, to a larger ellipsoidal region shown in red in~\cref{fig:poincare_compare}.
+            TO DO
         \end{correction}
@@
         \begin{itshape}
-            I previously mentioned my dissatisfaction with the interpretability of the control solution (which is plotted in planar components).
-On further consideration, I think a time series plot of the Jacobi energy would be beneficial to help vet whether the method is working reasonably.
-From intuition, one might expect an efficient solution to exhibit a span of monotonic energy increase, a relatively constant coasting span, and then a monotonic decrease back to the Poincare section value.
-As is, the two roughly constant thrusting phases visible in the plot suggest that this might be happening, but it's hard to be sure.
-Less consistent behavior might indicate, for example, that a longer-than-necessary transfer time was used.
+            The ydot/Jacobi treatment becomes much more problematic in Sec 4.2.  Its
+            second paragraph again describes achieving an "intersection" of the
+            reachability set and the target orbit, and an intermediate criterion of
+            minimizing the (x,xdot) distance toward the target. Page 22 line 15 again
+            describes the objective as enlarging the reachability set on the (x,xdot)
+            plane to include the intersections of the target manifold, and the
+            walk-through of Fig 11 discusses closing the gap in terms of (x,xdot)
+            alone.  In Sec 4.1, it was presuambly the case that the reachable set
+            projected onto (x,xdot) always had the Jacobi value of the target orbit.
+            This can't be the case for Figs 11-12.  Further, the uncontrolled dynamics
+            are in no way represented or accounted for (which, for example, cause a big
+            jump between stages 3 and 4 that is not due to the influence of control).
+            If I'm interpreting this right, the final maneuver is solely to make up for
+            the large remaining distance in Jacobi energy after (x,xdot) are already
+            matched.
+
+            What is really going on with this method is that we are traversing a
+            (x,xdot,E) cube from a point on the E=E0 side to a manifold on the opposite
+            side where E=Etgt, and at each crossing of y=0 the reachable set is a 3D
+            volume.  For Sec 4.1, we only care about the intersection of our reachable
+            set with the E=Etgt manifold, and so the 2D (x,xdot) reasoning makes sense.
+            But the multi-use approach is only reasoning about the (x,xdot) plane for a
+            single horizon with fixed transfer time, presumably with the E dimension
+            being reduced to whichever values maximize (x,xdot) distance, in which case
+            the red and blue points in Figs 11-12 might all have different energy
+            levels.  Fig 13c seems to disprove that the (x,xdot) distance is a good
+            heuristic for receding-horizon use of this scheme, since the resulting path
+            to the manifold is nothing at all like a straight line to the closest part
+            of it.  Furthermore, the spread of these points is about as much as the
+            spread of the reachable set in some individual stages such as stage 6.  Fig
+            13e is also not indicative of an efficient transfer; to me, the highly
+            nonmonotonic behavior suggest that a lot of fuel is being wasted to ensure
+            that y=0 is crossed at a rather arbitrary sequence of times.
+
+            As painful as it is to prolong the review process any more, I cannot
+            approve publication of the paper considering the extent of the problems
+            with the example of 4.2 (barring that the remaining reviewer and the editor
+            identify some way in which I am mistaken or have missed something).  The
+            authors have been very accommodating in making extensive updates to the
+            organization and writing of the paper, but have not adequately addressed
+            matters relating to the fundamental contributions. What it would take for
+            me to enthusiastically accept this manuscript is either for the mutli-rev
+            use of the method to be implemented in a better-reasoned way with a more
+            demonstrably good solution (I'm not sure how) or for example 4.2 to be
+            dropped and for the single-use case to be analyzed in more depth (some
+            steps were taken toward that in this revision). At a very minimum, all the
+            aspects I have mentioned need to at least be acknowledged as limitations
+            of the method/as being counter to some of the assumptions of the 
+            formulation.
         \end{itshape}
 
-        Additional plots have been included in the manuscript which shows the evolution of the Jacobi energy integral over the transfer horizon.
+        The authors agree with most of these comments, and we appreciate the thoughtful insights. 
+        As stated in the previous response, the intersection between the reachability set and the stable manifold to the desired periodic orbit is computed in the lower-dimensional space, and as such, the energy error in the last stage should be corrected by an additional optimization scheme.
 
-        The additional plot and some additional discussion for the first periodic orbit transfer is presented below.
+        However, the example in Section 4.2 illustrates that the proposed approach can be linked together to compute low-thrust transfers for more challenging maneuvers, with the requirement of correcting energy error at the last stage.
+        Such issue remains in the single stage example in Section 4.1. 
+        If needed in the multi stage examples, the Jacobi energy can be corrected in the intermediate steps to reduce the energy gap in the last stage.
+
+        This paper is motivated by the current common practice of computing low thrust orbital transfers, where a large set of initial guess is repeatedly optimized in a computationally intensive manner.
+        The main contribution is presenting an alternative framework of constructing low thrust orbital transfers in a more efficient and systematic fashion. 
+        And, the example in Section 4.2 demonstrates such capability.
+
+        In the revised manuscript, the following paragraph is added in Section 4.2 to explicitly describe the requirement of energy correction at the last stage, its implication and potential difficulties.
+
         \begin{correction}
-            \begin{figure} 
-                \centering 
-                \begin{subfigure}[htbp]{0.4\textwidth} 
-                    \includegraphics[width=\textwidth]{reach_trajectory} 
-                    \caption{Position space view of reachability set} \label{fig:reach_trajectory} 
-                \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
-                \begin{subfigure}[htbp]{0.4\textwidth} 
-                    \includegraphics[width=\textwidth]{poincare_compare} 
-                    \caption{\Poincare section view of reachability set} \label{fig:poincare_compare} 
-                \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
+            TODO
+        \end{correction}
 
-                \begin{subfigure}[htbp]{0.4\textwidth} 
-                    \includegraphics[width=\textwidth]{reach_transfer} 
-                    \caption{Transfer trajectory selected from reachability set viewed in the position space} \label{fig:reach_transfer} 
-                \end{subfigure}~ 
-                \begin{subfigure}[htbp]{0.4\textwidth} 
-                    \includegraphics[width=\textwidth]{control_input_l1} 
-                    \caption{Control input for the selected transfer trajectory} \label{fig:control_l1} 
-                \end{subfigure}~
-
-                \begin{subfigure}[htbp]{0.4\textwidth} 
-                    \includegraphics[width=\textwidth, keepaspectratio]{jacobi.pdf} 
-                    \caption{Jacobi Energy over transfer \label{fig:jacobi_l1}} 
-                \end{subfigure}
-
-                \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
-                    The reachability set is shown in the upper two figures in both the position and \Poincare space.
-                    From this reachability set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
-                The optimal control to achieve this transfer is shown in the lower right figure.}
-                \label{fig:reachability_set_transfer} 
-            \end{figure}
-
-            The Jacobi energy integral, computed using~\cref{eq:jacobi}, is shown in~\cref{fig:jacobi_l1}.
-            \Cref{fig:jacobi_l1} shows that the vehicle begins with an energy level equal to the periodic orbit and arrives at the target orbit with the appropriate energy.
-            The first half of the transfer is associated with and increase in energy as the control is used to transition towards the target orbit which is followed by a energy decrease to the target orbit.
-            This roughly corresponds with the expected optimal solution of a bang-coast-bang type orbital transfer.
-
-        \end{correction}
-        The additional plot for the geostationary orbit transfer is presented below.
-        \begin{correction}
-
-Each stage of the transfer serves to raise the energy level of the vehicle.
-After eight stages the reachability set intersects the stable manifold and a demonstrates that a transfer is achievable.
-The final optimal control drives the vehicle towards the target manifold with the appropriate energy level.
-\begin{figure} 
-        \centering 
-        \subcaptionbox{Transfer trajectory: Complete trajectory consisting of eight iterations (1-7 black and final in red) of the reachability sets with a final control-free coast (blue)  on the stable manifold (green).\label{fig:geo_transfer_full}}{\includegraphics[width=0.4\textwidth]{geo_transfer_full}}\hfill
-        \subcaptionbox{Detailed view of transfer: Centered at \(m_1\) the eight
-            iterations from the reachability analysis are shown in black.  The
-            final segment to the stable manifold is shown in red.  Each
-            iteration progressively approaches the stable manifold.  Once on
-        the stable manifold the vehicle can coast without thrust towards the
-    periodic orbit.\label{fig:geo_transfer_zoom}
-}{\includegraphics[width=0.4\textwidth]{geo_transfer_zoom}}
-
-\subcaptionbox{\Poincare section: Minimum distance states from each iteration of the reachability set analysis (red) with respect to the stable invariant manifold (green)\label{fig:geo_transfer_poincare}}{\includegraphics[width=0.4\textwidth]{poincare}}
-    \hfill
-    \subcaptionbox{Control Input: Combined control history for each stage of the reachability analysis. 
-    The control always remains within the maximum bound.\label{fig:control_input_geo}}{\includegraphics[width=0.4\textwidth]{control_input_geo.pdf}}
-
-    \subcaptionbox{Jacobi Energy: Energy history throughout the transfer.
-    Each stage of the transfer serves to increase the energy towards the stable manifold. 
-The final reachability set intersects the stable manifold and enables a large final manuever to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.4\textwidth]{jacobi_energy.pdf}}
-
-        \caption{Geostationary to \( L_1 \) periodic orbit transfer: Complete transfer from the geostationary orbit to the stable manifold.\label{fig:geo_transfer}}
-\end{figure}
-        \end{correction}
-    \item 
-        \begin{itshape}
-            Following that thought, I have serious concerns about chaining this Poincare-based approach, as done in the second problem and as I previously asked about.
-The authors' responded: "While the trajectories do return to the energy level of the Poincare section, there is a transfer of "energy" between the states of the vehicle." Frankly, this statement doesn't make sense to me.
-I think that you are actually getting a very inefficient solution by forcing these repeated (and at fixed intervals) returns to the same energy level (again, plot the Jacobi time series and see if you can square it with your intuition).
-Furthermore posing the problem as a transfer to the stable manifold kind of hamstrings the motivation of reducing transfer times.
-Honestly, I think this second example should simply be axed.
-If the first one is expanded enough to illustrate the problem fundamentals, it will be sufficient.
-Conducting the same kind of analysis at a couple different rev counts would
-            reveal the dimension that this second example added, and help reveal the limits of the approach under certain scenarios and thrust capabilities.
-        \end{itshape}
-
-		As pointed out by the reviewer, our prior response stating `..the trajectories do return to the energy level of the Poincare section..' is erroneous and misleading. 
-        Instead, the combination of several reachability sets does not force the system to return to a prescribed energy level.
-        Each reachability set computation serves to maximize a distance metric, defined in a reduced phase space corresponding to the \Poincare section.
-        The cost function is computed as the distance between the terminal states of the controlled and uncontrolled trajectories.
-        The additional plots shown in both examples highlight the fact that the Jacobi energy is continuously changing as the control input is used to maneuver the vehicle towards the desired target states.
-
-        The transfer to the stable manifold is presented as an example of the combination of existing control free  transfer methodologies with low thrust control inputs.
-        This example demonstrates one possible use of low-thrust control to enable a larger region of feasibility for the use of invariant manifold transfers in the three body problem.
-
-		With regards to the comments on the second example, the authors decided to keep them in the manuscript, as it illustrates the idea of multiple stages to construct orbital maneuvers with more challenging scenarios.
-		Nonetheless, the revised manuscript includes extended numerical results and discussions for the first example, for example to illustrate the effects of varying $t_f$ and $u_m$ and the evolution of the Jacobi constant.
-    \item 
-        \begin{itshape}
-            To summarize my thoughts on ISSUE 2:
-            As they exist, the examples illustrate that solutions can be found but they do not enlighten the reader as to the characteristics of the problem formulation/tool that should be understood and kept in mind when applying it or when planning related research efforts.
-Furthermore, I am very suspicious that Example 2 is not actually a good solution or a compelling use case for the approach.
-I do not necessarily ask or expect the authors to follow my above outline precisely, but I insist that some changes occur to make the analysis/illustrations better convey the general nature of the tool, not just that it gives you an answer.
-        \end{itshape}
-
-        Thank you for the comments.
-        The manuscript has been edited, as described previously, to incorporate several changes to illustrate the specific characteristics of this approach.
-        Also, an additional section is presented which highlights the effect of terminal time and control magnitude on the reachability set.
-
-        The second example presents a larger and more involved example of combining several iterations of the reachable set to achieve a larger and more complex transfer. 
-        This aspect of the proposed method, more specifically of combining several reachability sets to approach a target, is not demonstrated in the first example.
-        As a result, we feel that the second example does add some additional insight into the use of this method for a more complicated transfer.
-		Furthermore, the revised conclusions section summarizes the objective of the submitted manuscript, and the pros and the cons of the proposed approach.
+        In the current practice of academic publication, if a new idea is presented in a manuscript and both of the desirable properties and the potential issues are illustrated explicitly, it is considered to be worth of sharing.
+        The example in Section 4.2 demonstrates an interesting application of the proposed approach, and more importantly, it also illustrates a case where the limitation of the proposed approach can be problematic.
+        As such, it provides candid, explicit characterization of the proposed approach, as well as topics for prospective research. 
+        The authors do not agree with the comments that Section 4.2 should be excluded due to the deficiencies. 
+        Instead, the limitations of the proposed approach illustrated by Section 4.2 are explicitly stated in the revised manuscript.
 
@@
         \begin{itshape}
-            A more collected and focused discussion about how the parameter choices, assumptions, and limitations of the tool should inform its use is also warranted; currently remarks to this effect feel like scattered afterthoughts.
-One organized way to think about this is to distinguish all the dimensions of the problem and identify what is being solved rigorously, what is being set manually, what is being numerically explored, etc.
+
+            Given the significant remaining issues above, I am going to keep further
+            feedback a bit more brief:
+
+            * As mentioned before, at a high level, transferring onto the stable
+            manifold kind of undercuts the premise of using low thrust to reduce the
+            transfer time
         \end{itshape}
 
-        Thank you for the helpful comments to improve the manuscript.
-        The manuscript has been edited to reduce the length and repetition of the introduction.
-        In addition, an additional section is presented highlighting the effect of the variations of terminal time and control magnitude on the reachable set. 
-        Furthermore, the presented examples also include additional plots showing the change of the Jacobi energy during the transfer. 
+        In this paper, the natural, uncontrolled dynamics represented by stable manifold are considered as a good starting point for low thrust orbital transfers, as orbital maneuvers with low thrust over a short time period would not differ from those substantially.
+        But, as pointed out by the reviewer, the proposed approach does not impose any optimality in the constructed maneuver. 
+        The scope of this paper is constructing low thrust orbital maneuvers without relying on massive computations. 
+        To clarify these, the revised manuscripts states that there is no optimality guaranteed.
+
+    \item
+        \begin{itshape}
+
+            * The extra analysis behind Fig 8 is a step in the right direction, but in
+            practice it is hard to make sense of, and might require additional
+            figures/better color choices.  Why are all the reachable sets nearly
+            straight lines, when we expected some kind of closed path to arise from
+            sweeping 0 < theta < 360?  Why are some progressions of sets so irregular
+            (e.g. red cross --> blue cross --> green cross --> yellow cross signifying
+            increase in travel time with fixed large umax, which results in darting
+            back and forth in the x dimension)?  The 25 lines of discussion do not
+            acknowledge these oddities, nor do they seem to completely line up with the
+            figure.  There seem to remain unexplored/unexplained relationships that
+            would in practice need to be used to justify parameter choices by anyone
+            applying this method.   
+        \end{itshape}
+
+    \item
+        \begin{itshape}
+
+            * As mentioned before, the Poincare illustrations of Fig 2 and 3 would be
+            more helpful if they were more general.  The first could benefit from
+            showing nonperiodic dynamics, the second from showing a non-circular
+            reachable boundary (I know the text acknowledges this but please update the
+            figure).  
+        \end{itshape}
+
+    \item
+        \begin{itshape}
+
+            * I wonder under what circumstances the reachable set might intersect a
+            given $\theta_d$ in multiple places.  This is something that could be looked
+            into to expand the strength of a submission focusing only on the
+            single-horizon use of the method.
+        \end{itshape}
+
+    \item 
+        \begin{itshape}
+
+            * There are still some very long single paragraphs in the paper.
+        \end{itshape}
+
 \end{itemize}
-% \bibliography{library}
+
+\bibliography{library}
 % \bibliographystyle{spmpsci}

### 2018-08-31 Response.tex
Turn on bibliography
Updating manuscript for reviewer commetns

@@
         \end{itemize}
-        More specifically, the statements pointed out in the review are revised, and the following paragraph is added. 
+        More specifically, the statements pointed out in the review are revised, and the following paragraphs are modified in the manuscript. 
         \begin{correction}
-            TO DO
+            \Cref{fig:poincare_compare} shows that the reachable set and those of the descending target region intersect.
+            As both regions are discretely approximated a linear interpolation is used to determine the exact intersection state on the \Poincare section.
+            This intersection generates a partial target state of \( x_t, \dot{x}_t \) and \( y = 0 \) from the \Poincare section.
+            Using the energy level of the target region, defined by~\cref{eq:jacobi}, and the intersection state we can calculate the final component \( \dot{y} \). 
+            This results in a complete target state \( \vecbf{x}_t \) which lies on the reachable set and on the target orbit. 
+            Due to the use of low-thrust propulsion, there is a change in Jacobi energy during the transfer as the vehicle transitions between the initial and final periodic orbits. 
+            This behavior is demonstrated in~\cref{fig:jacobi_l1} as the vehicles experiences an increase in energy transitioning towards the target followed by a decrease to arrive at the target.
+
+            A final optimal trajectory is generated such that the \( \vecbf{x}(N) = \vecbf{x}_t \).
+            This transfer serves to satisfy a required boundary condition to ensure a continous trajectory to the target. 
+            The reachability set, and the intersection on the \Poincare section, are used to generate a partial target state for the transfer. 
+            Due to the difference in Jacobi energy, a final optimal transfer is computed to satisfy the desired target state.
+            This transfer trajectory is denoted by the green path in~\cref{fig:reach_transfer}.
+            The optimal control input is shown in~\cref{fig:control_l1}. 
+            The spacecraft achieves the desired target while satisfying the bounded control input.
+            The Jacobi energy integral, computed using~\cref{eq:jacobi}, is shown in~\cref{fig:jacobi_l1}.
+            \Cref{fig:jacobi_l1} shows that the vehicle begins with an energy level equal to the periodic orbit and arrives at the target orbit with the appropriate energy.
+            The first half of the transfer is associated with and increase in energy as the control is used to transition towards the target orbit which is followed by a energy decrease to the target orbit.
+            This roughly corresponds with the expected optimal solution of a bang-coast-bang type orbital transfer~\cite{kirk2012}.
+            Convergence statistics associated for this transfer are shown in~\cref{tab:l1_transfer_stats}.
         \end{correction}
@@
         As stated in the previous response, the intersection between the reachability set and the stable manifold to the desired periodic orbit is computed in the lower-dimensional space, and as such, the energy error in the last stage should be corrected by an additional optimization scheme.
+        The stable manifold is utilized a convienent target set for the example scenario. 
+        Furthermore, there is additional benefit for this type of transfer as the stable manifold does not pass sufficiently close to the Earth for a direct injection, i.e. from Low Earth Orbit.
 
         However, the example in Section 4.2 illustrates that the proposed approach can be linked together to compute low-thrust transfers for more challenging maneuvers, with the requirement of correcting energy error at the last stage.
-        Such issue remains in the single stage example in Section 4.1. 
+        This example demonstrates a possible approach to linking several computations of the reachability set to achieve a larger and more challenging transfer.
+        There is still a difference in Jacobi energy which is corrected in the final stage.
         If needed in the multi stage examples, the Jacobi energy can be corrected in the intermediate steps to reduce the energy gap in the last stage.
@@
         The main contribution is presenting an alternative framework of constructing low thrust orbital transfers in a more efficient and systematic fashion. 
-        And, the example in Section 4.2 demonstrates such capability.
+        The example in Section 4.2 demonstrates the use of several intersections of the reachability set to determine a transfer.
 
@@
         \begin{correction}
-            TODO
+            This numerical example demonstrates the ability to link several computations of the reachability set to enable a more general transfer.
+            We use eight iterations of computing the reachability set in order to transfer from the geostationary orbit to the stable manifold.
+            The example shows a transfer between an initial geostationary orbit and the stable manifold associated with a periodic orbit. 
+            This approach allows for a larger class of potential transfers which leverage the capabilities of low-thrust propulsion systems.
+            However, the presented approach does not offer a optimal solution in terms of energy/fuel usuage. 
+            The reachability sets are computed on the lower dimensional \Poincare section.
+            As a result, the repeated intersections of the trajectory with the \Poincare section results in a fuel inefficient transfer.
+            However, We are able to achieve a transfer that is not possible via the invariant manifold alone as manfiold does not pass sufficiently close to the Earth.
+            In addition, this example illustrates a straightforward method to depart from the natural dynamics and transfer to a large region of the phase space that is not accessible via invariant manifolds alone.
         \end{correction}
@@
 
+        The various reachable sets appear as straight lines due to the discretization used and the level of zoom of the figure.
+        In reality the reachable sets are highly elongated ellipses on the \Poincare section.
+        This is due to a combination of factors, namely that the low thrust propulsion cannot greatly deviate from the natural dynamics and that the nature of the optimization results in approximate numerical solutions subject to the choice of tolerance levels.
+        In addition, as the figure shows theres is a nontrivial relationship between terminal time and control magnitude. 
     \item
@@
 \bibliography{library}
-% \bibliographystyle{spmpsci}
+\bibliographystyle{spmpsci}
 \end{document}

### 2018-08-31 manuscript.tex
Updating manuscript for reviewer commetns

@@
 \title{Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem
-\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
+\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.
+The authors certify that they have no affiliations with or involvement in any organization or entity with any financial interest, or non-financial interest  in the subject matter or materials discussed in this manuscript.}
 }
@@
 As both regions are discretely approximated a linear interpolation is used to determine the exact intersection state on the \Poincare section.
-This intersection generates a partial target state of \( x_t \text{ and } \dot{x}_t \).
+This intersection generates a partial target state of \( x_t, \dot{x}_t \) and \( y = 0 \) from the \Poincare section.
 Using the energy level of the target region, defined by~\cref{eq:jacobi}, and the intersection state we can calculate the final component \( \dot{y} \). 
 This results in a complete target state \( \vecbf{x}_t \) which lies on the reachable set and on the target orbit. 
+Due to the use of low-thrust propulsion, there is a change in Jacobi energy during the transfer as the vehicle transitions between the initial and final periodic orbits. 
+This behavior is demonstrated in~\cref{fig:jacobi_l1} as the vehicles experiences an increase in energy transitioning towards the target followed by a decrease to arrive at the target.
+
 A final optimal trajectory is generated such that the \( \vecbf{x}(N) = \vecbf{x}_t \).
+This transfer serves to satisfy a required boundary condition to ensure a continous trajectory to the target. 
+The reachability set, and the intersection on the \Poincare section, are used to generate a partial target state for the transfer. 
+Due to the difference in Jacobi energy, a final optimal transfer is computed to satisfy the desired target state.
 This transfer trajectory is denoted by the green path in~\cref{fig:reach_transfer}.
@@
 We use eight iterations of computing the reachability set in order to transfer from the geostationary orbit to the stable manifold.
+The example shows a transfer between an initial geostationary orbit and the stable manifold associated with a periodic orbit. 
 This approach allows for a larger class of potential transfers which leverage the capabilities of low-thrust propulsion systems.
-We are able to achieve a transfer that is not possible via the standard invariant manifold approach. 
+However, the presented approach does not offer a optimal solution in terms of energy/fuel usuage. 
+The reachability sets are computed on the lower dimensional \Poincare section.
+As a result, the repeated intersections of the trajectory with the \Poincare section results in a fuel inefficient transfer.
+However, We are able to achieve a transfer that is not possible via the invariant manifold alone as manfiold does not pass sufficiently close to the Earth.
 In addition, this example illustrates a straightforward method to depart from the natural dynamics and transfer to a large region of the phase space that is not accessible via invariant manifolds alone.

### 2018-08-31 Response.tex
Update manuscript

@@
         This is due to a combination of factors, namely that the low thrust propulsion cannot greatly deviate from the natural dynamics and that the nature of the optimization results in approximate numerical solutions subject to the choice of tolerance levels.
-        In addition, as the figure shows theres is a nontrivial relationship between terminal time and control magnitude. 
+        In addition, as the figure shows there is a nontrivial relationship between terminal time and control magnitude. 
+        The resulting reachability set is heaviliy dependent on the specific problem, choice of initial conditions, terminal time, and control bound.
+        The figure is meant to illustrate some large scale, general properties of the approach, namely that an incrase in terminal time and/or control magnitude will have the effect of enlarging and moving the resulting reachable set.
+     The intention is not to present an exhaustive search/discussion of this specific situation but rather to illustrate the effect of variations in terminal time and maximum control magnitude.
+        The revised manuscript explicitly states this:
+        \begin{correction}
+            The reachable sets presented in~\cref{fig:varying_tf_reachability_sets} are highly dependent on the initial condition, terminal time, and maximum control bound. 
+            This example demonstrates that increasing the \( u_m \) results in a larger displacement between the controlled and uncontrolled trajectories.
+            The reachable set is enlarged from a single point, as shown in~\cref{fig:manifold_poincare} by the black point, to a larger region as shown in~\cref{fig:varying_tf_poincare}.
+            The choice of \( u_m\), \( t_f \) and initial condition all combine to change the resulting reachability set. 
+        \end{correction}
     \item
@@
         \end{itshape}
+        The illustrations of the \Poincare section match the typical depication in many other publications~\cite{koon2001}.
+        Furthermore, the illustrations are merely utilized as helpful visualizations of the concepts, specifically periodic solutions, lower dimensional \Poincare sections, and reachable sets, rather than specific depictions of a solution.
+        An additional comment is added to the caption specifically mentioning that the reachable set can be any arbitrary shape, and potentially non-continuous.
 
@@
 
+        Thank you for the comment.
+        This is an interesting question and a possible avenue for future research.
+
     \item 
@@
 
+        Thank you for the comment. 
+        The manuscript has been revised and the long single paragraphs have been divided.
 \end{itemize}

### 2018-08-31 manuscript.tex
Update manuscript

@@
 Both integration schemes result in trajectories that are initially nearly identical.
+
 \Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
@@
 In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
+
 This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
@@
             We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
-            Computation of the maximum distance, or reachability, for a variety angles gives a discrete approximation of the reachability set.\label{fig:reachability_set}}
+            Computation of the maximum distance, or reachability, for a variety angles gives a discrete approximation of the reachability set.
+        In general, the reachable set can be an arbitrary shape on the section rather than the circular set depicted.\label{fig:reachability_set}}
 \end{figure}

### 2018-09-04 Response.tex
tylee edit

@@
             This approach allows for a larger class of potential transfers which leverage the capabilities of low-thrust propulsion systems.
-            However, the presented approach does not offer a optimal solution in terms of energy/fuel usuage. 
+            However, the presented approach does not offer a optimal solution in terms of energy/fuel usage. 
             The reachability sets are computed on the lower dimensional \Poincare section.
             As a result, the repeated intersections of the trajectory with the \Poincare section results in a fuel inefficient transfer.
-            However, We are able to achieve a transfer that is not possible via the invariant manifold alone as manfiold does not pass sufficiently close to the Earth.
-            In addition, this example illustrates a straightforward method to depart from the natural dynamics and transfer to a large region of the phase space that is not accessible via invariant manifolds alone.
+            Furthermore, the issue of discrepancy in the Jacobi integral may become more serious when multiple reachability sets are interconnected. 
+            Nonetheless, we are able to apply the proposed approach for the challenging case where the reachability set is not sufficiently close to the invariant manifold to the target orbit. 
+            Consequently, this example illustrates a straightforward method to depart from the natural dynamics and transfer to a large region of the phase space that is not accessible via invariant manifolds alone.
+            The presented idea of connecting reachability sets can be further studied to address the aforementioned issues of optimality and mismatch of the Jacobi integral.
         \end{correction}
@@
         But, as pointed out by the reviewer, the proposed approach does not impose any optimality in the constructed maneuver. 
-        The scope of this paper is constructing low thrust orbital maneuvers without relying on massive computations. 
+        The scope of this paper is constructing low thrust orbital maneuvers without relying on massive computations that appear in the common practice. 
+        We aim to deliver a more systematic method for low thrust orbital design, which is to satisfy the given boundary conditions. 
         To clarify these, the revised manuscripts states that there is no optimality guaranteed.
@@
         \end{itshape}
+
         The illustrations of the \Poincare section match the typical depication in many other publications~\cite{koon2001}.

### 2018-09-04 Response.tex
Update response with tylee edits

@@
             Furthermore, the issue of discrepancy in the Jacobi integral may become more serious when multiple reachability sets are interconnected. 
-            Nonetheless, we are able to apply the proposed approach for the challenging case where the reachability set is not sufficiently close to the invariant manifold to the target orbit. 
+            Nonetheless, we are able to apply the proposed approach for the challenging case where the reachability set is not sufficiently close to the invariant manifold of the target orbit. 
             Consequently, this example illustrates a straightforward method to depart from the natural dynamics and transfer to a large region of the phase space that is not accessible via invariant manifolds alone.

### 2018-09-04 manuscript.tex
Update response with tylee edits

@@
 This approach allows for a larger class of potential transfers which leverage the capabilities of low-thrust propulsion systems.
-However, the presented approach does not offer a optimal solution in terms of energy/fuel usuage. 
+However, the presented approach does not offer a optimal solution in terms of energy/fuel usage. 
 The reachability sets are computed on the lower dimensional \Poincare section.
 As a result, the repeated intersections of the trajectory with the \Poincare section results in a fuel inefficient transfer.
-However, We are able to achieve a transfer that is not possible via the invariant manifold alone as manfiold does not pass sufficiently close to the Earth.
-In addition, this example illustrates a straightforward method to depart from the natural dynamics and transfer to a large region of the phase space that is not accessible via invariant manifolds alone.
+Furthermore, the issue of discrepancy in the Jacobi integral may become more serious when multiple reachability sets are interconnected. 
+Nonetheless, we are able to apply the proposed approach for the challenging case where the reachability set is not sufficiently close to the invariant manifold of the target orbit. 
+Consequently, this example illustrates a straightforward method to depart from the natural dynamics and transfer to a large region of the phase space that is not accessible via invariant manifolds alone.
+The presented idea of connecting reachability sets can be further studied to address the aforementioned issues of optimality and mismatch of the Jacobi integral.
 
